\documentclass[oneside]{article}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/nguyen/1_NQBH/reference/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{\small \textsc{Sect.} ~\thesection}
\rhead{\small \nouppercase{\leftmark}} % \nouppercase !
\renewcommand{\sectionmark}[1]{\markboth{#1}{}}
\cfoot{\thepage}
\def\labelitemii{$\circ$}

\title{Lundy Bancroft. \textit{Why Does He Do That?: Inside the Minds of Angry \& Controlling Men}}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
\tableofcontents

%------------------------------------------------------------------------------%

\begin{quotation}
    ``This fascinating investigation into what makes abusive men tick is alarming, but its candid handling of a difficult subject makes it a valuable resource for professionals and victims alike$\ldots$
    
    Jargon-free analysis is frequently broken up by interesting 1st-person accounts and boxes that distill in-depth information into simple checklists.
    
    Bancroft's book promises to be a beacon of calm for many storm-tossed families.'' - Publishers Weekly (starred review)
    \\
    
    ``Most books about abuse in relationships focus on women - how they're hurt, why they stay.
    
    As important as these questions are, they can also distract us from the heart of the problem.
    
    Bancroft boldly asks - and brilliantly answers - the most important questions of all: Why do so many men abuse women?
    
    What can be done about it?
    
    This book is desperately needed and long overdue.'' - Jackson Katz, creator of the award-winning video \textit{Tough Guise: Violence, Media \& the Crisis in Masculinity}
    
    ``Bancroft, a former codirector of Emerge, the 1st U.S. program for abusive men, and a 15-year veteran of work with abusive men, reminds readers that each year in this country, 2--4 million women are assaulted by their partners and that at least 1 out of 3 American women will be a victim of violence by a husband or boyfriend at some point in her life.
    
    His valuable resource covers early warning signs, 10 abusive personality types, the abusive mentality, problems with getting help from the legal system, and the long, complex process of change$\ldots$
    
    This is essential reading for those in the helping professions and highly recommended.'' - Library Journal (starred review)
    
    ``At last - the straight scoop on men who abuse women.
    
    This is a book not just for abused women and domestic violence professionals, but for everyone who wonders why there's so much violence in America.
    
    Read it.'' - Anne Jones, author of \textit{When Love Goes Wrong} \& \textit{Next Time She'll Be Dead}
    
    ``Bancroft helps women who feel trapped in unhealthy relationships make sense out of what is happening.'' - Sarah Buel, J.D., codirector, Domestic Violence Clinic, and lecturer, University of Texas Law School
    
    ``A compelling read about a tough topic.
    
    What you read here will come back to you long after you put the book down.'' - Angela Browne, author of \textit{When Battered Women Kill}
    
    ``An informative and necessary read.'' - Susan Weitzman, Ph.D., author of \textit{Not to People Like Us: Hidden Abuse in Upscale Marriages}
\end{quotation}
To the thousands of courageous women, many of them survivors of abuse themselves, who have created and sustained the movement against the abuse of women, and to the many men who have joined this struggle as allies.

\section*{Acknowledgments}
Bancroft has had many, many teachers along his path to understanding the mentality and behavior of abusive men.

The survivors of abuse have been Bancroft's greatest educators; if we could hear their voices much more, and the voices of the abusers and their allies much less, the world would move rapidly to eliminate the chronic mistreatment that so many women currently face in their intimate relationships.

%
Equally important to the growth of Bancroft's understanding of abusive men, and of their impact on their partners and children, was Carole Sousa, who simultaneously educated Bancroft at Emerge and kept them honest.

\fbox{Her criticisms of Bancroft's blind spots were often annoying, mostly because of how right they were.}

\section*{Note on Terminology}
In referring to angry and controlling men in this book, Bancroft has chosen to use in most cases the shorter terms \textit{abusive man} and \textit{abuser}.

Bancroft has used these terms for readability and not because Bancroft believes that every man who has problems with angry of controlling behaviors is abusive.

Bancroft needed to select a simple word Bancroft could apply to any man who has recurring problems with disrespecting, controlling, insulting, or devaluing his partner, whether or not his behavior also involves more explicit verbal abuse, physical aggression, or sexual mistreatment.

Any of these behaviors can have a serious impact on a woman's life and can lead her to feel confused, depressed, anxious, or afraid.

So even if your partner is not an abuser, you will find that much of what is described in the pages ahead can help to clarify for both of you the problems in your relationship and what steps you can take to head in a more satisfying, supportive, and intimate direction.

If you are not sure whether your partner's behavior should be called \textit{abuse} or not, turn to Chap. 5, which will help you sort out the distinctions.

%
At the same time, remember that even if your partner's behavior doesn't fit the definition of abuse, it may still have a serious affect on you.

Any coercion or disrespect by a relationship partner is an important problem.

Controlling men fall on a \textit{spectrum} of behaviors, from those who exhibit only a few of the tactics Bancroft describes in this book to those who use almost all of them.

Similarly, these men run a gamut in their attitudes, from those who are willing to accept confrontation about their behaviors and strive to change them, to those who won't listen to the woman's perspective at all, feel completely justified, and become highly retaliatory if she attempts to stand up for herself.

(In fact, as we see in Chap. 5, 1 of the best ways to tell how deep a man's control problem goes is by seeing how he reacts when you start demanding that he treats you better.

If he accepts your grievances and actually takes steps to change what he does, the prospects for the future brighten somewhat.)

The level of anger exhibited by a controlling man also shows wide variation, but unfortunately it doesn't tell us much in itself about how psychologically destructive he may be or how likely he is to change, as we will see.

%
In addition, Bancroft has chosen to use the terms \textit{he} to refer to the abusive person and \textit{she} to the abused partner.

Bancroft selected these terms for convenience and because they correctly describe the great majority of relationships in which power is being abused.

However, control and abuse are also a widespread problem in lesbian and gay male relationships, and the bulk of what Bancroft describes in this book as relevant to same-sex abusers.

\section*{Introduction}
Bancroft has been working with angry and controlling men for 15 years as a counselor, evaluator, and investigator, and have accumulated a wealth of knowledge from the 2000 or more cases with which Bancroft has been involved.

Bancroft has learned the warning signs of abuse and control that a woman can watch out for early in a relationship.

Bancroft has come to know what a controlling man is \textit{really} saying, the meaning that is hidden behind his words.

Bancroft has seen clues to recognizing when verbal and emotional aggression are heading toward violence.

Bancroft has found ways to separate out abusive men who are faking change from those who are doing some genuine work on themselves.

And Bancroft has learned that \fbox{the problem of abusiveness has surprisingly little to do with how a man \textit{feels}} - his clients actually differ very little from nonabusive men in their emotional experiences - and \fbox{everything to do with how he \textit{thinks}.}

\fbox{The answers are inside his mind.}

%
However, as delighted as Bancroft is to have had the opportunity to gain this insight, \textit{Bancroft is not 1 of the people who most needs it}.

The people who can best benefit from knowledge about abusers and how they think are \textit{women}, who can use what Bancroft has learned to help themselves recognize when they are being controlled or devalued in a relationship, to find ways to get free of abuse if it is happening, and to know how to avoid getting involved with an abusive man - or a controller or a user - next time.

The purpose of this book is to \fbox{equip women with the ability to protect themselves, physically and psychologically}, from angry and controlling men.

%
To prepare for writing this book, Bancroft 1st generated a list of the 21 questions that women most often ask him about their abusive partners, questions e.g.:
\begin{quotation}
    ``\textit{Is he really sorry?}''
    
    ``\textit{Why do so many of our friends side with him?}''
    
    ``\textit{Is he going to hit me some day?}''
\end{quotation}
and many others.

Bancroft then built his explanations around these concerns, to make sure that women would be able to look here to find the information they urgently need.

You will find these 21 questions highlighted as you go through this book; you might want to flip through the pages for a moment now just to grab a quick glimpse of where Bancroft has addressed the issues that are most pressing for you.

%
Another central goal of his is to offer assistance to each woman who is struggling with how she is being treated in a relationship, regardless of what label she may put on her partner's behavior.

Words like \textit{control} and \textit{abuse} can be loaded ones, and you may not feel that they fit your particular circumstances.

Bancroft has chosen to use the term \textit{abusers} to refer to men who use a wide range of controlling, devaluing, or intimidating behaviors.

In some cases Bancroft is talking about physical batterers and at other times about men who use or insult their partners but never frighten or intimidate them.

Some of the men Bancroft describes in the pages ahead change moods so drastically and so often that a woman could never feel sure \textit{what} they are like, much less attach a label.

Your partner may be arrogant, or may play mind games, or may act selfishly over and over again, but his better aspects may make you feel that he is miles away from being an ``abuser.''

Don't let Bancroft's language put you off; Bancroft has simply chosen the word \textit{abuser} as a shorthand way of saying ``men who chronically make their partners feel mistreated or devalued.''

You can adopt a different term if you know one that fits your partner better.

But whatever style of mistreated your partner uses, rest assured that you will find in these pages the answers to many questions that have perplexed you.

%
If the person you are involved with is the same sex you are, you have a place here too.

Lesbians and gay men who abuse their partners exhibit much of the same thinking, and most of the same tactics and excuses, that abusive heterosexual men do.

In this book Bancroft has used the term \textit{he} for the abuser and \textit{she} for the abused partner to keep his discussions simple and clear, but abused lesbians and gay men are very much in his thoughts, right alongside of abused straight women.

Of course, you will need to change the gender language to fit your relationship, for which Bancroft apologizes in advance.

You will also find a section in Chap. 6 where Bancroft speaks specifically about the similarities and differences in same-sex abusers.

%
Similarly, this book includes stories of men from a very wide range of racial and cultural backgrounds.

Although the attitudes and behaviors of controlling and abusive men vary somewhat from culture to culture, Bancroft has found that their similarities greatly outweigh their differences.

If your partner is a person of color or an immigrant, or if you are a member of 1 of these groups yourself, you will find that much of what this book discusses, or perhaps all of it, fits your experience quite well.

While Bancroft has not specified race or ethnicity in the cases Bancroft describes in these pages, roughly 1-3rd of the abusers whose stories Bancroft tells are men of color or men from nations outside of North America.

Bancroft further discusses some specific racial and cultural issues in Chap. 6.

\subsection{Bancroft's experience working with angry \& controlling men}
Bancroft began counseling abusive men individually and in groups in 1987, while working for a program called Emerge, the 1st agency in the United States to offer specialized services for men who abuse women.

Fo roughly the next 5 years Bancroft worked almost exclusively with clients who were coming to the program voluntarily.

They generally attended under heavy pressure from their female partners, who were either talking about leaving the relationship or had already done so.

In many cases, the woman had gone to court to seek a restraining order legally barring the man from the home and in many cases ordering him to stay away from the woman altogether.

The men's main motivation for seeking counseling was the hope of saving their relationships.

It was common for them to feel some guilt or discomfort about their abusive behavior, but they simultaneously believed strongly in the validity of their excuses and justifications, so their feelings of remorse would not have been enough in themselves to have kept them in Bancroft's program.

In those early years, the clients Bancroft worked with were men who used far more verbal and emotional abuse than physical violence, although most of them had been physically intimidating or assaultive on at least a few occasions.

%
During the 1990s the legal system became much more involved than it had been in the past in responding to domestic abuse, with the result that court-mandated clients started at 1st to trickle and then to pour in the doors of Bancroft's program.

These men often had a much greater propensity for physical violence than Bancroft's earlier clients, sometimes involving the use of weapons or vicious beatings resulting in the hospitalization of their partners.

Yet Bancroft observed that in other ways these men were generally not significantly different from Bancroft's verbally abusive clients: their attitudes and excuses tended to be the same, and they used mental cruelty side by side with their physical assaults.

Equally important was that the female partners of these battering men were largely describing the same distresses in their lives that Bancroft was hearing about from woman who had been psychologically abused, showing him that different forms of abuse have similar destructive impacts on women.

%
Throughout Bancroft's years of working with controlling and abusive men, his colleagues and he have been strict about always speaking to the woman whom their clients has mistreated, whether or not the couple is still together.

(And if he had started a new relationship, Bancroft talks with his current partner as well, which is part of how Bancroft became aware of the ways in which abusive men continue their patterns from 1 relationship to the next.)

It is through these interviews with women that Bancroft has received his greatest education about power and control in relationships.

The women's accounts also have taught Bancroft that abusive men present \textit{their own stories with tremendous denial, minimization, and distortion} of the history of their behaviors and that it is therefore otherwise impossible for Bancroft to get an accurate picture of what is going on in an abusive relationship without listening carefully to the abused woman.

%
\fbox{Counseling abusive men is difficult work.}

They are usually very reluctant to face up the damage that they have been causing women, and often children as well, and hold on tightly to their excuses and victim blaming.

As you will see in the pages ahead, they become attached to the various privileges they earn through mistreating their partners, and they have habits of mind that make it difficult for them to imagine being in a respectful and equal relationship with a woman.

%
Bancroft is sometimes asked: Why work with abusive men if it is so hard to get them to change?

The reasons are several.

1st, if even 1 man out of a 10-person group makes substantial and lasting changes, then Bancroft's time and energy have been invested well, because his partner and his children will experience a dramatic change in the quality of their lives.

2nd, Bancroft believes in holding abusers accountable for their actions.

If they attend an abuser program they can at least be called to task for the harm they have done, and Bancroft has hope (and see the signs) that cultural values can change over time if people find that men who chronically mistreat and degrade women are being made answerable.

3rd, and probably most important, is that Bancroft considers the woman that Bancroft's client has mistreated to be the person Bancroft is primarily serving, and Bancroft makes contact with her at least every few weeks.

Bancroft's goal is to give her emotional support, help her learn about counseling and legal services that exist for her in her community (usually for free), and help her get her mind untangled from the knot that her abusive partner has tied.

Bancroft can make it more difficult for him to manipulate her, and Bancroft may be able to warn her of underhanded maneuvers that he is planning or of escalation that Bancroft is observing.

As long as Bancroft stays focused on the woman and her children as those who are most deserving and in need of Bancroft's assistance, Bancroft can almost always make a positive contribution, whether or not Bancroft's abusive client decides to seriously face his own problem.

(In Chap. 14, Bancroft describes what actually goes on inside a counseling program for abusive men, and Bancroft explains how a woman can determine whether or not a particular program is being run properly.)

%
In recent years, through Bancroft's work as a child abuse investigator and as a custody evaluator for various courts, Bancroft has come to interact in a new way with families affected by abusive men.

Bancroft shares some of the insights Bancroft has gained through these experiences in Chap. 10, which examines the experiences of children who are exposed to abusive men - usually their fathers or stepfathers - and the ways in which some abusers continue their patterns of controlling and intimidating behavior through custody actions in the family courts.

\subsection{How to use this book}
1 of the prevalent features of life with an angry or controlling partner is that he frequently \textit{tells you what you should think} and tries to get you to doubt or devalue your own perceptions and beliefs.

Bancroft would not like to see your experience with this book re-create that unhealthy dynamic.

So the primary point to bear in mind as you read the pages ahead is to listen carefully to what Bancroft is saying, but always to think for yourself.

If any part of what Bancroft describes about abusers doesn't match your experience, cast it aside and focus on the parts that do fit.

You might even put the book down from time to time and ask yourself, ``How does this apply to \textit{my} relationship?

What are my own examples of how a controlling or cruel man thinks and behaves?''

If you come upon sections that don't speak to you - because you don't have children, e.g., or because your partner is never physically frightening - just skip ahead to the pieces that can help you more.

%
Some women will find that being alone with this book is too difficult because it awakens feelings and realizations that are overwhelming.

Bancroft encourages you to reach out for support from trusted friends and family as you go along.

While reading this book is likely to be clarifying for you, it may also awaken an awareness that can be painful or distressing.

%
If you can't find someone whom you know to talk to - or even if you can - call the abuse hotline in your area.

For many other options for assistance, look in the ``Resources'' section at the back of this book.

Again, don't be stymied by the word \textit{abuse}; the hot line staff is there to listen to you and to help you think about any relationship in which you are being treated in a way that is making you feel bad.

%
Bancroft understands how uncomfortable it can be to take the leap of talking with people you care about regarding the mistreatment you are experiencing in your relationship.

You may feel ashamed of having a partner who sometimes behaves in unkind or bullying ways, and you may fear that people will be critical of you for not leaving him right away.

Or you might have the opposite concern: that people around you are so fond of your partner that you question whether they will believe you when you describe how mean or abusive he can be.

But, regardless of these anxieties, it is essential \textit{not} to stay isolated with your distress or confusion about what is happening in your relationship.

Find someone whom you can trust - it might even be a person you have never considered opening up to before - and unburden yourself.

This is probably the single most critical step you can take toward building a life that is free from control or abuse.

%
If your partner's controlling or devaluing behavior is chronic, you no doubt find yourself about him a great deal of the time, wondering how to please him, how to keep him from straying, or how to get him to change.

As a result, you might find that you don't get much time to think about \textit{yourself} - except about what is wrong with you in his eyes.

1 of Bancroft's central reasons for writing this book is, \fbox{ironically, to help you think about him \textit{less}.}


Bancroft is hoping that by answering as many questions as possible and clearing away the confusion that abusive behavior creates, Bancroft can make it possible for you to escape the trap of preoccupation with your partner, so that you can put yourself - and your children if you are a mother - back in the center of your life where you belong.

\fbox{An angry and controlling man can be like a vacuum cleaner that sucks up a woman's mind and life}, but there are ways to get your life back.

The 1st step is to learn to identify what your partner is doing and why he does it, which is what the pages ahead will illuminate.

But when you have finished diving deeply into the abuser's mind, which this book will enable you to do, it is important to rise back to the surface and from then on \textit{try to stay out of the water as much as you can}.

Bancroft doesn't mean that you should necessarily leave your partner - that is a complex and highly personal decision that only you can make.

But whether you stay or go, the critical decision you can make is to \textit{stop letting your partner distort the lens of your life, always forcing his way into the center of the picture}.

You deserve to have your life be about \textit{you}; you are worth it.

\begin{center}
    Part I: The Nature of Abusive Thinking
\end{center}

\section{The Mystery}

\subsection{Listen to the voices of these women}
\begin{quotation}
    He's 2 different people. I feel like I'm living with Dr. Jekyll and Mr. Hyde
    
    He really doesn't mean to hurt me. He just loses control.
    
    Everyone else think he's great. I don't know what it is about me that sets him off.
    
    He's fine when he's sober. But when he's drunk, watch out.
    
    I feel like he's never happy with anything I do.
    
    He's scared me a few times, but he never touches the children. He's a great father.
    
    He calls me disgusting names, and then an hour later he wants sex. I don't get it.
    
    \fbox{He messes up my mind sometimes.}
    
    \fbox{The thing is, he really understands me.}
    
    Why does he \emph{do} that?
\end{quotation}
These are the words of women who are describing \fbox{their anxiety and inner conflict about their relationships.}

Each of these women knows that something is wrong - very wrong - but she can't put her finger on what it is.

Every time she thinks she's got her partner figured out, that she finally understands what is bothering him, something new happens, something changes.

The pieces refuse to fit together.

%
Each of these women is trying to make sense out of the roller-coaster ride that her relationship has become.

Consider Kristen's account:
\begin{quotation}
    When I 1st met Maury, he was the man I had dreamed of.
    
    \fbox{It seemed too good to be true.}
    
    He was charming, funny, and smart, and best of all, he was crazy about me.
    
    I opened up to him about hard things I'd been through over the previous few years, and he was so much on my side about it all.
    
    And he was so game for \textit{doing} things - whatever I wanted to do, he was up for it.
    
    The 1st year or so that we were together was great.
    
    %
    I can't say exactly when things started to change.
    
    I think it was around the time we started living together.
    
    It started with him saying he needed more space.
    
    I felt confused, because before that it had always seemed like \textit{he} was the one who wanted to be together every second.
    
    %
    Then he began to have more and more criticisms and complaints.
    
    %
    He would say that I talk on and on and that I'm \fbox{self-centered}.
    
    Maybe I am - it's true that I talk a lot.
    
    \fbox{But earlier it had seemed like he couldn't hear enough about me.}
    
    He started to say that I wasn't doing anything with my life.
    
    I know he has big ambitions, and maybe he's right that I should be more that way, but I'm happy with what I have.
    
    And then it was my weight.
    
    It started to seem like all the time he was saying that I needed to work out more, that I wasn't watching what I ate.
    
    \fbox{That hurt the most, to tell you the truth.}
    
    He seemed to want sex less and less often, and if I ever tried to be the one to initiate lovemaking, forget it.
    
    %
    We're still together, but I have a feeling he's going to leave me.
    
    I just can't seem to live up to what he needs.
    
    I'm trying, but he doesn't think so.
    
    And now when he's really angry or frustrated, he says things that cut me down.
    
    A few days ago he said, ``You're a lazy bitch, just looking for a man to live off of like your mother.''
    
    I don't get that; I've contributed a lot.
    
    I haven't worked the last 2 years since our baby was born, but I'm getting ready to go back to work soon.
    
    I don't think he really meant it, but still$\ldots$
    
    %
    He says I've changed a lot, but I'm not always so sure it's me.
    
    %
    Sometimes for a few days he seems like th guy I fell in love with, and I get hopeful, but then he slips away again into being so unhappy with me.
    
    I set him off somehow, but I don't know what I'm doing wrong.
\end{quotation}
Kristen was troubled by several questions.

What had happened to the man she had loved so much?

\fbox{Why was he always putting her down?}

What could she do to stop his explosions?

Why did he think \textit{she} was the one who had changed?

%
Other women tell stories that are quite distinct from Kristen's, but they feel just as confused as she does.

Here is what Barbara describes:
\begin{quotation}
    Fran is kind of quiet and shy.
    
    But he's cute as a button, and I got a crush on him the day I met him.
    
    I had to really go after him; it was hard to draw him out.
    
    We would go out and have great talks, and I couldn't wait to see him again.
    
    But 3 weeks would go by, and he'd say he hadn't been feeling well, or his sister was in town, or whatever.
    
    A couple of times he forgot dates we had.
    %
    
    Well, he finally opened up.
    
    It turned out he'd been really hurt before.
    
    He'd been cheated on a lot, and women had done some pretty mean things to him.
    
    He was afraid to get close again.
    
    %
    Little by little, he came around, but I was definitely the pursuer.
    
    I tried to show him that I wasn't like other women he's been with.
    
    I'm not flirtatious.
    
    I don't show my body off to other men; I'm just not that style.
    
    But Fran wouldn't believe it.
    
    He would always say that I was making eyes at a man at the next table, or that I was checking someone out who walked past us.
    
    I feel bad for him, he's so insecure.
    
    His mother cheated on his father when he was growing up, so I guess that's made it even worse.
    
    %
    I was eager to get married, because I thought then he'd feel secure that I was his, but he was very reluctant to commit.
    
    When we finally did tie the knot, he was more trusting for a while, but then the jealousy came back, and it's never left.
    
    I've asked him off and on for years to go see a therapist, but he gets really mad and says there's nothing wrong with him.
    
    %
    A few days ago we went to a birthday party for a friend of his, and I had this great conversation with his friend's brother.
    
    It was nothing but talking - I mean, the guy isn't even cute.
    
    Well, suddenly Fran was saying that we had to go home because he had a bad headache.
    
    On the drive home, it turned out the real reason was jealousy.
    
    He started yelling at me, saying he was sick of me humiliating him in front of other people, ``strutting your stuff,'' and on and on.
    
    He was pounding his fist on the dashboard, and 2 or 3 times he shoved me up against the car door.
    
    Each time that I told him it wasn't true he would go through the roof, so I stopped saying that.
    
    Our children were sitting in the backseat; it scared the daylights out of them.
    
    %
    At my age, it's hard to think about leaving him.
    
    Starting all over now seems so hard.
    
    I just wish he would get some help.
\end{quotation}
Barbara was struggling with issues different from Kristen's.

Why couldn't Fran trust her, and why was he isolating her from other people?

Why couldn't he see that he had a problem, and get help?

Was he going to hurt her badly some day?

Would her life ever get better?

%
At 1st look, Maury and Fran sound nothing like each other: One is young, popular, energetic, and assertive; the other is socially awkward, passive, and easily hurt.

Fran is physically violent sometimes, whereas Maury is not.

But are they as different as they seem?

Or do they both actually have the same set of issues under the surface, driving their behavior?

These are some of the questions for which we will find answers in the chapters ahead.

%
Consider 1 more account, from Laura:
\begin{quotation}
    Paul is a great guy.
    
    We dated for about 6 months, and now we've been living together for several more.
    
    We're engaged.
    
    I feel so bad for him.
    
    His ex-wife accused him of abusing her, and it's a total lie.
    
    He made 1 mistake, which is that he cheated on her, and she is determined to get him back for that.
    
    She will stop at nothing.
    
    Now she is even saying that he was \textit{violent}, claiming he slapped her a few times and broke her things.
    
    That's ridiculous!
    
    I've been with him for over a year now, and I can tell you, he's \textit{nothing} like that.
    
    Paul has never even raised a hand to me.
    
    In fact, he's tried to help me get my life together and has been really there for me.
    
    I was in a bad place when I met him, I was depressed and I was drinking too much, and I'm doing so much better now, because of him.
    
    I hate that bitch for accusing him of those things.
    
    We're going to work together on getting custody of his kids, because she's out of control.
\end{quotation}
Laura wondered how Paul's ex-wife could accuse such a delightful man of abuse.

She was so angry about it that she didn't notice several warning signs about her own relationship with Paul.

%
If Kristen, Barbara, and Laura were to sit down together and compare notes, they might decide that their partners couldn't be more different.

The personalities of the 3 men seem miles apart, and their relationships follow very separate paths.

Yet Maury, Fran, and Paul \fbox{actually have far more in common than meets the eye.}

Their moodiness, their excuses, their outlook, are all bubbling from the same source.

And all 3 are abusive men.

\subsection{The tragedy of abuse}
Abuse of women in relationships touches an unimaginable number of lives.

Even if we leave aside cases of purely verbal and mental abuse and just look at physical violence, the statistics are shocking: 2--4 million women are assaulted by their partners \textit{per year} in the United States.

The U.S. Surgeon General has declared that attacks by male partners are the \textit{number 1 cause of injury} to women between the ages of 15 and 44.

The American Medical Association reports that \textit{1 woman out of 3} will be a victim of violence by a husband or boyfriend at some point in her life.

The emotional effects of partner violence are a factor in more than 1-4th of female suicide attempts and are a leading cause of substance abuse in adult women.

Government statistics indicate that 1,500--2,000 women are murdered by partners and ex-partners per year, comprising more than 1-3rd of all female homicide victims, and that these homicides almost always follow a history of violence, threats, or stalking.

%
The abuse of women sends shock waves through the lives of children as well.

Experts estimate that 5 million children per year witness an assault on their mothers, an experience that can leave them traumatized.

Children exposed to violence at home show higher rates of \textit{school behavior and attention problems, aggression, substance abuse, depression}, and many other measures of childhood distress.

Abuse of women has been found to be a cause of roughly 1-3rd of divorces among couples with children and 1-half of divorces where custody is disputed.

%
As alarming as this picture is, we also know that physical assaults are just the beginning of the abuse that women may be subjected to.

There are millions more women who have never been beaten but who live with \textit{repeated verbal assaults, humiliation, sexual coercion}, and other forms of \textit{psychological abuse}, often accompanied by economic exploitation.

\fbox{The scars from mental cruelty can be as deep and long-lasting as wounds from punches or slaps but are often not as obvious.}

In fact, even among women who have experienced violence from a partner, half or more report that \textit{the man's emotional abuse is what is causing the greatest harm}.

%
The differences between the verbally abusive man and the physical batterer are not as great as many people believe.

The behavior of either style of abuser grows from the same roots and is driven by the same thinking.

Men in either category follow similar processes of change in overcoming their abusiveness - if they do change, which unfortunately is not common.

And \fbox{the categories tend to blur.}

\fbox{Physically assaultive men are also verbally abusive to their partners.}

Mentally cruel and manipulative men tend to gradually drift into using physical intimidation as well.

In this book you will meet abusers on a spectrum, ranging from those who never use violence to those who are terrifying.

The extend of their common ground may startle you.

%
1 of the obstacles to recognizing chronic mistreatment in relationships is that most abusive men simply don't \textit{seem} like abusers.

They have many good qualities, including times of kindness, warmth, and humor, especially in the early period of a relationship.

An abuser's friends may think the world of him.

He may have a successful work life and have no problems with drugs or alcohol.

He may simply not fit anyone's image of a cruel or intimidating person.

So when a woman feels her relationship spinning out of control, it is unlikely to occur to her that her partner is an abuser.

%
The symptoms of abuse are there, and the woman usually sees them: the \fbox{escalating frequency of put-downs.}

\fbox{Early generosity turning more and more to selfishness.}

Verbal explosions when he is irritated or when he doesn't get his way.

Her grievances constantly turned around on her, so that everything is her own fault.

His growing attitude that he knows what is good for her better than she does.

And, in many relationships, a mounting sense of fear or intimidation.

But the woman also sees that her partner is a human being who can be caring and affectionate at times, and she loves him.

She wants to figure out why he gets so upset, so that she can help him break his pattern of ups and downs.

She gets drawn into the complexities of his inner world, trying to uncover clues, moving pieces around in an attempt to solve an elaborate puzzle.

%
The abuser's mood changes are especially perplexing.

He can be a different person from day to day, or even from hour to hour.

At times he is aggressive and intimidating, his tone harsh, insults spewing from his mouth, ridicule dripping from him like oil from a drum.

When he's in this mode, nothing she says seems to have any impact on him, except to make him even angrier.

Her side of the argument counts for nothing in his eyes, and everything is her fault.

\fbox{He twisted her words around so that she always ends up on the defensive.}

As so many partners of Bancroft's clients have said to me, ``I just can't seem to do anything right.''

%
\fbox{At other moments, he sounds wounded and lost, hungering for love and for someone to take care of him.}

When this side of him emerges, he appears open and ready to heal.

He seems to let down his guard, his hard exterior softens, and he may take on the quality of a hurt child, difficult and frustrating but lovable.

Looking at him in this deflated state, his partner has trouble imagining that the abuser inside of him will ever be back.

The beast that takes him over at other times looks completely unrelated to the tender person she now sees.

%
Sooner or later, though, the shadow comes back over him, as if it had a life of its own.

Weeks of peace may go by, but eventually she finds herself under assault once again.

Then her head spins with the arduous effort of untangling the many threads of his character, until she begins to wonder whether she is the one whose head isn't quite right.

%
To make matters worse, everyone she talks to has a different opinion about the nature of his problem and what she should do about it.

Her clergyperson may tell her, ``Love heals all difficulties.

Give him your heart fully, and he will find the spirit of God.''

Her therapist speaks a different language, saying, ``He triggers strong reactions in you because he reminds you of your father, and you set things off in him because of his relationship with his mother.

You each need to work on not pushing each other's buttons.''

A recovering alcoholic friend tells her, ``He's a rage addict.

He controls you because he is terrified of his own fears.

You need to get him into a 12-step program.''

Her brother may say to her, ``He's a good guy.

I know he loses his temper with you sometimes - he does have a short fuse - but you're no prize yourself with that mouth of yours.

You 2 need to work it out, for the good of the children.''

And then, to crown her increasing confusion, she may hear from her mother, or her child's schoolteacher, or her best friend: ``He's mean and crazy, and he'll never change.

All he wants is to hurt you.

Leave him now before he does something even worse.''

%
All of these people are trying to help, and they are all talking about the same abuser.

But he looks different from each angle of view.

%
The woman knows from living with the abusive man that there are no simple answers.

Friends say: ``He's mean.''

But she knows many ways in which he has been good to her.

Friends say: ``He treats you that way because he can get away with it.

I would never let someone treat \textit{me} that way.''

But she knows that the times when she puts her foot down the most firmly, he responds by becoming his angriest and most intimidating.

When she stands up to him, he makes her pay for it - sooner or later.

Friends say: ``Leave him.''

But she knows it won't be that easy.

He will promise to change.

He'll get friends and relatives to feel sorry for him and pressure her to give him another chance

He'll get severely depressed, causing her to worry whether he'll be all right.

And, depending on what style of abuser he is, she may know that he will become dangerous when she tries to leave him.

She may even be concerned that he will try to take her children away from her, as some abusers do.

%
How is an abused women to make a sensible picture out of this confusion?

How can she gain enough insight into the causes of his problem to know what path to choose?

The questions she faces are urgent ones.

\subsection{5 puzzles}
Professionals who specialize in working with abusive and controlling men have had to face these same perplexing issues at work.

Bancroft was a codirector of the 1st counseling program in the United States - and perhaps in the world - for abusive men.

When Bancroft began leading group for abusers 15 years ago, they were as much of a mystery to Bancroft as they are to the women they live with.

Bancroft's colleagues and he had to put a picture together from the same strange clues faced by Kristen, Barbara, and Laura.

%
Several themes kept confronting us over and over again in our clients' stories, including:

\subsubsection{His version of the abuse if worlds apart from hers.}
A man named Dale in his mid-30s gave the following account when he entered Bancroft's group for abusive men:
\begin{quotation}
    My wife Maureen and I have been together for 11 years.
    
    The 1st 10 years we had a good marriage, and there was no problem with abuse or violence or anything.
    
    She was a great girl.
    
    Then about a year ago she started hanging around with this bitch she met named Eleanor who really has it in for me.
    
    Some people just can't stand to see anyone else happy.
    
    This girl was single and was obviously jealous that Maureen was in a good marriage, so she set out to wreck it.
    
    Nobody can get along with Eleanor, so of course she has no relationships that last.
    
    I just had the bad luck that she ran into my wife.
    
    %
    So this bitch started planting a lot of bad stuff about me in Maureen's head and turning her against me.
    
    She tells Maureen's head and turning her against me.
    
    She tells Maureen that I don't care about her, that I'm sleeping with other girls, all kinds of lies.
    
    And she's getting what she wants, because now Maureen and I have started having some wicked fights.
    
    This past year we haven't gotten along at all.
    
    I tell Maureen I don't want her hanging around with that girl, but she doesn't listen to me.
    
    She sneaks around and sees her behind my back.
    
    And, look, I'm not here to hide anything.
    
    I'll tell you straight out, it's true that 2 or 3 times this year I finally couldn't take all the accusations and yelling anymore, and I've hauled off and slapped her.
    
    I need help, I'm not denying it.
    
    I have to learn to deal with the stress better; I don't want her to get me arrested.
    
    And maybe I can still figure out how to persuade Maureen not to throw a great thing away, because at the rate we're going we'll be broken up in 6 months.
\end{quotation}
Bancroft always interviews the partner of each of Bancroft's clients as soon as possible after he enrolls in the program.

Bancroft reached Maureen by phone several days later, and heard her account:
\begin{quotation}
    Dale was great when I 1st met him, but by the time we got married something was already wrong.
    
    He had gone from thinking I was perfect to constantly criticizing me, and he would get in such bad moods over the littlest things.
    
    I wouldn't be able to figure out how to get him to feel better.
    
    Only a couple of months after the wedding he shoved me for the 1st time, and after that some explosion would happen about 2 or 3 times a year.
    
    Usually he would break something or raise a fist, but a few times he shoved me or slapped me.
    
    Some years he didn't do it at all, and I would think it was all over, but then it would happen again - it sort of came in waves.
    
    And he was always, always, putting me down and telling me what to do.
    
    I couldn't do anything right.
    
    %
    Anyhow, about a year ago I made a new friend, Eleanor.
    
    She started telling me that what Dale was doing was abuse, even though he had never punched me or injured me, and that I hadn't done anything to deserve it.
    
    At 1st I thought she was exaggerating, because I've known women that got it so much worse than me.
    
    And Dale can be really sweet and supportive when you least expect it.
    
    We've had a lot of good times, believe it or not.
    
    Anyhow, Eleanor kind of opened my eyes up.
    
    So I started standing up to Dale about how he talks to me, and told him I was thinking of moving out for a while.
    
    And what's happened is that he's gone nuts.
    
    I swear, something has happened to him.
    
    He's backhanded he twice in the last 8 months, and another time he threw me over a chair and my back went out.
    
    So I finally moved out.
    
    For now I'm not planning to get back with him, but I guess it depends partly on what he does in the abuser program.
\end{quotation}
Notice the striking contrasts.

Dale describes the 1st 10 years of his marriage as abuse-free, while Maureen remembers put-downs and even physical assaults during those years.

Maureen says that Eleanor helps and supports her, while Dale sees her as corrupting Maureen and turning her against him.

Dale says that they are still together, while Maureen reports that they have already broken up.

Each one thinks the other has developed a problem.

\fbox{How can their perceptions clash so strongly?}

In the chapters ahead, we will explore the thinking of abusive men to answer the question of why Dale's view contains such serious distortions.

\subsubsection{He gets insanely jealous, but in other ways he seems entirely rational.}
In a group session 1 day, a young client named Marshall was recounting a confrontation with his partner that had occurred in the previous week:
\begin{quotation}
    My wife and I had plans to meet in the lobby of the building where she works to go out for lunch.
    
    I was waiting around near the elevators, and when she finally came out I saw that she'd been alone on the elevator with this good-looking guy.
    
    He had a look on his face, and she did too, I can't really describe it, but I could tell something was up.
    
    I said, ``What was that all about?,'' and she pretended like she didn't know what I was talking about.
    
    That really pissed me off, and I guess I kind of blew up at her.
    
    I may have gotten a little louder than I should have.
    
    I was mad, though, and I was saying, ``You were making it with that guy on the elevator, weren't you?
    
    Don't lie to me, you slut, I'm not a fool.''
    
    But she kept on playing dumb, saying she doesn't even know him, which is a crock.
\end{quotation}
Marshall was extremely jealous, but Bancroft had worked with him long enough to know that he wasn't crazy.

He was lucid and logical in group, had a stable work history and normal friendships, and showed no signs of living in a world of fantasy or hallucination.

He simply did not have symptoms of the type of serious mental illness that could convince a man that his wife could have sex in an elevator, fully clothed and standing up, between floors of a busy office building.

Marshall had to know that his accusation wasn't true.

And when Bancroft confronted him, he admitted it.

%
\fbox{Given that even very jealous abusers turn out to have a reasonable grasp on reality}, why do they make these insane-seeming accusations?

\fbox{Is there something about acting crazy that they enjoy?}

What does this behavior accomplish for them?

(Bancroft answers these questions in Chap. 3, where we consider the issues of \textit{possessiveness}.)

\subsubsection{He succeeds in getting people to take his side against her.}
Martin, a man in this late 20s, joined Bancroft's abuser group while also seeing an individual therapist.

He told Bancroft the 1st day that he was confused about whether he had a problem or not, but that his long-time girlfriend Ginny was preparing to break up with him because she considered him abusive.

He went on to describe incidents of insulting or ignoring Ginny and of deliberating causing her emotional pain ``to show her how it feels when she hurts me.''

He also admitted to times of humiliating her in front of other people, being flirtatious with women when he was mad at her, and ruining a couple of recent important events in her life by causing big scenes.

He justified all of these behaviors because of ways he felt hurt by her.

%
As a routine part of my assessment of Martin, Bancroft contacted his private therapist to compare impressions.

The therapist turned out to have strong opinions about the case:
\begin{itemize}
    \item \textbf{Therapist}: I think it's a big mistake for Martin to be attending your abuser program.
    
    \fbox{He has very low self-esteem; he believes anything bad that anyone says about him.}
    
    If you tell him he's abusive, that will just tear him down further.
    
    His partner slams him with the word \textit{abusive} all the time, for reasons of her own.
    
    Ginny's got huge control issues, and she has obsessive-compulsive disorder.
    
    She needs treatment.
    
    I think having Martin in your program just gets her what she wants.
    \item \textbf{Bancroft}: So you have been doing couples counseling with them?
    \item \textbf{Therapist}: No, I see him individually.
    \item \textbf{Bancroft}: How many times have you met with her?
    \item \textbf{Therapist}: She hasn't been in at all.
    \item \textbf{Bancroft}: You must have had quite extensive phone contact with her, then.
    \item \textbf{Therapist}: No, I haven't spoken to her.
    \item \textbf{Bancroft}: You haven't spoken to her?
    
    You have assigned Ginny a clinical diagnosis based only on Martin's descriptions of her?
    \item \textbf{Therapist}: Yes, but you need to understand, we're talking about an unusually insightful man.
    
    Martin has told me many details, and he is perceptive and sensitive.
    \item \textbf{Bancroft}: But he admits to serious psychological abuse of Ginny, although he doesn't call it that.
    
    An abusive man is not a reliable source of information about his partner.
\end{itemize}
What Martin was getting from individual therapy, unfortunately, was an official seal of approval for his denial, and for his view that Ginny was mentally ill.

How had he shaped his therapist's view of his partner to get her to adopt this stance?

How can abusers be so adept at recruiting team members in this way, including sometimes ones with considerable status or influence, and why do they want to?

(These questions are the focus of Chap. 11, ``Abusive Men \& Their Allies.'')

\subsubsection{During some incidents he seems to lose control, but certain other controlling behaviors of his appear very calculated.}
Several years ago, a young man named Mark came to 1 of Bancroft's abuser groups.

When a client joins the program, Bancroft sets behavioral goals with him as soon as possible.

Bancroft often begins by asking, ``What are the top 3 or 4 complaints your partner has about you?''

Mark's response was:
\begin{quotation}
    1 of the things Eileen gets on me about the most is that she says I ignore her.
    
    She says I make her a low priority and always want to do other things instead of be with her, so she feels like she's nothing.
    
    I like to have time to myself a lot, or to relax and watch television.
    
    I guess I kind of tune her out.
\end{quotation}
Based on Mark's account, Bancroft wrote near the top of his Behavior Plan: \textbf{``Spend more time with Eileen. Make her a higher priority.''}

%
Eileen was very difficult to reach by phone, but 3 weeks later she finally called me, with a surprising story to tell:
\begin{quotation}
    A few weeks before Mark started your program, I told him that I needed a total break from the relationship.
    
    I just couldn't take it anymore, the yelling and the selfishness.
    
    He won't even let me sleep.
    
    So I didn't even want to talk to him for a while; I had to have time away to get myself together.
    
    I reassured him that the relationship wasn't over, and we'd work on getting back together in a couple of months, after a breather.
    
    %
    Then, a couple of weeks later, he called me and said that he had enrolled in an abuser program.
    
    He said that his counselor wants him to spend more time with me and had written it on his sheet, and that the program told him that being with me was part of how he needed to work on his issues.
    
    I wasn't ready for that yet at all, but I also didn't want to interfere with his program.
    
    So I started seeing him again.
    
    I want whatever is going to work best to help him change.
    
    I could have used a little more time apart, to tell you the truth, but if that's what your program recommends$\ldots$
\end{quotation}
Mark had succeeded in twisting the abuser program to suit his own purposes.

Bancroft explained to Eileen what had happened and apologized for the way Bancroft's program had added to the many difficulties she already had with him.

The high degree of manipulativeness that Mark used is not uncommon among abusive men, unfortunately.

\fbox{How can abusers be capable of such calculation yet at other times appear to be so out of control?}

What's the connection?

The answers can be found in Chap. 2, where we examine the excuses that abusive men use to justify their behavior.

\subsubsection{Sometimes he seems to be really changing, but it tends to vanish.}
Carl was a 26-year-old man who had been arrested repeatedly for domestic assaults and had finally served a few months in jail.

He said to me in a group session:
\begin{quotation}
    Going to jail was the last straw.
    
    I finally got it that I have to stop blaming my problems on everybody else and take a look at myself instead.
    
    People in jail said the same thing to me: If you don't want to be back in here, get real with yourself.
    
    I have a bad temper, and kind of a mean streak to tell you the truth, and I have to deal with it.
    
    I don't want to be back inside for anything.
\end{quotation}
At the end of each counseling session, Carl would make comments e.g., ``I can see that I've really got to work on my attitude'' and ``I learned a lot tonight about how excuses keep me from changing.''

1 night he looked at me and said, ``I'm really glad I met you, because I think if I wasn't hearing the things you are saying I would be headed straight back to being locked up.

You're helping me get my head on straight.''

%
Bancroft reached Carl's girlfriend, Peggy, by telephone and began to ask her about the history of Carl's problem with abusiveness.

She sounded noticeably distracted and uncomfortable.

Bancroft suspected strongly that Carl was listening to the conversation, so I made an excuse to wrap it up soon.

However, when Carl was at Bancroft's group the next week, Bancroft left his co-leader in charge of the session and slipped out to give Peggy another call, to see if she would feel freer to talk.

This time she gave Bancroft an earful:
\begin{quotation}
    Carl comes home from your program in a rage every week.
    
    I'm afraid to be around the house on Wednesday nights, which is when he has his group session.
    
    He says the program is total bullshit, and that he wouldn't have to be sitting there getting insulted by you people if I hadn't called the police on him, and he says that I know the fight that night was my fault anyhow.
    
    He says he especially hates that guy Lundy.
    
    A few nights ago I told him to stop blaming it on me that he has to go to counseling, and he slammed me up against the doorjamb and told me if I didn't shut up he'd choke me.
    
    I should call the police, but he'd get sent away for 2 years this time because he's on parole, and I'm afraid that would be enough to get him to kill me when he got out.
\end{quotation}
Peggy then went on to describe the history of beatings she had suffered at Carl's hands before he went to jail: the black eyes, the smashed furniture, the time he had held a knife to her throat.

He invariably had blamed each attack on her, no matter how brutal his abuse or how serious her injuries.

%
After speaking with Peggy, Bancroft returned to the group session, where Carl went through his usual routine of self-exploration and guilt.

Bancroft of course said nothing; if he knew Peggy had told Bancroft the truth, she would be in extraordinary danger.

Soon after this, Bancroft reported to his probation officer that he was not appropriate for Bancroft's program, without giving the real reason.

%
Carl created the appearance of learning a great deal at each session, and his comments suggested serious reflection on the issues, including the effects of his abuse on his partner.

What was happening each week inside his mind before he got home?

How can an abuser gain such insight into his feelings and still behave so destructively?

And how does real change happen?

(We'll return to these questions in Chap. 14, ``The Process of Change.'')

%
These are just a very few of the many confounding questions that face anyone - the partner of an abusive man, a friend, or a professional - who is looking for effective ways to respond to abusive behavior.

Bancroft came to realize, through his experience with over 2000 abusers, that the abusive man \textit{wants} to be a mystery.

To get away with his behavior and to avoid having to face his problem, he needs to convince everyone around him - and himself - that his behavior makes no sense.

He needs his partners to focus on everything \textit{except} the real causes of his behavior.

\fbox{To see the abuser as he really is, it is necessary to strip away layer after layer of confusion, mixed messages, and deception.}

\fbox{Like anyone with a serious problem, abusers work hard to keep their true selves hidden.}

%
Part of how the abuser escapes confronting himself is by convincing you that \textit{you} are the cause of his behavior, or that you at least share the blame.

But abuse is not a product of bad relationship dynamics, and you cannot make things better by changing your own behavior or by attempting to manage your partner better.

\fbox{Abuse is a problem that lies entirely within the abuser.}

%
Through years of direct work with abusers and their partners, Bancroft found that the realities behind the enigmatic abuser gradually came out into the bright light forming a picture that increasingly made sense to Bancroft.

The pages ahead will take you through the pieces that Bancroft watched fall into place 1 by 1, including:
\begin{itemize}
    \item Why abusers are charming early in relationships but don't stay that way
    \item What the early warning signs are that can tip you off that you may be involved with an abusive or controlling man
    \item Why his moods change at the drop of a hat
    \item What goes on inside his mind and how his thinking causes his behavior
    \item What role alcohol and drugs play - and don't play - in partner abuse
    \item Why leaving an abusive man doesn't always solve the problem
    \item How to tell whether an abuser is really changing - and what to do if he isn't
    \item How friends, relatives, and other community members can help to stop abuse
    \item Why many abusive men seem to be mentally ill - and why they usually aren't
\end{itemize}
We will explore answers to these questions on 3 levels.

The 1st level is the abuser's thinking - his attitudes and beliefs - in daily interactions.

The 2nd is his learning process, through which his thinking began to develop early in his life.

And the 3rd involves the rewards he reaps from controlling his partner, which encourage him to use abusive behavior over and over again.

As we clear away the abusive man's moke screen with these understandings, you will find that abusiveness turns out to be far less mysterious than it appears at 1st.

%
\fbox{Inside the abuser's mind, there is a world of beliefs, perceptions, \& responses that fits together in a surprisingly logical way.}

His behavior \textit{does} make sense.

Underneath the facade of irrationality and explosiveness, there is a human being with a comprehensible - and solvable - problem.

But he doesn't want you to figure him out.

%
\fbox{The abuser creates confusion because he has to.}

He can't control and intimidate you, he can't recruit people around him to take his side, he can't keep escaping the consequences of his actions, unless he can throw everyone off the track.

\fbox{When the world catches on to the abuser, his power begins to melt away.}

So we ar going to travel behind the abuser's mask to the heart of his problem.

This journey is critical to the health and healing of abused women and their children, for once you grasp how your partner's mind works, you can begin reclaiming control of your own life.

Unmasking the abuser also does \textit{him} a favor, because he will not confront - and overcome - his highly destructive problem as long as he can remain hidden.

%
The better we understand abusers, the more we can create homes and relationships that are havens of love and safety, as they should be.

\fbox{Peace really does begin at home.}

\section{The Mythology}
\begin{quotation}
    He's crazy.
    
    He feels so bad about himself. I just need to build up his self-image a little.
    
    He just loses it.
    
    \fbox{He's so insecure.}
    
    His mother abused him, and now he has a grudge against women and he takes it out one me.
    
    I'm so confused. I don't understand what's going on with him.
\end{quotation}
In 1 important way, \textit{an abusive man works like a magician: His tricks largely rely on getting you to look off in the wrong direction, distracting your attention so that you won't notice where the real action is}.

He draws you into focusing on the turbulent world of his feelings to keep your eyes turned away from the true cause of his abusiveness, which lies in how he \textit{thinks}.

He leads you into a convoluted maze, making your relationship with him a labyrinth of twists and turns.

He wants you to puzzle over him, to try to figure him out, as though he were a wonderful but broken machine for which you need only to find and fix the malfunctioning parts to bring it roaring to its full potential.

His desire, though he may not admit it even to himself, is that you wrack your brain in this way so that you don't notice the patterns and logic of his behavior, the consciousness behind the craziness.

%
To further divert your gaze, he may work to shape your view of his past partners to keep you from talking to them directly and to prepare you to disbelieve them should you happen to hear what they say.

If you could follow the thread of his conduct over a series of relationships, you would find out that his behavior isn't as erratic as it looks; in fact, it follows a fairly consistent pattern from woman to woman, except for brief relationships or ones he isn't that serious about.

%
\fbox{Above all, the abusive man wants to avoid having you zero is on his abusiveness itself.}

So he tries to fill your head up with excuses and distortions and keep you weighed down with self-doubt and self-blame.

And, unfortunately, much of the society tends to follow unsuspectingly along behind him, helping him to close your eyes, and his own, to his problem.

%
The mythology about abusive men that runs through modern culture has been created largely by the abusers themselves.

Abusive men concoct explanations for their actions which they give to their partners, therapists, clergypeople, relatives, and social researchers.

But it is a serious error to allow abusers to analyze and account for their own problems.

Would we ask an active alcoholic to tell us why he or she drinks, and then accept the explanation unquestioningly?

This is what we would hear:
\begin{quotation}
    ``I drink because I have bad luck in life.''
    
    ``I actually don't drink much at all - it's just a rumor that some people have been spreading about me because they don't like me.''
    
    ``I started to drink a lot because my self-esteem was ruined by all these unfair accusations that I'm alcoholic, which I'm hot.''
\end{quotation}
When we hear these kinds of excuses from a drunk, we assume they are exactly that - excuses.

We don't consider an active alcoholic a reliable source of insight.

So why would we let an angry and controlling man be the authority on partner abuse?

Our 1st task, therefore, is to remove the abusive man's smoke and mirrors, and then set about watching carefully to see what he is really doing.

\subsection{A brief exercise}
In Bancroft's public presentations on abuse, Bancroft often begins with a simple exercise.

Bancroft asks the audience members to write down everything they have ever heard, or ever believed, about where an abuser's problem comes from.

Bancroft invites you to close this book for 2 or 3 minutes now and make a similar list for yourself, so that you can refer to it as we go along.

%
Bancroft then asks people to call out items from their lists, and Bancroft writes them on the blackboard, organizing them into 3 categories: 1 for myths, 1 for partial truths, and 1 for accurate statements.

We usually end up with 20 or 30 myths, 4 or 5 half-truths, and perhaps 1 or 2 realities.

The audience members squint at Bancroft and fidgit in their seats, surprised to discover that the common beliefs about the causes of abuse contain several dollops of fantasy and misconception for each ounce of truth.

If you find as you go through this chapter that your own list turns out to contain mostly myths, you are not alone.

%
For the partner of an abusive or controlling man, having all of these mistaken theories pulled out from under you at once can be overwhelming.

But for each stick that we pull out of the structure of misconception about abusive men, a brick is waiting to take its place.

When we're finished, your partner will find it much harder than before to throw you off balance and confuse you, and your relationship will make sense to you in a way that it hasn't before.

\subsection{The myths about abusers}
\begin{enumerate}
    \item He was abused as a child.
    \item His previous partner hurt him.
    \item He abuses those he loves the most.
    \item He holds in his feelings too much.
    \item He has an aggressive personality.
    \item He loses control.
    \item He is too angry.
    \item He is mentally ill.
    \item He hates women.
    \item He is afraid of intimacy and abandonment.
    \item He has low self-esteem.
    \item His boss mistreats him.
    \item He has poor skills in communication an conflict resolution.
    \item There are as many abusive women as abusive men.
    \item His abusiveness is as bad for him as for his partner.
    \item He is a victim of racism.
    \item He abuses alcohol or drugs.
\end{enumerate}

\subsubsection{Myth \#1: He was abused as a child, \& he needs therapy for it.}
The partners of Bancroft's clients commonly believe that the roots of the man's abusiveness can be found in mistreatment that he suffered himself, and many professionals share the same misconception.

Bancroft hears explanations along the lines of:
\begin{example}
    ``He calls me all those horrible things because that is what his mother used to do to him.''
    
    ``His father used to get angry at him and beat him with a belt, so now if I get angry at all, he just freaks out and starts throwing things around the house.
    
    He says it's because deep down, he's really scared of my anger.''
    
    ``His stepmother was a witch.
    
    I've met her; she's vicious.
    
    So now he really has this thing against women.''
\end{example}commiserate

\subsubsection{Question 1: Is it because he was abused as a child?}
Multiple research studies have examined the question of whether men who abuse women tend to be survivors of childhood abuse, and the link has turned out to be weak; other predictors of which men are likely to abuse women have proven far more reliable, as we will see.

Notably, men who are violent toward other \textit{men} are often victims of child abuse - but the connection is much less clear for men who assault women.

The 1 exception is that those abusers who are brutally physically violent or terrifying toward women often do have histories of having been abused as children.

In other words, \textit{a bad childhood doesn't cause a man to become an abuser, but it can contribute to making a man who is abusive especially dangerous}.

%
\fbox{If abusiveness were the product of childhood emotional injury, abusers could overcome their problem through psychotherapy.}

But it is virtually unheard of for an abusive man to make substantial and lasting changes in \textit{his pattern of abusiveness} as a result of therapy.

(In Chap. 14, we'll examine the differences between psychotherapy and a specialized abuser program, because the latter sometimes can bring good results.)

He may work through other emotional difficulties, he may gain insight into himself, but his behavior continues.

In fact it typically gets worse, as he uses therapy to develop new excuses for his behavior, more sophisticated arguments to prove that his partner is mentally unstable, and more creative ways to make her feel responsible for his emotional distress.

Abusive men are sometimes masters of the hard-luck story, and may find that accounts of childhood abuse are 1 of the best ways to pull heartstrings.

%
For some abusive men, the blame-the-childhood approach has an additional reason for being appealing: By focusing on what his mother did wrong, he gets to blame a woman for his mistreatment of women.

This explanation can also appeal to the abused woman herself, since it makes sense out of his behavior and gives her someone safe to be angry at - since getting angry at \textit{him} always seems to blow up in her face.

The wider society, and the field of psychology in particular, has often jumped on this bandwagon instead of confronting the hard questions that partner abuse raises.

Abuse of women by men is so rampant that, unless people can somehow make it women's own fault, they are forced to take on a number of uncomfortable questions about men and about much of male thinking.

So it may seem easier to just lay the problem at the feet of the man's mother?

%
Bancroft's clients who have participated extensively in therapy or substance-abuse recovery programs sometimes sound like therapists themselves - and a few actually have been - as they adopt the terms of popular psychology or textbook theory.

1 client used to try to lure Bancroft into intellectual debates with comments e.g., ``Well, your group follows a cognitive-behavioral model, which has been shown to have limitations for addressing a problem as deep as this one.''

An abusive man who is adept in the language of feelings can make his partner feel crazy by turning each argument into a therapy session in which he puts her reactions under a microscope and assigns himself the role of ``helping'' her.

He may, e.g., ``explain'' to her the emotional issues she needs to work through, or analyze her reasons for ``mistakenly'' believing that he is mistreating her.

%
An abusive man may embellish his childhood suffering once he discovers that it helps him escape responsibility.

The National District Attorney's  Association Bulletin reported a revealing study that was conducted on another group of destructive men: child sexual abusers.

The researcher asked each man whether he himself had been sexually victimized as a child.

A hefty 67\% of the subjects said yes.

However, the researcher then informed the men that he was going to hook them up to a lie-detector test and ask them the same questions again.

Affirmative answers suddenly dropped to only 29\%.

In other words, abusers of all varieties tend to realize the mileage they can get out of saying, ``I'm abusive because the same thing was done to me.''

%
Although the typical abusive man works to maintain a positive public image, it is true that some women have abusive partners who are nasty or intimidating to \textit{everyone}.

How about that man?

Do his problems result from mistreatment by his parents?

The answer is both yes and no; it depends on \textit{which} problem we're talking about.

His hostility toward the human race may sprout from cruelty in his upbringing, but he abuses women because he has an abuse problem.

The 2 problems are related but distinct.

%
Bancroft is not saying that you should be unsympathetic to your partner's childhood suffering.

An abusive man deserves the same compassion that a nonabusive man dose, neither more nor less.

But a nonabusive man doesn't use his past as an excuse to mistreat you.

\fbox{Feeling sorry for your partner can be a trap, making you feel guilty for standing up to his abusiveness.}

%
Bancroft has sometimes said to a client: ``If you are so in touch with your feelings from your abusive childhood, then you should know what abuse feels like.

You should be able to remember how miserable it was to be cut down to nothing, to be put in fear, to be told that the abuse is your own fault.

You should be \textit{less} likely to abuse a woman, not more so, from having been through it.''

Once Bancroft makes this point, he generally stops mentioning his terrible childhood; \textbf{he only wants to draw attention to it if it's an excuse to stay the same, not if it's a reason to change.}

\subsubsection{Myth \#2: He had a previous partner who mistreated him terribly, \& now he has a problem with women as a result. He's a wonderful man, \& that bitch made him get like this.}
As we saw with Fran in Chap. 1, an abuser's bitter tale of emotional destruction by a past wife or girlfriend can have a powerful impact on his current partner.

In the most common version of this story, the man recounts how his ex-partner broke his heart by cheating on him, perhaps with several different men.

If you ask him how he found out, he answers that ``everybody'' knew about it or that his friends told him.

He also may say, ``I caught her cheating myself,'' but when you press him on what he actually \textit{saw}, it often turns out that he saw nothing, or that he saw her talking to some guy or riding in his car late at night, ``so I could tell.''

%
He may describe other wounds he received from a previous partner: She tried to control him; she wouldn't let him have any freedom; she expected him to wait on her hand and foot; she turned their children against him; she even ``had him arrested'' out of vindictiveness.

\fbox{What he is describing usually are his \textit{own} behaviors, but he attributes them to the woman so that he is the victim.}

He can gain sympathy from his new partner in this fashion, especially because so many women know what it is like to be abused - unfortunately - so they can connect with his distress.

%
The abusive or controlling man can draw a rich set of excuses from his past relationships.

For controlling his current partner's friendships and for accusing her of cheating on him: ``It's because my ex-partner hurt me so badly by cheating on me so many times, and that's why I'm so jealous and can't trust you.''

For throwing a temper tantrum when she asks him to clean up after himself: ``My ex-partner controlled me every move, and so now it makes me furious when I feel like you're telling me what to do.''

For having affairs of his own or keeping other love interests going on the side: ``I got so hurt last time that now I am really afraid of committing, so I want to keep having involvements with other people.''

\fbox{He can craft an excuse to fit any of his controlling behaviors.}

%
Bancroft recommends applying the following principle to assertions that an angry or controlling man makes about past women in his life:
\begin{center}
    \textbf{If it is an excuse for mistreating you, it's a distortion.}
\end{center}
A man who was genuinely mistreated in a relationship with a woman would not be using that experience to get away with hurting someone else.

%
Consider the reverse situation for a moment: Have you ever heard a woman claim that the reason why she is chronically mistreating her male partner is because a previous man abused her?

Bancroft has never run into his excuse in the 15 years Bancroft has worked in the field of abuse.

Certainly Bancroft has encountered cases where women had trouble trusting another man after leaving an abuser, but there is a critical distinction to be made: Her past experiences may explain how she \textit{feels}, but they are not an excuse for how she \textit{behaves}.

And the same is true for a man.

%
When a client of Bancroft blames a past relationship for his cruel or controlling behavior in the present, Bancroft jumps in with several questions: ``Did your ex-partner ever say that she felt controlled or intimidated by you?

What was her side of the story?

Did you ever put your hands on her in anger, or did she ever get a restraining order?''

By the time he has finished providing his answers, Bancroft usually can tell what happened: He abused that woman too.

%
It is fine to commiserate with a man about his bad experience with a previous partner, but the instant he uses her as an excuse to mistreat you, stop believing \textit{anything} he tells you about that relationship and instead recognize it as a sign that he has problems with relating to women.

Track down his ex-partner and talk with her as soon as possible, even if you hate her.

An abuser can mistreat partner after partner in relationships, each time believing that the problems are all the woman's fault and that he is the real victim.

%
Whether he presents himself as the victim of an ex-partner, or of his parents, the abuser's aim - though perhaps unconscious - is to play on your compassion, so that he can avoid dealing with his problem.

\subsubsection{Myth \#3: He's abusive because he feels so strongly about me. People cause those they care about most deeply the most pain.}

\section{The Abusive Mentality}

\section{The Types of Abusive Men}

\begin{center}
    Part II: The Abusive Man in Relationships
\end{center}

\section{How Abuse Begins}

\section{The Abusive Man in Everyday Life}

\section{Abusive Men \& Sex}

\section{Abusive Men \& Addiction}

\section{The Abusive Man \& Breaking Up}

\begin{center}
    Part III: The Abusive Man in the World
\end{center}

\section{Abusive Men as Parents}

\section{Abusive Men \& Their Allies}

\section{The Abusive Man \& the Legal System}

\begin{center}
    Part IV: Changing the Abusive Man
\end{center}

\section{The Making of an Abusive Man}

\section{The Process of Change}

\section{Creating an Abuse-Free World}

\section*{Resources}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}