\documentclass[oneside]{book}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/nguyen/1_NQBH/reference/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\addtolength{\headheight}{0pt}% obsolete
\lhead{\scriptsize \chaptername~\thechapter}
\rhead{\scriptsize \nouppercase{\leftmark}} %\nouppercase !
\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\cfoot{\thepage}
\def\labelitemii{$\circ$}

\title{Psychology}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\setcounter{secnumdepth}{6}
\tableofcontents

%------------------------------------------------------------------------------%

\chapter{Wikipedia's}

\section{\href{https://en.wikipedia.org/wiki/Psychology}{Wikipedia\texttt{/}Psychology}}


%------------------------------------------------------------------------------%

\section{\href{https://en.wikipedia.org/wiki/Narcissism}{Wikipedia\texttt{/}Narcissism}}
\textsf{\href{https://en.wikipedia.org/wiki/Narcissus_(Caravaggio)}{\textit{Narcissus}} (1597--99) by \href{https://en.wikipedia.org/wiki/Caravaggio}{Caravaggio}; the man in love with his own reflection.}

\textit{Narcissism} is the pursuit of gratification from \href{https://en.wikipedia.org/wiki/Vanity}{vanity} or \href{https://en.wikipedia.org/wiki/Egotistic}{egotistic} admiration of one's idealized \href{https://en.wikipedia.org/wiki/Self-image}{self-image} and attributes.

The term originated from \href{https://en.wikipedia.org/wiki/Greek_mythology}{Greek mythology}, where a young man named \href{https://en.wikipedia.org/wiki/Narcissus_(mythology)}{Narcissus} fell in love with his own image reflected in a pool of water.

Narcissism or pathological self-absorption was 1st identified as a disorder in 1898 by \href{https://en.wikipedia.org/wiki/Havelock_Ellis}{Havelock Ellis}[1] and featured in subsequent psychological models, e.g. in \href{https://en.wikipedia.org/wiki/Sigmund_Freud}{Freud}'s \href{https://en.wikipedia.org/wiki/On_Narcissism}{\textit{On Narcissism}} (1914).

The \href{https://en.wikipedia.org/wiki/American_Psychiatric_Association}{American Psychiatric Association} has listed the classification \href{https://en.wikipedia.org/wiki/Narcissistic_personality_disorder}{narcissistic personality disorder} in its \href{https://en.wikipedia.org/wiki/Diagnostic_and_Statistical_Manual_of_Mental_Disorders}{\textit{Diagnostic and Statistical Manual of Mental Disorders}} (DSM) since 1968, drawing on the historical concept of \href{https://en.wikipedia.org/wiki/Megalomania}{megalomania}.

%
It is distinct from concepts of distinguishing the self (\href{https://en.wikipedia.org/wiki/Egocentrism}{egocentrism} or \href{https://en.wikipedia.org/wiki/Egoism}{egoism}) and \href{https://en.wikipedia.org/wiki/Healthy_narcissism}{healthy forms of responsibility and care for oneself} (``\href{https://en.wikipedia.org/wiki/Primary_narcissism}{primary narcissism}'').

Narcissism by contrast is considered a problem for relationships with self and others and for maintaining a functional culture.

In \href{https://en.wikipedia.org/wiki/Trait_theory}{trait personality theory}, it features in several \href{https://en.wikipedia.org/wiki/Self-report_inventory}{self-report personality inventories} including the \href{https://en.wikipedia.org/wiki/Millon_Clinical_Multiaxial_Inventory}{Millon Clinical Multiaxial Inventory}.

It is 1 of the 3 \href{https://en.wikipedia.org/wiki/Dark_triad}{dark triadic} personality traits (the others being \href{https://en.wikipedia.org/wiki/Psychopathy}{psychopathy} and \href{https://en.wikipedia.org/wiki/Machiavellianism_(psychology)}{Machiavellianism}).

\subsection{History}
Main article: \href{https://en.wikipedia.org/wiki/History_of_narcissism}{History of narcissism.}

\textsf{The myth of \href{https://en.wikipedia.org/wiki/Sisyphus}{Sisyphus} tells about a man punished for his \href{https://en.wikipedia.org/wiki/Hubris}{hubristic} belief that his cleverness surpassed that of Zeus himself. He has to push a stone up a mountain each day, only to have to recommence the task on the next day.}

The term \textit{narcissism} comes from the Greek myth about \href{https://en.wikipedia.org/wiki/Narcissus_(mythology)}{Narcissus} (Greek: \textit{Narkissos}), a handsome Greek youth who, according to \href{https://en.wikipedia.org/wiki/Ovid}{Ovid}, rejected the desperate advances of the nymph \href{https://en.wikipedia.org/wiki/Echo_(mythology)}{Echo}.

This caused Narcissus to fall in love with his own reflection in a pool of water.

Unable to consummate his love, Narcissus ``lay gazing enraptured into the pool, hour after hour,'' and finally changed into a flower that bears his name, the \href{https://en.wikipedia.org/wiki/Narcissus_(plant)}{narcissus}.[2]

The concept of excessive \href{https://en.wikipedia.org/wiki/Selfishness}{selfishness} has been recognized throughout history.

In ancient Greece, the concept was understood as \href{https://en.wikipedia.org/wiki/Hubris}{hubris}.

It is only since the late 1800s that narcissism has been defined in psychological terms.[1]
\begin{itemize}
    \item In 1752 \href{https://en.wikipedia.org/wiki/Jean-Jacques_Rousseau}{Jean-Jacques Rousseau}'s play \textit{Narcissus: or the Self-Admirer} was performed in Paris.[3]
    \item In 1898 \href{https://en.wikipedia.org/wiki/Havelock_Ellis}{Havelock Ellis}, an English psychologist, used the term ``Narcissus-like'' in reference to excessive masturbation, whereby the person becomes his or her own sex object.[1]
    \item In 1899, Paul Näcke was the 1st person to use the term ``narcissism'' in a study of sexual perversions.[4]
    \item \href{https://en.wikipedia.org/wiki/Otto_Rank}{Otto Rank} in 1911 published the 1st psychoanalytical paper specifically concerned with narcissism, linking it to vanity and self-admiration.[1]
    \item \href{https://en.wikipedia.org/wiki/Sigmund_Freud}{Sigmund Freud} published a paper on narcissism in 1914 called ``\href{https://en.wikipedia.org/wiki/On_Narcissism}{On Narcissism: An Introduction}''.[5]
    \item In 1923, \href{https://en.wikipedia.org/wiki/Martin_Buber}{Martin Buber} published an essay ``\href{https://en.wikipedia.org/wiki/Ich_und_Du}{Ich und Du}'' (I and Thou), in which he pointed out that our narcissism often leads us to relate to others as objects instead of as equals.[6]
\end{itemize}

\subsection{Traits \& signs}
\begin{quotation}
    ``\textit{Life is a stage, and when the curtain falls upon an act, it is finished and forgotten. The emptiness of such a life is beyond imagination}.'' - \href{https://en.wikipedia.org/wiki/Alexander\_Lowen}{Alexander Lowen} describing the existence of a narcissist[7]
\end{quotation}
4 dimensions of narcissism as a personality variable have been delineated: leadership\texttt{/}authority, superiority\texttt{/}arrogance, self-absorption\texttt{/}self-admiration, and exploitativeness\texttt{/}entitlement.[8]

%
Numerous studies (Miller \& Campbell, 2008; Russ, Shedler, Bradley, \& Westen, 2008; Wink, 1991) have demonstrated that narcissism has 2 or more variants, grandiose narcissism and vulnerable narcissism.

The grandiose variant, which is usually measured using the \href{https://en.wikipedia.org/wiki/Narcissistic_Personality_Inventory}{NPI} is strongly linked to the \href{https://en.wikipedia.org/wiki/Diagnostic_and_Statistical_Manual_of_Mental_Disorders}{DSM-IV} conceptualization and is the variant associated with \href{https://en.wikipedia.org/wiki/Narcissistic_personality_disorder}{NPD}, a Cluster B personality disorder which reflect traits related to self-importance, entitlement, aggression, and dominance.

Vulnerable narcissism reflects a defensive and fragile grandiosity, which functions mainly as a cover for feelings of inadequacy.

Vulnerable narcissism characterized by hypersensitivity, defensiveness, and withdrawal and is strongly associated with \href{https://en.wikipedia.org/wiki/Borderline_personality_disorder}{BPD}, which is characterized by the fear of abandonment, interpersonal and affective instability, impulsivity, chronic feelings of emptiness, suicidal ideation, and self-mutilation.

Pathological narcissism is when both grandiose and vulnerable narcissism is concurrent, which is linked to poor self-esteem, lack of empathy, feelings of shame, interpersonal distress, aggression, and significant impairments in personality functioning across both clinical and non-clinical samples.

Pathological narcissism, as measured by the Pathological Narcissism Inventory (PNI), is embedded within the personality disorders organized at the borderline level, which include antisocial, borderline, narcissistic personality disorders.

PNI scales exhibited significant associations with \href{https://en.wikipedia.org/wiki/Parasuicidal}{parasuicidal} behavior, suicide attempts, homicidal ideation, and several aspects of psychotherapy utilization.[9]

Despite the phenomenological and empirical distinction between vulnerable and grandiose narcissism, there is experimental evidence from studies that have employed direct or indirect ego-threat to suggest that grandiose narcissists have also a fragile, vulnerable core.[10]

\subsection{Clinical \& research aspects}

\subsubsection{Narcissistic personality disorder}
Main article: \href{https://en.wikipedia.org/wiki/Narcissistic_personality_disorder}{Narcissistic personality disorder}.

Narcissistic personality disorder affects an estimated 1\% of the general population.[11][12]

Although most individuals have some narcissistic traits, high levels of narcissism can manifest themselves in a \href{https://en.wikipedia.org/wiki/Psychopathology}{pathological} form as narcissistic personality disorder (NPD), whereby the individual overestimates his or her abilities and has an excessive need for admiration and affirmation.

NPD was revised in the \href{https://en.wikipedia.org/wiki/DSM-5}{DSM-5}.

The general move towards a dimensional (personality trait-based) view of the Personality Disorders has been maintained.

Some narcissists may have a limited or minimal capability to experience emotions.[13]

\paragraph{Treatment \& management.} The Cochrane Collaboration has commissioned 2 reviews of the evidence for psychological and medical treatments for Narcissistic Personality Disorder (NPD).[14]

In both cases, they suspended their initiatives after the authors had made no progress in over a year.

There are no clear treatment strategies for NPD,[15] neither medication nor psychotherapy.[16] There is evidence that therapies effective in the treatment of other personality disorders do not generalize to NPD.[17]

\paragraph{Required element within normal development.} \href{https://en.wikipedia.org/wiki/Karen_Horney}{Karen Horney} saw the narcissistic personality as a temperament trait molded by a certain kind of early environment.

She did not see narcissistic needs and tendencies as inherent in human nature.[18]

%
Normal or \href{https://en.wikipedia.org/wiki/Healthy_narcissism}{healthy narcissism} is an essential component of mature \href{https://en.wikipedia.org/wiki/Self-esteem}{self-esteem} and basic self-worth.[19][20][21]

Dr. Craig Malkin describes a lack of healthy narcissism as echoism, a term inspired by the nymph \href{https://en.wikipedia.org/wiki/Echo_(mythology)}{Echo} in the mythology of \href{https://en.wikipedia.org/wiki/Narcissus_(mythology)}{Narcissus}.[22]

%
\href{https://en.wikipedia.org/wiki/Freud}{Freud} said that narcissism was an original state from which the individual develops the love object.[23]

He argued that healthy narcissism is an essential part of normal development.[5]

According to Freud, the love of the parents for their child and their attitude toward their child could be seen as a revival and reproduction of their own narcissism.[5]

The child has a megalomaniac \href{https://en.wikipedia.org/wiki/Omnipotence}{omnipotence} of thought;[23] the parents stimulate that feeling because in their child they see the things that they have never reached themselves.

Compared to neutral observers, parents tend to overvalue the qualities of their child.

When parents act in an extreme opposite style and the child is rejected or inconsistently reinforced depending on the mood of the parent, the self-needs of the child are not met.

%
Freud contrasted the natural development of active-egoistic and passive-altruistic tendencies in the individual with narcissism, in the former, and what Trevor Pederson referred to as echoism, in the latter.[24]
\begin{quotation}
    This is the place for 2 remarks.
    
    1st, how do we differentiate between the concepts of narcissism and egoism?
    
    Well, narcissism, I believe, is the libidinal complement to egoism.
    
    When we speak of egoism, we have in view only the individual's advantage; when we talk of narcissism we are also taking his libidinal satisfaction into account.
    
    As practical motives the 2 can be traced separately for quite a distance.
    
    It is possible to be absolutely egoistic and yet maintain powerful object-cathexes, in so far as libidinal satisfaction in relation to the object forms part of the ego's needs.
    
    In that case, egoism will see to it that striving for the object involves no damage to the ego.
    
    It is possible to be egoistic and at the same time to be excessively narcissistic -- that is to say, to have very little need for an object, whether, once more, for the purpose of direct sexual satisfaction, or in connection with the higher aspirations, derived from sexual need, which we are occasionally in the habit of contrasting with `sensuality' under the name of `love'.
    
    In all these connections egoism is what is self-evident and constant, while narcissism is the variable element.
    
    The opposite to egoism, altruism, does not, as a concept, coincide with libidinal object-cathexis, but is distinguished from it by the absence of longings for sexual satisfaction.
    
    When someone is completely in love, however, altruism converges with libidinal object-cathexis.
    
    As a rule the sexual object attracts a portion of the ego's narcissism to itself, and this becomes noticeable as what is known as the `sexual overvaluation' of the object.
    
    If in addition there is an altruistic transposition of egoism on to the sexual object, the object becomes supremely powerful; it has, as it were, absorbed the ego.'' (Freud, Introductory Lectures (1919), pp. 417--18)
\end{quotation}
Where the egoist can give up love in narcissism, the altruist can give up on the competition, or ``the will,'' in echoism.

The individual 1st has a non-ambivalent relations of \textit{fusion} with authority or love figures, which are characterized by the egoistic or altruistic drives.

2nd, the individual can move to \textit{defusion} from authority or love figures which leads to repetitions of ambivalent, narcissistic or echoistic relations.

In the 3rd movement, the individual becomes the dead or absent parental figure that never returned love to the echoist, or the perfect, grandiose parental figure in narcissism.[24]

While egoism and narcissism concern dynamics of power and inferiority/superiority, Pederson argues that altruism and echoism concern dynamics of belonging and inclusion/exclusion.

Pederson has 2 types of echoists: the ``subject altruist'' and the ``object altruist'', with the former being concerned with the belonging of others and loving them, and the latter being concerned with their own belonging and being loved.

The subject altruist is self-effacing, a people pleaser, and sacrifices her desire to help others who are outsiders become insiders, or to be the submissive helper of an insider.

The object altruist is gregarious, a people person, and wants to be interesting which is based on wanting to fit in and not be an outsider or wanting to be unique as an insider.

Both types of echoists show issues with being submissive, having problems saying no, and avoiding conflict.[24]

\paragraph{In relation to the pathological condition.} Freud's idea of narcissism described a pathology which manifests itself in the inability to love others, a lack of empathy, emptiness, boredom, and an unremitting need to search for power, while making the person unavailable to others.[25]

%
Healthy narcissism has to do with a strong feeling of ``own love'' protecting the human being against illness.

Eventually, however, the individual must love the other, ``the object love to not become ill''.

The individual becomes ill as a result of the frustration created when he is unable to love the object.[26]

In pathological narcissism such as the \href{https://en.wikipedia.org/wiki/Narcissistic_personality_disorder}{narcissistic personality disorder}, the person's libido has been withdrawn from objects in the world and produces \href{https://en.wikipedia.org/wiki/Megalomania}{megalomania}.

The clinical theorists \href{https://en.wikipedia.org/wiki/Kernberg}{Kernberg}, \href{https://en.wikipedia.org/wiki/Kohut}{Kohut} and \href{https://en.wikipedia.org/wiki/Theodore_Millon}{Theodore Millon} all saw pathological narcissism as a possible outcome in response to unempathic and inconsistent early childhood interactions.

They suggested that narcissists try to compensate in adult relationships.[27]

The pathological condition of narcissism is, as Freud suggested, a magnified, extreme manifestation of healthy narcissism.

%
Healthy narcissism has been suggested to be correlated with good psychological health.

\href{https://en.wikipedia.org/wiki/Self-esteem}{Self-esteem} works as a mediator between narcissism and psychological health.

Therefore, because of their elevated self-esteem, deriving from self-perceptions of competence and likability, high narcissists are relatively free of worry and gloom.[28]

%
Other researchers have suggested that healthy narcissism cannot be seen as `good' or `bad', but that it depends on the contexts and outcomes being measured.

In certain social contexts such as initiating social relationships, and with certain outcome variables, such as feeling good about oneself, healthy narcissism can be helpful.

In other contexts, such as maintaining long-term relationships and with outcome variables, such as accurate self-knowledge, healthy narcissism can be unhelpful.[29]

\subsubsection{Commonly used measures}

\paragraph{Narcissistic Personality Inventory.} Main article: \href{https://en.wikipedia.org/wiki/Narcissistic_Personality_Inventory}{Narcissistic Personality Inventory}.

The Narcissistic Personality Inventory (NPI) is the most widely used measure of narcissism in social psychological research.

Although several versions of the NPI have been proposed in the literature, a 40-item forced-choice version (Raskin \& Terry, 1988) is the one most commonly employed in current research.

Another shorter version, a 16-item version NPI-16 (Ames, Rose \& Anderson, 2013) is also present.

The NPI is based on the \href{https://en.wikipedia.org/wiki/Diagnostic_and_Statistical_Manual_of_Mental_Disorders}{DSM}-III clinical criteria for narcissistic personality disorder (NPD), although it was designed to measure these features in the general population.

Thus, the NPI is often said to measure ``normal'' or ``subclinical'' (borderline) narcissism (i.e., in people who score very high on the NPI do not necessarily meet criteria for diagnosis with NPD).

\paragraph{Millon Clinical Multiaxial Inventory.} Main article: \href{https://en.wikipedia.org/wiki/Millon_Clinical_Multiaxial_Inventory}{Millon Clinical Multiaxial Inventory}.

The Millon Clinical Multiaxial Inventory (MCMI) is a widely used diagnostic test developed by \href{https://en.wikipedia.org/wiki/Theodore_Millon}{Theodore Millon}.

The MCMI includes a scale for Narcissism.

The NPI and MCMI have been found to be well correlated, $r(146) = 0.55$, $p < 0.001$.[30]

Whereas the MCMI measures narcissistic personality disorder (NPD), the NPI measures narcissism as it occurs in the general population.

In other words, the NPI measures ``normal'' narcissism; i.e., most people who score very high on the NPI do not have NPD.

Indeed, the NPI does not capture any sort of narcissism taxon as would be expected if it measured NPD.[31]

\subsubsection{Empirical studies}
Within the field of \href{https://en.wikipedia.org/wiki/Psychology}{psychology}, there are 2 main branches of research into narcissism: (1) clinical and (2) social psychology.

%
These 2 approaches differ in their view of narcissism, with the former treating it as a disorder, thus as discrete, and the latter treating it as a \href{https://en.wikipedia.org/wiki/Personality_trait}{personality trait}, thus as a \textit{continuum}.

These 2 strands of research tend loosely to stand in a divergent relation to one another, although they converge in places.

%
Campbell and Foster (2007)[29] review the literature on narcissism.

They argue that narcissists possess the following ``basic ingredients'':
\begin{itemize}
    \item \textbf{Positive}: Narcissists think they are better than others.[32]
    \item \textbf{Inflated}: Narcissists' views tend to be contrary to reality.
    
    In measures that compare self-report to objective measures, narcissists' self-views tend to be greatly exaggerated.[33]
    \item \textbf{Agentic}: Narcissists' views tend to be most exaggerated in the \href{https://en.wikipedia.org/wiki/Agentic}{agentic} domain, relative to the communion domain.[32][33]
    \item \textbf{Special}: Narcissists perceive themselves to be unique and special people.[34]
    \item \textbf{Selfish}: Research upon narcissists' behavior in resource dilemmas supports the case for narcissists as being selfish.[35]
    \item \textbf{Oriented toward success}: Narcissists are oriented towards success by being, e.g., approach oriented.[36]
\end{itemize}
Narcissists tend to demonstrate a lack of interest in warm and caring interpersonal relationships.[29]

There are several ongoing controversies within narcissism literature, namely: whether narcissism is healthy or unhealthy; a \href{https://en.wikipedia.org/wiki/Personality_disorder}{personality disorder}; a discrete or continuous variable; defensive or offensive; the same across genders; the same across cultures; and changeable or unchangeable.

%
Campbell and Foster (2007) argue that \href{https://en.wikipedia.org/wiki/Self_control}{self-regulatory} strategies are of paramount importance to understanding narcissism.[29]

Self-regulation in narcissists involves such things as striving to make one's self look and feel positive, special, successful and important.

It comes in both intra-psychic, such as blaming a situation rather than self for failure, and interpersonal forms, such as using a relationship to serve one's own self.

Some differences in self-regulation between narcissists and non-narcissists can be seen with Campbell, Reeder, Sedikides \& Elliot (2000)[37] who conducted a study with 2 experiments.

In each experiment, participants took part in an achievement task, following which they were provided with false feedback; it was either bogus success or failure.

The study found that both narcissists and non-narcissists self-enhanced, but non-narcissists showed more flexibility in doing so.

Participants were measured on both a comparative and a non-comparative \href{https://en.wikipedia.org/wiki/Self-enhancement}{self-enhancement} strategy.

Both narcissists and non-narcissists employed the non-comparative strategy similarly; however, narcissists were found to be more self-serving with the comparative strategy, employing it far more than non-narcissists, suggesting a greater rigidity in their self-enhancement.

When narcissists receive negative feedback that threatens the self, they self-enhance at all costs, but non-narcissists tend to have limits.

%
Sorokowski et al. (2015) showed that narcissism is related to the frequency of posting \href{https://en.wikipedia.org/wiki/Selfie}{selfie}-type pictures on \href{https://en.wikipedia.org/wiki/Social_media}{social media}.

Sorokowski's study showed that this relationship was stronger among men than women.[38]

%
A 2020 study found that females scored significantly higher on vulnerable narcissism than males, but no gender differences were found for grandiose narcissism.[39]

%
Research indicates that being in a devalued social group can encourage narcissism in some members of that group, as said individuals attempt to compensate for their low social status (due to being a member of a stigmatized group) by exaggerating their own self-worth by engaging in narcissism, which may also help them psychologically cope with negative treatment at the hands of others, though it may also cause them to engage in behavior detrimental to themselves.[40][41]

\subsubsection{Heritability research using twin studies}
Livesley et al. concluded, in agreement with other studies, that narcissism, as measured by a standardized test, was a common inherited trait.[42]

Additionally, in similar agreement with those other studies, it was found that there exists a continuum between normal and disordered personality.

The study subjects were 175 volunteer twin pairs (90 identical, 85 fraternal) drawn from the general population.

Each twin completed a questionnaire that assessed 18 dimensions of personality disorder.

The authors estimated the \href{https://en.wikipedia.org/wiki/Heritability}{heritability} of each dimension of personality by standard methods, thus providing estimates of the relative contributions of \href{https://en.wikipedia.org/wiki/Genetics}{genetic} and \href{https://en.wikipedia.org/wiki/Natural_environment}{environmental} causation.
    
Of the 18 personality dimensions, narcissism was found to have the highest heritability (0.64), indicating that the \href{https://en.wikipedia.org/wiki/Concordance_(genetics)}{concordance} of this trait in the \href{https://en.wikipedia.org/wiki/Identical_twins}{identical twins} was significantly influenced by genetics.

Of the other dimensions of personality, only 4 were found to have heritability coefficients of greater than 0.5: \href{https://en.wikipedia.org/wiki/Callous_and_unemotional_traits}{callousness}, identity problems, oppositionality and social avoidance.

\subsubsection{Stigmatizing attitude towards psychiatric illness}
Arikan found that a \href{https://en.wikipedia.org/wiki/Social_stigma}{stigmatizing} attitude to \href{https://en.wikipedia.org/wiki/Psychiatric}{psychiatric} patients is associated with narcissistic personality traits.[43]

\subsubsection{In evolutionary psychology}
The concept of narcissism is used in \href{https://en.wikipedia.org/wiki/Evolutionary_psychology}{evolutionary psychology} in relation to the mechanisms of \href{https://en.wikipedia.org/wiki/Assortative_mating}{assortative mating}, or the non-random choice of a partner for purposes of procreation.

Evidence for assortative mating among humans is well established; humans mate assortatively regarding age, IQ, height, weight, nationality, educational and occupational level, physical and personality characteristics, and family relatedness.[44]

In the ``self seeking like'' hypothesis, individuals unconsciously look for a ``mirror image'' of themselves in others, seeking criteria of beauty or reproductive fitness in the context of self-reference.

Alvarez et al. found that facial resemblance between couples was a strong driving force among the mechanisms of assortative mating: human couples resemble each other significantly more than would be expected from random pair formation.[45]

Since facial characteristics are known to be inherited, the ``self seeking like'' mechanism may enhance reproduction between genetically similar mates, favoring the stabilization of genes supporting social behavior, with no kin relationship among them.

\subsection{Narcissistic supply}
Main article: \href{https://en.wikipedia.org/wiki/Narcissistic_supply}{Narcissistic supply}.

Narcissistic supply is a concept introduced into \href{https://en.wikipedia.org/wiki/Psychoanalysis#Theories}{psychoanalytic theory} by \href{https://en.wikipedia.org/wiki/Otto_Fenichel}{Otto Fenichel} in 1938, to describe a type of \href{https://en.wikipedia.org/wiki/Admiration}{admiration}, \href{https://en.wikipedia.org/wiki/Social_support}{interpersonal support} or sustenance drawn by an individual from his or her environment and essential to their \href{https://en.wikipedia.org/wiki/Self-esteem}{self-esteem}.[46]

The term is typically used in a negative sense, describing a pathological or excessive need for attention or admiration in \href{https://en.wikipedia.org/wiki/Codependents}{codependents} and the orally fixated, that does not take into account the feelings, opinions or preferences of other people.[47]

\subsection{Narcissistic rage \& narcissistic injury}
Main article: \href{https://en.wikipedia.org/wiki/Narcissistic_rage_and_narcissistic_injury}{Narcissistic rage \& narcissistic injury}.

Narcissistic rage is a reaction to narcissistic injury, which is a perceived threat to a narcissist's \href{https://en.wikipedia.org/wiki/Self-esteem}{self-esteem} or self-worth.

\textit{Narcissistic injury} and \textit{narcissistic scar} are terms used by \href{https://en.wikipedia.org/wiki/Sigmund_Freud}{Sigmund Freud} in the 1920s.

\textit{Narcissistic wound} and \textit{narcissistic blow} are other, almost interchangeable, terms.[48]

%
The term \textit{narcissistic rage} was coined by \href{https://en.wikipedia.org/wiki/Heinz_Kohut}{Heinz Kohut} in 1972.

Narcissistic rage occurs on a continuum from aloofness, to expressions of mild irritation or annoyance, to serious outbursts, including violent attacks.[49]

%
Narcissistic rage reactions are not limited to \href{https://en.wikipedia.org/wiki/Personality_disorders}{personality disorders}.

They may also be seen in \href{https://en.wikipedia.org/wiki/Catatonic}{catatonic}, \href{https://en.wikipedia.org/wiki/Paranoid_delusion}{paranoid delusion}, and depressive episodes.[49]

It has been suggested that narcissists have 2 layers of rage.

The 1st layer of rage can be thought of as a constant anger towards someone else, with the 2nd layer being a self-aimed anger.[50]

\subsection{Narcissistic defenses}
Main article: \href{https://en.wikipedia.org/wiki/Narcissistic_defenses}{Narcissistic defenses}.

%
Narcissistic defenses are those processes whereby the idealized aspects of the self are preserved, and its limitations denied.[51]

They tend to be rigid and totalitarian.[52]

They are often driven by feelings of \href{https://en.wikipedia.org/wiki/Shame}{shame} and \href{https://en.wikipedia.org/wiki/Guilt_(emotion)}{guilt}, conscious or unconscious.[53]

\subsection{Narcissistic abuse}
Main article: \href{https://en.wikipedia.org/wiki/Narcissistic_abuse}{Narcissistic abuse}.

Narcissistic abuse was originally just defined as a specific form of \href{https://en.wikipedia.org/wiki/Emotional_abuse}{emotional abuse} of children by \href{https://en.wikipedia.org/wiki/Narcissistic_parents}{narcissistic parents} - parents who require the child to give up their own wants and feelings in order to serve the parent's needs for esteem.[54]

The term emerged in the late 20th century due to the works of \href{https://en.wikipedia.org/wiki/Alice_Miller_(psychologist)}{Alice Miller} and other \href{https://en.wikipedia.org/wiki/Neo-Freudian}{Neo-Freudians}, rejecting psychoanalysis as being similar to the \href{https://en.wikipedia.org/wiki/Poisonous_pedagogy}{poisonous pedagogies}.[55]

%
\href{https://en.wikipedia.org/wiki/Self-help}{Self-help} culture assumes that someone abused by narcissistic parenting as a child likely struggles with \href{https://en.wikipedia.org/wiki/Codependency}{codependency} issues in adulthood.

An adult who is or has been in a relationship with a narcissist likely struggles with not knowing what constitutes a ``normal'' relationship.[56]

%
In recent years the term has been applied more broadly to refer to any abuse by a narcissist including in adult to adult relationships.[57] 

\subsection{Types}

\subsubsection{Masterson's subtypes (exhibitionist \& closet)}
In 1993, \href{https://en.wikipedia.org/wiki/James_F._Masterson}{James F. Masterson} proposed 2 categories for pathological narcissism, \href{https://en.wikipedia.org/wiki/Exhibitionist}{\textit{exhibitionist}} and \textit{closet}.[58]

Both fail to adequately develop an age- and phase- appropriate self because of defects in the quality of psychological nurturing provided, usually by the mother.

The exhibitionist narcissist is the one described in \href{https://en.wikipedia.org/wiki/DSM-IV}{DSM-IV} and differs from the closet narcissist in several important ways.

The closet narcissist is more likely to be described as having a deflated, inadequate self-perception and greater awareness of emptiness within.

The exhibitionist narcissist would be described as having an inflated, grandiose self-perception with little or no conscious awareness of the emptiness within.

Such a person would assume that this condition was normal and that others were just like him.

The closet narcissist seeks constant approval from others and appears similar to the \href{https://en.wikipedia.org/wiki/Borderline_Personality_Disorder}{borderline} in the need to please others.

The exhibitionist narcissist seeks perfect admiration all the time from others.

\subsubsection{Millon's variations}
In 1996 \href{https://en.wikipedia.org/wiki/Theodore_Millon}{Theodore Millon} identified 4 variations of narcissist.[11]

Any individual narcissist may exhibit none or 1 of the following:
\begin{itemize}
    \item \textbf{unprincipled narcissist}: including \href{https://en.wikipedia.org/wiki/Antisocial_personality_disorder}{antisocial} features.
    
    A charlatan - is a \href{https://en.wikipedia.org/wiki/Charlatan}{fraudulent}, \href{https://en.wiktionary.org/wiki/exploit}{exploitative}, \href{https://en.wikipedia.org/wiki/Amorality}{amoral}, disloyal, deceptive and unscrupulous individual.
    \item \textbf{amorous narcissist}: including \href{https://en.wikipedia.org/wiki/Histrionic_personality_disorder}{histrionic} features.
    
    The \href{https://en.wikipedia.org/wiki/Don_Juan}{Don Juan} or \href{https://en.wikipedia.org/wiki/Giacomo_Casanova}{Casanova} of our times - is erotic, tantalizing, enticing; clever, glib and pathological lying.
    \item \textbf{\href{https://en.wikipedia.org/wiki/Compensation_(psychology)}{compensatory} narcissist}: including negativistic (\href{https://en.wikipedia.org/wiki/Passive-aggressive_personality_disorder}{passive-aggressive}), \href{https://en.wikipedia.org/wiki/Avoidant_personality_disorder}{avoidant} features: admirable, exceptional, noteworthy; creating illusions of being superior.
    \item \textbf{\href{https://en.wikipedia.org/wiki/Elitism}{elitist} narcissist}: variant of pure pattern.
    
    Corresponds to \href{https://en.wikipedia.org/wiki/Wilhelm_Reich}{Wilhelm Reich}'s ``phallic narcissistic'' personality type.
\end{itemize}

\subsubsection{Other forms}

\paragraph{Acquired situational narcissism.} Acquired situational narcissism (ASN) is a form of narcissism that develops in late adolescence or adulthood, brought on by wealth, fame and the other trappings of \href{https://en.wikipedia.org/wiki/Celebrity}{celebrity}.

It was coined by \href{https://en.wikipedia.org/wiki/Robert_Millman}{Robert B. Millman}, professor of \href{https://en.wikipedia.org/wiki/Psychiatry}{psychiatry} at the \href{https://en.wikipedia.org/wiki/Weill_Cornell_Medical_College}{Weill Cornell Medical College} of \href{https://en.wikipedia.org/wiki/Cornell_University}{Cornell University}.

ASN differs from conventional narcissism in that it develops after childhood and is triggered and supported by the celebrity-obsessed society.

Fans, assistants and tabloid media all play into the idea that the person really is vastly more important than other people, triggering a narcissistic problem that might have been only a tendency, or latent, and helping it to become a full-blown personality disorder.

``Millman says that what happens to celebrities is that they get so used to people looking at them that they stop looking back at other people.''[59]

In its presentation and symptoms, it is indistinguishable from \href{https://en.wikipedia.org/wiki/Narcissistic_personality_disorder}{narcissistic personality disorder}, differing only in its late onset and its support by large numbers of others.

``The lack of social norms, controls, and of people telling them how life really is, also makes these people believe they're invulnerable,"[60] so that the person with ASN may suffer from unstable relationships, substance abuse and erratic behavior.

A famous fictional character with ASN is \href{https://en.wikipedia.org/wiki/Norma_Desmond}{Norma Desmond}, the main character of \href{https://en.wikipedia.org/wiki/Sunset_Boulevard_(film)}{Sunset Boulevard}.

\paragraph{Codependency.} Main article: \href{https://en.wikipedia.org/wiki/Codependency}{Codependency}.

Codependency is a tendency to behave in overly passive or excessively caretaking ways that negatively impact one's relationships and quality of life.

Narcissists are considered to be natural magnets for the codependent.

Rapport identifies codependents of narcissists as ``co-narcissists''.[61]

\paragraph{Collective or group narcissism.} Main article: \href{https://en.wikipedia.org/wiki/Collective_narcissism}{Collective narcissism}.

Collective narcissism (or group narcissism) is a type of narcissism where an individual has an inflated \href{https://en.wikipedia.org/wiki/Self-love}{self-love} of his or her own \href{https://en.wikipedia.org/wiki/Ingroup}{ingroup}, where an ``ingroup'' is a group in which an individual is personally involved.[62]

While the classic definition of narcissism focuses on the individual, collective narcissism asserts that one can have a similar excessively high opinion of a group, and that a group can function as a narcissistic entity.[62]

Collective narcissism is related to \href{https://en.wikipedia.org/wiki/Ethnocentrism}{ethnocentrism}; however, ethnocentrism primarily focuses on self-centeredness at an ethnic or cultural level, while collective narcissism is extended to any type of ingroup beyond just cultures and ethnicities.[62][63]

\paragraph{Conversational narcissism.} Conversational narcissism is a term used by sociologist \href{https://en.wikipedia.org/wiki/Charles_Derber}{Charles Derber} in his book, \textit{The Pursuit of Attention: Power and Ego in Everyday Life}.

Derber observed that the social support system in America is relatively weak, and this leads people to compete mightily for attention.

In social situations, they tend to steer the \href{https://en.wikipedia.org/wiki/Conversation}{conversation} away from others and toward themselves.

``Conversational narcissism is the key manifestation of the dominant attention-getting psychology in America,'' he wrote.

``It occurs in informal conversations among friends, family and coworkers.

The profusion of popular literature about listening and the etiquette of managing those who talk constantly about themselves suggests its pervasiveness in everyday life.''

What Derber describes as ``conversational narcissism'' often occurs subtly rather than overtly because it is prudent to avoid being judged an egotist.

Derber distinguishes the ``shift-response'' from the ``support-response,' as in the following 2 hypothetical conversation fragments:
\begin{quotation}
    John: I'm feeling really starved.
    
    Mary: Oh, I just ate. (shift-response)
    
    John: I'm feeling really starved.
    
    Mary: When was the last time you ate? (support-response)
\end{quotation}

\paragraph{Cultural narcissism.} In \href{https://en.wikipedia.org/wiki/The_Culture_of_Narcissism}{\textit{The Culture of Narcissism}}, \href{https://en.wikipedia.org/wiki/Christopher_Lasch}{Christopher Lasch} defines a narcissistic culture as one where every activity and relationship is defined by the \href{https://en.wikipedia.org/wiki/Hedonism}{hedonistic} need to acquire the symbols of wealth,[64] this becoming the only expression of rigid, yet covert, social \href{https://en.wikipedia.org/wiki/Hierarchy}{hierarchies}.

It is a culture where \href{https://en.wikipedia.org/wiki/Liberalism}{liberalism} only exists insofar as it serves a consumer society, and even \href{https://en.wikipedia.org/wiki/Art}{art}, \href{https://en.wikipedia.org/wiki/Sex}{sex} and \href{https://en.wikipedia.org/wiki/Religion}{religion} lose their liberating power.

In such a society of constant competition, there can be no allies, and little transparency.

The threats to acquisitions of social symbols are so numerous, varied and frequently incomprehensible, that defensiveness, as well as competitiveness, becomes a way of life.

Any real sense of community is undermined - or even destroyed - to be replaced by virtual equivalents that strive, unsuccessfully, to synthesize a sense of community.

\paragraph{Destructive narcissism.} Destructive narcissism is the constant exhibition of numerous and intense characteristics usually associated with the \href{https://en.wikipedia.org/wiki/Pathological}{pathological} narcissist but having fewer characteristics than pathological narcissism.[65]

\paragraph{Malignant narcissism.} Main article: \href{https://en.wikipedia.org/wiki/Malignant_narcissism}{Malignant narcissism}.

Malignant narcissism, a term 1st coined in a book by \href{https://en.wikipedia.org/wiki/Erich_Fromm}{Erich Fromm} in 1964,[66] is a \href{https://en.wikipedia.org/wiki/Syndrome}{syndrome} consisting of a cross breed of the \href{https://en.wikipedia.org/wiki/Narcissistic_personality_disorder}{narcissistic personality disorder}, the \href{https://en.wikipedia.org/wiki/Antisocial_personality_disorder}{antisocial personality disorder}, as well as \href{https://en.wikipedia.org/wiki/Paranoia}{paranoid} traits.

The malignant narcissist differs from one suffering from narcissistic personality disorder in that the malignant narcissist derives higher levels of psychological \href{https://en.wikipedia.org/wiki/Gratification}{gratification} from accomplishments over time (thus worsening the disorder).

Because the malignant narcissist becomes more involved in this psychological gratification, in the context of the right conditions, the narcissist is apt to develop the \href{https://en.wikipedia.org/wiki/Antisocial_personality_disorder}{antisocial}, the \href{https://en.wikipedia.org/wiki/Paranoid_personality_disorder}{paranoid}, and the \href{https://en.wikipedia.org/wiki/Schizoid}{schizoid} personality disorders.

The term \href{https://en.wikipedia.org/wiki/Malignant}{malignant} is added to the term \textit{narcissist} to indicate that individuals with this disorder have a severe form of narcissistic disorder that is characterized also by features of paranoia, psychopathy (anti-social behaviors), aggression, and sadism according to Kernberg and colleagues.[67]

\paragraph{Medical narcissism.} Medical narcissism is a term coined by John Banja in his book, \textit{Medical Errors and Medical Narcissism}.[68][69]

Banja defines ``medical narcissism'' as the need of health professionals to preserve their \href{https://en.wikipedia.org/wiki/Self-esteem}{self-esteem} leading to the compromise of error disclosure to patients.

In the book he explores the psychological, ethical and legal effects of \href{https://en.wikipedia.org/wiki/Medical_error}{medical errors} and the extent to which a need to constantly assert their competence can cause otherwise capable, and even exceptional, professionals to fall into narcissistic traps.

He claims that:
\begin{quotation}
    $\ldots$ most health professionals (in fact, most professionals of any ilk) work on cultivating a self that exudes authority, control, knowledge, competence and respectability.
    
    It's the narcissist in us all--we dread appearing stupid or incompetent.
\end{quotation}

\paragraph{In the workplace.} Main article: \href{https://en.wikipedia.org/wiki/Narcissism_in_the_workplace}{Narcissism in the workplace}.

Narcissism as a personality trait, generally assessed with the \href{https://en.wikipedia.org/wiki/Narcissistic_Personality_Inventory}{Narcissistic Personality Inventory}, is related to some types of behavior in the workplace.

E.g., individuals high in narcissism inventories are more likely to engage in \href{https://en.wikipedia.org/wiki/Counterproductive_work_behavior}{counterproductive work behavior} (CWB, behavior that harms organizations or other people in the workplace).[70]

Although individuals high in narcissism inventories might engage in more aggressive (and counterproductive) behaviors, they mainly do so when their self-esteem is threatened.[71]

Thus \fbox{narcissistic employees are more likely to engage in CWB when they feel threatened}.[72]

\fbox{Individuals high in narcissism have fragile self-esteem and are easily threatened.}

1 study found that employees who are high on narcissism are more likely to perceive the behaviors of others in the workplace as abusive and threatening than individuals who are low on narcissism.[73]

%
The narcissistic manager will have 2 main sources of \href{https://en.wikipedia.org/wiki/Narcissistic_supply}{narcissistic supply}: inanimate - \href{https://en.wikipedia.org/wiki/Status\_symbol}{status symbols} like company cars, company-issued smartphone or prestigious offices with window views; and animate - \href{https://en.wikipedia.org/wiki/Flattery}{flattery} and \href{https://en.wikipedia.org/wiki/Attention}{attention} from colleagues and subordinates.[74]:143

Teammates may find everyday offers of support swiftly turn them into enabling sources of permanent supply, unless they are very careful to maintain proper boundaries.[74]:143, 181

The need to protect such supply networks will prevent the narcissistic managers from taking objective decisions;[75] while long-term strategies will be evaluated according to their potential for attention-gaining for the manager themself.[74]:122

Organizational psychologist Alan Downs wrote a book in 1997 describing corporate narcissism.[76]

He explores high-profile corporate leaders (such as \href{https://en.wikipedia.org/wiki/Al_Dunlap}{Al Dunlap} and \href{https://en.wikipedia.org/wiki/Robert_Eugene_Allen}{Robert Allen}) who, he suggests, literally have only 1 thing on their minds: \href{https://en.wikipedia.org/wiki/Profit_(economics)}{profits}.

According to Downs, such narrow focus actually may yield positive short-term benefits, but ultimately it drags down individual employees as well as entire companies.

Alternative thinking is proposed, and some firms now utilizing these options are examined.

Downs' theories are relevant to those suggested by Victor Hill in his book, \textit{Corporate Narcissism in Accounting Firms Australia}.[77]

\paragraph{Primordial narcissism.} Psychiatrist \href{https://en.wikipedia.org/wiki/Ernst_Simmel}{Ernst Simmel} 1st defined primordial narcissism in 1944.[78]

Simmel's fundamental thesis is that the most primitive stage of libidinal development is not the oral, but the gastrointestinal one.

Mouth and anus are merely to be considered as the terminal parts of this organic zone.

Simmel terms the psychological condition of prenatal existence ``primordial narcissism.''

It is the vegetative stage of the pre-ego, identical with the id.

At this stage there is complete instinctual repose, manifested in unconsciousness.

Satiation of the gastrointestinal zone, the representative of the instinct of self-preservation, can bring back this complete instinctual repose, which, under pathological conditions, can become the aim of the instinct.

Contrary to \href{https://en.wikipedia.org/wiki/Christopher_Lasch}{Lasch}, \href{https://en.wikipedia.org/wiki/Bernard_Stiegler}{Bernard Stiegler} argues in his book, \href{https://en.wikipedia.org/wiki/Acting_Out_(book)}{\textit{Acting Out}}, that \href{https://en.wikipedia.org/wiki/Consumer_capitalism}{consumer capitalism} is in fact destructive of what he calls primordial narcissism, without which it is not possible to extend love to others.[79]

In other words, he is referring to the natural state of an infant as a \href{https://en.wikipedia.org/wiki/Human_fetus}{fetus} and in the 1st few days of its life, before it has learned that other people exist besides itself, and therefore cannot possibly be aware that they are human beings with feelings, rather than having anything to do with actual narcissism.

\paragraph{Sexual narcissism.} Sexual narcissism has been described as an \href{https://en.wikipedia.org/wiki/Egocentric}{egocentric} pattern of sexual behavior that involves an inflated sense of sexual ability and sexual entitlement.

In addition, sexual narcissism is the erotic preoccupation with oneself as a superb lover through a desire to merge sexually with a mirror image of oneself.

Sexual narcissism is an intimacy dysfunction in which sexual exploits are pursued, generally in the form of extramarital affairs, to overcompensate for low self-esteem and an inability to experience true intimacy.[80]

This behavioral pattern is believed to be more common in men than in women and has been tied to \href{https://en.wikipedia.org/wiki/Domestic_violence}{domestic violence} in men and sexual coercion in couples.[81][82]

Hurlbert argues that sex is a natural biological given and therefore cannot be deemed as an addiction.

He and his colleagues assert that any \href{https://en.wikipedia.org/wiki/Sexual_addiction}{sexual addiction} is nothing more than a misnomer for what is actually sexual narcissism or sexual compulsivity.[83]

While Hurlbert writes mainly of sexual narcissism in men, Schoenewolf (2013) describes what he calls ``gender narcissism"'' which occurs in both males and females who compensate for feelings of sexual inadequacy by becoming overly proud and obsessed with their masculinity or femininity.[84]

\subsubsection{Narcissistic parents}
Main article: \href{https://en.wikipedia.org/wiki/Narcissistic_parents}{Narcissistic parents}.

Narcissistic parents demand certain behavior from their children because they see the children as extensions of themselves, and need the children to represent them in the world in ways that meet the parents' emotional needs.

This parenting `style' most often results in estranged relationships with the children, coupled with feelings of resentment and self-destructive tendencies.[61]

They would intentionally create a situation to collect \href{https://en.wikipedia.org/wiki/Narcissistic_supply}{Narcissistic supply}.

\subsubsection{Narcissistic leadership}
Main article: \href{https://en.wikipedia.org/wiki/Narcissistic_leadership}{Narcissistic leadership}.

Narcissistic leadership is a common form of \href{https://en.wikipedia.org/wiki/Leadership}{leadership}.

The narcissism may be healthy or destructive although there is a continuum between the 2.

A study published in the journal \textit{Personality and Social Psychology Bulletin} suggests that when a group is without a leader, a narcissist is apt to take charge.

Researchers found that people who score high in narcissism tend to emerge as group leaders.[85]

\subsection{In culture \& society}
\begin{quotation}
    According to recent cultural criticism, Narcissus has replaced Oedipus as the myth of our time.
    
    Narcissism is now seen to be at the root of everything from the ill-fated romance with violent revolution to the enthralled mass consumption of state-of-the-art products and the `lifestyles of the rich and famous'. -- Jessica Benjamin (2000), ``The Oedipal Riddle,'' p. 233[86]
\end{quotation}
Some critics contend that \href{https://en.wikipedia.org/wiki/Popular_culture}{pop culture} has become more narcissistic in recent decades.[87]

This claim is supported by scholarship indicating some celebrities hire ``fake \href{https://en.wikipedia.org/wiki/Paparazzi}{paparazzi}'',[88] the frequency with which ``\href{https://en.wikipedia.org/wiki/Reality_TV}{reality TV}'' programs populate the television schedules,[87] and the growth of an online culture in which digital media, \href{https://en.wikipedia.org/wiki/Social_media}{social media} and the ``will-to-fame'' are generating a ``new era of public narcissism [that] is mutating with new media forms.''[89]

In this analysis, narcissism, rather than being the pathologized property of a discrete personality type, has been asserted as a constituent cultural feature of an entire generation since the end of World War II.[90][91][92]

%
Supporting the contention that \href{https://en.wikipedia.org/wiki/American_culture}{American culture} has become more narcissistic and that this is increasingly reflected in its cultural products is an analysis of US popular song lyrics between 1987 and 2007.

This found a growth in the use of 1st-person singular pronouns, reflecting a greater focus on the self, and also of references to antisocial behavior; during the same period, there was a diminution of words reflecting a focus on others, positive emotions, and social interactions.[93][94]

Similar patterns of change in cultural production are observable in other Western states.

A linguistic analysis of the largest circulation Norwegian newspaper found that the use of self-focused and individualistic terms increased in frequency by 69\% between 1984 and 2005 while collectivist terms declined by 32\%.[94]

References to narcissism and self-esteem in American popular print media have experienced vast inflation since the late 1980s.[94]

Between 1987 and 2007 direct mentions of self-esteem in leading US newspapers and magazines increased by 4,540\% while narcissism, which had been almost non-existent in the press during the 1970s, was referred to over 5,000 times between 2002 and 2007.[94]

%
Cross-cultural studies of differences in narcissism are rare.

Instead, as there is a positive association between narcissism and \href{https://en.wikipedia.org/wiki/Individualism}{individualism} and a negative one between it and \href{https://en.wikipedia.org/wiki/Collectivism}{collectivism}, these traits have been used as proxies for narcissism in some studies.[95]

This approach, however, risks the misapplication of the concepts of individualism and collectivism to create overly-fixed, ``caricature-like'',[96] oppositional categories.[97]

Nonetheless, 1 study looked at differences in advertising products between an individualistic culture, America, and a collectivist one, South Korea.

In American magazine advertisements, it found, there was a greater tendency to stress the distinctiveness and uniqueness of the person; conversely the South Korean ones stressed the importance of social conformity and harmony.[95]

This observation holds true for a cross-cultural analysis across a wide range of cultural outputs where individualistic national cultures produce more individualistic cultural products and collectivist national cultures produce more collectivist national products; these cultural effects were greater than the effects of individual differences within national cultures.[95]

\subsection{In fiction}
\begin{itemize}
    \item Maisie Farange, in \href{https://en.wikipedia.org/wiki/Henry_James}{Henry James}' novel \href{https://en.wikipedia.org/wiki/What_Maisie_Knew}{\textit{What Maisie Knew}} (1897), is neglected by her vain and self-absorbed parents.
        
    After her parents divorce, find new partners, and ultimately cheat again on their new partners, Maisie finally decides to move in with the morally strong family maid.
    \item \href{https://en.wikipedia.org/wiki/Young_Adult_(film)#Cast}{Mavis Gary}, \href{https://en.wikipedia.org/wiki/Charlize_Theron}{Charlize Theron}'s character in \href{https://en.wikipedia.org/wiki/Young_Adult_(film)}{\textit{Young Adult}} (2011), who conspires to return to her hometown to steal back her high school sweetheart, despite his being married with a child, embodies many narcissistic traits--including the emptiness she feels when she is not receiving attention or praise from her writing.
    \item \href{https://en.wikipedia.org/wiki/Jay_Gatsby}{Jay Gatsby}, the eponymous character of \href{https://en.wikipedia.org/wiki/F._Scott_Fitzgerald}{F. Scott Fitzgerald}'s novel \href{https://en.wikipedia.org/wiki/The_Great_Gatsby}{\textit{The Great Gatsby}} (1925), ``an archetype of self-made American men seeking to join \href{https://en.wikipedia.org/wiki/Upper_class}{high society}'', has been described as a ``pathological narcissist'' for whom the ``\href{https://en.wikipedia.org/wiki/Ego_ideal}{ego-ideal}'' has become ``inflated and destructive'' and whose ``grandiose lies, poor sense of reality, sense of entitlement, and exploitive treatment of others'' conspire toward his own demise.[98]
    \item \href{https://en.wikipedia.org/wiki/Gordon_Gekko}{Gordon Gekko}, the \href{https://en.wikipedia.org/wiki/Fictional_character}{fictional character} in the film \href{https://en.wikipedia.org/wiki/Wall_Street_(1987_film)}{\textit{Wall Street}} (1987) and its sequel \href{https://en.wikipedia.org/wiki/Wall_Street:_Money_Never_Sleeps}{\textit{Wall Street: Money Never Sleeps}} (2010),[99] has become a symbol in popular culture for unrestrained \href{https://en.wikipedia.org/wiki/Greed}{greed} and self-interest (with the signature line, ``Greed, for lack of a better word, is good''), often in fields outside \href{https://en.wikipedia.org/wiki/Corporate_finance}{corporate finance}.
    \item \href{https://en.wikipedia.org/wiki/Charles_Foster_Kane}{Charles Foster Kane}, a \href{https://en.wikipedia.org/wiki/Fictional_character}{fictional character} and the subject of \href{https://en.wikipedia.org/wiki/Orson_Welles}{\textit{Orson Welles}}' film \href{https://en.wikipedia.org/wiki/Citizen_Kane}{\textit{Citizen Kane}} (\href{https://en.wikipedia.org/wiki/1941_in_film}{1941}), which explores the life of the titular character, who is widely believed to be based on the publishing tycoon \href{https://en.wikipedia.org/wiki/William_Randolph_Hearst}{William Randolph Hearst}.
    
    Welles played Kane (receiving an Oscar nomination) in addition to producing, co-writing, and directing the film, while \href{https://en.wikipedia.org/wiki/Buddy_Swan}{Buddy Swan} played Kane as a child.
    
    In 1871, Kane's mother puts him under the guardianship of a New York City banker named Walter Parks Thatcher, who raises him in luxury.
    
    As an adult, Kane takes control of a \href{https://en.wikipedia.org/wiki/Newspaper}{newspaper}, which he uses to advance businesses in which Kane holds stock.
    
    Kane also hires staff members away from the rival \textit{Chronicle} newspaper, regarding them as collectibles.
    
    To finance the fledgling \textit{Inquirer}, Kane uses his personal resources, which allowed him to operate it, even at a million dollar annual loss, for decades.
    \item In \href{https://en.wikipedia.org/wiki/George_R._R._Martin}{George R. R. Martin}'s \href{https://en.wikipedia.org/wiki/A_Song_of_Ice_and_Fire}{\textit{A Song of Ice and Fire}} series and its television adaptation, \href{https://en.wikipedia.org/wiki/Game_of_Thrones}{\textit{Game of Thrones}}:
    \begin{itemize}
        \item The \href{https://en.wikipedia.org/wiki/Lannister}{Lannisters} have been deemed a ``family of narcissists''.[100]
        
        Licensed Mental Health Counselor (LMHC) Colleen Jordan has said the incestuous twins \href{https://en.wikipedia.org/wiki/Cersei_Lannister}{Cersei} and \href{https://en.wikipedia.org/wiki/Jaime_Lannister}{Jaime} have a combination of \href{https://en.wikipedia.org/wiki/Borderline_personality_disorder}{borderline personality disorder} and narcissistic personality disorder, and their younger brother Tyrion is an \href{https://en.wikipedia.org/wiki/Tyrion_Lannister}{alcoholic narcissist}.[101][100]
        
        Additionally, a clinical psychologist posted as Redditor Rain12913: ``People seem to be falling into the trap of thinking that Cersei really does genuinely love her brother and her (late) children.
        
        While she certainly says that she does quite a bit, and while her behavior may seem to suggest that she does, it is highly unlikely that such a narcissistic character is capable of true love.''[102]
        
        About the family's patriarch, Jordan observes that ``\href{https://en.wikipedia.org/wiki/Tywin_Lannister}{Tywin Lannister} is actually the worst of them''.[100]
        \item Of Lord \href{https://en.wikipedia.org/wiki/Petyr_Baelish}{Petyr Baelish} (nicknamed ``Littlefinger'') Jordan observes: ``If you look at Littlefinger, we know he's not remotely personally interested in Lysa, but he likes the attention.
        
        And he needs her.
        
        Narcissists use people for functions, which he does.''.[100]
    \end{itemize}
    \item \href{https://en.wikipedia.org/wiki/To_Die_For#Cast}{Suzanne Stone-Maretto}, \href{https://en.wikipedia.org/wiki/Nicole_Kidman}{Nicole Kidman}'s character in the film \href{https://en.wikipedia.org/wiki/To_Die_For}{\textit{To Die For}} (1995), wants to appear on television at all costs, even if this involves murdering her husband.
    
    A psychiatric assessment of her character noted that she ``was seen as a prototypical narcissistic person by the raters: on average, she satisfied 8 of 9 criteria for narcissistic personality disorder$\ldots$ had she been evaluated for personality disorders, she would receive a diagnosis of narcissistic personality disorder''.[103]
\end{itemize}

\subsection{See also}
\begin{itemize}
    \item Codependency
    \item Counterdependency
    \item Dark triad
    \item Dorian Gray syndrome
    \item Ego ideal
    \item Egotism
    \item Ethical egoism
    \item Gender differences in narcissism
    \item Illusory superiority
    \item Jointness (psychodynamics)
    \item Megalomania
    \item Narcissism of small differences
    \item Narcissistic elation
    \item Narcissistic mortification
    \item Narcissistic personality disorder
    \item Narcissistic withdrawal
    \item Narcissistic supply
    \item Optimism bias
    \item Pride
    \item Psychopathy
    \item Selfishness
    \item Superiority complex
    \item True self and false self
    \item Vulnerable narcissism
    \item Grandiose narcissism\hfill$\square$
\end{itemize}

%------------------------------------------------------------------------------%

\chapter{Shahida Arabi (2016). Becoming the Narcissist's Nightmare: How to Devalue \& Discard the Narcissist While Supplying Yourself}

\section*{Preface}
A few years ago on a survivor's forum, Arabi asked survivors what made them a narcissist's worst nightmare - the strengths, the talents, the assets, the attitudes, and the actions that made them absolutely deplorable ``supply'' for abusive and toxic people.

%
Survivors came up with epic lists of traits that annoyed and frustrated their narcissistic partners - this included having an incredible sense of humor and quick wit that could put a narcissist's covert put-down to shame, the psychological resilience they had developed over the course of the relationship, their ability to empathize with others, their ability to intelligently and astutely question the narcissist's false mask, their ability to emotionally connect with others, their thirst for knowledge, their confidence, their sensitivity, their compassion - endless lists of amazing qualities that narcissists feared and abhorred.

%
What occurred to Arabi was that our strengths - the ones that narcissists often convince us are weaknesses - are the very things that can \textit{save} us from narcissists, which is \textit{why narcissists work so very hard to diminish these strengths in the 1st place}.

Arabi also realized something even more incredible: that the techniques narcissists use against use can also be merged with those strengths to help us transcend and thrive after narcissistic abuse.

%
\fbox{The very same techniques that narcissists use on us are the very same ones we must use to get over them.}

Confused?

Let Arabi explain.

Arabi is not suggesting becoming a narcissist ourselves or being cruel - not at all.

These techniques will not be employed in the same way as a narcissist uses them - they will be adapted to detach from the narcissist altogether.

Read the following and see if you agree with Arabi here:

%
You once \textbf{idealized} the narcissist, put him or her on a pedestal after he or she did the same to you.

You saw them as the love of your life - at least, you saw their false self as the love of your life.

%
Now, you must \textbf{\textit{devalue \& discard}} the narcissist - both in your mind and in your physical reality, if you're still maintaining contact with him or her somehow.

In this book, these terms will not only be explained in the context of abuse, but be adapted to the survivor's own journey of detaching and healing from a narcissist.

%
\textbf{Devaluing the narcissist} means reconnecting with the reality of who the narcissist is, not who you wish him\texttt{/}her to be (the false self they once presented to you).

In essence, devaluing in this context means dismantling your perception of their false self and replacing it with the reality of their true, abusive self.

%
Chaps. 1--2 will give you the information, knowledge and tools to identify the covert manipulation tactics of narcissistic abusers as well as their motives and intentions.

%
\textbf{Discarding the narcissist} means going No Contact with the narcissist or Low Contact if you share children, have legal matters to sort out with the narcissist or have any other circumstances that prevent you from cutting all ties with this person.

This involves giving yourself closure by not giving the narcissist any.

In Chap. 5, you will learn more about the journey to No Contact.

%
\textbf{Re-idealizing \& supplying yourself} refers to using the tips and tools Arabi will mention Chap. 3 to release the toxic brainwashing from your narcissistic partner and begin to heal some of the biochemical and trauma bonds that have tethered you to your narcissistic abuser.

This also involves creating a healthier, positive \textit{reverse discourse} that substitutes self-defeating beliefs with empowering ones about yourself.

%
\textbf{Triangulation} in this case means welcoming new people into your life that will serve as a support network to you during this time.

All this time, the narcissist has triangulated you with others to validate their own sense of superiority.

Now you must ``triangulate'' your narcissist with a new support system to validate your experiences.

This is not meant to make the narcissist jealous - it is meant to give yourself the resources and empowerment you need to heal.

%
Due to your empathy and ability to emotionally connect, this triangulation has the added benefit of adding to your psychological resilience.

Having at least 1 person who has your back and can snap you back to reality, who knows what you have been through, is invaluable.

Having a whole community of other survivors who have your back?

Well, now you're bound to be invincible.

We will learn more about how to connect with these support networks while also getting rid of any defunct or toxic social networks (including the narcissist's own harem) that no longer serve your emotional well-being in Chap. 5.

%
\textbf{Reverse Gaslighting} means that whenever the narcissist tries to pull a fast one on you, being able to use the knowledge about their manipulative tactics which you will learn in the 1st 2 chapters to distance yourself.

This means \textit{going No Contact or Low Contact in response to their silent treatments, their stonewalling behavior, their petty put-downs and manipulation in order to prevent yourself from being pulled right back into their mind games}.

%
``Reverse Gaslight'' the narcissist by saying mentally to their accusations, projections and gaslighting tactics: \textit{I don't believe you}.

\textit{That's not what happened}.

\textit{I know my truth \& I own my truth}.

Reverse Gaslighting the narcissist is not actually gaslighting - it involves staying grounded in your reality and your own powerful truth while invalidating the lies of a narcissist.

This will help to relieve some of the cognitive dissonance you may have been experiencing as a result of being in a relationship with a narcissistic abuser who often distorts and manipulates your perception of reality.

%
\textbf{Create a false self for the time being that serves as armor if \& when you have to interact with the narcissist or even just start ruminating over them.}

This self is the strong, logical self that you must wear as you begin to detach and heal from a narcissist.

The one you may not feel like during No Contact or Low Contact, but the one you must be, in order to resist their attempts to retraumatize you and bring you back into the relationship be pressing the reset button.

Use this self to interact with the narcissist if you have to interact with them.

You will learn how to do this effectively through the methods we will discuss in Chaps. 3--6.

%
\textbf{Using this false self enables you to be self-protective, centered, \& unreactive.} Mirror them when you have to, much like they mirrored you during the idealization phase, by withdrawing from them when they withdraw, rather than pleading and begging for them.

Strive \textbf{not} to mirror them when they attempt to pull you back into the trauma of the relationship.

Do not share secrets and do not make them privy to your innermost feelings; in fact, Arabi encourages you not to disclose anything about your personal life once you have realized your partner may be a narcissist.

You are permitted to stop giving any information in the case where the narcissist may try to emotionally blackmail or manipulate you.

They played mind games all along - now it's their turn to not have the facts at hand.

%
\textbf{Finally, think of the narcissist as supply.} You do not need them because you have other sources of attention - healthier sources.

\fbox{Think of them as rotten supply.}

\fbox{They cannot ``work'' for you anymore, because they are not functional adults in relationships.}

\fbox{They are children in adult bodies, doing real harm to others.}

%
So you see, \textit{it's not about becoming the narcissist}, stooping down to their level or even counter-manipulating them - it's about conquering your own thoughts, beliefs and actions regarding the narcissist and adapting the very same techniques that he or she used to entrap you in the 1st place - except, in this case, you use them to set yourself free.

%
\fbox{Ultimately, it is not your ability to beat the narcissist at his or her own game that frightens them.}

\fbox{Playing games with someone with no remorse or empathy is sure to fail.}

Rather, it is your ability to seek your own validation and move forward into your success, channeling the experiences that were meant to destroy you into your greatest victories, that is appalling to these predators.

How dare their former victims become independent of their bullying and become even greater in spite of it?

Oh, but they can, and they will.

\section*{Introduction: My Story is Your Story}
\begin{quotation}\it
    ``I grew up with a narcissistic abusive father and then I spent 12 years of my adult life repeating that dysfunctional pattern when I married a narcissistic abuser.
    
    In the context of my 12-year long relationship, I never felt safe, emotionally, physically, financially or psychologically.
    
    I was never ``good enough'' to please him, even though I gave him the best years of my life and spent my life savings investing in his ``dream.''
    
    I didn't feel safe to eve have children with him$\ldots$
    
    I finally left him and 2 years later, our divorce has gone nowhere.
    
    I am not with him now, but we are still married and he's destroyed the business we built, wrecked my finances and credit.
    
    I am still struggling to find the courage and strength to end this nightmare.
    
    2 years of depression and darkness, living in fear of him and still attempting to support his dreams even though he never loved me or gave me anything in return.'' - Hallie, Survivor from Kentucky
\end{quotation}
For those of us who have experienced narcissistic abuse since childhood, many of us are familiar with the trauma bonds that can keep us locked in the vicious cycle of meeting and mating with narcissists.

Growing up with a narcissistic parent and witnessing narcissistic abuse was the precursor to the destructive, toxic relationships Arabi had with narcissists - from friends to relationship partners to acquaintances to co-workers.

Years later, Arabi can say that the experiences of narcissistic abuse were more of a gift than a curse, because Arabi learned to channel it into some of the greatest victories of Arabi's life.

%
Narcissistic abuse can be a vicious cycle, especially for those of us who have experienced it when we were particularly vulnerable as children.

Before Arabi met narcissistic romantic partners, Arabi experienced severe bullying in addition to witnessing domestic violence during Arabi's childhood.

These experiences essentially primed Arabi for narcissistic abuse.

Arabi's 1st ever boyfriend was a grandiose narcissist who would covertly and overtly put Arabi down, wounding Arabi with a sense of unworthiness that Arabi carried forward in her early adulthood.

Even since then, Arabi felt ``cursed'' to meet more narcissists and Arabi did.

Other dating and relationship partners as well as friends followed similar patterns of triangulation (manufacturing love triangles), stonewalling, emotional invalidation and verbal abuse.

It was not until Arabi discovered what narcissistic abuse was that Arabi came to realize that this pattern had been ingrained within Arabi ever since Arabi was a young child.

Although Arabi knew Arabi was not to blame for any of the abuse Arabi received, Arabi recognized how these patterns had come to be.

Back then, Arabi was not aware that Arabi's subconscious wounding was enabling Arabi to stay tethered to these toxic relationships.

Our subconscious mind is incredibly powerful; it carries forth the core wounds from childhood and does everything in its power to reinforce those wounds and prove them right.

Arabi was unintentionally gravitating towards people who reminded Arabi of Arabi's narcissistic parent, of Arabi's 1st boyfriend and of Arabi's bulling classmates; they were also gravitating towards Arabi as Arabi had not yet healed her wounds.

Arabi had not yet fully embraced herself as a whole, healthy divine being worthy of love, compassion and respect.

%
It was 1 of the worst narcissistic abusers Arabi had ever that finally led Arabi to break the cycle.

This relationship had her reeling.

Arabi had encountered emotionally abusive people before, but this particular relationship was special in how exaggerated it was, almost like a caricature of the narcissistic abuse cycle - idealization, devaluation and discard on steroids.

%
Everything about this relationship fit the criteria for narcissistic abuse to a tee - the idealization phase which was heavy with surveillance and an excessive amount of investment early on; the constant triangulation with other individuals (even complete strangers); chronic stonewalling during arguments and devaluation in the form of degrading, condescending comments that never seemed to end.

There were also sleepless nights filled with countless discussions initiated by this partner over seemingly irrelevant things, incredible blameshifting, terrifying narcissistic rage, lies, put-downs and projection as well as verbal abuse.

%
This devaluation was mixed in with the spurts of unexpected idealization and shallow flattery - what Arabi now knows to be intermittent reinforcement at its finest.

Anyone looking from the outside who understood narcissism could easily identify that it was psychological and emotional abuse.

Arabi had tried to deny and minimize this abuse throughout the relationship but by the end of the horrific discard, Arabi knew she had experienced something that was going to alter the course of her life.

%
\fbox{It was difficult, but Arabi knew she had to let the relationship go and never contact his person again.}

The way this partner had humiliated Arabi and put Arabi down made this a great deal easier.

Arabi now had no choice but to move forward with her life.

\fbox{The gift of this particular relationship was that it unraveled new knowledge and new networks.}

Shortly after going No Contact with Arabi's ex-partner, Arabi was lucky enough to find an online community through a post Arabi had read regarding abuse recovery during Arabi's tumultuous relationship.

On a hunch, Arabi googled the name of the community, not realizing that this would be the life-changing step that would transform everything Arabi knew to be true.

%
This same community led Arabi to read a book that would change Arabi's life forever, a book called \textit{Psychopath Free} by the community's founder, the brilliant and perceptive Jackson MacKenzie.

A survivor himself, he wrote about the manipulative tactics of abusive narcissists in a deeply personal yet eerily accurate way.

This book and its accompanying community of survivors opened Arabi's eyes to the fact that Arabi was not alone in her experiences; far from it.

The accounts of other survivors, their validation and their encouragement helped Arabi to regain hope in rebuilding a better life, one without abuse and disrespect.

%
This was when Arabi began to more readily identify the convert manipulation tactics of emotional abusers.

1 of the most incredible components of Arabi's healing journey was the powerful knowledge Arabi gained along the way.

Arabi was not necessarily new to narcissism; at the age of 18, while most people her age were reading Cosmopolitan, Arabi was reading Sam Vaknin to understand how manipulative, cruel and callous partners operated.

While Arabi had briefly read about the narcissistic personality and had begun to realize that Arabi's 1st boyfriend fit the criteria with his endless self-absorbed monologues, cruel comparisons, and degrading comments, Arabi did not learn or research the full complexity of narcissistic abuse until this other short-term relationship served as the catalyst in Arabi's journey to fully recognize and realize what Arabi had been dealing with.

%
A few years later, Arabi began to revisit the literature on sociopaths and narcissists in a way that was even more eye-opening, delving into the actual techniques that abusive individuals used.

The more and more books Arabi read, the more Arabi learned about how deliberate these tactics were and how many abusers were unlikely to change.

Arabi read books on narcissistic abusers and antisocial personalities by experts on this topic e.g. Dr. Robert D. Hare, Dr. George Simon, Dr. Martha Stout, and Lundy Bancroft, many of whom had worked with these personalities as clients.

%
In addition, Arabi tapped into the infinite wisdom of other survivors, those who had been through the abuse cycle themselves.

Arabi regularly visited blogs on emotional abuse, e.g. the information-rich After Narcissistic Abuse blog and the compassionate Healing from Complex and PTSD website.

Arabi frequently spoke with other survivors as they supported each other on their journey to healing.

The fact that they were all transforming themselves after this heartbreaking experience - or for people like Arabi, a set of heartbreaking experiences - was exhilarating, frightening and emotional.

It was also \textit{life-changing}.

%
Psychology was no stranger to Arabi; as fate would have it, Arabi had studied it extensively through Arabi's coursework in school and had even taken a graduate-level course on psychopathology where Arabi 1st learned the DSM definition of narcissism.

Ironically, however, Arabi had no idea at the time that Arabi was taking this class, Arabi was also in a relationship with a narcissistic abuser.

That is because many psychology books do not talk specifically about \textit{narcissistic abuse}.

%
Arabi also knew licensed mental health counselors with whom Arabi regularly spoke about these topics with; yet even \textit{they} had not been taught about the intricacies of these types of toxic relationships and how far these manipulative behaviors went.

It became a hobby of Arabi to consume psychology articles and books to learn more about this type of abuse.

Narcisstic abuse had never been fully explored in Arabi's psychology classes and Arabi developed a thirst to uncover the intentions behind the mask that many abusers presented to the world.

%
It is important to note that \fbox{knowledge alone, while powerful, cannot heal narcissistic abuse.}

To heal from narcissistic abuse, Arabi had to employ different healing modalities that targeted the mind, the body and the spirit.

Self-exploration and self-improvement was not new to Arabi, but narcissistic abuse led her to a cathartic path of healing that Arabi never would have undertaken otherwise.

Due to the anxiety and stress Arabi experienced from witnessing unhealthy family dynamics, as well as being a survivor of other traumas that took place outside of Arabi's household e.g. bullyin,g abuse and sexual assault, Arabi had spent years experimenting with different forms of mindfulness and self-care.

These helped Arabi to reconnect with a sense of wholeness Arabi felt Arabi had lost in the midst of lifelong trauma.

%
Arabi's self-care regimen included everything you could possibly think of: Arabi read hundreds of self-help books, created a daily practice of positive affirmations, cleaned up Arabi's diet, learned to do different types of yoga e.g. Hatha yoga, Vinyasa yoga and Bikram yoga; Arabi also tried Reiki healing, counseling, meditation, kickboxing, dance classes, art therapy and keeping a gratitude journal.

Arabi loved incorporating body and mind in activities geared towards greater healing and self-care.

Each part of Arabi's regimen helped to build a stronger foundation where Arabi was pulling herself towards self-compassion, self-respect, and self-love.

The ending of the toxic relationship, combined with the support of Arabi's fellow survivors, was the final push that made Arabi realize that Arabi could also help other survivors with what Arabi was learning.

%
That is when Arabi took an important step towards Arabi's own dreams: Arabi channeled all that Arabi had learned from Arabi's previous relationships, everything Arabi learned from Arabi's self-care activities, as well as the research Arabi had read about narcissistic abuse into writing a book Arabi had written pieces of but had never had the courage to finish.

It was a book about self-care for young women called \textit{The Smart Girl's Guide to Self-Care}.

Miraculously, this book would later become an Amazon Best Seller in 3 categories, its rank right next to 1 of Arabi's favorite authors in the category of Women's Personal Growth.

%
This is also the book Arabi wrote during No Contact with Arabi's narcissistic ex-partner and Arabi wouldn't have had it any other way.

Every day of No Contact, Arabi made a commitment: keep writing.

Instead of tethering herself to more toxic people, Arabi broke the cycle through her words and rewrote the script of her life.

Arabi kept writing and revising her way into rebuilding a better life while also sharing her story to help others find their way to self-care after abuse and trauma.

%
Life after No Contact was a challenging yet flourishing time for Arabi; it seemed that without a narcissistic partner, Arabi opened up a new space of healing and empowerment where love, community and miracles could enter.

1 of the biggest miracles Arabi experienced was writing an accompanying blog to Arabi's book called Self-Care Haven, which taught people about the warning signs of emotional abuse.

To Arabi's surprise, Arabi's blog went viral due to an entry Arabi had written called ``5 Powerful Ways Abusive Narcissists Get Inside Your Head.''

In this entry, Arabi used all the knowledge and life experiences Arabi had gained on Arabi's journey and poured in into an article that survivors and mental health practitioners alike would come to share all over the world.

Within a few months, Arabi's blog had achieved 1 million and a half views.

%
The letters from abuse survivors around the world began pouring in; both survivors and clinical psychologists began sharing the work through social media.

Most importantly, Arabi got comments from survivors who told Arabi that the article had changed their lives and that they were ready to take action to finally end the toxic relationships that had held them back for years.

Arabi had always known that the best revenge was moving forward and chasing her dreams, but never had she recognized the value of moving on and sharing her story as deeply as Arabi did then.

\textit{The story Arabi had was powerful because it was a story many others had longed to share}.

Had Arabi never shared her story, Arabi would have never been able to connect with those who were also looking to transform their lives and share their own stories.

%
Arabi's miraculous connection with a growing community of abuse survivors did not end there.

During No Contact, Arabi couldn't sleep at night, so Arabi also began making YouTube videos on how to maintain No Contact with toxic ex-partners.

As Arabi spoke about her insights on No Contact and narcissism, survivors began to connect with Arabi about their own No Contact journey and began subscribing to Arabi's channel.

Arabi was amazed to have gotten even 100 subscribers at the beginning of Arabi's No Contact journey, but by the end of that year, Arabi had achieved $> 4000$.

%
Deep within Arabi's soul, Arabi knew she was on the right path; for the 1st time in a very long time, things clicked into place and Arabi felt the aves of revolution as Arabi saw the community around Arabi growing exponentially - a community that fully resonated with Arabi's work and what Arabi had experienced.

A community of survivors and warriors who knew exactly what it felt like to be abused and the courage it took to tackle th numerous hardships after abuse.

\fbox{Life after narcissistic abuse meant the miracle of recognition.}

Arabi understood that part of Arabi's density was to help other survivors like Arabi - some of whom had also experienced emotional abuse in childhood that they later re-experienced in adulthood with their narcissistic partners.

%
Arabi had no idea what Arabi had started and Arabi was astounded at the feedback Arabi was receiving.

Survivors from all different backgrounds and experiences were reaching out to Arabi to share their stories of narcissistic abuse; even more powerfully, they were connecting with other survivors on Arabi's blog and supporting each other in loving, compassionate ways that filled Arabi with hope.

All Arabi knew was that Arabi was doing her best to channel the experiences life had given Arabi into the greater good and that the people who needed it the most - Arabi's fellow survivors, Arabi's fellow warriors, were finding healing too along with Arabi.

%
Arabi grew stronger; by the end of Arabi's 1 year of No Contact with that particular narcissist, Arabi knew Arabi stil had a great deal of learning but also more work to do.

Arabi regularly answered the questions of Arabi's readers regarding their narcissistic abuse experiences; Arabi invited survivors to share their stories on Arabi's blog posts and also rewrite the script of their lives, and it wasn't long before Arabi began a No Contact coaching program for survivors all over the world.

Arabi continued to apply what Arabi taught to others to her own personal life.

The next few times Arabi encountered narcissistic people, Arabi was able to detach from them more easily and quickly.

Arabi recognized the red flags, called them out, and eventually went on her way.

Arabi took more initiative in ending these relationships and friendships, going No Contact without further thought with people who were visibly toxic - people who did not support Arabi or her dreams.

%
Arabi refocused on the people who validated her and wanted her to rise rather than fall.

Arabi also made sure to validate herself and realized that while being among the few to recognize a narcissist was an alienating experience, it was also a liberating one.

There were many times Arabi saw behind the masks of toxic people, sociopaths or narcissists while others continued to believe in the false self they projected.

Instead of attempting to convince others of what Arabi observed, Arabi quietly turned the focus back onto herself and her own self-care.

Arabi stopped listening to the dark voices of others and began to reconnect with that divine light inside of her and other survivors.

Arabi knew th truth about toxic people and for the 1st time, Arabi's faith in herself was enough to break the spell.

%
It was by no means easy; sometimes it took longer for Arabi to detach from toxic people than Arabi felt it should have.

There were times when Arabi felt Arabi could have done better.

Yet Arabi treated herself compassionately and forgave herself for any failures, knowing that any types of ``relapse'' was simply an inevitable detour on the road to recovery.

So Arabi pushed forward and kept moving.

Arabi knew that each encounter with another narcissistic abuser, whether friend, foe or relationship partner, was simply a test - a test of how far Arabi's core wounds were still tethering her to toxic people.

%
The journey to healing is not simple; the destination is not some final end point at the end of some torturous maze; it is an exciting journey filled with ups and downs, with new learning experiences and challenges.

During life after narcissistic abuse, Arabi tried many different healing modalities to uncover Arabi's wounds.

Arabi went back far in childhook to better understand how those wounds have been created and how Arabi's subconscious had steered Arabi towards self-sabotage.

Arabi is happy to say that today, Arabi has much healthier boundaries, friendships and relationships precisely because Arabi has learned to set boundaries, recognize her self-worth and adopt a warrior mentality towards her own self-care.

No Contact with Arabi's toxic ex-partners has helped Arabi incredibly on her healing journey, but Arabi is not afraid to say that Arabi is always in the loving, evolving process of healing and becoming.

We all are.

The journey to healing is never really over, but that is the exciting thing about life after narcissistic abuse.

%
You get to heal in so many different ways.

Some powerful methods of healing for Arabi were through joining body and mind through meditating, writing, dancing, running, and stretching outwards in a yoga pose.

Arabi also stretched herself outwards in other ways - to the larger community of survivors, whose support was vital to Arabi's journey.

Arabi cannot stress enough the fact that the most important way Arabi healed was to reach out to other survivors through coaching, blogging, making videos and joining support groups.

\textit{Arabi discovered that when Arabi helped other people to heal, it reconnected her with that divine sense of love and purpose that was crucial for Arabi to rebuild her life}.

In this way, the experience of narcissistic abuse and all the lessons it has taught Arabi has prompted her to make miraculous shifts in her life.

%
Throughout Arabi's journey, Arabi learned that knowledge of narcissistic abuse was not something limited to licensed therapists and researchers - the accounts of real life abuse survivors provided crucial information regarding the patterns of these dysfunctional relationships.

\fbox{Those that offer a cure have often drunk the poison themselves.}

Experiencing narcissistic abuse yourself is quite different from learning about it from just a clinical perspective.

%
While there are people out there who have suffered narcissistic abuse without having ever met a narcissist previous to that singular experience, within the span of the last 7 years, Arabi has met and encountered many topic people - through dating, long-term relationships, school and the workplace.

Arabi has conducted - unintentionally - an ethnographic experiment of sorts in which Arabi has navigated and confronted different abusive styles, techniques that abusers and otherwise toxic people employ to manipulate their victims, and have observed these tactics in real-life settings.

All Arabi would have needed was a tape-recorder and a transcription of all these encounters to have made it a viable sociological experiment.

%
In addition to these real-life experiences, Arabi has had the privilege of reading hundreds of books on dating, relationships and pathological people from the experts.

Arabi has taken undergraduate and graduate-level courses in psychopathology.

All of these experiences and Arabi's research have been given Arabi the insight and knowledge to write this book and hopefully help other survivors of emotional abuse with what Arabi has learned on her journey.

From a spiritual perspective, Arabi feels her destiny was tied to learning more about abusive patterns and sharing her stories to strengthen and empower others who have undergone similar experiences.

You may not share the same perspective, and that's perfectly all right.

Some of you have undergone such severe suffering and trauma that placing this into a spiritual framework might be challenging.

All that matters is that you do what is most empowering for you and beneficial for your unique journey to healing.

That is why Arabi has written this book, to offer insights on how to spot emotional predators and ultimately, how to empower and begin to heal yourself after encountering them.

%
Throughout these encounters with emotional predators, Arabi has revolutionized her understanding of appropriate boundaries, reconnected with Arabi's authentic values, and empowered herself to recover from psychological abuse by using reverse discourse, self-care, and the channeling of her experiences into productive outlets.

Arabi's journey has not been perfect - it has been filled with adversity and the challenging of confronting and addressing unhealed wounds.

Yet Arabi has also written books, pursued her passions, cut toxic people out of her life, graduated from top universities, reached a larger community and have helped others on their own healing journeys.

Arabi knows 1sthand that a better life is possible and that you can always use what you have learned to fuel your success.

%
Arabi is so grateful for the opportunity to heal and to help others heal.

``\textit{I am grateful for every miracle that has allowed me to reach so many survivors}.''

We are all reaching for a greater life and this path can be paved when we begin to channel our pain into our healing, our crisis to our transformation, our crucifixion to our resurrection.

Arabi shares her story and the stories of other survivors in this book because Arabi wants all survivors of abuse and trauma, regardless of what their background is with narcissistic abuse, to know you should never give up.

%
That is why at the end of each chapter of this book, you will not only find articles by experienced therapists, writers, advocates, professionals and researchers on the topic of narcissistic abuse and the effects of trauma, you will also find a section called ``Survivor Insights,'' collected from hundreds of surveys responses Arabi has received from survivors all around the world.

These stories shed light on the complexities and intricacies of emotional and psychological abuse in a way that the clinical perspective often doesn't do justice to.

%
The accompanying articles in this book include works by bestselling authors, licensed therapists and popular bloggers and advocates on the topic.

Arabi is honored to include articles by therapist and author Andrea Schneider, LCSW, author and life coach Lisa A. Romano, award-winning blogger and founder of Healing Complex Trauma and PTSD Lilly Hope Lucario, Mental Health News Radio host Kristin Walker, founder of True Love Scam Jennifer Smith, therapist Ingrid Roekke, author and coach Kim Saeed and founder of Ending Abuse Media, Alison Soroka.

Arabi has also made sure to include the insights Arabi has gained from the thousands of survivor stories that Arabi has received on her platforms in addition to her personal experiences and research.

This is meant to help readers bridge the gap between the existing research on this topic and real-life experiences.

In Chap. 1, you will learn everything you need to know about narcissistic abuse, including more about the narcissistic abuse cycle, the manipulative tactics of abusive predators, the distinctions between narcissistic abuse and mistreatment by a selfish jerk, the origins of narcissism, ``translations'' of narcissistic behavior, common terms used in the realm of narcissistic abuse and clich\'es about normal relationships that do not work in the realm of abusive relationships.

You will also learn the victim-blaming myths and stereotypes that usually hold victims back from validating and understanding their own experiences.

%
Chap. 2 explores the biochemical, psychological and trauma bonds that keep abuse victims tethered to their abusers, long after they have discovered what they are dealing with, as well as the effects of trauma on the brain.

This is incredibly important to learn about, because many victims blame themselves for staying in the relationship too long without understanding how the effects of trauma keep us hooked to these toxic partners.

%
Chaps. 3--4 are arguably 2 of the most important chapters in this book, as they delve in-depth on how to combat each 1 these bonds, as well as the alternative and traditional methods you can use to heal from this type of abuse.

They also explore how you can tap into your warrior ``superpowers'' and strengths in order to detach from abusive people and rebuild an even more amazing life after narcissistic abuse.

%
Chap. 5 is helpful for those struggling with the journey of No Contact, as it contains resources, tips and methods to begin No Contact and maintain it.

In addition, it offers a list of 111 alternatives to breaking No Contact to prevent relapse.

%
Chap. 6 explores pathological narcissism in our society and culture, giving more insights on how to cope with narcissistic family members, friends, co-workers, and even online narcissists e.g. cyberbullies.

Chaps. 7--10 offer closing insights on what it means to be a survivor, how to deal from emotional abuse by owning our power and agency, as well as how to channel our traumatic experiences into victory by rewriting our stories and creating new narratives.

%
There will be a Frequently Asked Questions at the end of the book that you will not want to miss, as well as an additional resources list that will link to some of the most informative websites, articles and videos on narcissistic abuse.

%
Throughout the book, you will also find that there are links to ``Self-Care Haven'' meditations and reflections related to the themes of certain chapters that Arabi normally only offer to her coaching clients, which are now available to e-book readers as well.

%
As you read through this book, Arabi hopes you will recognize that your own story, when and if you are ready to share it, can help to heal others.

This experience, while alienating at 1st, can be channeled into a greater purpose and can be used to reconnect you with the larger world in ways you might never have thought possible.

Arabi knows there are moments when it seems hopeless and the pain seems unbearable, especially if you have a pattern of attaching to narcissistic people.

%
However, there is a whole other world that will open up for you in recovery and healing.

Be kind to yourself, even when you feel nobody is kind to you.

At the end of the day, you have to save yourself - over and over again, no matter what.

You and your story are needed.

You are worthy.

You are strong.

You are extraordinary.

You are a fighter.

You are a warrior.

%
Arabi encourages all survivors to take their own self-care journey after narcissistic abuse; you can empower yourself more fully by taking advantage of all the lessons it has to offer.

This experience can serve as a portal to healing and a challenge to rise above the ashes and resurrect to your best self.

%
No matter how much we learn and improve, we can always learn something new along the way.

The journey to healing is never really ``over'' - it is just beginning, and Arabi wants to share with you the empowering tools that have helped her towards her destination.

%
You will find that when you see the blessing in the darkness, many other blessings will also come to light - new support, new dreams to be fulfilled, new ways to connect to your divine inner guidance and God-given worth.

In you is something infinitely greater than whatever you are experiencing.

This light will always be within you, even during and after the most traumatic moments of your life.

%
Narcissists assume that they can break us down and destroy us, but the truth is, they can never destroy us.

Despite the damage they inflict, we are stronger than them, heart, mind, and soul.

As broken as you may feel right now, there is something within each and every 1 of you reading this book that is stronger than their abuse.

It is that part of you that knows you want to survive and thrive.

It is that part of you that you must connect to before you give up hope for a better life.

It is that part of you that will allow you to transcend their destruction and recreate yourself, more victorious than ever.

%
Your desire for knowledge and healing is just the beginning and it is a calling for something greater.

You do not belong in the shadows where you have been forced to hide.

\fbox{You deserve to be in the light with full force.}

Your density is tied to recognizing your worth and how utterly valuable you are to this world.

You have so much to give and you no longer have to sacrifice yourself for the needs of another person ever again.


You deserve to be seen and heard.

%
No matter who you are and where you are at this point of your life, you can always use the worst experiences of your life as the key to your best victories.

Life after narcissistic abuse is filled with miracles - you just have to be ready to reach out and let them in.

\section{Recognizing the Narcissist}
\begin{quotation}\it
    ``We are targeted by narcissists because we have everything that is beautiful and human.
    
    We can feel true joy and sadness.
    
    They want to live through us, and end up killing us.
    
    But we will rise, and we will be stronger and wiser.
    
    We need to love ourselves.
    
    We are not victims, we are survivors!
    
    This is the life lesson.'' - \emph{Dzana, Survivor from Stockholm}
\end{quotation}
In popular culture, the term ``narcissistic'' is thrown about quite loosely, usually referring to vanity and self-absorption.

This reduces narcissism to a common quality that everyone possesses and downplays the symptoms demonstrated by people with the actual disorder.

While narcissism does exist on a spectrum, narcissism as a full-fledged \textit{personality disorder} is quite different.

People who meet the criteria for Narcissistic Personality Disorder or those who have traits of Antisocial Personality Disorder can operate in extremely manipulative ways within the context of intimate relationships due to their deceitfulness, lack of empathy, and their tendency to be interpersonally exploitative.

Although Arabi will be focusing on narcissistic abusers in this book, due to the overlap of symptoms in these 2 disorders, this can potentially apply to interactions with those who have ASPD to an extent.

%
It is important in any kind of relationship that we learn to identify the red flags when interacting with people who display malignant narcissism and\texttt{/}or antisocial traits, so we can better protect ourselves from exploitation and abuse, set appropriate boundaries with others, and make informed decisions about who we keep in our lives.

Understanding the nature of these toxic interactions and how they affect us has an enormous impact on our ability to engage in self-care.

The Diagnostic \& Statistical Manual of Mental Disorders (5th ed.; DSM-5; American Psychiatric Association, 2013) defines a narcissistic person as someone who has a pervasive pattern of grandiosity (in fantasy or behavior), an excessive need for admiration, and lack of empathy, beginning by early adulthood and present in a variety of contexts.

This is indicated by 5 (or more) of the following:
\begin{itemize}
    \item Has a grandiose sense of self-importance (e.g., exaggerates achievements and talents, expects to be recognized as superior without commensurate achievements).
    \item Is preoccupied with fantasies of unlimited success, power, brilliance, beauty, or ideal love.
    \item Believes that he or she is ``special'' and unique and can only be understood by, or should associate with, other special or high-status people (or institutions).
    \item Requires excessive admiration.
    \item Has a sense of entitlement (i.e., unreasonable expectations of especially favorable treatment or automatic compliance with his or her expectations).
    \item Is interpersonally exploitative (i.e., takes advantage of others to achieve his or her own ends).
    \item Lacks empathy: is unwilling to recognize or identify with the feelings and needs of others.
    \item Is often envious of others or believes that others are envious of him or her.
    \item Show arrogant, haughty behaviors or attitudes.
\end{itemize}
That is the clinical definition, but the way a narcissist operates in relationships is stealthier and more complex than the visible signs of narcissism depicted on this list.

This book will help you gain knowledge about the red flags of narcissism and how it manifests in relationships, how to begin to heal and recover from narcissistic abuse as well as crucial insights drawn from both research and survivors.

You will learn healing strategies and tips to begin to detach from your narcissistic partner if you have not already, understand the reasons why you may still be addicted to your narcissistic partner, as well as the best ways to channel your experiences into constructive healing.

\subsection{What is Narcissistic Abuse?}
Narcissistic partners engage in chronic manipulation and devaluation of their victims, leaving victims feeling worthless, anxious and even suicidal.

This type of continual manipulation, which includes an idealization-devaluation-discard abuse cycle where the ``lovebomb'' their partners, devalue them, then discard them until the trauma begins again, is known as \textit{narcissistic abuse} -- abuse by a partner with NPD or on the far end of the narcissistic spectrum.

This type of abuse can leave psychological and emotional scars that can last a lifetime.

%
In fact, trauma expert Pete Walker (2013) suggests that chronic emotional abuse, especially if experienced in childhood and later re-experienced in adulthood, can cause symptoms of PTSD or Complex PTSD.

Psychotherapist Christine Louis de Canonville (2015) also discusses what she calls \href{https://narcissisticbehavior.net/narcissistic-victim-syndrome-what-the-heck-is-that/}{Narcissistic Victim Syndrome} in her book, \textit{The 3 Faces of Evil: Unmasking the Full Spectrum of Narcissistic Abuse}.

What makes narcissistic abuse so dangerous is that these individuals employ very covert and insidious methods to abuse their partners.

Due to the very nature of their abuse, they are able to escape accountability for the abuse because of the false persona they present to the outside world - which is usually a charming mask that hides their cruelty.

%
What is important to remember is that narcissists do not outright destroy you in the way physically abusive partners may do -- they plant the very seeds in your mind that will lead you to your own destruction.

They will cultivate doubts that never existed, manufacture insecurities that were never present, rub salt on the wounds they know you do not want opened -- all under the guise of ``helping'' you or gaslighting you into believing you deserved or provoked the abuse.

They also rewrite the abuse they have inflicted, making you look like you are the abuser.

%
As Arabi will describe more in-depth in Chap. 2, narcissists will subject you to a dizzying biochemical cocktail of traumatic highs and lows to keep you hooked on the drug that is their abuse.

This type of crazymaking is so covert that many victims do not even know they have been abused - some only learn to identify the tactics well within 10 years into a marriage with a narcissistic abuser, while others are lucky enough to learn about it early on in the relationship.

%
\textbf{The toxic behaviors in narcissistic abuse can include but are not limited to:}
\begin{itemize}
    \item Being overly critical and controlling towards their partners, covertly and overtly putting them down though cruel verbal abuse and manipulative behaviors that are employed to isolate them.
    
    This can include name-calling, harsh insults disguised as jokes, demeaning, contemptuous remarks about the victim's appearance, intelligence, line of work, lifestyle, skill sets, accomplishments, or other support networks the victim has outside of the relationship.
    \item Being physically and\texttt{/}or sexually abusive.
    
    This can include hitting the partner with objects, punching, shoving, slapping, choking or pushing the victim, forcing the victim to have sex without consent, and coercing the victim into sexual situations that they are not comfortable with.
    
    They may threaten to leave the victim or ruin the victim's life in some way if they do not comply with their wishes.
    \item Manufacturing hostile or aggressive situations where the victim is led to emotional distress, especially through the abuser's narcissistic rage over seemingly small or irrelevant things.
    
    The abuser creates an environment where the victim feels trapped, controlled, and limited in what he or she can say or do.
    \item Engaging in hot and cold behavior that switches quickly between a loving persona and an abusive one.
    
    This is an abuse cycle known as \textit{idealization, devaluation and discard}.
    
    It includes treating the victim cold and callously for no apparent reason, only to return to loving, affectionate behavior through a technique called \textit{intermittent reinforcement}.
    
    This conditions the victim into expecting less each and every time they interact and also programs the victim into associating love with unpredictability, distress and unease.
    
    The victim is later discarded in a humiliating, demeaning manner, often followed by a smear campaign so that the narcissist feels that they have ``won'' the break-up.
    \item Controlling every aspect of their partner's life to the point where they isolate them from family and friends; this includes sabotaging the victim's friendships, familial relationships, important life events or their goals and aspirations.
    \item Stonewalling their victims into silence should they bring up any concerns about the relationship; subjecting the victim to silent treatments and disappearances throughout the abuse cycle to create a sense of chronic insecurity in the victim, causing the victim to walk on eggshells and increase their efforts to please their abusive partner.
    \item Triangulating their victims with other love interests including their ex-partners; engaging in pathological lying and deceit while pursuing numerous affairs outside of the relationship; comparing the victim to other people regarding their appearance, personality, success and other attributes to instill in them a sense of worthlessness.
    
    This infidelity is not driven by dissatisfaction with the main partner but rather a sadistic need for narcissistic supply.
    
    This supply comes in the form of attention from multiple people as well as the victim's emotional distress in response to the triangulation.
    \item Gaslighting their partners into believing the abuse is not real by denying, minimizing or rationalizing the abuse.
    
    This includes deflecting any conversations about accountability using circular conversations and word salad in order to avoid being held accountable for their actions.
    \item Subjecting their victim to smear campaigns to slander their character and reputation so that the abuse victim is left with no support network.
    
    This includes projecting their own abusive behavior onto the victim so that no one believes the accounts of abuse.
    \item Blameshifting and projecting their malignant traits onto their partners during conversations while using a false charismatic self to make their victims look like the ``crazy'' ones.
    
    It is almost as if they hand off their own traits and shortcomings to their victims as if to say, ``Here, take my pathology. I do not want it.''
\end{itemize}
\textit{This} is what narcissistic abuse looks like -- and unfortunately, the full extent of narcissistic abuse is not taught in any psychology class or diagnostic manual, though these manipulative techniques can be found in numerous books by experts about narcissism, accounts from mental health professionals who have worked with survivors as clients as well as survivor accounts.

It is truly a narcissist's malignant behaviors and how they affect us that are the key to understanding if your partner is a narcissist.

Anyone can be a victim of narcissistic abuse, regardless of their gender or background.

A narcissistic abuser can lead survivors to feel depressed, suicidal, anxious, constantly on edge and worthless.

If your partner displays these types of toxic behaviors, at the very least, they are emotionally, verbally and psychologically abusive.

While malignant narcissists are certainly very dangerous, partners who display even some of these behaviors and refuse to change do not need the diagnosis of NPD in order for victims to recognize that they have a toxic relationship partner.

\subsection{What Causes Narcissism?}
The myth of Narcissus is a popular one to depict a narcissist's vanity and self-absorption, but a deeper dialogue about the origins of the full-fledged personality disorder is lacking.

It is true that narcissism is rising in our culture, but how exactly does it manifest, especially as a personality disorder?

There are many theories about how narcissism arises in the individual -- from a ``narcissistic wound'' in childhood, to a pattern of idealization and devaluation by the parent or even a neurological standpoint that focuses mainly on how a narcissist's brain has structural abnormalities related to compassion (Kernberg, 1975; Kohut, 1971; Lavendar, 2014; Schulze et. al, 2013).

Yet there is still no clear answer as to how Narcissistic Personality Disorder arises in an individual.

%
In ``The Cracked Mirror: Features of Narcissistic Personality Disorder in Children,'' Dr. Karen Bardenstein (2009) notes that there are certain risk factors when it comes to children developing the disorder.

The children who become narcissistic are often: the children of narcissistic parents, adopted children who are overindulged by their adoptive parent or face competition with their biological siblings, the children of successful parents, especially if the child lacks the same ability as their parents, overindulged, wealthy children, and the children of divorce, specifically, a divorce in which the child is vied over and is treated as an object to be ``won.''

%
Narcissists may have suffered something traumatic when they were a child - emotional neglect or physical neglect, as Kernberg (1975) suggests, by a critical parent.

\textit{This severe trauma in childhood, what psychologists call a ``narcissistic wound,'' may have led them to shut off the parts of themselves that could empathize with others, as a form of extreme self-protection} (Marks, 2012; Lavender, 2014; De Lisle 2015).

This may have been caused by a parent who neglected them, abused them and invalidated them, resulting in what Millon (1981) calls ``compensatory narcissism,'' a self-protective false self that enables a child to believe in an illusion of superiority in order to conceal feelings of low self-worth.

%
On the other end of the spectrum, there are theories that suggest that narcissism can be produced by \textit{overvaluation} of the child, allowing the child to remain like a child forever without any consequences or without any basis in objective reality for their presumed perfection.

This pattern of overvaluation rather than devaluation by a parent that may have led a child to arrested emotional development -- in other words, a child who is spoiled to the point where they develop an excessive sense of entitlement and disregard for the emotions of others.

%
A new research study in 2015 by the Proceedings of the National Academy of Sciences discovered that parents who overvalued their children by telling them how special they were produced narcissistic children.

In their words, ``Narcissism was predicted by parental overvaluation, not by lack of parental warmth.

Thus, children seem to acquire narcissism, in part, by internalizing parents' inflated views of them (e.g., `I am superior to others' and `I am entitled to privileges').'' (Brummelman et. al, 2015).

%
This new research aligns with what Randi Kreger (2012) calls the distinction between ``vulnerable'' narcissists and ``grandiose narcissists.''

According to Kreger, \textit{grandiose} narcissists are bred by parents who spoil them into thinking how superior and entitled they are, while \textit{vulnerable} narcissists tend to be the product of neglectful parenting.

%
As someone who has taught very young children of wealthier backgrounds, Arabi definitely saw this burgeoning sense of entitlement in the parents who enabled their children's misbehavior, allowing their children to trample upon the boundaries of others without so much of an apology.

The parents who taught their kids an excessive sense of entitlement had a very indifferent attitude and coddled their children to the extreme even when they overstepped boundaries, while the parents who taught their children healthy boundaries and respect were likely to show empathy, remorse and accountability for their behavior.

%
We also see narcissism arise when parents treat their children as trophies or live through them vicariously - essentially degrading and objectifying them, teaching them that they are objects - which teaches the narcissistic child to also view others as objects for their disposal, while also overvaluing them as highly important.

There can therefore be a dual ``overvaluation'' and ``neglect'' taking place in this circumstance that produces narcissism in a child; the parent may overvalue the child as perfect, but may also instill a sense of worthlessness in the child by not providing the child with healthier feedback that validates who they really are as a person.

This is because the narcissistic child is overvalued as ``perfect'' and this type of feedback is not balanced with realistic feedback.

It places pressure on the child to be a trophy, rather than a 3D human being.

It also teaches children a high degree of entitlement, instilling in them that they are deserving of things they have not earned.

%
Since this parental overvaluation is not balanced with acknowledgment of the child's imperfect humanity, it causes that child to develop a sense of grandiosity that vacillates between feelings of worthlessness and a hyperinflated ego - in other words, narcissism in the place of a healthy self-image (De Lisle, 2015).

It can also cause the child to highly emphasize 1 attribute over others - e.g. looks, intellectual ability or another talent, in place of healthy self-esteem and self-acceptance.

%
That being said, parents along may not be responsible for producing narcissism in their children and there are many parents with narcissistic children who do provide their children a loving, validating home.

That is where we have to consider biological predisposition being passed down in families.

Twin studies have shown that narcissism is a common, heritable trait (Livesley et. al, 1993).

There is also a biological and neurological standpoint that redirects us to consider how the brain of someone with NPD differs from someone who does not have NPD.

%
Research has discovered that there are in fact structural abnormalities related to compassion and empathy in narcissistic individuals, which is interesting to note considering that psychopaths can also demonstrate brain abnormalities; psychopaths not only show abnormalities in empathy, they also show abnormalities in the processing of pro-social emotions e.g. guide and moral reasoning (Schulze et. al, 2013; Sethi et. al, 2015).

%
While each theory is compelling and has its own set of evidence, Arabi needs to stress that clinicians are still not absolutely certain as to what causes NPD.

The answer is likely more complex than any individual theory.

In Arabi's opinion, psychopathology is often caused by an interaction between biological predisposition and environment.

There are also multicultural components that can make certain disorders more likely than others in specific countries or manifest differently across various contexts.

%
We have to consider that there are protective factors or risk factors that determine whether narcissism actually manifests in a full-fledged personality disorder or not as well as how it manifests in the individual.

It is usually not nature versus nurture but rather a combination of nature and nurture - most disorders are a result of the interaction between environment and biology rather than 1 or the other.

These factors include strong support network, access to therapy\texttt{/}medication, upbringing, religious beliefs, culture, media, as well as other experiences outside of the family unit like bullying, sexual assault, witnessing violence, or other traumas can all weaken or strengthen that predisposition towards pathology.

%
To put it simply, narcissists can probably be produced through a number of different ways and can come from a variety of backgrounds; their narcissism can be generated through an interaction between biology and environment.

Arabi personally has met narcissists from both types of backgrounds: traumatic ones as well as overvalued ones where their parents taught them an excessive sense of entitlement from very early on.

Arabi thinks it is important to acknowledge that narcissists can come from a variety of backgrounds, and that psychopathology in general is produced by this interaction between biological predisposition and the environment.

%
So while survivors cannot be 100\% certain what caused their partner's NPD, what survivors can be certain of is that being with a partner with NPD can be extremely dangerous and devastating due to their lack of empathy and tendency to be exploitive.

Arabi finds that the questions of ``why?'' and ``how?'' can often cause confusion in accountability that shames the victim rather than the abuser.

Just because someone has a disorder does not mean their abuse is not hurtful, that they are not responsible for their abusive behavior or that their abuse does not have an impact on the victim.

Quite the contrary.

Someone who has a disorder and does not get professional support in changing the behavior that harms others, and instead uses that disorder as an excuse to continue to abuse others, can still be accountable for their actions.

Yes, we can be compassionate towards the fact that our abuser may have been traumatized, but we cannot let that compassion blind us to the fact that if they are unwilling to change or receive treatment, as many narcissists are not, we need to make our own self-compassion and self-care a priority in order to detach from them.

%
In this type of 1-sided relationship, you will ultimately be used as an object to supply admiration, praise, attention, and whatever they may need at the moment; this is what Otto Fenichel (1945) calls a source of ``narcissistic supply.''

If you enter a relationship with a narcissist, be careful: the false self is often so charming and so different from the true self that you may fall prey to a vicious cycle of narcissistic abuse that can be very difficult to extricate yourself from.

The cognitive dissonance that results in being in love with an emotional predator is enhanced by the false mask the narcissist presents to the world.

A relationship with a narcissist often contains some degree of psychological, emotional and in some cases, physical and sexual violence depending on where the narcissistic person falls on the spectrum.

\subsection{Who is the Narcissist?}
You may be wondering if your partner is a narcissist, a sociopath or a psychopath.

There are distinctions among narcissists, sociopaths and psychopaths, but this book will be focusing on the overlapping symptoms, namely a lack of empathy and interpersonally exploitative behavior.

According to Dr. Scott Bonn (2015), a professor of criminology and sociology, sociopaths and psychopaths both both fall under the larger category of Antisocial Personality Disorder.

Both sociopaths and psychopaths share a heinous disregard for laws, social mores, the rights of others and have a tendency to demonstrate violent behavior.

According to the Psychopathy Checklist created by Robert Hare (1993), psychopaths tend to demonstrate callousness and lack of empathy, glib and superficial charm, a parasite lifestyle, failure to accept responsibility for their actions, shallow affect, and sexual promiscuity.

These are all traits that narcissists can share as well.

%
However, psychopaths often have juvenile delinquency in their history and criminal versatility.

Narcissists tend to commit more emotional and psychological ``crimes'' in the way they devalue, slander and sabotage the lives of their victims; they are theoretically capable of remorse and guilt unlike sociopaths and psychopaths, though more often than not, their self-absorption and lack of empathy prevents them from showing remorse for their behavior and being willing to change.

Both narcissists and psychopaths have been shown to have brain abnormalities - with narcissists, brain scans reveal abnormalities in structures of the brain related to compassion, while for psychopaths, brain scans revealed abnormalities in the areas of the brain related to moral reasoning and guilt (Schulze et. al, 2013; Sethi et. al, 2015).

%
While narcissists crave and need external validation, sociopaths and psychopaths often do not; sociopaths tend to exploit others for fun while narcissists tend to exploit others as a way to eliminate threat (Holbrook, 2013).

That being said, these individuals tend to be similar in the callous ways they exploit others and Arabi has witnessed firsthand that both narcissist and sociopaths tend to get a sadistic pleasure in harming others.

This is because narcissistic abusers thrive on power and control.

Due to their internal emptiness and inability to experience the full range of human emotions, they get their daily nourishment from inflicting trauma on their victims.

The lines can be blurred among these categories, so the ``label'' matters less than the fact that these individuals pose considerable harm to their victims, whether emotional, psychological, financial and\texttt{/} physical.

%
\textit{The 1st thing you must know about a narcissist is that you will truly never know anything about the narcissist}.

This is because a narcissist is someone who constructs what Sam Vaknin, a self-professed narcissist, calls ``false self'' to hide away from their true selves.

Their false self is the self they present to everyone else: the narcissist often presents the image of behind charming, well-loved, intelligent, successful, generous, and kind.

However, in private, and often to the most intimate partner the narcissist has (whether it is a spouse or boyfriend or girlfriend), they are selfish, cruel, exploitative, rageful, abusive, demonstrate an irrational sense of entitlement, are lacking in empathy, and are prone to being unfaithful.

They also manufacture situations where jealousy and insecurity is secretly provoked while outwardly disparaged to make the victim appear ``crazy'' and ``needy.''

%
Arabi has dated and befriended narcissists from a wide variety of fields.

The sociopathic social worker, the grandiose famous blogger, the disturbed award-winning director, the successful, contemptuous corporate attorney, the \fbox{entitled programmer}, the superficial banker - you name it, Arabi has encountered it.

These experiences were secret blessings, as they led Arabi to research narcissistic personality disorder extensively both inside and outside of the classroom, as well as connect with the survivor community in life-changing ways.

Due to the breadth of Arabi's real-life experiences with these predators, Arabi has noticed the commonalities among them as well as the quiet nuances that someone who has only read about this disorder but has never encountered it in the flesh flesh may miss.

%
1 thing Arabi would like to note is that while the image of the self-assured, successful, powerful narcissist is popular in the corporate world, not \textit{all} narcissists have reached their full potential in life.

1 misconception people have of narcissists is that they are all very successful people in power.

However, that is not \textit{always true}.

%
Narcissists come in all varieties and backgrounds.

They could be powerful CEO's of companies who dominate the little people or they could be waiter at restaurants who create a harem of women through all people they serve; they could live in mansions or life at home with their parents.

Some of them actually are not very successful and have not achieved many of their goals - these are likely to be ``vulnerable narcissists,'' though they can still put on an air of grandiosity to hide their feelings of failure (Kreger, 2012).

Narcissists can lie anywhere on the spectrum of incredible success or incredible poverty.

%
Narcissists can also inhabit positions that are seemingly charitable and philanthropic.

They can be in both helping professions (social workers or pastors at the church e.g.) as well as professions that are ideal for the narcissist to execute his or her power.

Now, there \textit{are} wonderful, validating and helpful counselors out there, but there are also counselors who abuse their power, just like in any other profession where power is granted.

There are mental health professionals who have NPD and masquerade as caring, empathic people to gain supply from their clients.

Unfortunately, abuse by a therapist is not uncommon, and it is likely that these meant to be a validating, therapeutic alliance to provoke, taunt and manipulate the victim according to their agenda.

%
Regardless of their professional background, prestige or status, the majority of narcissists share a devastating lack of empathy, a grandiose sense of entitlement and an unwillingness to change their destructive behavior regardless of their station in life.

This is what ultimately makes a relationship with them so damaging.

\subsection{The Narcissist's False Self \& True Self}
The narcissist does not feel empathy for others; he or she makes connections with other people for 1 purpose and 1 purpose only: \emph{narcissistic supply}.

\fbox{Narcissistic supply is the attention and admiration of the people the narcissist collects as trophies.}

It is anything that gives the narcissist a ``hit'' of praise, or even an emotional reaction to their ploys.

They need these sources of supply because they suffer from perpetual boredom, emotional shallowness and the inability to authentically emotionally connect to others who do have empathy.

%
A narcissist is prone to seeing you as an extension of himself or herself - a handbag or a piece of furniture rather than a complex individual with needs, wants and desires.

While genuine, non-narcissistic people wish to connect to others on a deep emotional level, narcissists fear and are disgusted by intimacy because intimacy would expose their true self rather than the false self they project.

%
Narcissists, you see, lack empathy.

They are unable to identify with the needs and feelings of others.

They can ``fake'' emotions quite well, and are charming chameleons that adapt to what they perceive others want and need for them.

They do so in order to get what they need and want - a harem, an audience, narcissistic supply in the form of attention, praise and superficial relationships which fulfill their need to be extolled while evading real intimacy.

%
This false self is a patchwork of different personas, qualities and traits they put together after years of learning to imitate other people as well as information from any media they observe.

Narcissists have what we call ``cold empathy'' - they are able to intellectually understand why someone may feel what they do, but they do not truly feel the same remorse or shame that others do for harming others (Vaknin, 2012).

\fbox{They may appear remorseful or cry crocodile tears}, but often it is when there is no other course of action to take to escape accountability for their actions.

%
However, narcissists do lure their victims with a \textit{false} sense of intimacy and security early on in the relationship with this false mask.

The majority of narcissists tend to be charming, charismatic and grandiose; they often have a magnetic pull that combines faux innocence with the seductive pull of ``kindness'' and edginess.

Many of Arabi's coaching clients and readers note that it was the narcissist's devil-may-care attitude that 1st attracted them to the narcissist; many narcissists appear unfazed by emotions and many are bold in the way they interact with others while others may seem ``off'' or emotionally disconnected.

%
It is no surprise, then, that a recent study noted that females with more romantic experience and a desire for marriage actually preferred narcissistic personalities due to their charming personality, social status and resource provision (Haslam \& Montrose 2015).

These women had presumably encountered narcissists before and yet still wanted them as long-term partners!

Despite their long-term destructive nature in relationships, narcissists can still appear to be desirable romantic partners and have a talent for projecting a sense of confidence and uniqueness that make them alluring to potential mates.

%
\fbox{To the outside world, narcissists can appear incredibly generous, even spiritually minded, and humane.}

\fbox{It is usually only to their closest and dearest, their victims, that their full-frontal cruelty is displayed.}

Much like sociopaths and psychopaths, narcissists hide behind this sweet, alluring, charming demeanor to hook their victims and escape accountability should they ever be threatened to be exposed - after all, who would believe the victim, if the perpetrator hides himself or herself in plain sight?

%
That is exactly how narcissists keep up the fa\c{c}ade - the mean and sweet\texttt{/}hot and cold cycle of abuse would not be as nearly traumatic without the incessant doubt about the narcissist's character - a doubt that is constantly reinforced every time we see their ``nice'' side and begin to wonder if we are the ones imagining things.

The truth is, we are simply witnessing the narcissist's mask during the sweet periods and their true self during the dark ones.

%
Narcissists can fool anyone and everyone with their false mask - including highly educated, successful, confident and attractive individuals.

Many intelligent people can be duped by narcissistic people, simply because they cannot imagine someone deliberately manipulating and hurting others in the way that narcissistic or antisocial personalities are prone to doing.

Let's get real: \textit{nobody wants to believe the person they care about is out to harm them - they would prefer believing in the false mask because the alternative is too horrifying to consider}.

%
It can take a great deal of time and effort before we can resolve that cognitive dissonance that this type of relationship evokes in us, as we are likely to develop conflicting beliefs, feelings and thoughts about an abusive partner who can switch masks so rapidly (Carver, 2004).

Resolving our cognitive dissonance involves grounding ourselves in the reality of the abuse rather than engaging in the denial, minimization or rationalization of the abuse, which is inevitable in trauma bonding with our abusers.

%
\fbox{\textit{The 1st truth that survivors must accept is that narcissists do not feel authentic empathy for their victims' emotions}.}

The closest narcissists get to feeling anything is a ``narcissistic injury,'' which is the rage and despair they feel when a person ``critiques'' them or poses a real or even false threat to their excessive sense of superiority and entitlement.

This makes them scramble to protect their false selves even more$\ldots$ so to attempt to make them remorseful is a lost cause.

They will always be protecting their true selves and this is why you can never win with the narcissist - they will always be defending the fortress of their false self.

%
How do narcissists get away with such shapeshifting?

The truth is that narcissists can embody \textit{multiply false selves} across various contexts and with difference people in order to gain supply (praise, attention, money, sex, etc.) from their audiences.

They are also proficient at gaslighting and projection, techniques they use to convince society that their victims are the crazy ones and to convince their victims that their perception of reality is inaccurate.

%
\fbox{\textit{Narcissists gaslight you so you begin to gaslight yourself}} into thinking what you are feeling, hearing, seeing and experiencing is not true.

A narcissistic partner can manipulate you into thinking that perhaps that hurtful comment really was just a joke and that their infidelity was just a 1-time thing.

Many of these partners engage in pathological lying and rewrite reality on a daily basis to suit their needs and to conceal their manipulative agenda.

%
\fbox{It is no wonder, then, that victims of narcissistic abuse often feel very isolated and invalidated in their experiences.}

It is easy for narcissists to convince the outside world that they are the sane, rational ones: pathological narcissists enjoy calling anyone who challenges their self-perception ``crazy.''

It is the word they will use to describe any valid emotional reaction victims have to their shady and inconsistent behavior.

It is gaslighting (invalidation of one's perception of reality) at its simplest form but over time becomes a complex type of psychological torture in which the victim starts to mistrust his or her perceptions of the covert abuse and feels unable to trust his or own reality.

\subsection{Who Does the Narcissist Target?}
Narcissists can target anyone and everyone for supply - so long as they have empathy.

It does not matter how beautiful, brilliant, successful, likable, funny, and strong you are - you can in fact still be duped by a narcissist, a sociopath or a psychopath because these people are wired differently and use manipulative tactics that are beyond our widest imagination.

Mental health professionals, especially couples' therapists who do not know much about narcissistic abuse, are fooled every day.

\fbox{Individuals with high levels of empathy tend to be targeted} because the narcissist is able to appeal to their sympathy and good-hearted nature to continue engaging in the toxic relationship cycle.

%
In ``\href{https://www.huffpost.com/entry/abusive-relationships-women_b_912513}{Why Do Smart Women Date Abusive Men},'' survivor and writer Hayley Rose Horzepa notes that anyone who understands the dynamics of an abusive relationship knows that being with and staying with an abusive partner has little to do with intelligence.

She notes that what makes victim susceptible to staying in an abusive relationship has more to do with their \textit{vulnerability}, rather than intelligence - a vulnerability that can be exacerbated by past traumas and losses.

Counter to the typical stereotypes about abuse victims, research indicates that women who are employed are twice as likely to be victims of domestic violence than those who do not; in fact, women who get involved with disordered personalities can also be very independent and lead active lives (Franklin \& Menaker, 2012; Brown, 2010).

This is proven by countless public figures who have been domestic violence victims - women who were smart, talented and successful.

Although the normative discourse is that these individuals had low self-esteem, the truth is that an abuser can also bring down someone from a place of high self-esteem and security - Arabi will discuss how later in this chapter.

%
This is not to dismiss or minimize anyone who has been with a narcissist and who feels dependent as a result, or feels they are in fact codependent, but it does reveal that our stereotypes of abuse survivors lack nuance, are outdated and need to be dismantled to include a wide variety of people who come from different backgrounds and have different personality traits than might be normally assumed or expected.

%
\fbox{Narcissists do not choose us because we are like them; they choose us because we are the light to their darkness}; regardless of any of our vulnerabilities, we exhibit the gorgeous traits of empathy, compassion, emotional intelligence and authentic confidence that their fragile egotism and false mask could never achieve.

%
While abuse survivors who have been traumatized before child abuse are undoubtedly more susceptible to being retraumatized, it is important to acknowledge that there are also survivors who were not traumatized in childhood that have encountered, been hooked by and have been abused by narcissists, sociopaths and psychopaths.

Arabi receives messages from survivors all around the world who come from a variety of trauma histories, who have been victimized.

%
Abuse survivors can come of \textit{any type of background} and many of them are strong, intelligent, talented and successful.

What puts them at risk is their empathy and their vulnerability - any cumulative trauma that has led them to have malleable boundaries and difficulties with self-esteem can certainly make individuals extra susceptible, but it has little to do with intelligence or strength and everything to do with the nature of trauma they have endured - both within abusive relationship and outside of it.

\subsection{The Abuse Cycle: Idealize, Devalue, Discard, Destroy, Hoover}
The abuse cycle with a narcissistic abuser includes an early idealization phase or honeymoon period, followed by an abusive incident or a series of abusive incidents before the cycle of idealization starts again, this time through the method of intermittent reinforcement which includes both devaluation of the victim as well as elements of excessive attention and adoration.

Narcissistic abusers 1st idealize their partners, flattering them excessively, giving them all sorts of attention in the form of constant texts and gifts.

They share secrets and stories with you to create a special bond; this technique also enables you to feel as if you can share your deepest insecurities and desires with them.

Later, they will use your disclosure as ammunition and pick at your weak spots to regain a sense of psychological control.

%
Yet it is during the honeymoon phase where these toxic partners reel you in and make them dependent upon them, look, line and sinker.

In a narcissist's idealization phase, you have never felt better: you feel beautiful, loved, cherished, and you feel like you have met your soulmate - this is a technique known widely in the survivor community as ``love-bombing.''

In the love-bombing or idealization phase, narcissists mirror their victims' values and interests to make them feel like they have met the ``one.''

%
The devaluation that follows, combined with periodic love-bombing, conditions the victim to walk on eggshells during a period of tension, hoping that the honeymoon period will last, only to inevitably experience yet another abusive incident that adds onto the cycle of trauma.

Narcissistic abusers rely on the victim's need for validation and approval, as well as the sense of worthlessness they have instilled in them throughout the abuse cycle, to remain dependent on their abusers.

%
These abusers will give you just the minimum amount of attention to keep you by their side, wishing you could go back to the idealization phase.

As the abuse gets incrementally worse, the abuse victim gets more and more conditioned to the abuse and invested in the abuse relationship due to the biochemical and trauma bonds that Arabi will discuss in Chap. 2.

%
While there is no time limit as to how long the cycle can continue, victims are likely to be discarded at some point, especially if they become aware of the narcissist's schemes or if they leave the narcissist themselves, although they are sometimes manipulated back into the relationship or quickly replaced with another source of supply.

%
Much like other survivors of narcissistic abuse, the pattern in Arabi's relationship with a narcissistic abuser went something like this: a period of sweet, excessive idealization which served as the honeymoon phase, where Arabi was properly ``groomed,'' complimented, flattered, placed on a pedestal, followed by incidents of devaluation littered throughout the idealization phase.

Arabi says littered because many covert narcissists are incredibly strategic about how they choose to devalue - they may leave essential clues and red flags throughout even the idealization phase, because they know you won't catch on; you will mistake their rudeness as awkward openness or constructive criticism; you will be tempted to exorcise the abuser inside of them and rescue them from their misconceptions of you.

%
\fbox{Narcissists are covert, underhanded and incredibly manipulative in the way they devalue their victims.}

They target both your insecurities as well as what you are most proud of.

It is a mix of both; if you tell them what you are most insecure about, they will make sure to sneak in a cutting remark at some point about it.

If you tell them what you are most proud of accomplishing, they will find ways to diminish your pride and make you feel as if you are not special - because, of course, only they are allowed to be special and unique.

%
This devaluation could be disguised as harsh ``jokes,'' or ``brutal honesty,'' when they in fact involved verbal abuse, name-calling, condescending sarcasm, a sudden withdrawal of affection, overt or covert put-downs or unhealthy comparisons to others to instill a sense of worthlessness.

The devaluation can also be overt and heinous.

In Arabi's own experiences, during more overt, full-fledged devaluing explosions, there was often a verbally and emotionally abusive argument that would last for hours, followed by false remorse and a glimpse of the narcissist's sweet false self to pull Arabi back in.

%
The cycle repeated over again until the temporary ``discard,'' where the narcissistic abuser would abandon Arabi through narcissistic rage and stonewalling if confronted, only to reappear again through ``hoovering,'' a tactic in which the abuser attempts to regain control over the victim by pretending he or she has changed.

The final discard, of course, gave Arabi the opportunity to go full No Contact with the abuser and end the cycle.

%
The narcissistic abuser cycle is difficult for survivors to extricate themselves from due to the immense amount of gaslighting, projection and psychological manipulation involved.

Abusers are able to convince their victims that the abuse they are suffering is not real, leading them to deny or minimize the abuse.

They may even convince their victim that the abuse is their fault by projecting their own flaws onto their victims or making the strengths of the victim appear to the weaknesses.

%
Victims are then mislead into thinking that the abusive incidents are just ``misunderstandings,'' and that the hot and cold behavior is due to something the victim did.

Unfortunately, for narcissists and their conscience-lacking cousins, those with antisocial personalities, the intent and pathology behind these actions are much more malicious, much more harmful and goes deeper than we will ever realize.

%
Survivors spend a lifetime waiting - waiting for their abusers to change, waiting for their kindness to turn into cruelty, waiting for the next blow or hit in an attempt to avoid it or escape it, waiting for the right time to end the relationship.

It is only when they stop waiting and start walking away that they begin to process what they have escaped and begin a journey back to wholeness.

It is challenging to walk away, even more difficult to heal, but more than possible with the right resources and support.

\subsubsection{Devaluation \& Discard}

\subsubsection{Hoovering}

\subsection{Victim-Blaming \& Stereotypes about Narcissistic Abuse}

\subsection{Are Narcissists Victims?}

\subsection{Should We Feel Compassion for the Narcissist?}

\subsection{Simultaneous Wounding \& Complex PTSD: How Our Past Wounds Can Make Us Susceptible to Toxic Narcissists}

\subsection{The Narcissist's False Mask Gaslights the Public, Enabling the Abuse to Continue}

\subsection{The Narcissist's Next Victim}

\subsection{1. The Idealization-Devaluation-Discard Phase}

\subsection{2. Gaslighting}

\subsection{3. Smear campaigns}

\subsection{4. Triangulation}

\subsection{5. The false self \& the true self}

\subsection{Learning the Language of Narcissists: How Abusers Use Anything \& Everything Against Their Victims}

\subsection{Dating Emotional Predators}

\subsection{A need for control}

\subsection{Addicted to provoking you}

\subsection{Inconsistent character \& behavior}

\subsection{Signs of a Toxic Person's Pathological Envy}

\subsection{Clich\'es That Fail Within Narcissistically Abusive Relationships}

\subsection{The Essential Dictionary to Understanding Narcissistic Abuse}

\subsection{The Narcissist Translator}

\subsection{How to Break Up with a Narcissist, Sociopath \& Toxic Partner Safely \& Securely}

\subsection{Discarding the Narcissist: In preparation for the final battle}

\subsection{Selfish with a Side of Disordered by Kristin Sunanta Walker}

\subsection{The Sociopath's Scam - It Isn't Personal, by Jennifer Smith}

\subsection{Survivor Insights: How Narcissistic Abuse Affects Survivors}

\section{Your Brain on Love, Sex, \& the Narcissist - Why Your Bond to Your Abuser is Difficult to Break}

\subsection{Your Brain on Trauma}

\subsection{Your Brain in the Abusive Relationship: The Biochemical Bonds That Get Us ``Hooked'' on Narcissists}

\subsection{Complex PTSD \& Narcissistic Abuse}

\subsection{Emotional \& Psychological Reasons Why Victims Stay}

\subsection{Narcissistic Abuse Can Cause Complex PTSD by Lilly Hope Lucario}

\section{How to Detach from the Narcissist \& Begin the Healing Journey from Narcissistic Abuse}

\subsection{11 Healing Steps}

\subsection{How to Combat Biochemical \& Trauma Bonds}

\subsection{Self-Care Warrior Pro-tip}

\subsection{Healing Trauma in the Brain: Traditional Healing Methods Explained}

\subsection{Cognitive Behavioral Therapy}

\subsection{Therapy for Complex PTSD}

\subsubsection{Prolonged Exposure Therapy}

\subsubsection{Cognitive Processing Therapy}

\subsection{Meditation}

\subsection{DBT}

\subsection{Group Therapy}

\subsection{Support Groups}

\subsection{Alternative Healing Methods Explained}

\subsection{EMDR}

\subsection{Emotional Freedom Technique (EFT)}

\subsection{Yoga}

\subsection{Mirror Work}

\subsection{Art Therapy}

\subsection{Reverse Discourse \& Narrative Therapy\texttt{/}NLP Reframing}

\subsection{Neurolinguistic Programming (NLP) \& Self-Hypnosis}

\subsection{Inner Child Work}

\subsection{Acupuncture}

\subsection{Chakra Balancing}

\subsection{Cord Cutting}

\subsection{Aromatherapy \& Incense}

\subsection{Spirituality\texttt{/}Faith\texttt{/}Prayer}

\subsection{The Power of Nature}

\subsection{Music Therapy as Catharsis}

\subsubsection{Self-Care Haven's Music Library for Survivors}

\subsection{Self-Care Warrior Pro Tip: Hybrid Techniques}

\subsection{Survivor Insights: Healing from Narcissistic Abuse}

\subsection{Distinguishing between a Validating vs. Invalidating Therapist}

\subsection{Interview with Andrea Schneider, MSW, LCSW}

\subsection{5 Powerful Self-Care Tips for Abuse \& Trauma Survivors}

\subsection{Healing the Childhood Programming that Makes You a Magnet for Narcissists by Lisa A. Romano}

\subsection{Re-Parenting the Abused Child by Ingrid Roekke}

\subsection{Sticks \& Stones \& Hurtful Words by Alison Soroka}

\section{The Game You Can Win - Detaching from the Narcissist by Becoming the Superhero or Superheroine of Your Life}

\subsection{Build Your Immunity}

\subsection{Eradicating People-Pleasing Habits}

\subsection{Recovery from Rejection}

\subsection{Let's Recap: 7 Inalienable Truths about Narcissistic Abuse}

\subsection{Supertechniques \& ``Power-Ups''}

\section{The Journey of No Contact}

\subsection{What No Contact is \& What It Isn't}

\subsection{Why We Establish No Contact in the Context of Abusive Relationships}

\subsection{How to Execute No Contact Effectively}

\subsection{Sticking to No Contact}

\subsection{No Contact's Healing Power}

\subsection{111 Alternatives to Breaking No Contact}

\subsection{Releasing Self-Judgment}

\subsection{Co-Parenting with a Narcissist by Kim Saeed}

\subsection{Mantras\texttt{/}Manifesto\texttt{/}Positive Affirmations for No Contact: Why Silence Can Be the Most Powerful Voice}

\subsection{Survivor Insights: No Contact}

\section{Pathological Narcissism in the Family, Friendships, Workplace, Society, \& Culture}

\subsection{Female Narcissists in Friendships}

\subsection{Narcissistic Family Members \& Parents}

\subsection{Narcissistic Co-Workers \& Bosses}

\subsection{How to Deal with Narcissists You Can't Avoid: 6 Ways to Skillfully Respond Cyberbullies}

\section{Healing from Emotional Trauma by Finding Your True Purpose \& Rebuilding Your Life}

\subsection{Survivor Insights: Tools to Transcend the Narcissistic Abuse Experience}

\section{10 Life-Changing Truths for Abuse Survivors}

\subsection{Survivor Insights: Chasing Dreams \& Rebuilding Your Life after Abuse}

\section{Owning Our Power \& Agency}

\subsection{The Distinction between Victim-Blaming \& Owning Our Agency}

\subsection{The 3 Steps to Owning Our Agency \& Power after Narcissistic Abuse}

\section{Writing Your Way to Recovery \& Closure}

\subsection{Survivor Insights: What What You Tell Other Survivors?}

\subsection{A Closing Love Letter to Survivors}

\subsection{FAQ: 50 Shades of Narcissism}

\subsection{Additional Resources on Narcissistic Abuse}

\section{About the Author}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}