\documentclass[oneside]{article}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/nguyen/1_NQBH/reference/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{\small \textsc{Sect.} ~\thesection}
\rhead{\small \nouppercase{\leftmark}} % \nouppercase !
\renewcommand{\sectionmark}[1]{\markboth{#1}{}}
\cfoot{\thepage}
\def\labelitemii{$\circ$}

\title{James Clear. \textit{Atomic Habits: An Easy \& Proven Way to Build Good Habits \& Break Bad Ones}. Tiny Changes, Remarkable Results}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
\tableofcontents

%------------------------------------------------------------------------------%

\noindent\textbf{a$\cdot$tom$\cdot$ic}
\begin{enumerate}
    \item an extremely small amount of a thing; the single irreducible unit of a larger system.
    \item the source of immense energy of power.
\end{enumerate}
\textbf{hab$\cdot$it}
\begin{enumerate}
    \item a routine or practice performed regularly; an automatic response to a specific situation.
\end{enumerate}

\section*{Introduction: My Story}

\subsection*{My Story}
On the final day of my sophomore year of high school, Clear was hit in the face with a baseball bat.

As Clear's classmate took a full swing, the bat slipped out of his hands and came flying toward Clear before striking Clear directly between the eyes.

Clear has no memory of the moment of impact.

%
The bat smashed into Clear's face with such force that it crushed Clear's nose into a distorted U-shape.

The collision sent the soft tissue of Clear's brain slamming into the inside of Clear's skull.

Immediately, a wave of swelling surged throughout Clear's head.

In a fraction of a second, Clear had a broken nose, multiple skull fractures, and 2 shattered eye sockets.

%
When Clear opened his eyes, he saw people starting at him and running over to help.

Clear looked down and noticed spots of red on his clothes.

1 of Clear's classmates took the shirt off his back and handed it to Clear.

Clear used it to plug the stream of blood rushing from Clear's broken nose.

Shocked and confused, Clear was unaware of how seriously Clear had been injured.

%
Clear's teacher looped his arm around Clear's shoulder and they began the long walk to the nurse's office: across the field, down the hill, and back into school.

Random hands touched Clear's sides, holding him upright.

They took their time and walked slowly.

\fbox{Nobody realized that every minute mattered.}

%
When they arrived at the nurse's office, she asked Clear a series of questions.

%
``What year is it?''

%
``1998,'' Clear answered.

It was actually 2002.

%
``Who is the president of the United States?''

%
``Bill Clinton,'' Clear said.

The correct answer was George W. Bush.

%
``What is your mom's name?''

``Uh. Um.'' Clear stalled.

10 seconds passed.

%
``Patti,'' Clear said casually, ignoring the fact that it had taken Clear 10 seconds to remembers his own mother's name.

%
That is the last question Clear remember.

Clear's body was unable to handle the rapid swelling in Clear's brain and Clear lost consciousness before the ambulance arrived.

Minutes later, Clear was carried out of school and taken to the local hospital.

%
Shortly after arriving, Clear's body began shutting down.

Clear struggled with basic functions like swallowing and breathing.

Clear had his 1st seizure of the day.

Then Clear stopped breathing entirely.

As the doctors hurried to supply Clear with oxygen, they also decided the local hospital was unequipped to handle the situation and ordered a helicopter to fly Clear to a larger hospital in Cincinnati.

%
Clear was rolled out of the emergency room doors and toward the helipad across the street.

The stretcher rattled on a bumpy sidewalk as 1 nurse pushed Clear along while another pumped each breath into Clear by hand.

Clear's mother, who had arrive at the hospital a few moments before, climbed into the helicopter beside him.

Clear remained unconscious and unable to breathe on his own as she held Clear's hand during the flight.

%
While Clear's mother rode with him in the helicopter, Clear's father went home to check on Clear's brother and sister and break the news to them.

He choked back tears as he explained to Clear's sister that he would miss her 8th-grade graduation ceremony that night.

After passing Clear's siblings off to family and friends, he drove to Cincinnatti to meet Clear's mother.

%
When Clear's mom and he landed on the roof of the hospital, a team of nearly 20 doctors and nurses sprinted onto the helipad and wheeled Clear into the trauma unit.

By this time, the swelling in Clear's brain had become so severe that Clear was having repeated post-traumatic seizures.

Clear's broken bones needed to be fixed, but Clear was in no condition to undergo surgery.

After yet another seizure - Clear's 3rd of the day - Clear was put into a medically induced coma and placed on a ventilator.

%
Clear's parents were no strangers to this hospital.

10 years earlier, they had entered the same building on the ground floor after Clear's sister was diagnosed with leukemia at age 3.

Clear was 5 at the time.

Clear's brother was just 6 months old.

After 2 and half years of chemotherapy treatments, spinal taps, and bone marrow biopsies, Clear's little sister finally walked out of the hospital happy, healthy, and cancer free.

And now, after 10 years of normal life, Clear's parents found themselves back in the same place with a different child.

%
While Clear slipped into a coma, the hospital sent a priest and a social worker to comfort Clear's parents.

It was the same priest who had met them a decade earlier on the evening they found out Clear's sister had cancer.

%
As day faded into night, a series of machines kept me alive.

Clear's parents slept restlessly on a hospital mattress - 1 moment they would collapse from fatigue, the next they would be wide awake with worry.

Clear's mother would tell me later, ``It was 1 of the worst nights I've ever had.''

\subsubsection*{My recovery}
Mercifully, by the next morning my breathing had rebounded to the point where the doctors felt comfortable releasing Clear from the coma.

\fbox{When Clear finally regained consciousness, Clear discovered that Clear had lost his ability to smell.}

As a test, a nurse asked Clear to blow Clear's nose and sniff an apple juice box.

Clear's sense of smell returned, but - to everyone's surprise - the act of blowing his nose forced air through the fractures in his eye socket and pushed his left eye outward.

Clear's eyeball bulged out of the socket, held precariously in place by his eyelid and the optic nerve attaching his eye to his brain.

%
The ophthalmologist said Clear's eye would gradually slide back into place as the air seeped out, but it was hard to tell how long this would take.

Clear was scheduled for survey 1 week later, which would allow Clear some additional time to heal.

Clear looked like he had been on the wrong end of a boxing match, but Clear was cleared to leave the hospital.

Clear returned home with a broken nose, half a dozen facial fractures, and a bulging left eye.

%
The following months were hard.

It felt like everything in his life was on pause.

Clear had double vision for weeks; Clear literally couldn't see straight.

It took more than a month, but Clear's eyeball did eventually return to its normal location.

Between the seizures and Clear's vision problems, it was 8 months before Clear could drive a car again.

At physical therapy, Clear practiced basic motor patterns like walking in a straight line.

Clear was determine not to let his injury get him down, but there were more than a few moments when Clear felt depressed and overwhelmed.

%
Clear became painfully aware of how far Clear had to go when he returned to the baseball field 1 year later.

Baseball had always been a major part of Clear's life.

Clear's dad had played minor league baseball for the St. Louis Cardinals, and Clear had a dream of playing professionally, too.

After months of rehabilitation, what Clear wanted more than anything was to get back on the field.

%
But Clear's return to baseball was not smooth.

When the season rolled around, Clear was the only junior to be cut from the varsity baseball team.

Clear was sent down to play with sophomores on junior varsity.

Clear had been playing since age 4, and for someone who had spent so much time and effort on the sport, getting cut was humiliating.

Clear vividly remembers the day it happened.

Clear sat in his car and cried as he flipped through the radio, desperately searching for a song that would make him feel better.

%
After a year of self-doubt, Clear managed to make the varsity team as a senior, but Clear rarely made it on the field.

In total, Clear played 11 innings of high school varsity baseball, barely more than a single game.

%
Despite Clear's lackluster high school career, Clear still believes he could become a great player.

And Clear knew that if things were going to improve, Clear was the one responsible for making it happen.

The turning point came 2 years after Clear's injury, when Clear began college at Denison University.

It was a new beginning, and it was the place where Clear could discover the surprising power of small habits for the 1st time.

\subsubsection*{How I learned about habits}
Attending Denison was 1 of the best decisions of his life.

Clear earned a spot on the baseball team and, although Clear was at the bottom of the roster as a freshman, Clear was thrilled.

Despite the chaos of his high school years, Clear had managed to become a college athlete.

%
Clear wasn't going to be starting on the baseball team anytime soon, so Clear \fbox{focused on getting his life in order.}

While Clear's peers stayed up late and played video games, Clear build good sleep habits and went to bed early each night.

In the messy world of a college dorm, Clear made a point to keep his room neat and tidy.

\fbox{These improvements were minor, but they gave Clear a sense of control over his life.}

Clear started to feel confident again.

And this growing belief in himself rippled into the classroom as Clear improved his study habits and managed to earn straight A's during his 1st year.

%
\fbox{A habit is a routine or behavior that is performed regularly - and, in many cases, automatically.}

As each semester passed, Clear accumulated \fbox{small but consistent habits that ultimately led to results that were unimaginable} to Clear when Clear started.

E.g., for the 1st time in his life, Clear made it a habit to lift weights multiple times per week, and in the years that followed, Clear's 6-foot-4-inch frame bulked up from a featherweight 170 to a lean 200 pounds.

%
When his sophomore season arrived, Clear earned a starting role on the pitching staff.

By his junior year, Clear was voted team captain and at the end of the season, Clear was selected for the all-conference team.

But it was not until his senior season that his sleep habits, study habits, and strength-training habits really began to pay off.

%
6 years after Clear had been hit in the face with a baseball bat, flown to the hospital, and placed into a coma, Clear was selected as the top male athlete at Denison University and named to the ESPN Academic All-America Team - an honor given to just 33 players across the country.

By the time Clear graduated, Clear was listed in the school record books in 8 different categories.

That same year, Clear was awarded the university's highest academic honor, the President's Medal.

%
Clear hope you'll forgive him if this sounds boastful.

To be honest, there was nothing legendary or historic about his athletic career.

Clear never ended up playing professionally.

However, looking back on these years, Clear believes he accomplished something just as rare: \fbox{Clear fulfilled his potential.}

And Clear believes the concepts in this book can help you fulfill your potential as well.

%
\fbox{We all face challenges in life.}

This injury was 1 of Clear's, and the experience taught Clear a critical lesson: \textit{changes that seem small and unimportant at 1st will compound into remarkable results if you're willing to stick with them for years}.

\fbox{We all deal with setbacks but in the long run, the quality of our lives often depends on the quality of our habits.}

\fbox{With the same habits, you'll end up with the same results.}

\fbox{But with better habits, anything is possible.}

%
Maybe there are people who can achieve incredible success overnight.

Clear doesn't know any of them, and Clear is certainly not 1 of them.

There wasn't 1 defining moment on his journey from medically induced coma to Academic All-American; there were many.

\fbox{It was a gradual evolution, a long series of small wins and tiny breakthroughs.}

The only way Clear made progress - the only choice Clear had - was to start small.

And Clear employed this same strategy a few years later when he started his own business and began working on this book.

\subsubsection*{How \& why I wrote this book}
In Nov 2012, Clear began publishing articles at \url{jamesclear.com}.

For years, Clear had been keeping notes about his personal experiments with habits and Clear was finally ready to share some of them publicly.

Clear began by publishing a new article every Monday and Thursday.

Within a few months, this simple writing habit led to his 1st 1000 email subscribers, and by the end of 2013 that number had grown to more than 30,000 people.

%
In 2014, Clear's email list expanded to over 100,000 subscribers, which made it 1 of the fastest-growing newsletters on the internet.

Clear had felt like an impostor when he began writing 2 years earlier, but now Clear was becoming known as an expert on habits - a new label that excited him but also felt uncomfortable.

Clear had never considered himself a master of the topic, but rather someone who was experimenting alongside his readers.

%
In 2015, Clear reached 200,000 email subscribers and signed a book deal with Penguin Random House to begin writing the book you are reading now.

As Clear's audience grew, so did Clear's business opportunities.

Clear was increasingly asked to speak at top companies about the science of habit formation, behavior change, and continuous improvement.

Clear found himself delivering keynote speeches at conferences in the United States and Europe.

%
In 2016, Clear's articles began to appear regularly in major publications like \textit{Time, Entrepreneur}, and \textit{Forbes}.

Incredibly, Clear's writing was read by over 8 million people that year.

Coaches in the NFL, NBA, and MLB began reading his work and sharing it with their teams.

%
At the start of 2017, Clear launched the Habits Academy, which became the premier training platform for organizations and individuals interested in building better habits in life and work.\footnote{Interested readers can learn more at \url{habitsacademy.com}.}

Fortune 500 companies and growing start-ups began to enroll their leaders and train their staff.

In total, over 10,000 leaders, managers, coaches, and teachers have graduated from the Habits Academy, and Clear's work with them has taught him an incredible amount about what it takes to make habits work in the real world.

%
As Clear put the finishing touches on this book in 2018, \url{jamesclear.com} is receiving millions visitors per month and nearly 500,000 people subscribe to Clear's weekly email newsletter - a number that is so far beyond Clear's expectations when he began that he's not even sure what to think of it.

\subsubsection*{How this book will benefit you}
The entrepreneur and investor Naval Ravikant has said, \fbox{``To write a great book, you must 1st become the book.''}

Clear originally learned about the ideas mentioned here because Clear had to live them.

Clear had to rely on small habits to rebound from his injury, to get stronger in the gym, to perform at a high level on the field, to become a writer, to build a successful business, and simply to develop into a responsible adult.

Small habits helped me fulfill Clear's potential, and since you picked up this book, Clear is guessing you'd like to fulfill yours as well.

%
In the pages that follow, Clear will share a step-by-step plan for building better habits - not for days or weeks, but for a lifetime.

While science supports everything Clear has written, \fbox{this book is not an academic research paper; it's an operating manual.}

You'll find wisdom and practical advice front and center as Clear explains the science of how to create and change your habits in a way that is easy to understand and apply.

%
The fields Clear draws on - biology, neuroscience, philosophy, psychology, and more - have been around for many years.

What Clear offers you is a synthesis of the best ideas smart people figured out a long time ago as well as the most compelling discoveries scientists have made recently.

Clear's contribution, he hopes, is to find the ideas that matter most and connect them in a way that is highly actionable.

Anything wise in these pages you should credit to the many experts who preceded him.

\fbox{Anything foolish, assume it is Clear's error.}

%
The backbone of this book is \fbox{Clear's 4-step model of habits - cue, craving, response, and reward} - and the 4 laws of behavior change that evolve out of these steps.

Readers with a psychology background may recognize some of these terms from operant conditioning, which was 1st proposed as ``stimulus, response, reward'' by B. F. Skinner in the 1930s and has been popularized more recently as ``cue, routine, reward'' in \textit{The Power of Habit} by Charles Duhigg.

%
Behavioral scientists like Skinner realized that if you offered the right reward or punishment, you could get people to act in a certain way.

But while Skinner's model did an excellent job of explaining how external stimuli influenced our habits, it lacked a good explanation for \textit{how our thoughts, feelings, and beliefs impact our behavior}.

\fbox{Internal states - our moods and emotions - matter, too.}

In recent decades, scientists have begun to determine the connection between our thoughts, feelings, and behavior.

This research will also be covered in these pages.

%
In total, the framework Clear offers is an integrated model of the cognitive and behavioral sciences.

Clear believes it is 1 of the 1st models of human behavior to accurately account for both the influence of external stimuli and internal emotions on our habits.

While some of the language may be familiar, Clear is confident that the details - and the applications of the 4 Laws of Behavior Change - will offer a new way to think about your habits.

%
\fbox{Human behavior is always changing: situation to situation, moment to moment, second to second.}

But this book is about what \textit{doesn't} change.

It's about the \fbox{fundamentals of human behavior}.

The lasting principles you can rely on year after year.

The ideas you can build a business around, build a family around, build a life around.

%
There is no one right way to create better habits, but this book describes the best way Clear knows - an approach that will be effective regardless of where you start or what you're trying to change.

The strategies Clear covers will be relevant to anyone looking for a step-by-step system for improvement, whether your goals center on health, money, productivity, relationships, or all of the above.

As long as human behavior is invoked, this book will be your guide.

\begin{center}\huge
    \textbf{The Fundamentals}:
    
    Why Tiny Changes Make a Big Difference
\end{center}

\section{The Surprising Power of Atomic Habits}

\section{How Your Habits Shape Your Identity (\& Vice Versa)}

\section{How to Build Better Habits in 4 Simple Steps}

\begin{center}\huge
    \textbf{The 1st Law}:
    
    Make It Obvious
\end{center}

\section{The Man Who Didn't Look Right}

\section{The Best Way to Start a New Habit}

\section{Motivation Is Overrated; Environment Often Matters More}

\section{The Secret to Self-Control}

\begin{center}\huge
    \textbf{The 2nd Law}:
    
    Make It Attractive
\end{center}

\section{How to Make a Habit Irresistible}

\section{The Role of Family \& Friends in Shaping Your Habits}

\section{How to Find \& Fix the Causes of Your Bad Habits}

\begin{center}\huge
    \textbf{The 3rd Law}:
    
    Make It Easy
\end{center}

\section{Walk Slowly, but Never Backward}

\section{The Law of Least Effort}

\section{How to Stop Procrastinating by Using the 2-Minute Rule}

\section{How to Make Good Habits Inevitable \& Bad Habits Impossible}

\begin{center}\huge
    \textbf{The 4th Law}:
    
    Make It Satisfying
\end{center}

\section{The Cardinal Rule of Behavior Change}

\section{How to Stick with Good Habits Every Day}

\section{How an Accountability Partner Can Change Everything}

\begin{center}\huge
    \textbf{Advanced Tactics}:
    
    How to Go from Being Merely Good to Being Truly Great
\end{center}

\section{The Truth About Talent (When Genes Matter \& When They Don't)}

\section{The Goldilocks Rule: How to Stay Motivated in Life \& Work}

\section{The Downside of Creating Good Habits}

\section{Conclusion: The Secret to Results That Last}

\begin{center}\huge
    \textbf{Appendix}
\end{center}

\section{What Should You Read Next?}

\section{Little Lessons from the 4 Laws}

\section{How to Apply These Ideas to Business}

\section{How to Apply These Ideas to Parenting}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}