\documentclass[onsided]{book}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/nguyen/1_NQBH/math/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\addtolength{\headheight}{0pt}% obsolete
\lhead{\itshape \scriptsize \chaptername~\thechapter}
\rhead{\itshape \scriptsize \nouppercase{\leftmark}} %\nouppercase !
\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\cfoot{\thepage}

\title{Marie Sk\l odowska-Curie Actions}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
\tableofcontents

%------------------------------------------------------------------------------%

\part{Wikipedia's}

\chapter{Wikipedia/Marie Sk\l odowska-Curie Actions}

\textit{Marie Skłodowska-Curie Actions (MSCA)} are a set of major research fellowships created by the \href{https://en.wikipedia.org/wiki/European_Union}{European Union}/\href{https://en.wikipedia.org/wiki/European_Commission}{European Commission} to support research in the \href{https://en.wikipedia.org/wiki/European_Research_Area}{European Research Area} (ERA).

The Marie Skłodowska-Curie Actions are among Europe's most competitive and prestigious research and innovation fellowships.[1][2]

%
Established in 1996 as Marie Curie Actions and known since 2014 as Marie Skłodowska-Curie Actions, they aim to foster the career development and further training of researchers at all career stages.

The Marie Skłodowska-Curie Actions promote interdisciplinary research and international collaborations, supporting scientists from not only within Europe but also across the globe.

%
Marie Skłodowska-Curie Actions are currently financed through the 8th \href{https://en.wikipedia.org/wiki/Framework_Programmes_for_Research_and_Technological_Development}{Framework Programme for Research and Technological Development} (called \href{https://en.wikipedia.org/wiki/Horizon_2020}{Horizon 2020}) and belong to the so-called `first pillar' of Horizon2020: ``Excellent Science.''

Through this funding scheme, the \href{https://en.wikipedia.org/wiki/Research_Executive_Agency}{Research Executive Agency} (REA) has devoted over €6 billion to the Marie Skłodowska-Curie Actions between 2014 and 2020.

%
Since the launch of the programme in 1996, over 100,000 researchers had received MSCA grants by Mar 2017.[3][4]

To mark this milestone, the European Commission selected thirty highly-promising researchers (who achieved the highest evaluation scores in 2016)[3] to showcase the EU's actions dedicated to excellence and worldwide mobility in research.[5]

%
\href{https://en.wikipedia.org/wiki/Marie_Curie}{Marie Skłodowska-Curie} is the \href{https://en.wikipedia.org/wiki/Poles_in_France}{Polish-French} namesake of the programme and was the \href{https://en.wikipedia.org/wiki/List_of_female_Nobel_laureates}{1st female Nobel prize winner}.

The only person \href{https://en.wikipedia.org/wiki/Nobel_Prize#Multiple_laureates}{to win a Nobel Prize for contributions in 2 different sciences} (\href{https://en.wikipedia.org/wiki/Nobel_Prize_in_Physics}{physics} and \href{https://en.wikipedia.org/wiki/Nobel_Prize_in_Chemistry}{chemistry}), she was also the 1st - and only woman - to have been awarded a Nobel Prize twice.

\section{Types of funding}
Fellowships are awarded by the European Commission across scientific disciplines within the framework of Horizon 2020.

%
MSCA are grouped into the following schemes:
\begin{itemize}
    \item Research Networks (ITN),
    \item Individual Fellowships (IF),
    \item Research and Innovation Staff Exchanges (RISE),
    \item Co-funding of regional, national and international programs involving mobility (COFUND),
    \item European Researchers' Night (NIGHT).
\end{itemize}
Within the framework of Horizon 2020, which runs from 2014 through 2020, MSCA will award €6.16 billion in funding.

\section{External links}
\begin{itemize}
    \item \href{http://ec.europa.eu/research/mariecurieactions/}{Main Marie Sk\l odowska-Curie Actions website}
    \item \href{http://cordis.europa.eu/fp7/people/}{People programme (FP7)}
    \item \href{https://www.mariecuriealumni.eu/}{Marie Curie Alumni Association}\hfill$\square$
\end{itemize}

%------------------------------------------------------------------------------%

\chapter{\href{https://ec.europa.eu/research/mariecurieactions/actions/individual-fellowships_en}{European Commission/Marie Skłodowska-Curie Actions}}

\section{\href{https://ec.europa.eu/research/mariecurieactions/}{Home}}

\subsection{Marie Skłodowska-Curie Actions - Research Fellowship Programme}
The Marie Skłodowska-Curie actions support researchers at all stages of their careers, regardless of age and nationality.

Researchers working across all disciplines are eligible for funding.

The MSCA also support cooperation between industry and academia and innovative training to enhance employability and career development.

\section{\href{https://ec.europa.eu/research/mariecurieactions/msca-actions_en}{Actions}}

\subsection{MSCA Actions}
\begin{quotation}
    Along with individual fellowships, the MSCA help develop training networks, promote staff exchanges and fund mobility programmes with an international flavor.
\end{quotation}
Grants provided by Marie Skłodowska-Curie Actions are available for all stages of a researcher's career.

Fellows include PhD candidates and those carrying out more advanced research.

%
Because they encourage individuals to work in other countries, the MSCA make the whole world a learning environment.

They encourage collaboration and sharing of ideas between different industrial sectors and research disciplines - all to the benefit of the wider European economy.

MSCA also back initiatives that break down barriers between academia, industry and business.

In addition, they reach out to the public with events that promote the value - and fun side - of science.

%
These inter-related goals are reflected in the various types of MSCA actions.

%
It all adds up to a huge investment.

In fact, the EU has set aside EUR 6.16 billion to be spent by 2020 on researcher training and career development.

\textit{Choose your action:}

\subsubsection{Innovative Training Networks (\href{https://ec.europa.eu/research/mariecurieactions/actions/research-networks_en}{ITN})}
ITN drive scientific excellence and innovation.

They bring together universities, research institutes and other sectors from across the world to train researchers to doctorate level.

\subsubsection{Individual Fellowships (\href{https://ec.europa.eu/research/mariecurieactions/actions/individual-fellowships_en}{IF})}
IF are great opportunities if you are an experienced researcher looking to give your career a boost by working abroad.

They offer exciting new learning opportunities and a chance to add some sparkle to your CV.

\subsubsection{Research and Innovation Staff Exchange (\href{https://ec.europa.eu/research/mariecurieactions/actions/staff-exchange_en}{RISE})}
RISE funds short-term exchanges of personnel between academic, industrial and commercial organizations throughout the world.

%
It helps people develop their knowledge, skills and careers, while building links between organizations working in different sectors of the economy, including universities, research institutes and SMEs.

\subsubsection{Co-funding of regional, national and international programmes (\href{https://ec.europa.eu/research/mariecurieactions/actions/co-funding-programmes_en}{COFUND})}
COFUND provides organizations with additional financial support for their own researcher training and career development programmes.

%
The extra funds are available for new or existing schemes for training researchers abroad and across various sectors.

\subsubsection{European Researchers' Night (\href{https://ec.europa.eu/research/mariecurieactions/actions/european-researchers-night_en}{NIGHT})}
This is a Europe-wide public event dedicated to popular science and fun learning. It takes place each year in September. Around 30 countries and over 300 cities are involved.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/actions/research-networks_en}{Research Networks}}
\begin{center}
    Innovative Training Networks (ITN) drive scientific excellence and innovation.
\end{center}

\begin{flushright}
    Funding scheme: \textbf{MSCA-ITN}
\end{flushright}
Innovative Training Networks (ITN) drive scientific excellence and innovation.

They bring together universities, research institutes and other sectors from across the world to train researchers to doctorate level.

\subsubsection{Types of ITN}
There are 3 types of Research networks:
\begin{itemize}
    \item European Training Networks;
    \item European Industrial Doctorates; and
    \item European Joint Doctorates.
\end{itemize}

\paragraph{European Training Networks} help researchers gain experience of different working environments while developing transferable skills.

%
They must involve at least 3 partners from inside and outside academia.

Organizations managing such a network should be established in at least 3 different \href{http://ec.europa.eu/research/participants/data/ref/h2020/other/wp/2016-2017/annexes/h2020-wp1617-annex-a-countries-rules_en.pdf}{EU or associated countries}, though additional participants can join from across the world.

\paragraph{European Industrial Doctorates} help PhD candidates step outside academia and develop skills in industry and business.

%
This type of network is provided by \textit{at least 1 academic partner} and \textit{partners from the business world}.

Individuals are enrolled in a doctoral programme and jointly supervised by the academic and non-academic partners.

%
Organizations offering Industrial Doctorates should be established in at least 2 different \href{http://ec.europa.eu/research/participants/data/ref/h2020/other/wp/2016-2017/annexes/h2020-wp1617-annex-a-countries-rules_en.pdf}{EU or associated countries}, though others are welcome to join.

\paragraph{European Joint Doctorates} promote international collaboration that cuts across different business and research sectors.

%
They are run by \textit{a minimum of 3 academic organizations} forming a network to supervise joint, double or multiple degrees.

These organizations should be from different \href{http://ec.europa.eu/research/participants/data/ref/h2020/other/wp/2016-2017/annexes/h2020-wp1617-annex-a-countries-rules_en.pdf}{EU or associated countries}.

Other bodies can join from anywhere across the world, including the non-academic sector.

\subsubsection{What can be funded?}
\textit{All research areas} can be funded - from molecular biology to urban development.

%
Proposals should:
\begin{itemize}
    \item reflect existing or planned \textit{research cooperation} among the partners;
    \item involve the researchers through \textbf{personalized research projects}
    \item explain how the \textit{recruitment of the researchers}, who must come from another country, will be open and transparent
\end{itemize}

\subsubsection{What does the funding cover?}
Grants cover:
\begin{itemize}
    \item \textit{recruitment and training} of each researcher for up to \textit{3 years}.
    
    The researcher is hired under an employment contract and benefits from a monthly living allowance, social security cover, plus a mobility and family allowance.
    \item \textit{research costs} including the organization of joint activities and conferences.
    \item \textit{management and indirect costs}.
    \item The maximum duration of an ITN project is \textit{4 years}.
\end{itemize}

\subsubsection{How do we apply?}
\href{https://ec.europa.eu/research/mariecurieactions/actions/get-funding_en}{Apply Now}.

Here you will find details of all active calls for Marie Skłodowska-Curie Actions.

The information is linked directly to the EU's Participant Portal, which provides all you need to start your application.

See also their page on ``\href{https://ec.europa.eu/research/mariecurieactions/how-to/apply_en}{How to get funding}''.

%
Advice and information can also be found through the \href{http://ec.europa.eu/research/participants/portal/desktop/en/support/national_contact_points.html}{National Contact Points} and \href{http://ec.europa.eu/research/index.cfm?pg=enquiries}{Research Enquiry Helpdesk}.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/actions/individual-fellowships_en}{Individual Fellowships}}
\begin{quotation}
    Individual Fellowships (IF) are a great option if you are an experienced researcher looking to give your career a boost by working abroad.
\end{quotation}
\textit{Individual Fellowships (IF)} are a great option if you are an experienced researcher looking to give your career a boost by working abroad.

They offer exciting new learning opportunities and a chance to add some sparkle to your CV.

\subsubsection{Types of Individual Fellowships}
There are 2 types of Individual Fellowships:
\begin{itemize}
    \item European Fellowships
    \item Global Fellowships
\end{itemize}

\paragraph{European Fellowships.}
\begin{itemize}
    \item are open to researchers moving within Europe, as well as those coming in from other parts of the world.
    \item can restart a research career after a break, e.g. parental leave.
    \item can help researchers coming back to Europe find a new position.
\end{itemize}
These Fellowships are held in the EU or associated countries and last for 1--2 years.

\paragraph{Global Fellowships.}
\begin{itemize}
    \item fund positions outside Europe for researchers based in the \href{http://ec.europa.eu/research/participants/data/ref/h2020/other/wp/2016-2017/annexes/h2020-wp1617-annex-a-countries-rules_en.pdf}{EU or associated countries}.
    \item last between 2 and 3 years.
    \item the researcher has to come back for one year to an organization based in the EU or associated countries.
\end{itemize}
Both types of Fellowship can also include a \textit{secondment period} of up to 3 or 6 months in another organization in Europe.

\subsubsection{Who can apply?}
This action is for \textit{experienced researchers} from across the world.

%
Applicants need a doctoral degree or at least 4 years' full-time research experience by the time of the call deadline.

\subsubsection{What can be funded?}
\textit{All research areas} can be funded.

MSCA Fellows come from a wide variety of disciplines - from physics to linguistics, and from health-sciences to mathematical modeling.

\subsubsection{What the funding covers}
The grant provides an \textit{allowance to cover living, travel and family costs}.

In addition, the EU contributes to the training, networking and research costs of the fellow, as well as to the management and indirect costs of the project.

The grant is awarded to the host organization, usually a university, research center or a company in Europe.

\subsubsection{How do I apply?}
You submit a \textit{research proposal}, including your CV.

The proposal is \textit{written jointly with your chosen host organisation(s)}.

\href{https://ec.europa.eu/research/mariecurieactions/actions/get-funding_en}{Apply Now}.

Here you will find details of all active calls for Marie Skłodowska-Curie Actions.

The information is linked directly to the EU's Participant Portal, which provides all you need to start your application.

%
Advice and information can also be found through the \href{http://ec.europa.eu/research/participants/portal/desktop/en/support/national_contact_points.html}{National Contact Points} and \href{http://ec.europa.eu/research/index.cfm?pg=enquiries}{Research Enquiry Helpdesk}.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/actions/staff-exchange_en}{Staff Exchange}}
\begin{quotation}
    Research and Innovation Staff Exchange (RISE) funds short-term exchanges of personnel between academic, industrial and commercial organizations throughout the world.
\end{quotation}
\textit{Research and Innovation Staff Exchange (RISE)} funds short-term exchanges of personnel between academic, industrial and commercial organizations throughout the world.

%
It helps people develop their knowledge, skills and careers, while building links between organizations working in different sectors of the economy, including universities, research institutes and SMEs.

\subsubsection{Who can apply?}
\textit{Proposals must include at least 3 partners}, which can be academic or non-academic organizations from \textit{3 different countries}.

At least \textit{2 of these should be from the} \href{http://ec.europa.eu/research/participants/data/ref/h2020/other/wp/2016-2017/annexes/h2020-wp1617-annex-a-countries-rules_en.pdf}{EU or associated countries}.

Partners from further afield can also join.

%
Exchanges between organizations in \href{http://ec.europa.eu/research/participants/data/ref/h2020/other/wp/2016-2017/annexes/h2020-wp1617-annex-a-countries-rules_en.pdf}{EU or associated countries} must be between different sectors.

In worldwide partnerships, exchanges within the same sector are possible.

For available local funding for participants from non-EU countries (not eligible for automatic funding) please refer to the \href{http://ec.europa.eu/research/participants/docs/h2020-funding-guide/cross-cutting-issues/international-cooperation_en.htm}{Horizon 2020 Online Manual}.

%
Proposals should highlight:
\begin{itemize}
    \item \textit{knowledge creation},
    \item \textit{sharing know-how}, and
    \item \textit{skills development}.
\end{itemize}
Research staff of \textit{any nationality} and at \textit{any career level} - from PhD preparation to experienced researchers - can undertake a secondment.

Staff members working in managerial, technical or administrative roles can also take part.

\subsubsection{What can be funded?}
This action is open to \textit{all research areas}, whether you are examining social issues like youth unemployment, or creating new solar power applications.

\subsubsection{How does it work?}
The grant supports the secondment of staff members for \textit{1 month to 1 year}.

They must be engaged in or linked to research and innovation activities at their home organization for a certain time prior to the secondment.

%
They return to their home organization after the secondment, to pass on their knowledge.

%
Funding for a RISE project can last up to 4 years.

\subsubsection{How to apply}
\href{https://ec.europa.eu/research/mariecurieactions/actions/get-funding_en}{Apply Now}.

Here you will find details of all active calls for Marie Skłodowska-Curie Actions.

The information is linked directly to the EU's Participant Portal, which provides all you need to start your application.

%
Advice and information can also be found through the \href{http://ec.europa.eu/research/participants/portal/desktop/en/support/national_contact_points.html}{National Contact Points} and \href{http://ec.europa.eu/research/index.cfm?pg=enquiries}{Research Enquiry Helpdesk}.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/actions/co-funding-programmes_en}{Co-funding}}
\begin{center}
    of regional, national and international programmes.
\end{center}
\textit{Co-funding of regional, national and international programmes (COFUND)} provides organizations with additional financial support for their own researcher training and career development programmes.

%
The extra funds are available for new or existing schemes for \textit{training researchers abroad and across various sectors}.

%
COFUND supports \textit{doctoral programmes} for \textit{PhD candidates, as well as fellowship programmes} for \textit{experienced researchers}.

\subsubsection{Who can apply?}
Each COFUND proposal should have \textit{1 main participant}, such as a:
\begin{itemize}
    \item government ministry
    \item regional authority
    \item funding agency
    \item university
    \item research organization
    \item enterprise.    
\end{itemize}
\textit{Experienced researchers} and \textit{doctoral candidates} apply directly to the relevant funding organization through vacancies for co-funded programmes on \href{http://ec.europa.eu/euraxess/index.cfm/jobs/index}{Euraxess}, which is a jobs portal for researchers.

\subsubsection{What can be funded?}
\textit{All research areas} can be funded.

\subsubsection{What does the funding cover?}
\textit{Participating organizations} receive a fixed amount for each supported researcher, as a contribution to their living allowance and for the programme's management costs.

%
Selected programmes will receive co-funding for up \textit{to 5 years}, for a maximum total amount of EUR 10 million.

\subsubsection{How do we apply?}
\href{https://ec.europa.eu/research/mariecurieactions/actions/get-funding_en}{Apply Now}.

Here you will find details of all active calls for Marie Skłodowska-Curie Actions.

The information is linked directly to the EU's Participant Portal, which provides all you need to start your application.

%
Advice and information can also be found through the \href{http://ec.europa.eu/research/participants/portal/desktop/en/support/national_contact_points.html}{National Contact Points} and \href{http://ec.europa.eu/research/index.cfm?pg=enquiries}{Research Enquiry Helpdesk}.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/actions/european-researchers-night_en}{European Researcher's Night}}
\begin{center}
    Get funding to organize an event, promote research and channel its impact to improve citizens' daily lives
\end{center}
\textsf{Source: European Union 2020}.

\subsubsection{European Researchers' Night}
The European Researchers' Night, funded under the Marie Skłodowska-Curie actions (MSCA), is a Europe-wide public event that brings researchers closer to the public.

The Night provides researchers the opportunity to showcase the diversity of science and its impact on citizens' daily lives, and to stimulate interest in research careers - especially among young people.

The events highlight how researchers contribute to our society by displaying their work in an interactive and engaging forum.

%
In 2019, 55 projects were implemented. The projects took place in 433 cities from 27 countries across Europe and beyond.

Over 1.6 million visitors attended the event and over 36,000 researchers took part - including 955 MSCA fellows!

%
In 2020, the European Researchers' Night will take place on Friday, Nov 27.

More information about this year's event will be available soon.

\subsubsection{Who can apply?}
European Researchers' Night grants may be awarded to any legal entity established in an EU Member State or \href{http://ec.europa.eu/research/participants/data/ref/h2020/grants_manual/hi/3cpart/h2020-hi-list-ac_en.pdf}{Associated Country}.

Legal entities involve the coordination of activities between local, regional, national or international partners.

%
Possible beneficiaries' profiles may include private and public research organizations, companies, public authorities, schools, science museums, parent-teacher organizations, EU mobility centers for researchers, foundations or the media.

\subsubsection{What can be funded?}
Any activity or event that boosts public \textit{awareness of the positive role of research in society} - particularly among \textit{young people} - is eligible for funding.

%
EU-funded researchers are encouraged to contribute and interact as much as possible with visitors by displaying their projects and promoting the use of research.

\subsubsection{What does funding cover?}
The funding covers \textit{any expenses linked to the organization of a research outreach event}.

It can cover the preparation of such event, on an awareness campaign, pre-events, the event itself, and the assessment of its impact.

These activities may include:
\begin{itemize}
    \item \textit{hands-on experiments} conducted by researchers
    \item \textit{science shows} with public participation
    \item \textit{debates}
    \item ``\textit{researchers' speed dating}'' (meet researchers and ask them questions)
    \item \textit{competitions} (e.g., science quizzes, games, puzzles, photo and art contests, etc.)
    \item \textit{workshops} for children
    \item \textit{science slams}
    \item \textit{guided visits} of labs, research institutes, and other relevant places that are usually closed to the public
\end{itemize}
The value of grants varies according to the scale of the events proposed.

\subsubsection{Who decides?}
All projects are selected through an \textit{open, independent and transparent evaluation}.

\textit{The competition uses a series of pre-determined criteria} (listed in the call for proposals).

\subsubsection{How do we apply?}
The call for the 2021 edition of the European Researchers' Night was published on Oct 1, 2020.

The deadline for applications is Jan 12, 2021.

All the information to apply is available on the \href{https://ec.europa.eu/info/funding-tenders/opportunities/portal/screen/opportunities/topic-details/msca-night-2020bis;freeTextSearchKeyword=;typeCodes=0,1;statusCodes=31094501,31094502,31094503;programCode=null;programDivisionCode=null;focusAreaCode=null;crossCuttingPriorityCode=null;callCode=Default;sortQuery=submissionStatus;orderBy=asc;onlyTenders=false;topicListKey=topicSearchTablePageState}{Funding \& Tender Opportunities Portal}.

\subsubsection{Visit the events!}
$\ldots$\textit{as 1.6 million visitors did in 2019!}

In 2020, the European Researchers' Night will take place on Friday, 27 November.

%
See \href{https://ec.europa.eu/research/mariecurieactions/events/european-researchers-night-2020_en}{what's happening on the night}.

%
Learn how research is fighting cancer, stopping global warming, preventing hunger and drought, improving technology to assist disabled persons and tackling challenges for life in space.

Meet researchers and discover the fascinating work of science in a fun and interactive environment - with family, friends, your school or on your own.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/actions/get-funding_en}{Funding Opportunities}}
\begin{quotation}
    To receive funding, you and/or your organization need to apply to an open call. They are listed here, along with the deadline for application.
\end{quotation}
Click on the link for the call and access the European Union's Participant Portal to obtain all the information and documents you will need.

%
Keep checking this page, as new calls are added on a regular basis.

\href{https://ec.europa.eu/research/mariecurieactions/how-to/apply\_en}{How to apply}.

\section{How to$\ldots$}

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/how-quick-guide_en}{How to$\ldots$ a quick guide}}
\begin{quotation}
    A quick and easy-to-use guide to applying for funding, framing proposals, getting jobs and managing projects under Marie Skłodowska-Curie Actions.
\end{quotation}

\subsubsection{How to get started}
To get funding, either respond to a job vacancy or submit a proposal, together with an organization, to 1 of the calls published every year.

Calls are published on the \href{http://ec.europa.eu/research/participants/portal/desktop/en/home.html}{Participant Portal}, the single-entry point for research and innovation programmes run by the European Union.

%
Find more information in the \href{https://ec.europa.eu/research/mariecurieactions/how-to/get-started_en}{Get started} section.

\subsubsection{How to find a job}
Use the \href{https://euraxess.ec.europa.eu/}{EURAXESS} portal to find posts related to Marie Skłodowska-Curie Actions fellowships.

%
EURAXESS holds details about thousands of jobs and researcher opportunities in an easy-to-search database.

\textit{You can also post your CV on the site}.

Whether it is a PhD fellowship in the US, an electronics engineer for an SME in Malta, or investigating the archaeology of historical buildings in Italy - there really is something for everyone.

%
Find more information in the \href{https://ec.europa.eu/research/mariecurieactions/how-to/find-job_en}{Find a job} section.

\subsubsection{How to apply}
The \textit{Participant Portal} will guide you through the process of \textit{preparing and submitting} a proposal.

%
The EU's \href{https://ec.europa.eu/info/departments/research-executive-agency_en}{Research Executive Agency} (REA) \textit{evaluates all eligible proposals}, identifying those good enough to receive funding.

The evaluation is based on a peer-review carried out by independent experts.

%
REA then \textit{establishes grant agreements with} applicants whose proposals have passed the evaluation stage.

\textit{The process is designed to work for you! Thousands of individuals and organizations have secured grant funding} - If you want to know more about the application process, you can soon check out our infographic.

%
Find more information in the \href{https://ec.europa.eu/research/mariecurieactions/how-to/apply_en}{Apply} section.

\subsubsection{How to manage your project}
When managing a project that has received EU funding, you will be expected to:
\begin{itemize}
    \item keep good, up-to-date \textit{records} throughout the project's lifetime;
    \item provide \textit{progress reports} and information about \textit{payments};
    \item submit a list of \textit{deliverables};
    \item devise a plan for the \textit{dissemination and exploitation} of results; and
    \item allow the EU to carry out any necessary \textit{checks, audits, reviews} and \textit{investigations}.
\end{itemize}
Project coordinators should also make steps to \textit{acknowledge funding} from the EU.

%
More information is available on the `How to manage your project' page [link to page] on this site and through the \href{http://ec.europa.eu/research/participants/docs/h2020-funding-guide/grants/grant-management_en.htm}{Participant Portal}.

Find more information in the \href{https://ec.europa.eu/research/mariecurieactions/how-to/manage-your-project_en}{Manage your project} section

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/how-to/get-started_en}{How to get started}}
\begin{center}
    How to get started and see which of the Marie Skłodowska-Curie Actions on offer is most suited to you.
\end{center}
\textit{Check out MSCA's \href{https://ec.europa.eu/research/mariecurieactions/msca-actions_en}{dedicated pages} to see which of the Marie Skłodowska-Curie Actions on offer is most suited to you}.

%
If you are an experienced researcher looking to carry out your research in or outside the EU, the Individual fellowships might be for you.

Working in an organization seeking to expand their network?

Then check out Research networks.

Once you know which of the 5 actions fits your needs, you can take a closer look at what costs are covered by the EU on the action page.

%
Need inspiration? Check out the \href{https://ec.europa.eu/research/mariecurieactions/resources/testimonials_en}{testimonials} and \href{https://ec.europa.eu/research/mariecurieactions/resources/success-stories_en}{success} stories from the MSCA alumni.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/how-to/apply_en}{How to apply}}
\begin{quotation}
    To get funding you and/or the research organisation have to submit a proposal to 1 of the calls published on the Participant Portal, the single-entry point for the EU research and innovation programmes.
\end{quotation}
\textit{To get funding you and/or the research organization have to submit a proposal to 1 of the calls published on the \href{http://ec.europa.eu/research/participants/portal/desktop/en/home.html}{Participant Portal}, the single-entry point for the EU research and innovation programmes}.

%
The \href{https://ec.europa.eu/research/mariecurieactions/actions/get-funding_en}{Get Funding} page gives you an overview of the open calls and their deadlines.

Find out more about the \textit{application process} below.

\subsubsection{Who can apply?}
\begin{itemize}
    \item Individual researchers together with the host organizations apply for Individual fellowships (IF)
    \item International networks of research organizations from the academic and non-academic sectors apply to run research networks (ITN) and staff exchanges (RISE)
    \item Organizations funding or managing doctoral or fellowship programmes apply for co-funding  programmes (COFUND)
    \item Any organization able  to organize a public event to promote science and research can apply for the European Researchers' Night (NIGHT)
\end{itemize}
Check out the \href{https://ec.europa.eu/research/mariecurieactions/sites/mariecurie2/files/dgeac-msca-infographic-how-to-apply_en.pdf}{6 steps to MSCA funding}.

\subsubsection{Need more help \& guidance?}
Check out these guides for in-depth advice about making a proposal and applying for MSCA funding:
\begin{itemize}
    \item \href{https://ec.europa.eu/research/participants/portal/desktop/en/support/faq.html#c,faqs=categoryMachineName/t/pp_roles_and_rights/0/1/1&categoryMachineName/t/p_registration/0/1/1&categoryMachineName/t/work_programme_calls/0/1/1&categoryMachineName/t/p_submission_eval/1/1/1&categoryMachineName/t/ethics_research_integrity/0/1/1&category}{FAQs on Proposal submission and evaluation}
    \item PDF Guides for Applicants:
    \begin{itemize}
        \item \href{http://ec.europa.eu/research/participants/data/ref/h2020/other/guides_for_applicants/h2020-guide-appl-msca-if-2018-20_en.pdf}{Individual Fellowships (IF)}
        \item \href{http://ec.europa.eu/research/participants/data/ref/h2020/other/guides_for_applicants/h2020-guide-appl-msca-rise_en.pdf}{Staff Exchanges (RISE)}
        \item \href{http://ec.europa.eu/research/participants/data/ref/h2020/other/guides_for_applicants/h2020-guide-appl-msca-itn_en.pdf}{Research Networks (ITN)}
        \item \href{http://ec.europa.eu/research/participants/data/ref/h2020/other/guides_for_applicants/h2020-guide-appl-msca-night-2018_en.pdf}{European Researcher's Night (NIGHT)}
    \end{itemize}
\end{itemize}
The \href{http://ec.europa.eu/research/participants/portal/desktop/en/support/national_contact_points.html}{National Contact Point} in your country can advise you on administrative matters.

Services provided may vary.

Check the \href{http://ec.europa.eu/research/participants/portal/desktop/en/home.html}{Participant Portal} for more information.

\subsubsection{Looking for a job?}
Go to the \href{https://ec.europa.eu/research/mariecurieactions/how-to/find-job_en}{How to find a job} page on this website.

All jobs in projects financed by MSCA are published on the EU's \href{https://euraxess.ec.europa.eu/}{EURAXESS} jobs portal.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/how-to/find-job_en}{How to find a job}}
\begin{center}
    As a professional researcher, you may also be looking for work.
\end{center}
\textit{As a professional researcher, you may also be looking for work. The \href{https://euraxess.ec.europa.eu/}{EURAXESS - Researchers in Motion} portal holds details about thousands of jobs and opportunities for researchers, both inside and outside the EU. Register and use the portal to find posts on existing MSCA-funded projects}.

%
The database is easy to use.

You can \textit{search by research field, sector or even by EU programme} - allowing you to narrow your search down to MSCA posts only.

Job offers come from across the EU and beyond, and you can even post your CV on the site.

%
The portal is more than just a job service: you can find advice about \textit{visas, pensions, recognition of qualifications} and other issues related to working as a researcher abroad.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/how-to/manage-your-project_en}{How to manage your project}}
\begin{quotation}
    If you are about to start managing a Marie Skłodowska-Curie Actions project, there are certain activities you must undertake as part of your grant agreement.
\end{quotation}
\textit{If you are about to start managing a Marie Skłodowska-Curie Actions project, there are certain activities you must undertake as part of your grant agreement. Following the steps below will assist you in your work and help you deliver great results}.

\subsubsection{Keeping records}
Adequate, up-to-date \textit{records} of your project should be kept.

This shows the EU that the project is being properly rolled out, and proves that funding is being spent in accordance with the rules.

\subsubsection{Reports \& payments}
You must submit \textit{technical and financial reports} - including requests for payments.

Specifically, these include:
\begin{itemize}
    \item \textit{periodic reports} within 60 days of the end of each project reporting period; and
    \item a \textit{final report} within 60 days of the end of the last reporting period.
    
    This should include a summary for publication.
\end{itemize}

\subsubsection{Submitting deliverables}
\textit{Deliverables} are produced throughout the project's lifetime and can include:
\begin{itemize}
    \item special reports;
    \item technical brochures;
    \item lists of extra information;
    \item software milestones; and
    \item any other output that adds to the project's body of work.    
\end{itemize}

\subsubsection{Dissemination and exploitation of results}
Good dissemination plans make it easier for new knowledge to reach its audience.

They also show taxpayers that their money is being well spent.

%
\textit{Dissemination} activities feed into \textit{exploitation}, which is about using the scientific results in the commercial, industrial and policy-making worlds.

%
Try to take account of some of the following issues when developing your dissemination and exploitation strategies:
\begin{itemize}
    \item Link your project to the \textit{policy context} in the call for proposals;
    \item Involve \textit{potential end-users} and \textit{stakeholders} in your proposal;
    \item Say how you expect the \textit{results} of your project to be \textit{applied};
    \item Show you understand the \textit{barriers to any application} of your results.    
\end{itemize}

\subsubsection{Communicating your project}
Project partners will also be expected to \textit{promote the project} and its results to the media and the general public.

Communication activities \textit{must} be part of any project proposal.

%
A comprehensive \textit{communication plan} should therefore define clear objectives adapted to a range of target audiences.

\subsubsection{Acknowledgement of EU funding}
All \textit{communication and dissemination activities} - as well as any \textit{equipment and infrastructure} (e.g., a new building) - must display references to EU funding and carry the \textit{EU emblem}, e.g., on a piece of equipment or on any promotional material of your training network.

Whatever you choose, \textit{display prominently} so people can see how the EU has helped make it happen!

\subsubsection{Checks, audits, reviews \& investigations}
The EU might carry out a number of checks, audits and reviews, both during the project and after it is completed.

It may, therefore:
\begin{itemize}
    \item \textit{check} the project is being properly run and in compliance with the grant agreement.
    
    That could mean checking documentation, assessing whether work plans are being followed, and searching for double-funding and even plagiarism.
    \item \textit{carry out reviews} into project management, including assessing deliverables and reports.
    \item \textit{conduct investigations} in case fraud is suspected.
    \item \textit{launch audits} of your MSCA project during its lifetime or up to 2 years after the final payment.
\end{itemize}
For detailed information on these issues, please consult the Horizon 2020 \href{http://ec.europa.eu/research/participants/portal/desktop/en/home.html}{Participant Portal} where you will find the \href{http://ec.europa.eu/research/participants/data/ref/h2020/grants_manual/amga/h2020-amga_en.pdf}{Annotated Model Grant Agreement}, the \href{http://ec.europa.eu/research/participants/docs/h2020-funding-guide/grants/grant-management_en.htm}{Grant Management Manual} and the \href{http://ec.europa.eu/research/participants/data/ref/h2020/other/gm/h2020-guide-comm_en.pdf}{Guide on Communication Aspects}, as well as all \href{http://ec.europa.eu/research/participants/portal/desktop/en/support/faq.html#c,faqs=categoryMachineName/t/pp_roles_and_rights/0/1/1/category&tagList/t/MSCA/1/1/1/tag}{FAQs}.

\section{News \& Events}

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/news_en}{News}}
\begin{center}
    Items related to Marie Skłodowska-Curie Actions
\end{center}
See \href{https://ec.europa.eu/research/mariecurieactions/news_en}{News} for a massive list of news.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/events_en}{Events}}
\begin{center}
    Events related to Marie Skłodowska-Curie Actions
\end{center}
See \href{https://ec.europa.eu/research/mariecurieactions/events_en}{Events} for a massive list of events.

\section{Resources}

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/resources/corona-information-MSCA_en}{Coronavirus: information for the MSCA Community}}
\begin{quotation}
    The Marie Skłodowska-Curie Actions are providing up-to-date, responsive information related to the COVID-19 crisis to MSCA fellows and project managers.
\end{quotation}
The COVID-19 outbreak has caused major disruptions for many researchers, projects and organizations, including those working with EU funding.

The Marie Skłodowska-Curie Actions (MSCA) are therefore seeking to provide responsive, up-to-date information for all MSCA researchers and projects.

%
Researchers, project managers and interested stakeholders can access information on relevant news and initiatives, call updates or frequently asked questions, and also learn about inspiring stories on how MSCA researchers and projects are working to tackle the spread of coronavirus and its wider effects on our society.

\subsubsection{MSCA projects relevant to COVID-19 \& related topics}
The current COVID-19 outbreak has not caught EU-funded research off guard.

The Marie Skłodowska-Curie Actions are supporting outstanding researchers in finding solutions to challenges posed by COVID-19 and other infectious diseases and their consequences on our societies.

%
Read more about these MSCA projects on the \href{https://ec.europa.eu/research/mariecurieactions/msca-projects-researching-covid-19-sars-cov-2-and-related-topics_en}{dedicated page}.

\subsubsection{Information for MSCA Fellows \& Project Managers}
\begin{itemize}
    \item \textbf{Extension Information for MSCA Individual Fellowships.}
    
    On 21/04/2020 the European Commission published \href{https://ec.europa.eu/research/mariecurieactions/news/coronavirus-extensions-msca-projects_en}{an information note} for MSCA Individual Fellowships (IF) projects and fellows impacted by confinement measures due to the COVID-19 outbreak.
    \item \textbf{Open Access guidelines for H2020 projects working on COVID-19, SARS-CoV-2 and related topics.}
    
    The European Commission is calling upon all H2020 projects working on COVID-19, SARS-CoV-2 and related topics to provide immediate open access to their related publications, data and any other output.
    
    To support the projects, a \href{https://ec.europa.eu/research/participants/docs/h2020-funding-guide/cross-cutting-issues/open-access-data-management/open-access_en.htm}{guidance document} has been developed that includes specific guidelines regarding aspects such as FAIR data, open access or data management plans.
    \item \textbf{COVID-19 Data Portal.}
    
    The \href{https://www.covid19dataportal.org/}{COVID-19 Data Portal} was launched in Apr 2020 to bring together relevant datasets for sharing and analysis in an effort to accelerate coronavirus research.
    
    It enables researchers to upload, access and analyze COVID-19 related reference data and specialist datasets as part of the wider European COVID-19 Data Platform.
    \item \textbf{Horizon Results Platfrom – COVID-19 relevant results.}
    
    The \href{https://ec.europa.eu/info/funding-tenders/opportunities/portal/screen/opportunities/horizon-results-platform}{Horizon Results Platform} has created a dedicated section with \href{https://ec.europa.eu/info/funding-tenders/opportunities/portal/screen/opportunities/horizon-results-platform/search;campaigns=1}{results relevant to COVID-19} developed by H2020 projects.
    
    MSCA projects can upload their results and related information, e.g., their needs to develop them further and relevance for policy making.
    
    MSCA projects can find guidance on how to upload their results to the platform in the \href{https://webgate.ec.europa.eu/funding/display/ECResearchGMS/Creating+a+Project+Result}{guidelines} available online.
    \item \textbf{Research/Innovation Corona Platform FAQs.}
    
    The European Research Area (ERA) operates a central \href{https://ec.europa.eu/info/funding-tenders/opportunities/portal/screen/covid-19}{Corona Platform} for all information on how the current crisis may effect research and innovation projects.
    
    Some of the most frequently asked questions (FAQs) by MSCA project coordinators and fellows can be found on the \href{https://ec.europa.eu/info/funding-tenders/opportunities/portal/screen/support/faq;type=1;categories=;programme=null;actions=;keyword=covid}{platform's FAQ page}.
    
    New MSCA FAQs have been added on Apr 17, 2020, additional FAQs might be published in the future.
    \item \textbf{REA News.}
    
    \href{https://ec.europa.eu/info/news/covid-19-implications-your-msca-project-2020-mar-16_en?fbclid=IwAR1GgsbNJkI7jLCHa18VDIXX_OAlR0_soWsN3LvzzYdulnT_Im5umE_GPGE}{The Research Executive Agency (REA)'s news} on the coronavirus can offer advice to MSCA project coordinators and fellows whose projects and activities may be affected by the coronavirus outbreak.
    
    In order to minimize disruptions caused by COVID-19 on business continuity, The REA has also published a \href{https://ec.europa.eu/info/news/covid-19-research-executive-agency-business-continuity-measures-2020-mar-18_en}{news item detailing the latest decisions} taken at the European Commission level that have an impact on its research management activities.
    
    The news item also contains relevant links and information on calls for proposals, evaluations and experts, grant management and links to FAQs for project managers and fellows.
    \item \textbf{Call extensions.}
    
    In light of the COVID-19 outbreak and the difficulties applicants may have experienced in submitting their proposals, the Commission has decided to \href{https://ec.europa.eu/research/mariecurieactions/news/RISE-2020-deadline-extended_en}{Extend the RISE 2020 call} deadline by 2 weeks, until 12 May 2020, 17:00 Brussels time.    
\end{itemize}

\subsubsection{Related information}
\begin{itemize}
    \item \textbf{Common European Response.}
    
    The European Commission (EC) is coordinating a \textit{Common European Response} to the COVID-19 outbreak.
    
    For more information about the EC's work to mitigate its effects across Europe, visit the \href{https://ec.europa.eu/info/live-work-travel-eu/health/coronavirus-response_en}{EC's Coordinated Response page}.
    \item \textbf{Coronavirus: online learning resources.}
    
    The Education and Training website of the European Commission (EC) has prepared a selection of \href{https://ec.europa.eu/education/resources-and-tools/coronavirus-online-learning-resources_en}{online resources and tools} for learners, teachers and educators during the outbreak of COVID-19.
    \item \textbf{Commission-funded coronavirus and pandemic research.}
    
    The European Commission has been at the forefront of supporting research and coordination of European and global efforts to tackle the spread of coronavirus, including preparedness for pandemics and other outbreaks.
    
    Some of this amazing research can be referenced via the \href{https://ec.europa.eu/info/research-and-innovation/research-area/health-research-and-innovation/coronavirus-research-and-innovation\_en}{coronavirus research and innnovation pages}.
    
    Information on other research and innovation actions of the European Commission can also be found on the \href{https://ec.europa.eu/info/research-and-innovation/research-area/}{research areas support page}.
    \item \textbf{EUvsVirus Pan-European Hackathon.}
    
    The European Commission, in close collaboration with EU member states, will host the \href{https://ec.europa.eu/research/mariecurieactions/news/commission-launches-euvirus-hackathon_en}{Pan-European hackathon \#EUvsVirus} to work in developing innovative solutions for coronavirus-related challenges.
    
    The event will take place from Apr 24--26, 2020.
    
    Applications can be sent through the \href{https://euvsvirus.org/}{hackathon's website}.
\end{itemize}

\subsubsection{The MSCA on Social Media}
The MSCA social media are continuously updated with relevant information for MSCA project coordinators and fellows, as well as testimonials of top-notch MSCA fellows and projects working to find solutions to challenges posed by coronaviruses and other infectious diseases.

%
The MSCA can found on social media at:
\begin{itemize}
    \item \href{https://twitter.com/MSCActions}{MSCA on Twitter}
    \item \href{https://www.facebook.com/Marie.Curie.Actions}{MSCA Facebook page}
    \item \href{https://twitter.com/i/events/1250381718831411200}{Testimonials of leading MSCA fellows}
\end{itemize}

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/resources/document-library_en}{Library}}
\begin{center}
    Legal documents, Policy documents, Reports, Publications, Guides, Factsheets and more.
\end{center}
See \href{https://ec.europa.eu/research/mariecurieactions/resources/document-library_en}{Library} for a massive list.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/resources/success-stories_en}{Success Stories}}
\begin{center}
    Related to Marie Skłodowska-Curie Actions
\end{center}
See \href{https://ec.europa.eu/research/mariecurieactions/resources/success-stories_en}{Success Stories}.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/resources/testimonials_en}{Testimonials}}
\begin{center}
    From Fellows and Researchers
\end{center}
See \href{https://ec.europa.eu/research/mariecurieactions/resources/testimonials_en}{Testimonials}.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/funded-projects_en}{Projects funded by Marie Skłodowska-Curie Actions}}
\begin{center}
    This is just a small sample of the projects funded
\end{center}
See \href{https://ec.europa.eu/research/mariecurieactions/funded-projects_en}{Projects funded by Marie Skłodowska-Curie Actions}.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/msca-numbers_en}{MSCA in Numbers}}
\begin{center}
    Facts and figures about Marie Skłodowska-Curie Actions
\end{center}
See \href{https://ec.europa.eu/research/mariecurieactions/msca-numbers_en}{MSCA in Numbers} for a full list.

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/resources/photos_en}{Photo gallery}}
\begin{center}
    Photos related to Marie Skłodowska-Curie Actions
\end{center}
See \href{https://ec.europa.eu/research/mariecurieactions/resources/photos/msca-satellite-event_en}{MSCA Satellite Event - ESOF 2018}. 

\subsection{\href{https://ec.europa.eu/research/mariecurieactions/resources/videos_en}{Video gallery}}
\begin{center}
    Videos related to Marie Skłodowska-Curie Actions
\end{center}
See \href{https://ec.europa.eu/research/mariecurieactions/resources/videos_en}{Video gallery}.

\section{\href{https://ec.europa.eu/research/mariecurieactions/contact_en}{Contact}}
\begin{center}
    How to find out more about the Marie Skłodowska-Curie Actions
\end{center}

\subsection{Marie Skłodowska-Curie Actions - General contact}
If you have any questions about the Marie Skłodowska-Curie Actions (\textit{calls for proposals, application procedures, etc.}) please submit it to the \href{http://ec.europa.eu/info/research-and-innovation/contact/research-enquiry-service-and-participant-validation_en}{Research Enquiry Service}.

%
The service is the main contact point for future and current fellows, universities, research organizations, public institutions and businesses to answer all questions about their participation in the MSCA and in the European Union's Research Framework Programmes.

\subsection{Marie Skłodowska-Curie Actions - Local contacts}
If you have any questions about the Marie Skłodowska-Curie Actions and need further assistance, practical information, guidance and assistance on all aspects of participation in the MSCA are available \textit{in your local language}.

Submit your query to your \textit{Marie Skłodowska-Curie Actions National Contact Point} (MSCA NCP).

%
Visit the \href{https://ec.europa.eu/info/funding-tenders/opportunities/portal/screen/support/ncp}{Funding and tender opportunities portal} for the complete list of MSCA NCPs.

To find your NCP, you should select your country and ``Marie Skłodowska-Curie'' in the list of functions.

\subsection{Marie Skłodowska-Curie Actions - Frequently asked questions}
You can also find a list of \href{https://ec.europa.eu/info/funding-tenders/opportunities/portal/screen/support/faq}{FAQs on the Marie Skłodowska-Curie Actions} on the \href{https://ec.europa.eu/info/funding-tenders/opportunities/portal/screen/home}{Funding and tender opportunities portal}.

\subsection{Press}
The Marie Skłodowska-Curie Actions are managed by \href{https://ec.europa.eu/info/departments/education-youth-sport-and-culture\_en}{DG Education, Youth, Sport and Culture} of the European Commission.

The Commissioner responsible for the programme is Mariya Gabriel.

%
If you are a journalist, contact the European Commission \href{https://ec.europa.eu/info/about-european-commission/contact/press-services/press-contacts/press-contacts-eu-countries_en}{press officer in your country}.\hfill$\square$

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}