\documentclass{book}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/nguyen/1_NQBH/math/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\addtolength{\headheight}{0pt}% obsolete
\lhead{\itshape \scriptsize \chaptername~\thechapter}
\rhead{\itshape \scriptsize \nouppercase{\leftmark}} %\nouppercase !
\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\cfoot{\thepage}

\title{Psychological Manipulation}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
\tableofcontents

%------------------------------------------------------------------------------%

\part{Wikipedia's}

\chapter{\href{https://en.wikipedia.org/wiki/Psychological_manipulation}{Wikipedia/Psychological Manipulation}}

\textsf{Examples of \href{https://en.wikipedia.org/wiki/Television}{televised} manipulation can be found in \href{https://en.wikipedia.org/wiki/News_programs}{news programs} that can potentially influence mass audiences. Pictured is the infamous \href{https://en.wikipedia.org/wiki/Dziennik_Telewizyjny}{\textit{Dziennik}} (Journal) news cast, which attempted to slander \href{https://en.wikipedia.org/wiki/Capitalism}{capitalism} in \href{https://en.wikipedia.org/wiki/Polish_People's_Republic}{then-communist} Poland using \href{https://en.wikipedia.org/wiki/Loaded_language}{emotive and loaded language}.}

\textit{Psychological manipulation} is a type of \href{https://en.wikipedia.org/wiki/Social_influence}{social influence} that aims to change the behavior or \href{https://en.wikipedia.org/wiki/Perception}{perception} of others through indirect, \href{https://en.wikipedia.org/wiki/Deceptive}{deceptive}, or underhanded tactics.[``Definition of 'Manipulate'''. www.merriam-webster.com. Retrieved 2019-02-24.]

By advancing the interests of the manipulator, often at another's expense, such methods could be considered exploitative and devious.

%
\href{https://en.wikipedia.org/wiki/Social_influence}{Social influence} is not necessarily negative.

E.g., people such as friends, family and doctors, can try to \href{https://en.wikipedia.org/wiki/Persuade}{persuade} to change clearly unhelpful habits and behaviors.

Social influence is generally perceived to be harmless when it respects the right of the influenced to accept or reject it, and is not unduly coercive.

Depending on the context and \href{https://en.wikipedia.org/wiki/Motivations}{motivations}, social influence may constitute underhanded manipulation.

\section{Theories on successful manipulation}
According to psychology author \href{https://en.wikipedia.org/wiki/George_K._Simon}{George K. Simon}, successful psychological manipulation primarily involves the manipulator:[Simon, George K (1996). \textit{In Sheep's Clothing: Understanding and Dealing with Manipulative People}. ISBN 978-1-935166-30-6.]
\begin{itemize}
    \item Concealing \href{https://en.wikipedia.org/wiki/Aggression}{aggressive} intentions and behaviors and being affable.
    \item Knowing the psychological \href{https://en.wikipedia.org/wiki/Vulnerabilities}{vulnerabilities} of the victim to determine which tactics are likely to be the most effective.
    \item Having a sufficient level of ruthlessness to have no qualms about causing harm to the victim if necessary.
\end{itemize}
Consequently, the manipulation is likely to be accomplished through covert aggressive means.[Simon, George K (1996). In Sheep's Clothing: Understanding and Dealing with Manipulative People. ISBN 978-1-935166-30-6.]

\subsection{According to Braiker}
Harriet B. Braiker (2004) identified the following ways that manipulators \href{https://en.wikipedia.org/wiki/Abusive_power_and_control}{control} their victims:[Braiker, Harriet B. (2004). \textit{Who's Pulling Your Strings? How to Break The Cycle of Manipulation}. ISBN 978-0-07-144672-3.]
\begin{itemize}
    \item \href{https://en.wikipedia.org/wiki/Positive_reinforcement}{Positive reinforcement}: includes \href{https://en.wikipedia.org/wiki/Praise}{praise}, \href{https://en.wikipedia.org/wiki/Superficial_charm}{superficial charm}, superficial \href{https://en.wikipedia.org/wiki/Sympathy}{sympathy} (\href{https://en.wikipedia.org/wiki/Crocodile_tears}{crocodile tears}), excessive apologizing, money, approval, gifts, attention, facial expressions such as a forced laugh or smile, and public recognition.
    \item \href{https://en.wikipedia.org/wiki/Negative_reinforcement}{Negative reinforcement}: involves removing one from a negative situation as a reward.
    \item \href{https://en.wikipedia.org/wiki/Reinforcement#Intermittent_reinforcement_schedules}{Intermittent or partial reinforcement}: Partial or intermittent negative reinforcement can create an effective \href{https://en.wikipedia.org/wiki/Climate_of_fear}{climate of fear} and doubt.
    
    Partial or intermittent positive reinforcement can encourage the victim to persist.
    \item \href{https://en.wikipedia.org/wiki/Punishment_(psychology)}{Punishment}: includes \href{https://en.wikipedia.org/wiki/Nagging}{nagging}, yelling, the \href{https://en.wikipedia.org/wiki/Silent_treatment}{silent treatment}, \href{https://en.wikipedia.org/wiki/Intimidation}{intimidation}, threats, swearing, \href{https://en.wikipedia.org/wiki/Emotional_blackmail}{emotional blackmail}, \href{https://en.wikipedia.org/wiki/Guilt_trip}{guilt trips}, sulking, crying, and \href{https://en.wikipedia.org/wiki/Playing_the_victim}{playing the victim}.
    \item Traumatic 1-trial learning: using verbal abuse, explosive anger, or other intimidating behavior to establish dominance or superiority; even 1 incident of such behavior can \href{https://en.wikipedia.org/wiki/Classical_conditioning}{condition} or train victims to avoid upsetting, confronting or contradicting the manipulator.
\end{itemize}

\subsection{According to Simon}
Simon identified the following manipulative techniques:[Simon, George K (1996). \textit{In Sheep's Clothing: Understanding and Dealing with Manipulative People}. ISBN 978-1-935166-30-6.]
\begin{itemize}
    \item \textbf{Lying (by commission)}: It is hard to tell if somebody is lying at the time they do it, although often the truth may be apparent later when it is too late.
    
    1 way to minimize the chances of being lied to is to understand that some personality types (particularly \href{https://en.wikipedia.org/wiki/Psychopaths}{psychopaths}) are experts at lying and cheating, doing it frequently, and often in subtle ways.
    \item \href{https://en.wikipedia.org/wiki/Lie#Lying_by_omission}{\textbf{Lying by omission}}: This is a subtle form of lying by withholding a significant amount of the truth.
    
    This technique is also used in \href{https://en.wikipedia.org/wiki/Propaganda}{propaganda}.
    \item \href{https://en.wikipedia.org/wiki/Denial}{\textbf{Denial}}: Manipulator refuses to admit that they have done something wrong.
    \item \href{https://en.wikipedia.org/wiki/Rationalization_(psychology)}{\textbf{Rationalization}}: An excuse made by the manipulator for inappropriate behavior.
    
    Rationalization is closely related to \href{https://en.wikipedia.org/wiki/Spin_(public_relations)}{spin}.
    \item \href{https://en.wikipedia.org/wiki/Minimisation_(psychology)}{\textbf{Minimization}}: This is a type of denial coupled with rationalization.
    
    The manipulator asserts that their behavior is not as harmful or irresponsible as someone else was suggesting.
    \item \textbf{Selective inattention or selective \href{https://en.wikipedia.org/wiki/Attention}{attention}}: Manipulator refuses to pay attention to anything that may distract from their agenda.
    \item \href{https://en.wikipedia.org/wiki/Distraction}{\textbf{Diversion}}: Manipulator not giving a straight answer to a straight question and instead being diversionary, steering the conversation onto another topic.
    \item \href{https://en.wikipedia.org/wiki/Evasion_(ethics)}{\textbf{Evasion}}: Similar to diversion but giving irrelevant, rambling, or vague responses.
    \item \textbf{Covert \href{https://en.wikipedia.org/wiki/Intimidation}{intimidation}}: Manipulator putting the victim onto the defensive by using veiled (subtle, indirect or implied) threats.
    \item \textbf{Guilt trip}: A special kind of intimidation tactic.
    
    A manipulator suggests to the \href{https://en.wikipedia.org/wiki/Conscientious}{conscientious} victim that they do not care enough, are too selfish or have it too easy.
    
    This can result in the victim feeling bad, keeping them in a \href{https://en.wikipedia.org/wiki/Doubt}{self-doubting}, \href{https://en.wikipedia.org/wiki/Anxious}{anxious} and \href{https://en.wikipedia.org/wiki/Submissive}{submissive} position.
    \item \href{https://en.wikipedia.org/wiki/Shaming}{\textbf{Shaming}}: Manipulator uses \href{https://en.wikipedia.org/wiki/Sarcasm}{sarcasm} and put-downs to increase fear and self-doubt in the victim.
    
    Manipulators use this tactic to make others feel unworthy and therefore defer to them.
    
    Manipulators can make one feel ashamed for even daring to challenge them.
    
    It is an effective way to foster a sense of inadequacy in the victim.
    \item \href{https://en.wikipedia.org/wiki/Victim_blaming}{\textbf{Vilifying the victim}}: This tactic is a powerful means of putting the victim on the defensive while simultaneously masking the aggressive intent of the manipulator, while the manipulator falsely accuses the victim as being an abuser in response when the victim stands up for or defends themselves or their position.
    \item \textbf{\href{https://en.wikipedia.org/wiki/Victim_playing}{Playing the victim} role}: Manipulator portrays themselves as a victim of circumstance or of someone else's behavior in order to gain \href{https://en.wikipedia.org/wiki/Pity}{pity}, \href{https://en.wikipedia.org/wiki/Sympathy}{sympathy} or evoke \href{https://en.wikipedia.org/wiki/Compassion}{compassion} and thereby get something from another.
    
    Caring and conscientious people often cannot stand to see anyone suffering and the manipulator often finds it easy to play on sympathy to get cooperation.
    \item \textbf{Playing the servant role}: Cloaking a self-serving agenda in the guise of a service to a more noble cause.
    \item \href{https://en.wikipedia.org/wiki/Seduction}{\textbf{Seduction}}: Manipulator uses \href{https://en.wikipedia.org/wiki/Superficial_charm}{charm}, praise, flattery or overtly supporting others in order to get them to lower their defenses and give their \href{https://en.wikipedia.org/wiki/Trust_(social_sciences)}{trust} and \href{https://en.wikipedia.org/wiki/Loyalty}{loyalty} to the manipulator.
    
    They will also offer help with the intent to gain trust and access to an unsuspecting victim they have charmed.
    \item \textbf{\href{https://en.wikipedia.org/wiki/Psychological_projection}{Projecting} the \href{https://en.wikipedia.org/wiki/Blame}{blame} (blaming others)}: Manipulator \href{https://en.wikipedia.org/wiki/Scapegoats}{scapegoats} in often subtle, hard-to-detect ways.
    
    Often, the manipulator will project their own thinking onto the victim, making the victim look like they have done something wrong.
    
    Manipulators will also claim that the victim is the one who is at fault for believing lies that they were conned into believing, as if the victim forced the manipulator to be deceitful.
    
    All blame, except for the part that is used by the manipulator to accept false guilt, is done in order to make the victim feel guilty about making healthy choices, correct thinking and good behaviors.
    
    It is frequently used as a means of psychological and emotional manipulation and control. Manipulators lie about lying, only to re-manipulate the original, less believable story into a ``more acceptable'' truth that the victim will believe.
    
    Projecting lies as being the truth is another common method of control and manipulation.
    
    Manipulators may falsely accuse the victim of ``deserving to be treated that way''.
    
    They often claim that the victim is crazy or abusive, especially when there is evidence against the manipulator.
    \item \textbf{Feigning \href{https://en.wikipedia.org/wiki/Innocence}{innocence}}: Manipulator tries to suggest that any harm done was unintentional or that they did not do something that they were accused of.
    
    Manipulator may put on a look of surprise or indignation.
    
    This tactic makes the victim question their own \href{https://en.wikipedia.org/wiki/Judgment}{judgment} and possibly their own sanity.
    \item \textbf{Feigning \href{https://en.wikipedia.org/wiki/Confusion}{confusion}}: Manipulator tries to play dumb by pretending they do not know what the victim is talking about or is confused about an important issue brought to their attention.
    
    The manipulator intentionally confuses the victim in order for the victim to doubt their own accuracy of perception, often pointing out key elements that the manipulator intentionally included in case there is room for doubt.
    
    Sometimes manipulators will have used cohorts in advance to help back up their story.
    \item \textbf{Brandishing anger}: Manipulator uses anger to brandish sufficient emotional intensity and \href{https://en.wikipedia.org/wiki/Rage_(emotion)}{rage} to shock the victim into submission.
    
    The manipulator is not actually angry, they just put on an act.
    
    They just want what they want and get ``angry'' when denied.
    
    Controlled anger is often used as a manipulation tactic to avoid confrontation, avoid telling the truth or to further hide intent.
    
    There are often threats used by the manipulator of going to the police, or falsely reporting abuses that the manipulator intentionally contrived to scare or intimidate the victim into submission.
    
    Blackmail and other threats of exposure are other forms of controlled anger and manipulation, especially when the victim refuses initial requests or suggestions by the manipulator.
    
    Anger is also used as a defense so the manipulator can avoid telling truths at inconvenient times or circumstances.
    
    Anger is often used as a tool or defense to ward off inquiries or suspicion.
    
    The victim becomes more focused on the anger instead of the manipulation tactic.
    \item \textbf{Bandwagon effect}: Manipulator comforts the victim into submission by claiming (whether true or false) that many people already have done something, and the victim should as well.
    
    Such manipulation can be seen in \href{https://en.wikipedia.org/wiki/Peer_pressure}{peer pressure} situations, often occurring in scenarios where the manipulator attempts to influence the victim into trying drugs or other substances.
\end{itemize}

\subsection{Vulnerabilities exploited by manipulators}
According to Braiker's self-help book,[Braiker, Harriet B. (2004). \textit{Who's Pulling Your Strings? How to Break The Cycle of Manipulation}. ISBN 978-0-07-144672-3.] manipulators exploit the following vulnerabilities (buttons) that may exist in victims:
\begin{itemize}
    \item the desire to please
    \item addiction to earning the approval and acceptance of others
    \item emotophobia (fear of negative emotion; i.e. a fear of expressing anger, frustration or disapproval)
    \item lack of \href{https://en.wikipedia.org/wiki/Assertiveness}{assertiveness} and ability to say no
    \item blurry sense of identity (with soft \href{https://en.wikipedia.org/wiki/Personal_boundaries}{personal boundaries})
    \item low \href{https://en.wikipedia.org/wiki/Self-sufficiency}{self-reliance}
    \item external \href{https://en.wikipedia.org/wiki/Locus_of_control}{locus of control}
\end{itemize}
According to Simon,[Simon, George K (1996). \textit{In Sheep's Clothing: Understanding and Dealing with Manipulative People}. ISBN 978-1-935166-30-6.] manipulators exploit the following vulnerabilities that may exist in victims: 
\begin{itemize}
    \item \textbf{Naïveté} - victim finds it too hard to accept the idea that some people are cunning, devious and ruthless or is ``in \href{https://en.wikipedia.org/wiki/Denial}{denial}'' if they are being victimized.
    \item \textbf{Over-\href{https://en.wikipedia.org/wiki/Conscientiousness}{conscientiousness}} - victim is too willing to give manipulator the benefit of the doubt and see their side of things in which they blame the victim.
    \item \textbf{Low self-\href{https://en.wikipedia.org/wiki/Confidence}{confidence}} - victim is \href{https://en.wikipedia.org/wiki/Doubt}{self-doubting}, lacking in confidence and \href{https://en.wikipedia.org/wiki/Assertiveness}{assertiveness}, likely to go on the defensive too easily.
    \item \textbf{Over-\href{https://en.wikipedia.org/wiki/Intellectualization}{intellectualization}} - victim tries too hard to understand and believes the manipulator has some understandable reason to be hurtful.
    \item \textbf{Emotional \href{https://en.wikipedia.org/wiki/Codependent}{dependency}} - victim has a \href{https://en.wikipedia.org/wiki/Submissive}{submissive} or dependent personality.
    
    The more emotionally dependent the victim is, the more vulnerable they are to being exploited and manipulated.
\end{itemize}
Manipulators generally take the time to scope out the characteristics and vulnerabilities of their victims.

%
Kantor advises in his book \textit{The Psychopathology of Everyday Life: How Antisocial Personality Disorder Affects All of Us}[Kantor, Martin (2006). \textit{The Psychopathology of Everyday Life: How Antisocial Personality Disorder Affects All of Us.} ISBN 978-0-275-98798-5.] that vulnerability to \href{https://en.wikipedia.org/wiki/Psychopathic}{psychopathic} manipulators involves being too: 
\begin{itemize}
    \item \href{https://en.wikipedia.org/wiki/Dependent_personality_disorder}{\textbf{Dependent}} - dependent people need to be loved and are therefore gullible and liable to say yes to something to which they should say no.
    \item \href{https://en.wikipedia.org/wiki/Maturity_(psychological)}{\textbf{Immature}} - has impaired judgment and so tends to believe exaggerated advertising claims.
    \item \href{https://en.wikipedia.org/wiki/Na%C3%AFve}{\textbf{Naïve}} - cannot believe there are dishonest people in the world, or takes it for granted that if there are any, they will not be allowed to prey on others.
    \item \textbf{Impressionable} - overly seduced by \href{https://en.wikipedia.org/wiki/Superficial_charm}{charmers}.
    \item \href{https://en.wikipedia.org/wiki/Trusting}{\textbf{Trusting}} - people who are honest often assume that everyone else is honest.
    
    They are more likely to commit themselves to people they hardly know without checking credentials, etc., and less likely to question so-called experts.
    \item \href{https://en.wikipedia.org/wiki/Carelessness}{\textbf{Carelessness}} - not giving sufficient amount of thought or attention on harm or errors.
    \item \href{https://en.wikipedia.org/wiki/Loneliness}{\textbf{Lonely}} - lonely people may accept any offer of human contact.
    
    A psychopathic stranger may offer human companionship for a price.
    \item \href{https://en.wikipedia.org/wiki/Narcissistic}{\textbf{Narcissistic}} - narcissists are prone to falling for unmerited flattery.
    \item \href{https://en.wikipedia.org/wiki/Impulsivity}{\textbf{Impulsive}} - make snap decisions.
    \item \href{https://en.wikipedia.org/wiki/Altruistic}{Altruistic} - the opposite of psychopathic: too honest, too fair, too empathetic.
    \item \href{https://en.wikipedia.org/wiki/Altruistic}{\textbf{Altruistic}} - the opposite of psychopathic: too honest, too fair, too empathetic.
    \item \href{https://en.wikipedia.org/wiki/Frugal}{\textbf{Frugal}} - cannot say no to a bargain even if they know the reason it is so cheap.
    \item \href{https://en.wikipedia.org/wiki/Materialistic}{\textbf{Materialistic}} - easy prey for loan sharks or get-rich-quick schemes.
    \item \href{https://en.wikipedia.org/wiki/Greed}{\textbf{Greedy}} - the greedy and dishonest may fall prey to a psychopath who can easily entice them to act in an immoral way.
    \item \href{https://en.wikipedia.org/wiki/Masochistic_personality_disorder}{\textbf{Masochistic}} - lack self-respect and so unconsciously let psychopaths take advantage of them.
    
    They think they deserve it out of a sense of guilt.
    \item \textbf{The \href{https://en.wikipedia.org/wiki/Elderly}{elderly}} - the elderly can become fatigued and less capable of multi-tasking.
    
    When hearing a sales pitch they are less likely to consider that it could be a con.
    
    They are prone to giving money to someone with a hard-luck story.
    
    See \href{https://en.wikipedia.org/wiki/Elder_abuse}{elder abuse}.
\end{itemize}

\section{Motivations of manipulators}
Manipulators can have various possible motivations, including but not limited to:[Braiker, Harriet B. (2004). \textit{Who's Pulling Your Strings? How to Break The Cycle of Manipulation}. ISBN 978-0-07-144672-3.]
\begin{itemize}
    \item the need to advance their own purposes and personal gain at virtually any cost to others
    \item a strong need to attain feelings of \href{https://en.wikipedia.org/wiki/Abusive_power_and_control}{power} and superiority in relationships with others
    \item a want and need to feel in \href{https://en.wikipedia.org/wiki/Abusive_power_and_control}{control}
    \item a desire to gain a feeling of power over others in order to raise their perception of \href{https://en.wikipedia.org/wiki/Self-esteem}{self-esteem}
    \item boredom, or growing tired of their surroundings, seeing it as a game more than hurting others
    \item covert agenda, criminal or otherwise, including financial manipulation (often seen when the elderly or unsuspecting, unprotected wealthy are intentionally targeted for the sole purpose of obtaining a victim's financial assets)
    \item not identifying with underlying emotions, commitment phobia, and subsequent rationalization (offender does not manipulate consciously, but rather tries to convince themselves of the invalidity of their own emotions)
    \item lack of self control over impulsive and anti social behavior thus pre-emptive or reactionary manipulation to maintain image
\end{itemize}

\section{Psychopathy}
Main article: \href{https://en.wikipedia.org/wiki/Psychopathy}{Psychopathy}.

%
Being manipulative appears in Factor 1 of the \href{https://en.wikipedia.org/wiki/Hare_Psychopathy_Checklist}{Hare Psychopathy Checklist} (PCL).[Skeem, J. L.; Polaschek, D. L. L.; Patrick, C. J.; Lilienfeld, S. O. (2011). "Psychopathic Personality: Bridging the Gap Between Scientific Evidence and Public Policy". Psychological Science in the Public Interest. 12 (3): 95--162. doi:10.1177/1529100611426706. PMID 26167886. S2CID 8521465.]

\subsection{In the workplace}
Main article: \href{https://en.wikipedia.org/wiki/Psychopathy_in_the_workplace}{Psychopathy in the workplace}.

%
1 approach to \href{https://en.wikipedia.org/wiki/Management}{management} in general identifies a very fine, almost non-existent dividing line between management and manipulation.[Frank, Prabbal (2007). \textit{People Manipulation: A Positive Approach} (2 ed.). New Delhi: Sterling Publishers Pvt. Ltd (published 2009). pp. 3--7. ISBN 978-81-207-4352-6. Retrieved 2019-11-09.]

%
The workplace psychopath may often rapidly shift between emotions - used to manipulate people or to cause high anxiety.[Faggioni M \& White M. \textit{Organizational Psychopaths - Who Are They and How to Protect Your Organization from Them} (2009)]

%
The authors of the book \href{https://en.wikipedia.org/wiki/Snakes_in_Suits:_When_Psychopaths_Go_to_Work}{Snakes in Suits: When Psychopaths Go to Work} describe a 5-phase model of how a typical workplace psychopath climbs to and maintains \href{https://en.wikipedia.org/wiki/Power_(social_and_political)}{power}.

In phase 3 (manipulation) the psychopath will create a scenario of \href{https://en.wikipedia.org/wiki/Cognitive_distortion}{``psychopathic fiction''} where positive information about themselves and negative \href{https://en.wikipedia.org/wiki/Disinformation}{disinformation} about others will be created, where one's role as a part of a network of \href{https://en.wikipedia.org/wiki/Flying_monkeys_(psychology)}{pawns or patrons} will be utilized and one will be \href{https://en.wikipedia.org/wiki/Operant_conditioning}{groomed} into accepting the psychopath's agenda.[Baibak, P; Hare, R. D. \textit{Snakes in Suits: When Psychopaths Go to Work} (2007).]

\section{Antisocial, borderline \& narcissistic personality disorders}
According to \href{https://en.wikipedia.org/wiki/Otto_Kernberg}{Kernberg}, \href{https://en.wikipedia.org/wiki/Antisocial_Personality_Disorder}{antisocial}, \href{https://en.wikipedia.org/wiki/Borderline_Personality_Disorder}{borderline}, and \href{https://en.wikipedia.org/wiki/Narcissistic_Personality_Disorder}{narcissistic} personality disorders are all organized at a borderline level of personality organization,[Kernberg, O (1975). \textit{Borderline Conditions and Pathological Narcissism}. New York: Jason Aronson. ISBN 978-0-87668-205-0.] and the 3 share some common characterological deficits and overlapping personality traits, with deceitfulness and exceptional manipulative abilities being the most common traits among antisocial and narcissism.

Borderline is emphasized by unintentional and dysfunctional manipulation, but stigma towards borderlines being deceitful still wrongfully persists.[Aguirre, Blaise (2016). ``Borderline Personality Disorder: From Stigma to Compassionate Care''. Stigma and Prejudice. Current Clinical Psychiatry. Humana Press, Cham. pp. 133--143. doi:10.1007/978-3-319-27580-2\_8. ISBN 9783319275789.]

Antisocials, borderlines, and narcissists are often \href{https://en.wikipedia.org/wiki/Pathological_lying}{pathological liars}.[Kernberg, O (1975). \textit{Borderline Conditions and Pathological Narcissism}. New York: Jason Aronson. ISBN 978-0-87668-205-0.]

Other shared traits may include pathological narcissism,[9] consistent irresponsibility, \href{https://en.wikipedia.org/wiki/Machiavellianism_(psychology)}{Machiavellianism}, lack of \href{https://en.wikipedia.org/wiki/Empathy}{empathy},[Baron-Cohen, S (2012). \textit{The Science of Evil: On Empathy and the Origins of Cruelty}. Basic Books. pp. 45--98. ISBN 978-0-465-03142-9.] cruelty, \href{https://en.wikipedia.org/wiki/Meanness}{meanness}, impulsivity, proneness to self-harm and addictions,[Casillas, A.; Clark, L.A.k (October 2002). ``Dependency, impulsivity, and self-harm: traits hypothesized to underlie the association between cluster B personality and substance use disorders''. Journal of Personality Disorders. 16 (5): 424--36. doi:10.1521/pedi.16.5.424.22124. PMID 12489309.] interpersonal exploitation, hostility, anger and rage, vanity, emotional instability, rejection sensitivity, perfectionism, and the use of primitive \href{https://en.wikipedia.org/wiki/Defence_mechanisms}{defence mechanisms} that are pathological and narcissistic.

Common \href{https://en.wikipedia.org/wiki/Narcissistic_defences}{narcissistic defences} include \href{https://en.wikipedia.org/wiki/Splitting_(psychology)}{splitting}, \href{https://en.wikipedia.org/wiki/Denial}{denial}, \href{https://en.wikipedia.org/wiki/Psychological_projection}{projection}, \href{https://en.wikipedia.org/wiki/Projective_identification}{projective identification}, \href{https://en.wikipedia.org/wiki/Idealization_and_devaluation}{primitive idealization and devaluation}, \href{https://en.wikipedia.org/wiki/Cognitive_distortion}{distortion} (including \href{https://en.wikipedia.org/wiki/Exaggeration}{exaggeration}, \href{https://en.wikipedia.org/wiki/Minimisation_(psychology)}{minimization} and \href{https://en.wikipedia.org/wiki/Lie}{lies}), and \href{https://en.wikipedia.org/wiki/Omnipotence}{omnipotence}.[13]

%
Psychologist \href{https://en.wikipedia.org/wiki/Marsha_M._Linehan}{Marsha M. Linehan} has stated that people with borderline personality disorder often exhibit behaviors which are not truly manipulative, but are erroneously interpreted as such.[``On Manipulation with the Borderline Personality''. ToddlerTime Network. Retrieved Dec 28, 2014.]

According to her, these behaviors often appear as unthinking manifestations of intense pain, and are often not deliberate as to be considered truly manipulative.

In the \href{https://en.wikipedia.org/wiki/DSM-V}{DSM-V}, manipulation was removed as a defining characteristic of borderline personality disorder.[Aguirre, Blaise (2016). ``\textit{Borderline Personality Disorder: From Stigma to Compassionate Care}''. Stigma and Prejudice. Current Clinical Psychiatry. Humana Press, Cham. pp. 133--143. doi:10.1007/978-3-319-27580-2\_8. ISBN 9783319275789.]

%
Manipulative behavior is intrinsic to narcissists, who use manipulation to obtain power and \href{https://en.wikipedia.org/wiki/Narcissistic_supply}{narcissistic supply}.

Those with antisocial personalities will manipulate for material items, power, revenge, and a wide variety of other reasons.[\href{https://en.wikipedia.org/wiki/Psychological_manipulation#CITEREFAmerican_Psychiatric_Association2000}{American Psychiatric Association 2000}]

\section{Histrionic personality disorder}
Main article: \href{https://en.wikipedia.org/wiki/Histrionic_personality_disorder}{Histrionic personality disorder}.

%
People with histrionic personality disorder are usually high-functioning, both socially and professionally.

They usually have good \href{https://en.wikipedia.org/wiki/Social_skills}{social skills}, despite tending to use them to manipulate others into making them the center of attention.[``Histrionic Personality Disorder''. The Cleveland Clinic. Archived from the original on 3 October 2011. Retrieved 23 November 2011.]

\section{Machiavellianism}
\href{https://en.wikipedia.org/wiki/Machiavellianism_(psychology)}{Machiavellianism} is a term that some \href{https://en.wikipedia.org/wiki/Social_psychology}{social} and \href{https://en.wikipedia.org/wiki/Personality_psychology}{personality} psychologists use to describe a person's tendency to be unemotional, uninfluenced by conventional morality and more prone to \href{https://en.wikipedia.org/wiki/Deceive}{deceive} and manipulate others.

In the 1960s, Richard Christie and Florence L. Geis developed a test for measuring a person's level of Machiavellianism (sometimes referred to as the \textit{Machiavelli test}).[Christie, R., and F. L. Geis. (1970) ``How devious are you? Take the Machiavelli test to find out.'' Journal of Management in Engineering 15.4: 17.] 

\section{See also}
\begin{itemize}
    \item Advertising
    \item Appeal to emotion
    \item Blackmail
    \item Brainwashing
    \item Bullying
    \item Culture of fear
    \item Coercion
    \item Coercive persuasion
    \item Confidence trick
    \item Crowd manipulation
    \item Covert hypnosis
    \item Covert interrogation
    \item Dark triad
    \item Deception
    \item Demagogy
    \item Discrediting tactic
    \item DISC assessment
    \item Dumbing down
    \item Fear mongering
    \item Gaslighting
    \item Half-truth
    \item Internet manipulation
    \item Isolation to facilitate abuse
    \item List of confidence tricks
    \item List of fallacies
    \item Lying
    \item Master suppression techniques
    \item Media manipulation
    \item Mind control
    \item Mobbing
    \item Psychological abuse
    \item Psychological warfare
    \item Sheeple
    \item Social engineering (political science)
    \item Social engineering (security)
    \item Social influence
    \item Whispering campaign
    \item Project MKUltra
\end{itemize}

\section{Further reading}

\subsection{Books}
\begin{itemize}
    \item Admin, AB. \textit{30 Covert Emotional Manipulation Tactics: How Manipulators Take Control In Personal Relationships} (2014).
    \item Alessandra, Tony. \textit{Non-Manipulative Selling} (1992).
    \item Barber, Brian K. \textit{Intrusive Parenting: How Psychological Control Affects Children and Adolescents} (2001).
    \item Bowman, Robert P.; Cooper, Kathy; Miles, Ron; \& Carr, Tom. \textit{Innovative Strategies for Unlocking Difficult Children: Attention Seekers, Manipulative Students, Apathetic Students, Hostile Students} (1998).
    \item Bursten, Ben. \textit{Manipulator: A Psychoanalytic View} (1973).
    \item Crawford, Craig. \textit{The Politics of Life: 25 Rules for Survival in a Brutal and Manipulative World} (2007).
    \item Forward, Susan. \textit{Emotional Blackmail} (1997).
    \item Klatte, Bill \& Thompson, Kate. \textit{It's So Hard to Love You: Staying Sane When Your Loved One Is Manipulative, Needy, Dishonest, or Addicted} (2007).
    \item McCoy, Dorothy. \textit{The Manipulative Man: Identify His Behavior, Counter the Abuse, Regain Control} (2006).
    \item McMillan, Dina L. \textit{But He Says He Loves Me: How to Avoid Being Trapped in a Manipulative Relationship} (2008).
    \item Sasson, Janet Edgette. \textit{Stop Negotiating With Your Teen: Strategies for Parenting Your Angry, Manipulative, Moody, or Depressed Adolescent} (2002).
    \item Stern, Robin. \textit{The Gaslight Effect: How to Spot and Survive the Hidden Manipulation Others Use to Control Your Life} (2008).
    \item Swihart, Ernest W. Jr. \& Cotter, Patrick. \textit{The Manipulative Child: How to Regain Control and Raise Resilient, Resourceful, and Independent Kids} (1998).
\end{itemize}

\subsection{Academic journals}
\begin{itemize}
    \item Aglietta M, Reberioux A, Babiak P. ``Psychopathic manipulation in organizations: pawns, patrons and patsies'', in Cooke A, Forth A, Newman J, Hare R (Eds), \textit{International Perspectives and Psychopathy}, British Psychological Society, Leicester, pp. 12--17. (1996).
    \item Aglietta, M.; Reberioux, A.; Babiak, P. ``Psychopathic manipulation at work'', in Gacono, C.B. (Ed), \textit{The Clinical and Forensic Assessment of Psychopathy: A Practitioner's Guide}, Erlbaum, Mahwah, NJ, pp. 287--311. (2000).
    \item Bursten, Ben. ``The Manipulative Personality'', \textit{Archives of General Psychiatry}, Vol 26 No 4, 318--321 (1972).
    \item Buss DM, Gomes M, Higgins DS, Lauterback K. ``Tactics of Manipulation'', \textit{Journal of Personality and Social Psychology}, Vol 52 No 6 1219--1279 (1987).
    \item Fischer, Alexander/Christian Illies. ``Modulated Feelings: The Pleasurable-Ends-Model of Manipulation'', \textit{Philosophical Inquiries IV/2}, 25-44 (2018).
    \item Hofer, Paul. ``The Role of Manipulation in the Antisocial Personality'', \textit{International Journal of Offender Therapy and Comparative Criminology}, Vol. 33 No 2, 91--101 (1989).
\end{itemize}

\section{External links}
\begin{itemize}
    \item Sarah Strudwick (Nov 16, 2010) \href{https://www.youtube.com/watch?v=PwWBHRKFYCA}{Dark Souls - Mind Games, Manipulation and Gaslighting}\hfill$\square$
\end{itemize}

%------------------------------------------------------------------------------%

\chapter{\href{https://en.wikipedia.org/wiki/Psychological_projection}{Wikipedia/Psychological Projection}}

\textit{Psychological projection} is a \href{https://en.wikipedia.org/wiki/Defense_mechanism}{defense mechanism} in which the \href{https://en.wikipedia.org/wiki/Ego_(Freudian)}{ego} defends itself against unconscious impulses or qualities (both positive and negative) by \href{https://en.wikipedia.org/wiki/Denial}{denying} their existence in themselves by \href{https://en.wikipedia.org/wiki/Attribution_(psychology)}{attributing} them to others.[Sigmund Freud, \textit{Case Histories II} (PFL 9) p. 132]

E.g., a bully may project their own feelings of vulnerability onto the target, or a person who is confused will project their own feelings of confusion and inadequacy on other people.

%
Projection incorporates \href{https://en.wikipedia.org/wiki/Blame_shifting}{blame shifting} and can manifest as \textbf{shame dumping}.[Hotchkiss, Sandy; foreword by Masterson, James F. \textit{Why Is It Always About You?: The Seven Deadly Sins of Narcissism} (Free Press, 2003)]

%
Projection has been described as an early phase of \href{https://en.wikipedia.org/wiki/Introjection}{introjection}.[Malancharuvil JM (December 2004). ``Projection, introjection, and projective identification: a reformulation'' (PDF). Am J Psychoanal. 64 (4): 375--82. doi:10.1007/s11231-004-4325-y. PMID 15577283. S2CID 19730486.]

\section{Historical precursors}
A prominent precursor in the formulation of the projection principle was \href{https://en.wikipedia.org/wiki/Giambattista_Vico}{Giambattista Vico}.[4][5]

In 1841, \href{https://en.wikipedia.org/wiki/Ludwig_Feuerbach}{Ludwig Feuerbach} was the 1st \href{https://en.wikipedia.org/wiki/Age_of_Enlightenment}{enlightenment} thinker to employ this concept as the basis for a systematic critique of religion.[6][7][8]

%
The \href{https://en.wikipedia.org/wiki/Babylonian_Talmud}{Babylonian Talmud} (500 AD) notes the human tendency toward projection and warns against it: ``Do not taunt your neighbor with the blemish you yourself have.''[9]

Religious people of the Christian faith believe that in the \href{https://en.wikipedia.org/wiki/New_Testament}{New Testament}, Jesus also warned against projection:

\begin{quotation}
    ``\textit{Why do you look at the speck of sawdust in your brother's eye and pay no attention to the plank in your own eye? How can you say to your brother, `Let me take the speck out of your eye,' when all the time there is a plank in your own eye? You hypocrite, first take the plank out of your own eye, and then you will see clearly to remove the speck from your brother's eye}.''[10]
\end{quotation}

\section{Psychoanalytic developments}
Projection (\href{https://en.wikipedia.org/wiki/German_language}{German}: \textit{Projektion}) was conceptualized by \href{https://en.wikipedia.org/wiki/Sigmund_Freud}{Sigmund Freud} in his letters to \href{https://en.wikipedia.org/wiki/Wilhelm_Fliess}{Wilhelm Fliess},[11] and further refined by \href{https://en.wikipedia.org/wiki/Karl_Abraham}{Karl Abraham} and \href{https://en.wikipedia.org/wiki/Anna_Freud}{Anna Freud}.

Freud considered that, in projection, thoughts, motivations, desires, and feelings that cannot be accepted as one's own are dealt with by being placed in the outside world and attributed to someone else.[Case Studies II p. 210.]

What the ego repudiates is \href{https://en.wikipedia.org/wiki/Splitting_(psychology)}{split off} and placed in another.[Otto Fenichel, \textit{The Psychoanalytic Theory of Neurosis} (London 1946) p. 146.]

%
Freud would later come to believe that projection did not take place arbitrarily, but rather seized on and \href{https://en.wikipedia.org/wiki/Exaggerated}{exaggerated} an element that already existed on a small scale in the other person.[Sigmund Freud, On Psychopathology (PFL 10) pp. 200--01.]

(The related defence of \href{https://en.wikipedia.org/wiki/Projective_identification}{projective identification} differs from projection in that the other person is expected to become identified with the \href{https://en.wikipedia.org/wiki/Impulse_(psychology)}{impulse} or desire projected outside,[Patrick Casement, \textit{Further Learning from the Patient} (1997) p. 177.] so that the self maintains a connection with what is projected, in contrast to the total repudiation of projection proper.)[Otto F. Kernberg, \textit{Borderline Conditions and Pathological Narcissism} (London 1990) p. 56.]

%
\href{https://en.wikipedia.org/wiki/Melanie_Klein}{Melanie Klein} saw the projection of good parts of the self as leading potentially to over-\href{https://en.wikipedia.org/wiki/Idealisation_and_devaluation}{idealization} of the object.[17]

Equally, it may be one's conscience that is projected, in an attempt to escape its control: a more benign version of this allows one to come to terms with outside authority.[R. Wollheim, On the Emotions (1999) pp. 217--18.]

\section{Theoretical examples}
Projection tends to come to the fore in normal people at times of personal or political \href{https://en.wikipedia.org/wiki/Crisis}{crisis}[Erik Erikson, \textit{Childhood and Society} (1973) p. 241.] but is more commonly found in \href{https://en.wikipedia.org/wiki/Narcissistic_personality_disorder}{narcissistic personality disorder} or \href{https://en.wikipedia.org/wiki/Borderline_personality_disorder}{borderline personality disorder}.[Glen O. Gabbard, \textit{Long-Term Psychodynamic Psychotherapy} (London 2010) p. 33.]

%
\href{https://en.wikipedia.org/wiki/Carl_Jung}{Carl Jung} considered that the unacceptable parts of the personality represented by the \href{https://en.wikipedia.org/wiki/Shadow_(psychology)}{Shadow} archetype were particularly likely to give rise to projection, both small-scale and on a national/international basis.[Carl G. Jung ed., \textit{Man and his Symbols} (London 1978) pp. 181--82.]

\href{https://en.wikipedia.org/wiki/Marie-Louise_Von_Franz}{Marie-Louise Von Franz} extended her view of projection, stating that ``wherever known reality stops, where we touch the unknown, there we project an \href{https://en.wikipedia.org/wiki/Archetype}{archetypal} image''.[22]

%
Psychological projection is 1 of the \href{https://en.wikipedia.org/wiki/Medical_explanations_of_bewitchment}{medical explanations of bewitchment} used to explain the behavior of the afflicted children at \href{https://en.wikipedia.org/wiki/Salem_witch_trials}{Salem} in 1692.

The historian \href{https://en.wikipedia.org/wiki/John_Demos}{John Demos} wrote in 1970 that the symptoms of bewitchment displayed by the afflicted girls could have been due to the girls undergoing psychological projection of \href{https://en.wikipedia.org/wiki/Repression_(psychology)}{repressed} aggression.[Demos, John (1970). ``Underlying Themes in the Witchcraft of Seventeenth-Century New England''. American Historical Review. 75 (5): 1311--26 [p. 1322]. doi:10.2307/1844480. JSTOR 1844480. PMID 11609522.]

\section{Practical examples}
\begin{itemize}
    \item \href{https://en.wikipedia.org/wiki/Victim_blaming}{\textbf{Victim blaming}}: The victim of someone else's actions or bad luck may be offered criticism, the theory being that the victim may be at fault for having attracted the other person's hostility.
    
    In such cases, the psyche projects the experiences of weakness or vulnerability with the aim of ridding itself of the feelings and, through its disdain for them or the act of blaming, their conflict with the ego.[\textit{The Pursuit of Health}, June Bingham \& Norman Tamarkin, M.D., Walker Press. ]
    \item \textbf{Projection of marital guilt}: Thoughts of \href{https://en.wikipedia.org/wiki/Infidelity}{infidelity} to a partner may be \href{https://en.wikipedia.org/wiki/Unconscious_mind}{unconsciously} projected in self-defence on to the partner in question, so that the \href{https://en.wikipedia.org/wiki/Guilt_(emotion)}{guilt} attached to the thoughts can be repudiated or turned to \href{https://en.wikipedia.org/wiki/Blame}{blame} instead, in a process linked to \href{https://en.wikipedia.org/wiki/Denial}{denial}.[25]
    
    E.g., a person who is having a sexual affair may fear that their spouse is planning an affair or may accuse the innocent spouse of \href{https://en.wikipedia.org/wiki/Adultery}{adultery}.
    \item \href{https://en.wikipedia.org/wiki/Bullying}{\textbf{Bullying}}: A bully may project their own feelings of \href{https://en.wikipedia.org/wiki/Vulnerability}{vulnerability} onto the target(s) of the bullying activity.
    
    Despite the fact that a bully's typically denigrating activities are aimed at the bully's targets, the true source of such negativity is ultimately almost always found in the bully's own sense of personal \href{https://en.wikipedia.org/wiki/Emotional_security}{insecurity} or vulnerability.[Paul Gilbert, \textit{Overcoming Depression} (1999) pp. 185--86.]
    
    Such aggressive projections of displaced negative emotions can occur anywhere from the micro-level of \href{https://en.wikipedia.org/wiki/Interpersonal_relationships}{interpersonal relationships}, all the way up through to the macro-level of international politics, or even international armed conflict.[Carl G. Jung ed., \textit{Man and his Symbols} (London 1978) pp. 181--82.]
    \item \textbf{Projection of general guilt}: Projection of a severe conscience[Patrick Casement, \textit{Further Learning from the Patient} (1990) p. 142.] is another form of defense, one which may be linked to the making of \href{https://en.wikipedia.org/wiki/False_accusations}{false accusations}, personal or political.[Carl G. Jung ed., \textit{Man and his Symbols} (London 1978) pp. 181--82.]
    \item \textbf{Projection of hope}: Also, in a more positive light, a patient may sometimes project their feelings of \href{https://en.wikipedia.org/wiki/Hope}{hope} onto the therapist.[Patrick Casement, \textit{Further Learning from the Patient} (1990) p. 122.]
\end{itemize}

\section{Counter-projection}
Jung wrote, ``All projections provoke counter-projection when the object is unconscious of the quality projected upon it by the subject.''[\textit{General Aspects of Dream Psychology}, CW 8, par. 519.]

Thus, what is unconscious in the recipient will be projected back onto the projector, precipitating a form of mutual \href{https://en.wikipedia.org/wiki/Acting_out}{acting out}.[30]

%
In a rather different usage, \href{https://en.wikipedia.org/wiki/Harry_Stack_Sullivan}{Harry Stack Sullivan} saw counter-projection in the therapeutic context as a way of warding off the \href{https://en.wikipedia.org/wiki/Compulsive_behavior}{compulsive}\footnote{NQBH: \textit{Why do I always eat my finger nails unconsciously?}} re-enactment of a \href{https://en.wikipedia.org/wiki/Psychological_trauma}{psychological trauma}, by emphasizing the difference between the current situation and the projected \href{https://en.wikipedia.org/wiki/Fixation_(psychology)}{obsession} with the perceived perpetrator of the original trauma.[F. S. Anderson ed., \textit{Bodies in Treatment} (2007) p. 160.]

\section{Clinical approaches}
Drawing on \href{https://en.wikipedia.org/wiki/Gordon_Allport}{Gordon Allport}'s idea of the expression of self onto activities and objects, projective techniques have been devised to aid personality assessment, including the \href{https://en.wikipedia.org/wiki/Rorschach_test}{Rorschach} ink-blots and the \href{https://en.wikipedia.org/wiki/Thematic_Apperception_Test}{Thematic Apperception Test} (TAT).[Semeonoff, B. (1987). ``Projective Techniques''. In Gregory, Richard (ed.). \textit{The Oxford Companion to the Mind}. New York: Oxford University Press. p. 646. ISBN 0-19-866124-X.]

%
Projection may help a fragile \href{https://en.wikipedia.org/wiki/Id,_ego,_and_super-ego}{ego} reduce \href{https://en.wikipedia.org/wiki/Anxiety}{anxiety}, but at the cost of a certain \href{https://en.wikipedia.org/wiki/Dissociation_(psychology)}{dissociation}, as in \href{https://en.wikipedia.org/wiki/Dissociative_identity_disorder}{dissociative identity disorder}.[\href{http://www.medscape.com/viewarticle/408692}{Trauma \& Projection}]

In extreme cases, an individual's personality may end up becoming critically \href{https://en.wikipedia.org/wiki/Ego\_depletion}{depleted}.[34]
    
In such cases, therapy may be required which would include the slow rebuilding of the personality through the ``taking back'' of such projections.[35]

%
The method of managed projection is a type of projective technique.

The basic principle of the method is that a subject is presented with their own verbal portrait named by the name of another person, as well as with a portrait of their fictional opposition (V. V. Stolin, 1981).

%
The technique is suitable for application in psychological counseling and might provide valuable information about the form and nature of their self-esteem Bodalev, A (2000). ``\textit{General psychodiagnostics}''.

\section{Criticism}
Some studies were critical of Freud's theory.

Research on \href{https://en.wikipedia.org/wiki/Social_projection}{social projection} supports the existence of a \href{https://en.wikipedia.org/wiki/False-consensus_effect}{false-consensus effect} whereby humans have a broad tendency to believe that others are similar to themselves, and thus ``project'' their personal traits onto others.[Robbins, Jordan M.; Krueger, Joachim I. (2005). ``Social Projection to Ingroups and Outgroups: A Review and Meta-Analysis''. \textit{Personality and Social Psychology Review}. 9 (1): 32--47. doi:10.1207/s15327957pspr0901\_3. ISSN 1088-8683.]

This applies to good traits as well as bad traits and is not a defense mechanism for denying the existence of the trait within the self.[Baumeister, Roy F.; Dale, Karen; Sommer, Kristin L. (1998). ``Freudian Defense Mechanisms and Empirical Findings in Modern Social Psychology: Reaction Formation, Projection, Displacement, Undoing, Isolation, Sublimation, and Denial''.\textit{Journal of Personality}. 66 (6): 1090--92. doi:10.1111/1467-6494.00043.]

%
Instead, Newman, Duff, and Baumeister (1997) proposed a new model of defensive projection.

In this view, \href{https://en.wikipedia.org/wiki/Psychological_repression}{repressors} try to \href{https://en.wikipedia.org/wiki/Thought_suppression}{suppress thoughts} of their undesirable traits, and these efforts make those trait categories highly accessible - so that they are then used all the more often when forming impressions of others.

The projection is then only a by-product of the real defensive mechanism.[Newman, Leonard S.; Duff, Kimberley J.; Baumeister, Roy F. (1997). ``A new look at defensive projection: Thought suppression, accessibility, and biased person perception''. \textit{Journal of Personality and Social Psychology}. 72 (5): 980–1001. doi:10.1037/0022-3514.72.5.980. PMID 9150580.]

\section{See also}
\begin{itemize}
    \item Animism
    \item Anthropology of religion
    \item Displacement
    \item Double standard
    \item Giambattista Vico
    \item Hypocrisy
    \item Hostile attribution bias
    \item Identified patient
    \item Introjection
    \item Narcissistic abuse
    \item Narcissistic rage and narcissistic injury
    \item Participation mystique
    \item Psychoanalytic theory
    \item Psychodynamics
    \item Rationalization
    \item Reaction formation
    \item Regression
    \item Repression
    \item Scapegoating
    \item Transference
    \item The pot calling the kettle black
    \item \textit{Tu quoque}\hfill$\square$
\end{itemize}

%------------------------------------------------------------------------------%



%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}