\documentclass{book}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/nguyen/1_NQBH/math/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\addtolength{\headheight}{0pt} % obsolete
\lhead{\scriptsize \chaptername~\thechapter}
\rhead{\itshape \scriptsize \nouppercase{\leftmark}} %\nouppercase !
\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\cfoot{\thepage}

\title{\href{https://terrytao.wordpress.com/}{Terence Tao's Blog: What's New}}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
\tableofcontents

%------------------------------------------------------------------------------%

\chapter{Home}

%------------------------------------------------------------------------------%

\chapter{About}

%------------------------------------------------------------------------------%

\chapter{\href{https://terrytao.wordpress.com/career-advice/}{Career Advice}}

\begin{quotation}
    \textit{Advice is what we ask for when we already know the answer but wish we didn't}. - \href{http://en.wikipedia.org/wiki/Erica_Jong}{Erica Jong}
\end{quotation}
Here is Terence Tao's collection of various pieces of advice on academic career issues in mathematics, roughly arranged by the stage of career at which the advice is most pertinent (though of course some of the advice pertains to multiple stages).

\begin{remark}[Disclaimer]
    The advice here is very generic in nature; Terence Tao does not pretend to have any sort of ``silver bullet'' that will solve all career issues.
    
    You will of course need to evaluate many factors, contexts, and needs specific to your own situation, as well as employing a healthy dose of common sense, before making any important career decisions.
    
    Terence Tao would in particular recommend \href{https://terrytao.wordpress.com/career-advice/talk-to-your-advisor/}{discussing such decisions with your advisor} if you have one, as he or she will be familiar with your situation and will likely be able to provide pertinent advice.
    
    Also, it should be clear that most of this advice is targeted towards academic careers in mathematics; of course, there are many other career options available besides this, but Terence Tao has no particularly informed advice to offer for such alternatives.
\end{remark}

\section{\href{https://terrytao.wordpress.com/career-advice/talk-to-your-advisor/}{Talk to your advisor}}
\begin{quotation}
    \textit{It is the province of knowledge to speak and it is the privilege of wisdom to listen}. - \href{https://en.wikipedia.org/wiki/Oliver_Wendell_Holmes_Sr.}{Oliver Wendell Holmes}, ``The Poet at the Breakfast Table''
\end{quotation}
Your advisor is 1 of the best sources of guidance you have; not only in directly assisting you with your \textit{research topic}, but in directing you (both explicitly and implicitly) to \textit{relevant researchers, conferences, publications, open problems, folklore}, or \textit{other pieces of good mathematics}.

Your advisor also knows your situation well and can give career advice which is tailored to your specific strengths and weaknesses (unlike the \href{https://terrytao.wordpress.com/career-advice/}{generic} advice in these pages).

%
If things get to the point that you are actively avoiding your advisor (or vice versa), that is a very bad sign. In particular, you should be aware of your advisor's schedule, and conversely your advisor should be aware of when you will be available in the department, and what you are currently working on.

%
For similar reasons, you should give your advisor some advance warning if you want to take a long period of time away from your studies.

%
If your advisor is unavailable, you should regularly discuss mathematical issues with at least 1 other mathematician instead, preferably an experienced one.

[Also, it is not uncommon for a student to have both a formal advisor, who handles all the official paperwork, and an informal advisor, with which you discuss research and career issues.]

%
Of course, you should not rely \textit{purely} on your advisor; you also need to \href{https://terrytao.wordpress.com/career-advice/take-the-initiative/}{take the initiative} when it comes to your mathematical career.

\section{\href{https://terrytao.wordpress.com/career-advice/take-the-initiative/}{Take the initiative}}
\begin{quotation}
    \textit{The best teacher is the one who suggests rather than dogmatizes, and inspires his listener with the wish to teach himself}. - \href{https://en.wikipedia.org/wiki/Edward_Bulwer-Lytton}{Edward Bulwer-Lytton}
\end{quotation}
While you should talk to your advisor, you should not be completely reliant on him or her; after all, you are going to have to do mathematics primarily on your own once you graduate!

%
If you feel like you want to learn something, do something, or write something, you do not have to clear it with your advisor - just go ahead and do it (though in some cases other priorities, such as writing your thesis, may be temporarily more important, and you should of course keep your advisor updated as to what you are doing mathematically).

Research your library or the internet, talk with other graduate students or faculty, read papers and books on your own (both \href{https://terrytao.wordpress.com/career-advice/learn-and-relearn-your-field/}{in your field} and in \href{https://terrytao.wordpress.com/career-advice/don't-be-afraid-to-learn-things-outside-your-field/}{nearby fields}), \href{https://terrytao.wordpress.com/career-advice/attend-talks-and-conferences-even-those-not-directly-related-to-your-work/}{attend conferences}, and so forth. (See also ``\href{https://terrytao.wordpress.com/career-advice/ask-yourself-dumb-questions-and-answer-them/}{ask yourself dumb questions}''.)

%
1 specific suggestion Terence Tao has is to subscribe (either \href{http://www.arxiv.org/help/rss}{by RSS}, or \href{http://arxiv.org/help/subscribe}{by email}) to be notified of new papers which appear on the \href{http://www.arxiv.org/}{arXiv} in the subject areas that you are interested in.

%
In a somewhat related spirit, while it is certainly acceptable to have mathematical role models, one should not try to mimic them too slavishly; you need to \fbox{\href{https://terrytao.wordpress.com/advice-on-writing-papers/write-in-your-own-voice/}{develop your own personal style}}, \fbox{\textit{exploiting your own strengths}} and \textit{mitigating your own weaknesses}, which will not be identical to those of your role models.

\textit{Ultimately, it is \fbox{better to follow the mathematics than to follow a mathematician}.}

\section{\href{https://terrytao.wordpress.com/career-advice/learn-and-relearn-your-field/}{Learn \& relearn your field}}
\begin{quotation}
    \it
    Even fairly good students, when they have obtained the solution of the problem and written down neatly the argument, shut their books and look for something else.
    
    Doing so, they miss an important and instructive phase of the work. $\ldots$
    
    A good teacher should understand and impress on his students the view that no problem whatever is completely exhausted.
    
    1 of the 1st and foremost duties of the teacher is not to give his students the impression that mathematical problems have little connection with each other, and no connection at all with anything else.
    
    We have a natural opportunity to investigate the connections of a problem when looking back at its solution. - \href{http://en.wikipedia.org/wiki/George_P%C3%B3lya}{\emph{George Pólya}}, ``\href{http://en.wikipedia.org/wiki/How_to_Solve_It}{How to Solve It}''
\end{quotation}
Learning never really stops in this business, even in your chosen specialty; e.g. Terence Tao is still learning surprising things about basic harmonic analysis, more than 10 years after writing his thesis in the topic.

%
Just because you know a statement and proof of Fundamental Lemma X, you should not take that lemma for granted; instead, you should dig deeper until you really understand what the lemma is all about:
\begin{itemize}
    \item \textit{Can you find alternate proofs?}
    \item \textit{If you know 2 proofs of the lemma, do you know to what extent the proofs are equivalent?}
    
    \textit{Do they generalize in different ways?}
    
    \textit{What themes do the proofs have in common?}
    
    \textit{What are the other relative strengths and weaknesses of the 2 proofs?}
    \item \textit{Do you know why each of the hypotheses are necessary?}
    \item \textit{What kind of generalizations are known/conjectured/heuristic?}
    \item \textit{Are there weaker and simpler versions which can suffice for some applications?}
    \item \textit{What are some model examples demonstrating that lemma in action?}
    \item \textit{When is it a good idea to use the lemma, and when isn't it?}
    \item \textit{What kind of problems can it solve, and what kind of problems are beyond its ability to assist with?}
    \item \textit{Are there analogues of that lemma in other areas of mathematics?}
    \item \textit{Does the lemma fit into a wider paradigm or program?}
\end{itemize}
It is particularly useful to lecture on your field, or \href{https://terrytao.wordpress.com/career-advice/write-down-what-youve-done/}{write lecture notes or other expository material}, even if it is just for your own personal use.

You will eventually be able to internalize even very difficult results using efficient mental shorthand; this not only allows you to use these results effortlessly, and \href{https://terrytao.wordpress.com/career-advice/continually-aim-just-beyond-your-current-range/}{improve your own ability} in the field, but also \textit{frees up mental space to learn even more material}.

%
Another useful way to learn more about one's field is to take a \fbox{key paper} in that field, and perform a citation search on that paper (i.e. search for other papers that cite the key paper).

There are many tools for citation searches nowadays; e.g., \href{http://www.ams.org/mathscinet/search.html}{MathSciNet} offers this functionality, and even a general-purpose web search engine can often give useful ``hits'' that one might not have previously been aware of.

\section{\href{https://terrytao.wordpress.com/career-advice/write-down-what-youve-done/}{Write down what you've done}}
\begin{quotation}
    \textit{Every composer knows the anguish and despair occasioned by forgetting ideas which one had no time to write down}. - \href{https://en.wikipedia.org/wiki/Hector_Berlioz}{Hector Berlioz}
\end{quotation}
There were many occasions early in his career when Terence Tao read, heard about, or stumbled upon some neat mathematical trick or argument, and thought he understood it well enough that he didn't need to write it down; and then, say 6 months later, when he actually needed to recall that trick, he couldn't reconstruct it at all.

Eventually he resolved to write down (preferably on a computer) a sketch of any interesting argument he came across - not necessarily at a publication level of quality, but detailed enough that he could then safely forget about the details, and readily recover the argument from the sketch whenever the need arises.

%
Terence Tao recommend that you do this also, as it serves several useful purposes:
\begin{enumerate}
    \item It makes the argument permanently available to you in the future, and may eventually be helpful in your later research papers, lecture notes, teaching, or research proposals.
    \item It gives you practice in mathematical writing, both at the technical level (e.g. in learning how to use TeX) and at an expository or pedagogical level.
    \item It tests whether you have really understood the argument on more than just a superficial level.
    \item It frees up mental space; you no longer have to remember the exact details of the argument, and so can devote your memory to learning newer topics.
\end{enumerate}
Once you have written up such a sketch, you might consider \href{https://terrytao.wordpress.com/career-advice/make-your-work-available/}{making it available} (e.g. on your web site), even if it does not rise to the level of originality and depth required for a publishable paper.

%
For somewhat similar reasons, if you have an incomplete (or otherwise unsatisfactory) argument for a problem that you are working on, and you are \href{https://terrytao.wordpress.com/career-advice/use-the-wastebasket/}{planning to abandon it}, you may still wish to write an informal sketch of it just for yourself (giving barely enough details to allow you to readily reconstruct the whole thing later on), and store it somewhere on your computer, just in case you find you have need for it some time in the future.

(Of course, Terence Tao also recommends some way of backing up your computer files, e.g. by using a cloud-based file storage system such as Dropbox.)

%------------------------------------------------------------------------------%

\chapter{On Writing}

%------------------------------------------------------------------------------%

\chapter{Books}

%------------------------------------------------------------------------------%

\chapter{Applets}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}