\documentclass{article}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\usepackage{soul}
\addbibresource{/home/nguyen/1_NQBH/math/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}

\title{\href{https://www.facebook.com/EducatedMindsPage}{Facebook/Educated Minds}}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\begin{abstract}
    \textbf{mind} [n] a beautiful servant, a dangerous master.
\end{abstract}
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
%\tableofcontents

%------------------------------------------------------------------------------%

\begin{enumerate}
    \item \textsc{Successful adulting}: Not stabbing someone when you really want to.
    \item \textbf{Check yourself}
    
    Sometimes you are the toxic person.
    
    Sometimes you are the mean, negative person you're looking to push away.
    
    Sometimes the problem is you.
    
    And that doesn't make you less worthy.
    
    Keep on growing.
    
    Keep on checking yourself.
    
    Keep on motivating yourself.
    
    Mistakes are opportunities.
    
    Look at them, own them. Grow from them and move on.
    
    Do better.
    
    Be better.
    
    You're human.
    
    It's okay.
    \item ``\textit{You stop explaining yourself when you realize people only understand from their level of perception.}'' - Jim Carrey
    \item ``It's impossible,'' said pride.
    
    ``It's risky,'' said experience.
    
    ``It's pointless,'' said reason.
    
    ``Give it a try,'' whispered the heart.
    \item Relationship status:
    \begin{itemize}
        \item Single
        \item Taken
        \item IDK, cursed or something
    \end{itemize}
    \item Talk with people who make you see the world differently.
    \item ``\textit{It's fucked up how people get judged for being real and how people get loved for being fake.}'' - 2Pac
    \item Don't treat people as bad as they are, treat them as good as you are.
    \item People who tolerate me on a daily basis$\ldots$ they're the real heroes.
    \item Biggest lesson learned this year is probably to not give so much of yourself to people who will not do the same for you.
    \item Real friends talk shit to your face, and say nice things behind your back.
    \item Not everyone likes me, but not everyone matters.
    \item Learn to be along.
    
    Not everyone will stay forever.
    \item I want my daughter to be kind but I also want her to know that she can throat punch someone if she needs to.
    \item How many girls can proudly say, \textit{I have original eyebrows}.
    \item If we date, there will be moments when \textit{I will just stare at you and smile}, know that in those moments I'm appreciating everything about you.
    \item Rule number 1: Fuck what they think.
    \item Sad truth: some of your friends secretly hate you.
    \item Dear best friend,
    
    I love you more daily.
    
    I wish you could see yourself the way I see you and I wish you could love yourself the way I love you.
    
    And above all, I wish your life is everything you deserve because, in my opinion, you deserve the world.
    
    I will stand by you forever.
    
    My heart will always belong to you.
    \item According to research, people who are always late are usually more creative.
    \item If a woman says `do what you want'
    
    Do not do what you want.
    
    Stand still, do not blink, do not answer, don't even breathe, just play dead.
    \item Don't be afraid of losing people.
    
    Be afraid of losing yourself by trying to please everyone around you.
    \item Just because I \textit{can't} sing, doesn't mean I \textit{won't} sing.
    \item \textsc{Be brave}.
    
    Even if you're not, pretend to be.
    \item ``\textit{Life is like riding a bicycle. To keep your balance, you must keep moving.}'' - Albert Einstein, in a letter to his son Eduard, Feb 5, 1930
    \item My attitude is a result of your actions, so if you don't like my attitude, blame yourself!
    \item The oldest sibling is usually the most well behaved.
    \item Do you have a friend who is far away from you but is still your bestfriend?
    \item You are the \textit{books} you read, the \textit{movies} you watch, the \textit{music} you listen to, the \textit{people} you spend time with, the \textit{conversations} you engage in.
    
    Choose wisely what you feed your mind.
    \item Every now and then, a person with no agenda, no ulterior motive and no self-interest will take pleasure in helping you to succeed, grow and live your purpose.
    
    This person will operate in love, will seek no praise and will want nothing in return.
    
    This person is a gift.
    \item I hope to arrive to my death \textit{late, in love} and a little drunk.
    \item Don't confuse my personality with my attitude.
    
    My personality is who I am.
    
    My attitude depends on who you are.
    \item Everything is temporary; emotions, thoughts, people and scenery.
    
    Do not become attached, just flow with it.
    \item Both facing poverty, but the daughter still sees her \textit{dad} as a king and he sees her as his whole world.
    
    This is what money can't buy.
    \item Who else loves that earthy smell when rain hits the ground?
    \item Current attitude towards work
    \begin{itemize}
        \item Fuck this job, I'm out!
        \item Lol just kidding, I've got bills to pay.
    \end{itemize}
    \item People come and go in your life, but the right ones will always stay.
    \item Talking to you, laughing with you, being with you, changes my whole mood.
    \item When you meet me, you think I'm quiet.
    
    When you talk to me, you wish I was quiet.
    
    When you know me, you get scared when I am quiet.
    \item I like looking at you when you are not paying attention because I think that is when you are the most beautiful.
    
    Because you are not trying, you are just being you and to to me, that is better than any picture you or anyone else could ever take.
    \item Don't be too confident when someone tells you they like you.
    
    The real question is, \textit{until when?}
    
    Because just like seasons, people \textit{change}.
    \item Love is the answer.
    \item Easy to make 10 friends in a year.
    
    But 1 friend for 10 years is special.
    \item No one is coming to save you.
    
    This life is 100\% your \textit{responsibility}.
    \item Never \textit{hide} your bad side to make someone stay, \textit{show} your bad side and see who can stay.
    \item \textit{My goal} is to build a life I don't need a vacation from.
    \item Sometimes you don't get closure.
    
    You just move on.
    \item Work hard, stay disciplined and be patient.
    
    You time will come.
    \item If overthinking situations burned calories, I'd be dead.
    \item Karma comes after everyone eventually.
    
    You can't get away with screwing people over your whole life, I don't care who you are.
    
    What goes around comes around.
    
    That's how it works.
    
    Sooner or later the universe will serve you the revenge that you deserve.
    \item ``\textit{Having a soft heart in a cruel world is courage, not weakness.}'' - Katherine Henson
    \item When someone tells me to do something that I was already planning on doing, well now I am not doing it.
    \item ``\textit{She is rare because she is real}.'' - Mark Anthony
    \item If you expect the world to be fair with you because you are fair, you're fooling yourself.
    
    That's like expecting the lion not to eat you because you didn't eat him.
    \item Win in your mind and you will win in your reality.
    \item Love all, trust few.
    
    Not everything is real \& not everyone is true.
    \item \textsc{Suggest a book} for others to read.
    \item \textsc{We were all humans} until race disconnected us, religion separated us, politics divided us and wealth classified us.
    \item ``\textit{She likes you?}
    
    \textit{How do you know?}''
    
    ``\textit{The eyes, Chico}.
    
    \textit{They never lie}.''
    \item \textbf{I'll always encourage the reckless texts confessing your feelings.}
    
    The kind where you throw your phone after hitting Send.
    
    I'll always encourage the horribly straightforward conversations at 3am when conversations get deep and you can't always put how you feel into words.
    \item True love has no expiry date.
    \item To people who reply late: If I wanted a reply after 2 days then I would have sent the message via a pigeon.
    \item There's difference between being liked and being valued.
    
    A lot of people like you.
    
    Not many value you.
    
    Be valued.
    \item The same boiling water that softens the potato hardens the egg.
    
    It's about what you're made of, not the circumstances.
    \item Most of my misunderstandings are because of my tone of voice.
    
    People think I'm angry even if I'm just explaining my point.
    \item ``\textit{Yesterday I was clever, so I wanted to change the world}.
    
    \textit{Today I am wise, so I am changing myself.}'' - Rumi
    \item I literally love being at home.
    
    In my own space.
    
    Comfortable.
    
    Not surrounded by people.
    \item I love places that make you realize how tiny you and \textsc{Your} problems are.
    \item Things we can't buy: Time, Happiness, Friends, Dreams, Hope, Love, Health.
    \item When you pray for someone, you are offering them the purest kind of love.
    \item Cheating is a choice, not a mistake.
    
    Loyalty is a responsibility, not a choice.
    \item A time will come in your life when some people will regret why they treated you wrong.
    
    Trust me, it will definitely come.
    \item The ego wants quantity but the soul wants quality.
    \item Yes I'm old school.
    
    I have good manners.
    
    I show others respect and I will always help those who need me.
    
    It's not because I'm old fashioned, it's because I was raised properly.
    \item You are rich, when you are content and happy with what you have.
    \item 1 of my greatest accomplishments is keeping my mouth shut even when I'm pissed off.
    \item Family is anyone who loves you unconditionally.
    \item Saying what you feel.
    
    It's not being rude.
    
    It's called being real.
    \item Deep conversation with the right people are priceless.
    \item Always cherish the girl who is with you when you have nothing.
    \item I need a coffee, a vacation, and a bag full of cash.
    
    That's all.
    \item It's nice when someone remembers small details about you.
    \item God removes people from your life because he heard conversations that you didn't hear.
    \item What's the most important thing you've done this year?
    
    \textit{Survived}$\ldots$
    \item Before you judge me, make sure you're perfect.
    
    If you're not, then shut up.
    \item What is salary?
    
    Something which comes like Tortoise and goes like Rabbit.
    \item You get what you focus on, so focus on what you want.
    \item The wrong one will find you in peace and leave you in pieces, but the right one will find you in pieces and lead you to peace.
    \item If you love someone, let them sleep.
    \item I'm not picky.
    
    I just know what I want.
    \item The best thing about the worst time of your life is that you get to see the \textit{true colors} of everyone.
    \item ``\textit{Exterior beauty, without the depth of a kind soul is merely decoration}.'' - Vanessa Quintero
    \item Every master was once a beginner.
    \item The older I get the more I appreciate being home doing absolutely nothing.
    \item I still watch cartoons and I don't care how old I am.
    \item - Wooo! It's Friday! You excited?!
    
    - Any excitement for `Friday' only proves that you're an unwitting slave to this social construct we call `the week'.
    
    Friday is nothing but a false hope, designed to distract you from noticing the machine silently oppressing you.
    
    You're just a hamster on a wheel linda, too excited by `Friday' to notice you're in a cage.
    
    - This is why nobody likes you Steve.
    \item No matter how nice your pictures are or how real your quotes are$\ldots$ there are some people who will never hit the like button just because it's you.
    \item The smartest thing a woman can ever learn, is to never need a man.
    \item If you find someone who makes you smile, who checks up on you often to see if you're okay, who watches out for you and wants the very best for you, \textbf{don't let them go}.
    
    Keep them close and don't take them for granted.
    
    People like that are hard to find.
    \item ``\textit{If my eyes could show my soul, everyone would cry when they saw me smile}.'' - Kurt Cobain
    \item People who are always Online, are the loneliest people.
    \item Sometimes to go far you must go alone.
    \item Obstacles do not block the path.
    
    They are the path.
    \item Never argue with someone who believes their own lies.
    \item Do you have a friend who is far away from you but is still your bestfriend?
    \item ``\textit{Go for someone who is proud to have you}.'' - Frank Ocean
    \item ``\textit{I have absolutely no desire to fit in}.'' - Coach Tim
    \item \textbf{Askhole} [n] A person who constantly asks for your advice, yet always does the exact the opposite of what you told them.
    \item Sometimes you have to keep your good news to yourself.
    
    Everybody is not genuinely happy for you.
    \item Fall in love with someone who wants you, who waits for you, who understands you even in the madness: someone who helps you, and guides you, someone who is your support, your hope.
    
    Fall in love with someone who talks with you after a fight.
    \item Sometimes I feel like giving up, and then I remember I have a lot of motherfuckers to prove wrong.
    \item I like people I can have comfortable silences with.
    \item We're all bad in someone's story.
    \item Dear music, thanks for always clearing my head, healing my heart, and lifting my spirit.
    \item I don't care what anyone thinks of me.
    
    Except dogs, I want dogs to like me.
    \item Be Honest: Are you missing someone right now?
    \item The number 1 mistake many women make is wasting years of their life waiting for a man to change.
    \item I had to forgive a person who wasn't even sorry$\ldots$ that's strength.
    \item Let them hate.
    
    Just make sure they spell your name right.
    \item The best 6 doctors: Sunshine. Water. Rest. Air. Exercise.
    \item ``\textit{To all the people who are loving and kind to me.}
    
    \textit{Thank you for all the sunshine you bring into my life}.'' - Brigitte Nicole
    \item It's easy to take off your clothes and have sex.
    
    People do it all the time.
    
    But opening up your soul to someone, letting them into your spirit, thoughts, fears, future, hopes, dreams$\ldots$ that is being naked.
    \item I truly appreciate kindness.
    
    I appreciate people checking up on me.
    
    I appreciate a quick message, I appreciate those who ask if I'm okay, I appreciate every single person in my life who has tried to brighten my days.
    
    \textit{It's the little things that matter the most}.
    \item Do you have someone with whom you have completed more than 5 years of friendships?
    \item Life is more beautiful when you meet the right hairdresser.
    \item Yes I am a girl, and I'm not interested in makeup.
    
    Yes, We exist!
    \item Have you ever been too nice and ended up in a situation that could've been avoided if you just would've been an asshole.
    \item ``\textit{Holding a grudge doesn't make you strong; it makes you bitter.}
    
    \textit{Forgiving doesn't make you weak; it sets you free}.'' - Dave Willis
    \item ``\textit{Sometimes you have to love people from a distance}.'' - Robert Tew
    \item The world is not full of assholes, but they are strategically placed so that you'll come across one every day.
    
    Every. Day.
    \item ``\textit{The most important thing in communication is hearing what isn't said}.'' - Peter F. Drucker
    \item \textsc{Know your worth}: You must find the courage to leave the table if respect is no longer being served.
    \item The best conversations are from 12--3am.
    \item No one wants to support you at the start of your journey but once they see your success they all want to be friends again.
    \item I'm naturally irritated when I 1st wake up.
    
    You have to give me a few minutes to adjust.
    \item Always love your friends from your heart not from your Mood or Need.
    \item Real is rare and fake is everywhere.
    \item Life is weird.
    
    1st you want to grow up, then you want to be a kid again.
    \item You know too much psychology when you can't get mad because you understand everyone's reasons for doing everything.
    \item It's blessing to have a friend with same level of dirty mind.
    \item Have you ever looked at someone and hoped they stay in your life forever?
    \item Shit happens.
    
    Every day.
    
    To everyone.
    
    The difference is in how people deal with it.
    \item I either keep it all inside, or say exactly how I feel with no filter.
    
    There is no in between.
    \item I don't have time to google Lyrics.
    
    I sing what I hear.
    \item If a cat crosses road it doesn't mean bad luck.
    
    It means that the cat is going somewhere.
    \item Weak people revenge.
    
    Strong people forgive.
    
    Intelligent people ignore.
    \item ``I want you back'' - Me talking to the money that I spent for nothing
    \item A person becomes 10 times more attractive not by their looks but by their acts of kindness, love, respect, honesty, and loyalty they show.
    \item Lucky are those who find \textbf{a true loyal friend} in this fake world.
    \item Stop following the crowd$\ldots$
    
    They are lost.
    \item Everyone has a friend during each stage of life.
    
    But only lucky ones have the same friend in all stages of life.
    \item You can say sorry.
    
    But you can't change the story.
    \item ``\textit{The hardest decisions in life is whether to ``walk away'' or ``try harder''.}'' - Ziad K. Abdelnour
    \item Sometimes \textit{home} isn't 4 walls, it's 2 eyes and a heartbeat.
    \item Every pain gives a lesson and every lesson changes a person.
    \item If you're funny, you're automatically more attractive.
    
    Beauty fades but \textit{sarcasm is forever}.
    \item ``\textit{Soon you realize that many people will love the idea of you but will lack the maturity to handle the reality of you}.'' - Reyna Biddy
    \item A moment of silence for those who hate us but can't unfriend us because they afraid of not knowing what's happening in our lives.
    \item ``\textit{There will be very painful moments in your life that will change your entire world in a matter of minutes.}
    
    \textit{These moments will change you}.
    
    \textit{Let them make you \textsc{stronger, smarter}, and \textsc{kinder}.}
    
    \textit{But don't you go and become someone that you're not}.
    
    \textit{Cry}.
    
    \textit{Scream if you have to}.
    
    \textit{Then you straighten out that crown and keep moving}.'' - Erin van Vuren
    \item I just want to
    \begin{itemize}
        \item take photos
        \item drink coffee
        \item travel the world
        \item meet new people
        \item be happy
    \end{itemize}
    \item Re-set, Re-adjust, Re-start, Re-focus$\ldots$
    
    As many times as you need to.
    \item If you have just 3 tickets for world tour.
    
    Choose that 2 friends with whom you want to travel the world.
    \item Learn when to be aggressive and when to be patient.
    \item \textsc{Educate yourself}.
    
    When a question about a certain topic pops up, google it.
    
    Watch movies and documentaries.
    
    When something sparks your interest, read about it.
    
    Read read read.
    
    Study, learn, stimulate your brain.
    
    Don't just rely on the school system, educate that beautiful mind of yours.
    \item And then it happens$\ldots$
    
    1 day you wake up and you're in this place.
    
    You're in this place where everything feels right.
    
    You heart is calm.
    
    Your soul is lit.
    
    Your thoughts are positive.
    
    Your vision is clear.
    
    You're at peace, at peace with where you've been, at peace with what you've been though and at peace with where you're headed.
    \item I never cared about the material things someone could give me.
    
    I care about time, attention, honesty, loyalty, and effort.
    
    Those gifts mean more than anything money could buy.
    \item A year changes you a lot.
    \item With great power comes great electricity bill.
    \item Some talk to you in their free time, and some free their time to talk to you.
    
    Learn the difference.
    \item ``Dead people receive more flowers than the living ones because regret is stronger than gratitude.'' - Anne Frank
    \item - Where did you find that?
    
    I've been search for it everywhere.
    
    - I created it [happiness] myself.
    \item No more expectations, just gonna go with the flow and whatever happens, happens.
    \item Be the reason someone feels, seen, heard, and supported.
    \item Karma said: If you break someone's heart and they still talk to you with the same excitement and respect, believe me, they really love you.
    \item Birds born in a cage think flying is an illness.
    \item An English professor wrote the words:
    
    ``A woman without her man is nothing'' on the chalkboard and asked the students to punctuate it correctly.
    
    All of the makes in the class wrote: ``A woman, without her man, is nothing.''
    
    All of the females in the class wrote: ``A woman: without her, man is nothing.''
    
    Punctuation is powerful.
    \item ``To be is to do.'' - Socrates
    
    ``To do is to be.'' - Kant
    
    ``Do be do be do'' - Scooby Doo
    \item Be a good person, but don't waste time to prove it.
    \item Friends come and go like the waves of the ocean, but true ones stay like an octopus on your face.
    \item People come and go in your life, but the right ones will always stay.
    \item Have you ever \textit{lost all respect} for someone?
    
    Like, you don't hate them but you don't feel the need to associate yourself or say anything to them anymore.
    \item Sometimes in life we just need a hug$\ldots$ no words, no advice, just a bug to make you feel you matter.
    \item Raise your hand if you need a break from life and go on a trip far away from society and worries.
    \item Normal People: 1 episode per night
    
    Me: 1 season per night
    \item Unplanned trips are the best because planned trips never happen.
    \item You know what's beautiful?
    
    \textit{A real conversation}.
    \item ``\textit{Mondays are fine.}
    
    \textit{It's your life that sucks}.'' - R. Gervais
    \item You love her because her skin is white and soft?
    
    It's not Love, it's Dove.
    \item Every girl has male best friend in her life with whom she can share everything, without any fear.
    \item Not responding is also a response.
    \item Stupid is Knowing the truth, seeing the truth, but still believing the lies.
    \item Who is your 2 am friend?
    \item Life is too short to be normal.
    
    Stay weird.
    \item Boys who can cook are husband materials.
    \item That 1 girl who looks shy and innocent from outside, but she is big naughty from inside.
    \item - Hello, I'm here for a job interview.
    
    - Great, and do you have experience?
    
    - Yes, this is my 20th interview.
    \item \textbf{Don't complicate life}
    \begin{itemize}
        \item Missing somebody? \textbf{Call}
        \item Wanna meet up? \textbf{Invite}
        \item Wanna be understood? \textbf{Explain}
        \item Have questions? \textbf{Ask}
        \item Don't like something? \textbf{Say it}
        \item Like something? \textbf{State it}
        \item Want something? \textbf{Ask for it}
        \item Love someone? \textbf{Tell them}
    \end{itemize}
    Keep your life SIMPLE.
    \item A Daughter is 1 of the most beautiful gifts the world has to give.
    \item You don't always need a logical reason for doing everything in your life.
    
    Do it because you want to; because it's fun; because it makes you happy.
    \item Never forget who helped you out while everyone else was making excuses.
    \item Silence is better than unnecessary drama.
    \item Getting married after 30 is still beautiful.
    
    Starting a family after 35 is still possible.
    
    Buying your 1st House after 40 is still a boss move.
    
    Don't let people rush you with their timelines.
    \item I try to be a nice person, but sometimes my mouth doesn't cooperate.
    \item 1 of my dream is to go on a night camping trip with my favorite people.
    \item If everybody likes you, you have a serious problem.
    \item Jobs fill your pocket.
    
    Adventures fill your soul.
    \item Everything in life is temporary.
    
    So if things are going good, enjoy it because it won't last forever.
    
    And if things are going bad, don't worry.
    
    It can't last forever either.
    \item ``\textit{I can respect any person who can put their ego aside and say, I made a mistake, ``I apologize, and I am correcting the behavior''}.'' - sylvester mcnutt III
    \item Happiness is$\ldots$ Not setting alarm for the next day.
    \item Laughter is the best medicine, but if you're laughing for no reason you need medicine.
    \item Remember that you are \textsc{Water}.
    
    Cry. Cleanse. Flow. Let Go.
    
    Remember that you are \textsc{Fire}
    
    Burn. Tame. Adapt. Ignite.
    
    Remember that you are \textsc{Air}
    
    Observe. Breath. Focus. Decide.
    
    Remember that you are \textsc{Earth}
    
    Ground. Give. Build. Heal.
    
    Remember that you are \textsc{Spirit}
    
    Connect. Listen. Know. Be Still.
    \item Stay low key.
    
    Not everyone needs to know everything about you.
    \item I'm in love with this quote:
    
    ``\textit{When you get what you want, that's God's direction, when you don't get what you want, that's God's protection}.''
    \item Once you have matured, you realize silence is more important than proving a point.
    \item - Morning: Tired.
    
    - Afternoon: Dying for a rest.
    
    - Night: Can't sleep.
    \item Sometimes, the nicest people you meet are covered in tattoos, and sometimes the most judgmental people you meet go to church on Sundays.
    \item ``\textit{You don't have a soul.}
    
    \textit{You are a soul}.
    
    \textit{You have a body}.'' - C. S. Lewis
    \item Choose people who choose you.
    \item ``\textit{The problem with the world is that the intelligent people are full of doubts and the stupid ones are full of confidence}.'' - Charles Bukowski
    \item The older I get, the more I realized I don't want to be around drama, conflict or stress.
    
    I want a cozy home, good food, and to be surrounded by happy people.
    \item Don't \textbf{become} who hurt you.
    \item A beautiful face will age and a perfect body will change, but a beautiful soul will always be a beautiful soul.
    \item All good things are wild \& free.
    \item If you focus on the hurt, you will continue to suffer.
    
    If you focus on the lesson, you will continue to grow.
    \item Blanket on: too hot
    
    Blanket off: too cold
    
    1 leg out: perfect
    \item ``\textit{Understand me}.
    
    \textit{I'm not like an ordinary world}.
    
    \textit{I have my madness, I live in another dimension and I do not have time for things that have no soul}.'' - Charles Bukowski
    \item Communicate.
    
    Even when it's uncomfortable or uneasy.
    
    1 of the best ways to heal, is simply getting everything out.
    \item Accepting your nose is the 1st step to self love.
    \item Stop looking for a partner.
    
    Focus on your goals and rebuilding your life.
    
    The right person will eventually find their way to you.
    \item Do more things that make you forget to check your phone.
    \item Hello, my name is human.
    
    And I came down from the stars.
    \item I suffer from that disorder where I speak the truth and it pissed people off.
    \item ``\textit{She fell}
    
    \textit{She crashed}
    
    \textit{She broke}
    
    \textit{She cried}
    
    \textit{She crawled}
    
    \textit{She hurt}
    
    \textit{She surrender}
    
    \textit{And then$\ldots$ She rose again} - Nausicaa Twila
    \item There are 2 reasons why we don't trust people.
    
    1st, we don't know them.
    
    2nd, we know them.
    \item ``\textit{I didn't wish him all the best, because that would be a lie, but I didn't wish him all the worst either}.
    
    \textit{I simply wished him whatever it was he deserved}.
    
    \textit{Now whether he deserved good things or bad things, was none of my business - that was between him, and karma}.'' - CiCi B.
    \item ``\textit{The child who is not embraced by the village will burn it down to feel its warmth}.'' - African Proverb
    \item Every time I trust somebody, they show me why I shouldn't.
    \item In a room full of art I'd still stare at you.
    \item \textsc{A serious life tip}:
    
    Even the nicest people have their limits.
    
    Don't push them too far and don't try to reach those limits, because the nicest people can also be the scariest assholes once they've had enough.
    \item ``\textit{Life has taught me that you can't control someone's loyalty}.
    
    \textit{No matter how good you are to them, doesn't mean they'll treat you the same}.
    
    \textit{No matter how much they mean to you, doesn't mean they'll value you the same}.
    
    \textit{Sometimes the people you love the most, turn out to be the people you can trust the least}.'' - Trent Shalton
    \item Love is Temporary.
    
    Food is Permanent.
    \item When you become really close to someone, you can hear their voice in your head when you read their texts.
    \item Ever locked yourself in the bathroom, cried, washed your face, and came back like nothing happened?
    \item Don't you just wanna sit on the roof of your house with someone, watch the night sky, and talk about how crazy life is.
    \item I lost many friends just because I stop texting them 1st.
    \item \textbf{How to be happy}
    
    Ignore people who think they know more about you than you do.
    \item \textbf{Did you know?}
    
    Every day, the heart creates enough energy to drive a truck 20 miles.
    
    In a lifetime, that is equivalent to driving to the moon and back.
    
    So, when you tell someone you love them ``to the moon and back'' you're essentially saying you will love them with all the blood your heart pumps your whole life, which I think is equally as meaningful.
    \item \textit{Fall in love} with someone who wants you, who waits for you, who understands you.
    
    Someone who helps you, and guides you, someone who is your support, your hope.
    
    Fall in love with someone who talks with you after a fight.
    
    Fall in love with someone who misses you and wants to be with you.
    
    Do not fall in love only with a body or with a face; or with the idea of being in love.
    \item Never fuck with someone who is not afraid to be alone.
    
    You will lose every single  time.
    \item ``\textit{People who are spiritually minded tend to suffer from anxiety and depression more}.
    
    \textit{You know why?}
    
    \textit{Because their eyes are open to a world that is in need of repair}.
    
    \textit{They literally have an increased ability to feel the emotions of people around them}.'' - Osho
    \item When something good happens, \textbf{travel} to celebrate.
    
    If something bad happens, \textbf{travel} to forget it.
    
    If nothing happens, \textbf{travel} to make something happen.
    \item \textsc{Positive life}
    
    If you can't find true love, work hard, make money and enjoy your single life in peace.
    
    Nobody has ever died from being Single, but so many have died for being with the wrong partner$\ldots$
    
    Life is too short to be wasting your time with the wrong person.
    \item Good girls love bad boys
    \item When you see your Ex posting status about true love, loyalty \& emotions.
    \item Be good to people for no reason.
    \item Fill your life with adventures, not things.
    
    Have stories to tell, not stuff to show.
    \item Did you know?
    
    Chocolate comes from Cocoa, which is a tree.
    
    That makes it a plant$\ldots$
    
    So chocolate is a salad.
    \item Mr. Bean taught me 1 thing in life.
    
    Enjoy your own company instead of expecting someone else to make you happy.
    \item Girls be like$\ldots$ I don't have anything to wear.
    \item A person who trusts no one now, once trusted someone too much.
    \item What's a queen without her king?
    
    Well, historically speaking, more powerful.
    \item Maybe if we tell people the \textit{brain} is an \textit{app}, they'll start using it.
    \item If you don't do stupid things while you're young, you'll have nothing to laugh about when you're old.
    \item ``\textit{If you're pretty, you're pretty}.
    
    \textit{But the only way to be beautiful is \textbf{to be loving}}.
    
    \textit{Otherwise, it's just ``Congratulations about your face.''}.'' - John Mayer
    \item When life puts you in tough situations, don't say ``Why me'' say ``Try me''.
    \item If you kick me when I'm down, you better pray I don't get up.
    \item You learn nothing from life if you think you're right all the time.
    \item \textbf{A Perfect Man}
    \begin{itemize}
        \item wakes up at 5 am everyday
        \item exercises everyday
        \item makes his own bed
        \item cleans his room
        \item works sincerely
        \item does not touch alcohol
        \item helps in the kitchen
        \item does not indulge in night life
        \item is always punctual
        \item prays daily
        \item reads
        \item hits the bed at 9 pm sharp.
    \end{itemize}
    Where do you find such Perfect Men?
    
    Answer: Jail.
    \item Being a music lover with a terrible voice is the saddest part of my life.
    \item 98\% of my problems would be solved if I stopped overthinking things and calmed the hell down.
    \item ``\textit{Everyone you meet always asks if you have a \textsc{Career}, are \textsc{Married} or own a \textsc{House} as if life was some kind of grocery list}.
    
    \textit{But no one ever asks you if you are \textsc{Happy}}.'' - Heath Ledger
    \item Birthplace: Earth
    
    Race: Human
    
    Politics: Freedom
    
    Religion: Love
    \item \textbf{Signs of maturity}:
    \begin{itemize}
        \item[1.] Small talks no longer excite you.
        \item[2.] Sleep is better than a Friday night out.
        \item[3.] You forgive more.
        \item[4.] You become more open-minded.
        \item[5.] You respect differences.
        \item[6.] You don't force love.
        \item[7.] You accept heartaches.
        \item[8.] You don't judge easily.
        \item[9.] You sometimes prefer to be silent than to engage in a nonsense fight.
        \item[10.] Your happiness don't depend on people but on your inner self.
    \end{itemize}
    \item What's coming is better than what's gone.
    \item If you don't like me please don't pretend that you do.
    
    Ever.
    \item Imagine being loved the way you love.
    \item Someone once asked me: ``Why do you love music so much?''.
    
    I replied: Because it's the only thing that stays when everything and everyone is gone.
    \item ``\textit{Just because you don't share it on social media, doesn't mean you're not up to big things}.
    
    \textit{Live it and stay low key}.
    
    \textit{Privacy is everything}. - Denzel Washington
    \item Deleting your story after the target audience saw it.
    \item In life, what you really want will never come easy.
    \item If you're not losing friends, you're not growing up.
    \item To that friend who has never been tired in Listening to my dramas.
    
    Thank you I owe you a lot!
    \item You act like it's you against the world but it's really just you against yourself.
    \item What if you met the right person but at the wrong time.
    \item We all have that 1 friend who eats a lot but never gets fat.
    \item \textit{\textbf{Stop being offended} by a Facebook post, by a piece of art, by people displaying affection, or by what someone said to you}.
    
    \textit{\textbf{Be offended} by war, poverty, greed and injustice}. - Sue Fitzmaurice
    \item Have less.
    
    Be more.
    
    Do more.
    \item \textbf{Brother} [n] (bruhth-er) The person who is there when you need him; someone who picks you up when you fall; a person who sticks up for you when no one else will; a brother is always a friend.
    \item Be Honest: Do you still love your ex?
    \item ``\textit{I think people spend \textbf{too much time staring at screens} and not enough time drinking wine, tongue kissing, and dancing under the moon}.'' - Rachel Wolchin
    \item The only way to win with a toxic person is not to play.
    \item \textbf{Keep going.}
    
    No matter how stuck you feel.
    
    No matter how bad things are right now.
    
    No matter how many days you've spent crying.
    
    No matter how hopeless and depressed you feel.
    
    No matter how many days you've spent wishing things were different.
    
    I promise you won't feel this way forever.
    
    \textbf{Keep going}.
    \item \textbf{Earthing}: The process of absorbing earths free flowing electrons from it's surface through the soles of ones feet.
    \item \textbf{How to save your heart?}
    
    \textsc{Should}:
    \begin{itemize}
        \item Never expect.
        \item Never demand.
        \item Never assume.
    \end{itemize}
    \item \textsc{Know}:
    \begin{itemize}
        \item Your limits.
        \item Where you stand.
        \item Your role.
    \end{itemize}
    \item \textsc{Don't}:
    \begin{itemize}
        \item Get affected.
        \item Get jealous.
        \item Get paranoid.
    \end{itemize}
    \item \textsc{Just}: Go with the flow and stay happy.
    \item Thins are not always what they seem.
    
    The deer isn't crossing the road.
    
    The road is crossing the forest.
    \item Surround yourself with people who push you to do and be better.
    
    No drama or negativity.
    
    Just higher goals and higher motivation.
    
    Good times and positive energy.
    
    No jealousy or hate.
    
    Simply bringing out the absolute best in each other.
    \item Move in silence, only speak when it's time to say \textsc{Checkmate}.
    \item Once you start loving someone, it's very hard to stop$\ldots$
    \item Home is not a place, it's a feeling.
    \item Do you know that awesome feeling when you get into bed, fall right to sleep, stay asleep all night, and wake up feeling refreshed?
    
    Me neither.
    \item So much of our happiness depends on how we choose to look at the world.
    \item It was never the way he looked.
    
    Always the way she was.
    
    I could've fallen in love with my eyes closed.
    \item Robin William's friends: ``He was always happy.
    
    Everyone adored him.''
    
    Kate Spade's Dad: ``I just talked to her an hour before and she was planning a trip.
    
    She was just like her brand - happy, cheerful and full of color.''
    
    Anthony Bourdain's best friend: ``He loved his life and had this extraordinary ability to just connect with people.''
    
    So, let me say this really loud so the people in the back of the room can hear me$\ldots$
    
    \textsc{Sometimes you need to check on those who seem the strongest}.
    \item ``\textit{There is a misconception that Buddhism is a religion \& that you worship Buddha}.
    
    \textit{Buddhism is a practice, like Yoga, you can be a Christian \& practice Buddhism}.
    
    \textit{I met a Catholic priest who lives in a Buddhist Monastery in France}.
    
    \textit{He told me that Buddhism makes him a better Christian}.
    
    \textit{I love that}.'' - Thich Nhat Hanh
    \item Never judge people by their past.
    
    People learn.
    
    People change.
    
    People move on.
    \item Dear Girls:
    
    Before going into a relationship with any guy in this century.
    
    Take the boy's picture \& post it on Facebook, Twitter, \& Instagram with caption: ``\textit{Whose boyfriend is this?}''
    
    If no one claims him in a week than you're free to date him.
    \item Be careful who you push away$\ldots$
    
    Some of us don't come back.
    \item If it \textit{excites you} and \textit{scares you} at the same time, it might be a good thing to try.
    \item Classy is when you have a lot to say but you choose to remain silent in front of fools.
    \item Wearing unbranded cheap clothes doesn't mean you're poor.
    
    Remember: You have a family to feed.
    
    Not a community to impress.
    \item You're lucky if you've found a person who never gets tired of understanding your nonsense attitude.
    \item \textbf{I miss you}.
    
    I miss your voice.
    
    I miss your smile.
    
    I miss your smell.
    
    I miss your hugs.
    
    I miss your jokes.
    
    I miss how you made me feel.
    
    \textbf{I miss your everything}$\ldots$
    \item Tell my mistakes to me not to others.
    
    Because my mistakes are to be corrected by me, not by others.
    \item Stop editing your pics, what if you go missing?
    
    How can we find you if you look like Angelina Jolie on Instagram \& potato in real life.
    \item When you build in silence, people don't know what to attack.
    \item Sometimes you gotta play the tool, to fool the fool, who thinks they're fooling you.
    \item I'm actually a very nice person, until you piss me off.
    \item Make sure you check your spelling and grammar before post anything.
    
    Because there are so many jobless English professor on social media.
    \item I wish I met some people earlier, some a little later and some never at all.
    \item Being good is a mistake nowadays, people think you're Fake.
    \item To my best friends,
    
    1 day, all of us will get separated from each other.
    
    We will miss our conversations of everything and nothing and the dreams we had.
    
    Days, months and years will pass until this contact becomes rare.
    
    1 day, our children will see our pictures and ask: ``Who are these people?''
    
    And we will smile with invisible tears because a heart is touched with a strong word and you will say:
    
    ``It was them that I had the best days of my life with.''
    \item I've only met 3 or 4 people that understand me.
    
    Everyone else assumes I'm either angry, sarcastic or just an asshole.
    \item Never forget 2 people in your life.
    \begin{itemize}
        \item[1.] The person who lost everything just to make you win. (Your Father)
        \item[2.] The person who was with you in every pain. (Your Mother)
    \end{itemize}
    \item Stand up for what you believe in, even if it means standing alone.
    \item Romeo died because of Juliet, Jack died because of Rose, stay single if you want to live.
    \item Train your mind to stay calm in every situation.
    \item \textit{Go for it}.
    
    Whether it ends good or bad, it was an experience.
    \item Never attribute to malice what can be explained by ignorance.
    
    People are far more stupid than they are evil.
    \item Me risking my job, career and future to get an extra 15 minutes of sleep.
    \item Mean World Syndrome is a phenomenon whereby violence-related content of mass media makes viewers believe that the world is more dangerous than it actually is.
    \item Find someone who enjoys your Madness.
    
    Not an Idiot who forces you to be normal.
    \item People will come and go in life, but the person in the mirror will be there forever: So be good to yourself.
    \item Old saying - Think before you speak.
    
    New saying - Google, before you post.
    \item I think \textit{1 of the greatest feelings} in the world is when someone openly tells you how much you mean to them.
    
    Stuff like that is so rare.
    \item I don't care if I'm selfish.
    
    After putting people 1st for the longest time and being disappointed I deserve to do whatever makes me feel happy.
    \item My 3 wishes:
    \begin{itemize}
        \item[1.] To earn money without working.
        \item[2.] To love without being hurt.
        \item[3.] To eat without getting fat.
    \end{itemize}
    \item I just wanna sit outside at night and talk about life with someone.
    \item 1 of the best feelings in the world is when you hug someone you love and they hug you back even tighter.
    \item \textbf{Petrichor} [n] the smell of earth after rain.
    \item Post an unpopular opinion you have and only comment \textit{agree} or \textit{disagree} under others.
    
    \textit{No arguing}.
    \item Don't consider my kindness as my weakness.
    
    The best in me is sleeping, not dead.
    \item It has been said that everlasting friends go long periods of time without speaking and never question the friendship.
    
    These friends pick up like they just spoke yesterday, regardless of how long it has been or how far away they live, and they don't hold grudges.
    
    They understand that life is busy, and you will always love them.
    \item \textbf{A beautiful day begins with a beautiful mindset.}
    
    When you wake up, take a second to think about what privilege it is to simply be alive and healthy.
    
    The moment you start acting like \textit{life is a blessing}, it will start to feel like one.
    \item In the end, we all just want \textit{someone that chooses us}.
    
    Over everyone else.
    
    Under any circumstances.
    \item When a man realizes that his woman is a teammate not an opponent, that's when his life changes.
    
    Don't break her down.
    
    Build her up and let her help you win.
    \item Not all storms come to disrupt your life, some come to clear your path.
    \item Be strong enough to stand alone, be yourself enough to stand apart, but be wise enough to stand together when the time comes.
    \item Some husbands hold their wife's hand in malls because if they have her hand, she'll go for shopping.
    
    It looks ``Romantic'' but it's actually ``Economic''.
    \item \textit{You are so missed and loved}.
    
    There isn't a day that goes by that I don't think about you.
    
    Although I can't see you, you are always in my heart.
    
    \textit{Rest in Peace}.
    \item Don't ignore the effort of a person who tries to keep in touch.
    
    It's not all the time someone cares.
    \item Be kind to unkind people, they need it the most.
    \item Never hurt that person whose only intention was to see you happy.
    \item Silence is the best reply to a fool.
    \item Don't love too deeply until you're sure that the other person loves you with the same depth.
    
    Because the depth of your love today is the depth of your wound tomorrow.
    \item ``\textit{I have no special talent}.
    
    \textit{I am only passionately curious}.'' - Albert Einstein
    \item \textbf{Be happy} in front of people who don't like you - it kills them.
    \item Spend time with people who are good for your mental health.
    \item A real man knows that 1 woman is enough.
    \item You can't go back and change the beginning, but you can start where you are and change the ending.
    \item Not everyone is meant to be in your future.
    
    Some people are just passing through to teach lessons in life.
    \item If all religions teach peace, why can't all religions achieve peace?
    \item \textit{Angels} exist but sometimes they don't have wings and are called \textit{friends}.
    \item Veni. Vidi. Amavi.
    
    We came. We saw. We \textit{loved}.
    \item I'm innocent but my bestfriend taught me about dirty stuffs.
    \item To all the people who make me happy: thank you for being in my life!!!
    \item Stuck between ``I need to save money'' and ``You only live one''.
    \item ``\textit{People will teach you how to love by not loving you back}.
    
    \textit{People will teach you how to forgive by not apologizing}.
    
    \textit{People will teach you kindness by their judgment}.
    
    \textit{People will teach you how to grow by remaining stagnant}.
    
    \textit{Pay attention when you're going through painful and mysterious times}.
    
    \textit{Listen to the wisdom life is trying to teach you}. - Meredith Marple
    \item \textit{Keep your distance} from people who will never admit they are wrong and always try to make you feel like it's all your fault.
    \item I have no time to battle egos and small minds.
    \item If you aren't happy single, you won't be happy taken.
    
    Happiness comes from within, NOT from men.
    \item A religious person will do what he is told$\ldots$ no matter what is right$\ldots$ whereas a spiritual person will do what is right$\ldots$ no matter what he is told.
    \item Leadership is not about being the best.
    
    Leadership is about making everyone else better.
    \item ``\textit{How amazing it is to find someone who wants to hear about all the things that go on in your head}.'' - Nina LaCour
    \item I think as you grow older you look for very different things in people.
    
    \textbf{Honesty. Loyalty. Integrity.}
    
    But most of all, you look for someone who will stand right by your side when the walls start crumbing and the fires rage within.
    
    They are right there, and at that moment, you know they've got you. - The better man project
    \item Sometime those who don't socialize much aren't actually anti-social.
    
    They just have no tolerance for drama, stupidity, and fake people.
    \item I like being at home in my own little world.
    
    The real world has too many assholes.
    \item Does anyone else tell their pets, ``I'll be back soon'' when they leave the house or is that just me?
    \item \textbf{If a door closes, quit banging on it.}
    
    Whatever was behind it was not meant for you.
    
    Consider that perhaps the door was closed because you're worth so much more than what was on the other side.
    \item Not everyone will understand your journey.
    
    That's okay.
    
    You're here to live your life, not to make everyone understand.
    \item I asked an old man, ``\textit{Even after 95 years, you still call your wife Darling, Honey and Love. What's the secret?}''
    
    Old man: ``\textit{I forgot her name 10 years ago and I'm scare to ask her}.''
    \item Whatever you do, good or bad, people will always have something negative to say about you and that's life.
    \item Me: \textit{I want to travel}.
    
    Bank account: \textit{Where? To work?}
    \item Can I just be a little kid again?
    
    No stress, no worries, just fun.
    \item \textbf{Never beg for love.}
    
    \textbf{Never beg} someone to be with you.
    
    \textbf{Never beg} someone to come back to stay.
    
    \textbf{Never beg} for attention, commitment, affection, time and effort.
    
    You should never have to ask to feel wanted.
    
    Begging is demanding and degrading.
    
    If someone doesn't willingly give you these things, with their arms wide open, they aren't worth it.
    
    No one, under any circumstances, is ever worth begging for.
    \item The world is gonna judge you no matter what you do, so live your life the way you want to.
    \item ``\textit{You could give some people a drop of water, and they'd still appreciate you}.
    
    \textit{You could give other people the entire ocean, and they'd still take you for granted}. - Yasmin Mogahed
    \item Silence isn't empty, it's full of answers.
    \item That moment when your best friend sends you a voice message \& you're running everywhere to search earphones cause you don't trust that idiot.
    \item When someone calls me ``lazy'': Thank you. I know.
    \item The best thing about being single is you can have crush on anyone, anywhere, anytime.
    \item 1 of the best things about getting older: knowing someone is an asshole before they even speak.
    \item Don't wait until you are rich to be happy.
    
    Happiness is free!
    \item The fact that jellyfish have survived for 650 million years without brains, gives hope to many people.
    \item \textit{The most beautiful things in life are not things}.
    
    They are people, places, memories and pictures.
    
    They're feelings and moments, and smiles, and laughter.
    \item \textit{Psychology says}: When you focus on problem, you will have more problems.
    
    When you focus on possibilities, you will have more opportunities.
    \item ``\textit{Don't let anyone tell you that your independence is the reason for you being single}.
    
    \textit{Your strength as a woman isn't the cause for your loneliness}.
    
    \textit{You're alone because you're rather not entertain a weak man}.'' - r.h. Sin
    \item ``\textit{Look 1st for honesty in a smile, not beauty}.'' - J.H. Hard
    \item ``\textit{Beware of false knowledge; it is more dangerous than ignorance}.'' - George Bernard Shaw
    \item Only you and you alone can change your situation.
    
    Don't blame it on anything or anyone.
    \item \textit{Parked car conversations} are low key therapy sessions.
    \item \textit{So don't hurt her, don't change her, don't analyze and don't expect more than she can give}.
    
    \textit{Smile when she makes you happy, let her know when she makes you mad, and miss her when she's not there}.'' - Bob Marley
    \item Don't wait for things to get easier, simpler, better.
    
    Life will always be complicated.
    
    Learn to be happy right now.
    
    Otherwise, you'll run out of time.
    \item Car rides by yourself with loud music are good for the soul.
    \item ``\textit{How beautiful it is to find someone who asks for nothing but your company}.'' - Brigitte Nicole
    \item The best things in life aren't things.
    \item 4 things you can't get back:
    \begin{itemize}
        \item The \textit{stone} after it's \textit{thrown}.
        \item The \textit{word} after it's \textit{said}.
        \item The \textit{occasion} after it's \textit{missed}.
        \item The \textit{time} after it's \textit{gone}.
    \end{itemize}
    \item ``Someone told me the other day that he felt bad for single people because they are lonely all the time.
    
    I told him that's not true I'm single and I don't feel lonely.
    
    I take myself out to eat, I buy myself clothes.
    
    I have great times by myself.
    
    Once you know how to take care of yourself company becomes an option and not a necessity.'' - Keanu Reeves
    \item To care for those who once cared for us is 1 of the highest honors.
    \item She was simple like quantum physics.
    \item If you feel like you're losing everything, remember that trees lose their leaves every year and they still stand tall and wait for better days to come.
    \item Fuck your shoe collection.
    
    Show me your book collection.
    \item Sometimes God sends an ex back into your life to see if you are still stupid.
    \item You don't need to be liked by everyone, not everyone has a good taste.
    \item ``\textit{To me, Nature is a place where you retreat whenever you feel exhausted or sad}.
    
    \textit{I'm more happy when I'm surrounded by the sound of birds than sound of people}.'' - Keanu Reeves
    \item When you are angry, be silent.
    \item Sometimes I just agree with people so that they can stop talking.
    \item If ``Plan A'' didn't work.
    
    The alphabet has 25 more letters!
    
    Stay Cool.
    \item It is important to have friends who are proud of you when you get a new job or learn to bake or do big things, but it is also important to have friends who are proud of you when you \textit{get out of bed and take a shower}.
    \item When you die and hear your wife saying ``I've never made him sad.''
    \item What makes you different makes you beautiful. - Book of Prosperity
    \item Small steps in the right direction are better than big ones in the wrong direction.
    \item ``\textit{$\ldots$ and oh, how they \textbf{love the woman who won't take any shit} from anyone, until they realize that she won't take any shit from them, either}.'' - CiCi B.
    \item 1 day, you will be at the place you always wanted to be.
    
    Don't stop believing.
    \item You can't force people to stay in your life.
    
    Stay is a choice, so be thankful to the people who chose you.
    \item Stay strong, make them wonder how you are still smiling.
    \item ``\textit{My life isn't perfect but I'm thankful for everything I have}.'' - Note from Life
    \item Only once you have carried your own water, will you learn the value of every drop.
    \item ``Nothing stronger than a broken man rebuilding himself.'' - Via (Dark Secrets)
    \item If you fail, never give up because F.A.I.L means ``First Attempt In Learning''; End is not the end, in fact E.N.D. means ``Effort Never Dies''.
    
    If you get No as an answer, remember N.O. means ``Next Opportunity''.
    \item Don't wait until you've reached your goal to be proud of yourself, be proud of every step you take toward teaching that goal.
    \item Family 1st no matter what.
    \item When my circle got smaller, my vision got clearer.
    
    There's strength in loyalty, not numbers.
    \item People who do not understand your silence will never understand your words.
    \item Behind every successful man there is a woman. - Fact
    
    Because women don't follow unsuccessful men. - Universal Truth
    \item \textit{What's your relationship status?}
    
    \textsc{Single}
    
    \textsc{In a relationship}
    
    \textsc{Crushing}
    
    \textsc{Married}
    
    \textsc{It's complicated}
    
    \textsc{Broken hearted}
    \item You will never speak to anyone more than you speak to yourself in your head, be kind to yourself.
    \item Find happiness in simple things.
    \item Happiness is not about getting all you want, it is about enjoying all you have.
    \item ``\textit{That woman has loved me skinny, she's loved me fat}.
    
    \textit{She's love me bald, she's loved me hairy}.
    
    \textit{That woman, I know, loves me}.
    
    \textit{So, I'm a lucky man}. - Tom Hanks about his wife, Rita Wilson
    \item \textbf{Be afraid of the quiet ones.}
    
    They are the ones who actually think.
    \item Never stop being a good person because of bad people.
    \item \textit{Good sense of humor}, \textit{dirty mind} and \textit{beautiful heart}.
    
    Deadly combination.
    \item I just want to
    \begin{itemize}
        \item take photos
        \item drink coffee
        \item travel the world
        \item meet new people
        \item be happy
    \end{itemize}
    \item \textit{Admit it}, sometimes you say your true feeling through jokes.
    \item ``\textit{1 man's \emph{I'm not ready} is another man's \textbf{I knew the second I saw her}.}'' - Meredith Marple
    \item Give$\ldots$ Yourself$\ldots$ Time.
    \item A relationship should be about helping each other deal with the stress the world brings.
    
    Not adding unnecessary stress to each others lives.
    \item I am not open to many people.
    
    I'm usually quiet and I don't really like attention.
    
    So if I like you enough to show you the real me, you must be very special.
    \item Why am I the type of person that still believes someone's a good person even when they've shown me in every way that they're not?
    \item No matter how good a person you are, you are evil in someone's story.
    \item ``\textit{I was raised to treat the janitor with the same respect as the CEO}.'' - Tom Hardy
    \item Next time a stranger talks to me when I am alone, I will just look at them shocked and just whisper quietly ``You can see me?''
    \item Sometimes I wish I was a bird, so I could fly over certain people and shit on their heads.
    \item \textsc{Listen} and \textsc{Silent} are spelled with the same letters.
    
    Think about it.
    \item I'm not bossy!
    
    I have skills$\ldots$
    
    leadership skills!!
    
    Understand?
    \item Don't be afraid to be open-minded.
    
    Your brain isn't going to fall out.
    \item Just because I give you advice, it doesn't mean I know more than you.
    
    It just means I've done more stupid shit.
    \item When you're laughing so hard \& you try to stop, but you look at the person and laugh again.
    \item I hold in a lot.
    
    When I'm in pain, I don't like to worry other people.
    
    No matter how hard I cry, or how much somebody asks, my answers will be, I'm fine.
    
    Even if it's not true.
    
    Do you do this too?
    \item I hate when people ask me ``\textit{Why are you so quiet?}''.
    
    Because I am.
    
    That's how I function.
    
    I don't ask others ``\textit{Why are you so noisy?}
    
    \textit{Why do you talk so much?}''
    
    It's rude.
    \item If you're not amazed by the stars on a clear night, then we won't work.
    \item ``The 2 most important days in your life are the day you are born and the day you find out why.'' - Mark Twain (1835--1910)
    \item Me: \textit{I'm blessed, I have someone who loves me \& fights for me}.
    
    Friend: \textit{Who?}
    
    Me: \textit{My white blood cells}$\ldots$
    \item ``\textit{Sometimes life has a cruel sense of humor, giving you the thing you always wanted at the worst time possible}.'' - Lisa Kleypas
    \item Loyal people get hurt the most.
    
    True or False?
    \item ``\textit{We all have inner demons to fight}.
    
    \textit{We call these demons `fear', and `hatred', and `anger'}.
    
    \textit{If you don't conquer them, then a life of a hundred years$\ldots$ is a tragedy}.
    
    \textit{If you do, a life of a single day can be a triumph}.'' - Yip Man
    \item Sometimes beauty depends on phone quality
    \item Sometimes people pretend you're a bad person so they don't feel guilty about the things they did to you.
    \item Me driving home from work knowing I'm only going home to eat and sleep so I can do it all again tomorrow.
    \item I said I'm able to work under pressure, not that I will die for your company.
    \item There's a message in the way a person treats you.
    
    \textbf{Just listen$\ldots$}
    \item Find someone who will \textit{love your soul} more than your body.
    \item They laugh at me because I'm different I laugh at them because they are all same.
    \item ``\textit{When a person tells you that you hurt them, you don't get to decide that you didn't}.'' - Louis C.K.
    \item Don't tell anyone ``I hate you'' directly, say ``You are the Monday of my life''.
    \item What the fuck$\ldots$!
    
    Psychotherapist is 1 word! 1 word!!! [Psycho the rapist]
    \item Imagine if instead of putting tombstones, we planted trees on grave, we would never have to worry about deforestation.
    \item Who else loves that earthy smell when rain hits the ground?
    \item ``\textit{If we fully understood the power of sexual energy we would refrain from casual sex}.
    
    \textit{Sex is sacred and to be shared with authentic purity of both partners}.
    
    \textit{Sexual energy is intense and can heal the universe through the vibration when 2 people join together and share their souls together}.
    
    \textit{DNA is exchanged during sex}.
    
    \textit{You imprint yourself on another}.
    
    \textit{Mindful sex is important}.'' - Diane L
    \item Half the world is starving.
    
    The other half is trying to lose weight.
    \item \textbf{Life is the most difficult exam.}
    
    Many people fail because they try to copy others, not realizing that everyone has a different question paper.
    \item Live a healthy lifestyle.
    
    Choose drinks with minimal amounts of sugar.
    \item I see humans but no humanity.
    \item ``\textit{I don't love studying}.
    
    \textit{I hate studying}.
    
    \textit{I like learning}.
    
    \textit{Learning is beautiful}.'' - Natalie Portman
    \item \textbf{Do things} that make you forget to check your phone.
    \item Note to self: stop missing people who don't even waste a second thinking about you.
    \item Dirty minded friends make life so much funnier.
    \item Yes, I overthink but I also \textit{over-love}$\ldots$
    \item It's never too late for a new beginning in your life.
    \item Do you ever think about going somewhere nobody knows you and starting a new life?
    \item I don't have a short temper, I just have a quick reaction to bullshit.
    \item Things take time.
    
    So just be patient.
    \item You learn nothing from life if you think you're right all the time.
    \item Constantly torn between ``if it's meant to be, it will be'' and ``if you want it, go and get it.''
    \item If you are not doing what you love, you are wasting your time.
    \item ``\textit{The problem is people are being hated when they are real, and are being loved when they are fake}.'' - Bob Marley
    \item Always love your friends from your heart, not from your mood or need.
    \item Followers will never know how hard the leader tries to create path.
    \item 1 day, you will be at the place you always wanted to be.
    
    Don't stop believing.
    \item 6 months from now you can be in a completely different space, mentally, spiritually and financially.
    
    Keep working and believing in yourself.
    \item Life's too short to tolerate shit that doesn't make you happy.
    \item The best sign of a \textit{happy relationship} is no sign of it on social media.
    \item Dear Beach, I think of you all the time$\ldots$
    \item If you have a friend with whom you can speak and open your heart anytime.
    
    Trust me you're blessed.
    \item \textbf{Hugs are actually so underrated}, especially those hugs that are so tight you can literally feel the other person's heartbeat and for a moment everything feels so calm and safe like nothing can hurt you.
    \item I never make the same mistake twice, I make it like 5 or 6 times, just to be sure.
    \item Don't chase, don't beg, don't be desperate, don't stress, just relax.
    
    When you relax, it will come to you.
    
    \textbf{Make you wants want you.}
    \item A \textbf{mistake} that makes you \textbf{humble} is better than an \textbf{achievement} that makes you \textbf{arrogant}.
    \item 1 loyal friend is worth more than 1000 fakes ones.
    \item Dear life, I've always been strong ever since.
    
    But can you please be easy on me sometimes?
    \item ``\textit{I've learned that home isn't a place, it's a feeling}.'' - Cecelia Ahern
    \item A soulmate is a friend who never leaves.
    \item Don't believe everything you think.
    \item 1 of the best pleasures in life$\ldots$ is to read a book in total silence.
    \item A good teacher is like a candle - it consumes itself to light the way for others.
    \item To have someone understand your mind is a different kind of intimacy.
    \item Listen, smile, agree, and then do whatever the fuck you were gonna do anyway.
    \item 1 of the best lessons you can learn in life is to master how to \textbf{remain calm.}
    \item ``\textit{When it hurts - observe}.
    
    \textit{Life is trying to teach you something}.'' - Anita Krizzan
    \item Take being called crazy as a Compliment.
    
    It means you've found the courage to be yourself when so many others have not.
    \item If a friendship lasts longer than 7 years, psychologists say it would probably last a lifetime.
    \item It's better to wait long than to marry wrong.
    \item ``\textit{If tomorrow, women woke up and decided they really like their bodies, just think how many industries would go out of business}.'' - Dr. Gail Dines
    \item Having someone who can handle all your moods is such a blessing.
    \item I'm at a point in my life where I just want to be very quiet.
    \item Some people act like they are trying to help you.
    \item \textit{You know what's attractive?}
    
    Seeing people change for that 1 person.
    
    Whether it's cutting down their drinking habits or to stop doing drugs and maybe even something small like to stop swearing.
    
    It just proves that when it comes down to it, they would do whatever it takes to see that person happy.
    
    Stopping habits is a hard thing to do and to see people actually doing that for someone's love is really attractive to me.
    \item Sisters are the 2nd version of Mother.
    
    She will give you advice, protect you and scold you, but will never tolerate if someone hurts you.
    \item If you have at least 1 person genuinely supporting you$\ldots$ \textbf{you're blessed.}
    \item \textit{12 things to remember.}
    \begin{itemize}
        \item[1.] The past cannot be changed.
        \item[2.] Opinions don't define your reality.
        \item[3.] Everyone's journey is different.
        \item[4.] Things always get better with time.
        \item[5.] Judgments are a confession of character.
        \item[6.] Overthinking will lead to sadness.
        \item[7.] Happiness is found within.
        \item[8.] Positive thoughts create positive things.
        \item[9.] Smiles are contagious.
        \item[10.] Kindness is free.
        \item[11.] You only fail if you quit.
        \item[12.] What goes around comes around.
    \end{itemize}
    \item I wish my life had background music so I could understand what the hell is going on.
    \item Sometimes the people with the greatest potential often take the longest to find their path because their sensitivity is a double edged sword - it lives at the heart of their brilliance, but it also makes them more susceptible to life's pains.
    
    Good thing we aren't being penalized for handing in our purpose late.
    
    The soul doesn't know a thing about deadlines.
    \item May your vibes shift the whole damn frequency of the room when you walk in.
    \item Scientists have found that babies sometimes cry at night to prevent their parents from making another baby.
    \item A strong woman will automatically stop trying if she feels unwanted.
    
    She won't fix it or beg, she'll just walk away.
    \item I found the key to happiness.
    
    Stay away from idiots.
    \item We are energy and energy never dies, it just changes form.
    \item As you get older, you'll realize that a \$300 watch and a \$30 watch both tell the same time.
    
    A Miachel Kors wallet and a Forever 21 wallet hold the same amount of money.
    
    A \$300,000 house and a \$100,000 house host the same loneliness.
    
    A Ford will also drive you as far as a Bentley.
    
    True happiness is not found in materialistic things.
    
    It comes from the love an laughter found with each other.
    
    Stay humble$\ldots$ the holes dug for us in the ground are all the same size.
    \item Once you feel you're avoided by someone, never disturb them again.
    \item ``\textit{There's an herb for every system, every organ, every gland, and every tissue of our body}.
    
    \textit{Mother nature has put medicine in our food}.'' - Bob Marley
    \item Her intuition was her favorite superpower.
    \item Facebook friends: 600
    
    Real friends: 20
    
    Hard time friends: 2
    \item If you're single
    \begin{itemize}
        \item[1.] You're either talking to someone
        \item[2.] Stuck on an ex
        \item[3.] Chasing someone who's taken
        \item[4.] Ignoring someone who wants you
        \item[5.] Too happy on your own
    \end{itemize}
    \item Can a boy and girl be a bestfriend?
    \item You don't always need a plan.
    
    Sometimes you just need to breathe.
    
    Trust.
    
    Let go, and see what happens.
    \item Everything depends on the way \textsc{You} see!
    \item Some people are so Fake that if you look behind their necks you'll see: \texttt{Made in China}.
    \item ``\textit{The coolest people I've ever met have the most colorful pasts, they've lived lives of risk, made bad choices, learned lessons, explored, and they're not afraid of being real}.
    
    \textit{Tattered tapestries woven of similar threads, they've my kind of people}.
    
    \textit{My favorite shades of crazy}.'' - Stephen L. Lizotte
    \item ``\textit{Take a \textbf{lover} who looks at you like maybe you are \textbf{magic}.}'' - Frida Kahlo
    \item A Japanese legend says that if you \textbf{can't sleep at night} it's because you're awake in someone else's dream.
    \item Here's what's cool:
    \begin{itemize}
        \item[1.] Say ``thank you.''
        \item[2.] Apologizing when you're wrong.
        \item[3.] Showing up on time.
        \item[4.] Being nice to strangers.
        \item[5.] Listening without interrupting.
        \item[6.] Admitting you were wrong.
        \item[7.] Following your dreams.
        \item[8.] Being a mentor.
        \item[9.] Learning and using people's names.
        \item[10.] Holding doors open for others.
    \end{itemize}
    \item \textit{Some say I'm too sensitive}, but truth is I just feel too much.
    
    Every word, every action and every energy goes straight to my heart.
    \item My favorite thing is when people remember little things I told them.
    
    Like seriously?
    
    You actually listened to me thank you.
    \item This is Jane.
    
    Jane is in a relationship.
    
    Jane doesn't post on social media about how much she loves her partner.
    
    She tells him this in person.
    
    She doesn't mention every little insignificant thing they do.
    
    Jane knows that nobody gives a damn.
    
    Jane is smart.
    
    Be like Jane.
    \item No one is more real than your mom.
    
    She'll talk shit on your face but pray for you behind your back.
    \item \textsf{A picture taken somewhere on the Indian highway: 2 elephants reach out in a brief moment of love and bonding before being taken away from each other for a lifetime of serving man.}
    \item I've always been someone who looks `\textbf{too deep}' into something or someone.
    
    That's because I realized from a young age that there's always more than what meets the eye.
    \item ``\textit{Lie your life for you not for anyone else}.
    
    \textit{Don't let the fear of being judged, rejected or disliked stop you from being yourself}.'' - Sonya Parker
    \item ``\textit{I don't trust someone who is nice to me but rude to the waiter, because they would treat me the same way if I were in that position}.'' - Muhammad Ali
    \item 2 constant issues of my life:
    \begin{itemize}
        \item[1.] I have nothing to wear.
        \item[2.] I have no space to keep my clothes.
    \end{itemize}
    \item Maybe you don't see people looking at you because you aren't looking at them.
    
    Maybe you don't hear all the good things people say about you because you're too focused on the bad.
    
    \textbf{Maybe you're a lot more} wonderful, beautiful and special than you ever give yourself credit for.
    \item Beautiful people are bad at math.
    \item A mistake which makes you Humble is much better than an Achievement that makes you Arrogant.
    \item ``\textit{Jobs fill your pocket}.
    
    \textit{Adventures fill your soul}.'' - Jaime Lyn Beatty
    \item It's time to start living the life you've imagined.
    \item My mind is more talkative than my mouth.
    \item Why do people call?
    
    Why can't they just text!?!
    \item ``\textit{You think you've see her naked because she took her clothes off?}
    
    \textit{Tell me about her dreams}.
    
    \textit{Tell me what breaks her heart}.
    
    \textit{What is she passionate about, and what makes her cry?}
    
    \textit{Tell me about her childhood}.
    
    \textit{Better yet, tell me 1 story about her that you're not in}.
    
    \textit{You've seen her skin, and you've touched her body}.
    
    \textit{But you still know as much about her as a book you once found, but never got around to opening}.'' - Dominic Matthew Jackson
    \item I don't know how to be anything other than intense.
    
    I don't know how to experience without feeling too much and thinking too much.
    
    I don't know how to sit still and quiet my mind and just be.
    
    I am always searching, always questioning, struggling to find the meaning in everything.
    
    I am passionate and I am deep, and even if I am misunderstood, I am finally okay with that.
    \item And when he asked what it was that he could give to her that she'd never had before, her answer was so simple.
    
    \textbf{Consistency}, she said.
    
    ``If you want to give me something that no man has ever given to me, then don't give me mixed signals nor mixed emotions that leave me wondering.
    
    I'm tired of wondering.
    
    If you're going to be here with me, then be here, if you ever feel the need to leave me, then stay gone.
    
    \textbf{All I need at this point from someone is consistency.}''
    \item ``\textit{I'm starved for connection, not attention}.'' - Bridgett Devoue
    \item ``If you have been brutally broken but still have the courage to be gentle to other living beings, then you're a badass with a heart of an angel.'' - Keanu Reeves
    \item Nothing teaches better than this trio: the fears, the tears, the years, \textbf{\{the golden trio\}}.
    \item Ask a woman what kind of a man she wants then sit and listen how she explains characteristics of non living things.
    \item \textit{7 reasons you should be thankful for having a \textbf{big sister}}
    \begin{itemize}
        \item[1.] She's your go-to person for everything.
        \item[2.] You may hate each other at times.
        
        But your love is unimaginable.
        \item[3.] You know she will always protect you with everything she has.
        \item[4.] She's your best secret keeper.
        \item[5.] Her wardrobe is your wardrobe.
        \item[6.] She has seen the craziest side of you but still loves you.
        \item[7.] Nothing can replace the beautiful bond you 2 share.
    \end{itemize}
    \item To my best friends,
    
    1 day, all of us will get separated from each other.
    
    We will miss our conversations of everything and nothing and the dreams we had.
    
    Days, months and years will pass until this contact becomes rare.
    
    1 day, our children will see our pictures and ask: ``\textit{Who are these people?}''.
    
    And we will smile with invisible tears because a heart is touched with a strong word and you will say:
    
    ``\textit{It was them that I had the best days of my life with}.''
    \item The true mark of maturity is when somebody hurts you, and you try to \textbf{understand their situation} instead of hurting them back.
    \item Dear stress: let's break up.
    \item ``\textit{Expectation is the root of all heartache}.'' - William Shakespeare
    \item Fake people have an image to maintain.
    
    Real people just don't give a fuck.
    \item You want to come into my life, the door is open.
    
    You want to get out of my life, the door is open.
    
    Just 1 request$\ldots$ \textbf{don't stand at the door}, you're blocking the traffic.
    \item ``\textit{So, if you are too tired to speak, sit next to me for I, too, am fluent in silence}.'' - R. Arnold
    \item Fall in love with the person who enjoys your madness.
    
    Not an idiot who forces you to be normal.
    \item Don't be afraid to love again.
    
    Not everyone is like your ex.
    \item I hate remembering good times I had with people who ended up being really shitty to me.
    \item The most beautiful part to loving a guarded girl is this: when she lets you in, it's not because she needs you.
    
    She stopped needing people a long time ago.
    
    It's because she wants you.
    
    And that - that is the purest love of all.
    \item You don't always need a plan.
    
    Sometimes you just need to breathe, trust, let go, and see what happens.
    \item What consumes your mind, controls your mind.
    \item I have 3 different personalities:
    \begin{itemize}
        \item the one where I'm out-going and loud,
        \item the one where I'm shy and quiet as fuck,
        \item the one where I hate everyone and every little thing bothers me.
    \end{itemize}
    \item People are the pretties when they talk about something they really love, with passion in their eyes.
    \item Be happy.
    
    Not because everything is good, but because you can see the good side of everything.
    \item I'm 100\% OK with having like 3 friends because people suck.
    \item \textbf{Brother}: a person who is there when you need him; someone who picks you up when you fall; a person who sticks up for you when no one else will;
    
    \textbf{a brother is always a friend.}
    \item I don't care if I'm selfish.
    
    After putting people 1st for the longest time and being disappointed I deserve to do whatever makes me feel happy.
    \item Sometimes, having coffee with your best friend is all the therapy you need.
    \item Even the smallest lie can break the biggest trust.
    \item Because at the end of the day you're the person I want to come home to.
    
    You're the person I want to tell how my day went.
    
    You're the person I want to share my happiness, sadness, frustration, and success with.
    \item Don't expect to get what you give.
    
    Not everyone has a heart like you.
    \item You don't really know someone until you say \textsc{No} to them.
    \item So you want to be happy?
    
    Then stop letting the smallest things ruin your whole entire day.
    
    If you're bored with your daily routine, do something unexpected.
    
    Stop complaining about how alone you are when you're surrounded by people who actually care about you.
    
    Forget all the drama and let go of all the grudges you've been holding.
    
    Stop wasting time lingering over all that you could have, should have and would have done.
    
    Stop spending your days thinking of how much better you could do; stop longing for something that has been and always will be out of your reach.
    
    Just live the days as they come.
    
    Wake up every morning and smile at the wonderful day that awaits you.
    
    Take a risk for once.
    
    Let yourself be happy, because you deserve it.
    \item ``\textit{1 day you will wake up and there won't be any more time to do the things you've always wanted}.
    
    \textbf{\textit{Do it now}.}'' - Paulo Coelho
    \item Fall in love with the person who enjoys your madness.
    
    Not an idiot who forces you to be normal.
    \item Always be in love with a soul, not a face.
    \item And if they want to leave, hold the door open for them.
    \item Never lie to girls with big foreheads.
    
    They never forget easily.
    
    They're hiding 64GB drive in there.
    \item ``\textit{Before you diagnose yourself which \textbf{depression} or \textbf{low self-esteem}, 1st make sure that you are not, in fact, just \textbf{surrounded by assholes}.}'' - William Gibson
    \item I need to spend a good night \textit{laughing too loud} and \textit{eating too much} with my best friends.
    \item ``\textit{Travel and tell no one}.
    
    \textit{Live a true love story and tell no one}.
    
    \textit{Live happily and tell no one}.
    
    \textit{People ruin beautiful things}.'' - Kahlil Gibran
    \item Psychology says people who love black color have the most colorful minds.
    \item Some day we will find what we are looking for.
    
    Or maybe not, maybe we'll find something much greater than that.
    \item Stop looking for a partner.
    
    Focus on your goals and rebuilding your life.
    
    The right person will eventually find their way to you.
    \item Without rain nothing grows.
    
    Learn to embrace the storms in your life.
    \item Raise your hand: if you have a bad habit of laughing at serious moments.
    \item Never reply when you are angry.
    
    Never make a promise when you are happy.
    
    Never make a decision when you are sad.
    \item Some people will only ``love you'' as much as they can use you.
    
    Their ``loyalty'' ends where the benefits stop.
    \item Brain Cell $\sim$ The Universe
    
    What if we're living in a brain cell of another creature?
    \item If you try and fail, congratulations because most people don't even try.
    \item I'm not impressed by money, followers, degrees and titles.
    
    I'm impressed by kindness, integrity, humility and generosity.
    \item Instead of wiping away your tears, wipe away the people who made you cry!
    \item Girls, if a guy
    \begin{itemize}
        \item remembers your birthday
        \item knows what you enjoy
        \item saves your pictures
        \item understands your family \& friends.
    \end{itemize}
    This guy is not your man.
    
    This guy is Mark Zuckerberg.
    \item Lucky are those who find a true loyal friend in this fake world.
    \item I am at the stage of my life where I don't force someone to stay in life.
    
    If you wanna stay you can stay.
    
    If you wanna leave you can leave.
    \item I feel like water solves all problems.
    
    Want to lose weight?
    
    Drink water.
    
    Clear face?
    
    Use water.
    
    Tired of a person?
    
    Drown them.
    \item Do you ever listen to a song and remember exactly what life was like when you 1st heard it?
    \item It is not the \textit{lie} that bothers me.
    
    It's the \textit{insult to my intelligence} that I find offensive.
    \item ``\textit{Okay I know that there are terrible terrible people out there but listen:}
    
    \textit{I also know that there are people who stop and smile at tiny plants growing out of sidewalk cracks, people who laugh so loud they snort, people who compliment others randomly, people who take pictures of their friends because they love seeing their friends happy, people who ramble about things that they're passionate about, people who blush and stutter, people who are kind, people who are warm, people who love and love and love and love.}'' - queens-bees
    \item You know what your problem is?
    
    You get attached, fast.
    And once you're attached to someone, you do everything you can to please them and make them happy.
    
    It's never been about what you want, it's always everyone's needs before your own.
    
    You give out too many chances to people, who quite frankly, do not deserve them.
    
    They take advantage of you, and you become a pushover.
    
    But you're okay with that, because they're in your life and that's all you ever really wanted.
    
    And even if they screw you over, you'll still be there for them.
    
    Because that's you, that who you are.
    
    Once you get attached to someone, they capture your heart and they always have a place there.
    
    And that is why it's so hard for you to let him go.
    \item Strength does not come from winning.
    
    Your struggles develop your strengths.
    
    When you go through hardships and decide not to surrender, that is strength.
    \item Sometimes, the girl who is always there for everyone else, needs someone to be there for her.
    \item \textsc{Thought} for the day: A person who feels appreciated will always do more than is expected.
    \item It's easy to make 15 friends in 1 year.
    
    But keeping 1 friend for 15 years is special.
    \item \textsc{Value Yourself!}
    
    Losing your love is okay.
    
    Losing your best friend is okay.
    
    But losing yourself in the process of getting them back to you is not okay.
    
    Value yourself, You are important.
    \item ``\textit{Once a year, go somewhere you're never been before}.'' - Dalai Lama
    \item Everybody isn't your friend.
    
    Just because they hang out around you and laugh with you doesn't mean they're there for you.
    
    Just because they say they got your back, doesn't mean they won't stab you in it.
    
    People pretend well.
    
    Jealousy sometimes doesn't live far.
    
    So know your circle.
    
    At the end of the day, real situations expose fake people, so pay attention.
    \item Be yourself.
    
    People don't have to like you, and you don't have to care.
    \item I've found the key to happiness.
    
    Stay away from assholes.
    \item ``\textit{I often think that the night is more alive than the day}.'' - Vincent Van Gogh
    \item Life is too short to be \textit{normal}.
    
    Stay \textit{weird}.
    \item You can tell a lot about a person by what's on their playlist.
    \item I'd rather be alone, than around chaos and confusion.
    
    \textbf{Silence beats drama any day.}
    \item ``Have the courage to be disliked.'' - Bruce Lee
    \item I think weekends are made in China: they don't last long.
    \item When you're trying to love people but you're also an introvert and have boundaries.
    \item Your brother will never say he loves you but he loves you more than anyone else in this world.
    \item I honestly love being around positive people.
    
    You're not judged, there's no drama, everyone just wants to relax and have a nice time.
    \item When you truly don't care what the fuck anyone thinks of you, you have reached a dangerously awesome level of freedom$\ldots$!!
    \item So many years of education yet nobody ever taught us how to love ourselves and why it's so important.
    \item You have to train your mind to be stronger than your emotions or else you'll lose yourself every time.
    \item Girls are so extra dramatic when they get mad.
    
    Everything is ``Bye Delete my number'' ``I'm done. I hate you. You won't hear from me no more. Have a blessed life. Leave me alone'' and then they sit there waiting for a reply.
    \item Stop explaining yourself to other people.
    
    You owe no one any explanation of what you do.
    
    Your life is yours, not theirs.
    \item ``\textit{Great minds discuss ideas}.
    
    \textit{Average minds discuss events}.
    
    \textit{Small minds discuss people}.'' - Eleanor Roosevelt
    \item Sir William Golding regarding women:
    \begin{quotation}
        I think women are foolish to pretend they are equal to men, they are far superior and always have been.
        
        Whatever you give a woman, she will make it greater.
        
        If you give her sperm, she'll give you a baby.
        
        If you give her groceries, she'll give you a meal.
        
        If you give her a smile, she'll give you heart heart.
        
        She multiplies and enlarges what is given to her.
        
        So, if you give her any crap, be ready to receive a ton of shit.
    \end{quotation}
    \item I truly appreciate kindness.
    
    I appreciate a quick message.
    
    I appreciate those who ask if I'm okay.
    
    I appreciate people checking up on me.
    
    I appreciate every single person in my life who has tried to brighten my days.
    
    \textit{It's the little things that matter the most}.
    \item An arrow can only be shot by pulling it backward.
    
    When life is dragging you back with difficulties, it means it's going to launch you into something great.
    
    So just focus, and keep aiming.
    \item Too much perfection is a mistake.
    \item \textbf{Left handers:} are creative, smart and have a big heart.
    
    Who do you know is a lefty?
    \item Time won't make you forget, it will make you grow and understand things.
    \item ``\textit{Stop thinking that just because you have a college degree it makes you smart}.
    
    \textit{I know a lot of people who have a driver's license but they can't drive for shit}.'' - Ray Velcoro
    \item Sometimes the hardest battle is against yourself.
    \item Being raised right doesn't mean you don't drink, party, smoke or use curse words.
    
    Being raised right is how you treat people, your manners and respect.
    \item A group of Engineering students and their teacher were given free airplane tickets to go on a holiday.
    
    Once on the plane the Captain announced that they were on the plane the students had built.
    
    Everyone freaked out and rushed out of the plane, except for the teacher who stayed there with calm.
    
    When the flight attendant asked why he hadn't left he responded ``I know the abilities of my students quite well, this shit won't even start''.
    \item If you have a friend with whom you can speak and open your heart anytime$\ldots$
    
    Trust me, you're blessed.
    \item Shout out to the women who fix another woman's crown without telling the world it was crooked.
    \item I respect people with real intentions who tell me the truth.
    \item Never be afraid to say what you feel.
    \item I'll choose honesty over perfection every single time.
    \item Enjoy every moment with your partner$\ldots$
    
    Because you're gonna block each other 1 day.
    \item Find a woman with a \textit{brain}.
    
    They all have \textit{vaginas}.
    \item Even the strongest feelings expire when ignore and taken for granted.
    \item \textit{The best advice}$\ldots$
    
    Save money every week!
    
    It doesn't matter how much.
    
    Just save!
    
    Listen to your parent's advice$\ldots$ at the end of the day they are the only ones who want the best for you.
    
    Choose your friends wisely as you are the product of your environment.
    
    Learn to be alone and independent.
    
    It's a skill few master.
    
    Educate yourself - read, read, read.
    
    Be healthy \& look after your body.
    
    Don't wait for someone to love you; learn to love \textsc{Yourself} 1st.
    
    You'll be OK.
    \item I need a rest.
    
    A long, long, long, long, long, long long rest.
    \item I am strong, but I am tired.
    \item Always remember, someone's effort is a reflection of their interest in you.
    \item ``\textit{Education is not the learning of facts but the training of the mind to think}.'' - Albert Einstein
    \item In life, what you really want; will never come easy.
    \item Always hope but never expect.
    \item Never let your friends feel lonely disturb them all the time.
    \item The problem with closed minded people is their mouth is always open.
    \item Please don't be an asshole to me, because then I may have to be an asshole to you, and I'm way better at being an asshole than you are.
    \item A well read woman is a dangerous creature.
    \item Even devil, ghosts, monsters, and witches are less dangerous than fake human beings.
    \item Never change to be accepted by others.
    
    Stay weird.
    \item Take the risk or lose the chance.
    \item 1 day, someone will love you the way you deserve to be loved and you won't have to fight for it.
    \item The best revenge is no revenge.
    
    Move on.
    
    Be happy.
    \item It's not that I can't see what they see.
    
    It's that I see what they can't.
    \item Respect people's privacy$\ldots$
    
    Don't ask them why they aren't married yet, when they going to have a baby, why they are separated or divorced, why they left their job, why they like to watch movies alone, why they cry listening to someone's sad story, why they don't go out more$\ldots$
    
    Unless they decide to share it with you, it's none of your business$\ldots$!
    \item A woman is \textbf{unstoppable} after she realizes she deserves better.
    \item Time has a way of showing us what really matters in life.
    \item A new study links drinking coffee to a longer lifespan.
    
    Awesome!
    
    I'm gonna be immortal.
    \item People who have been single for too long are the hardest to love.
    
    They have become so used to being single, independent and self-sufficient that it takes something extraordinary to convince them that they need you in their life.
    \item You don't have to be crazy to be my friend.
    
    \textit{I'll train you}.
    \item When you're wrong admit it.
    
    When you're right be quiet.
    \item A good wife can bring balance to your life.
    \item Twinkle twinkle little star.
    
    Singles are the superstar.
    \item ``\textit{There are rare people who will show up at the right time, help you through the hard times and stay into your best time$\ldots$}
    
    \textit{Those are the keepers}.'' - Nausicaa Twila
    \item When a storm is coming, all other birds seek shelter.
    
    The Eagle alone, avoids the storm by flying above it.
    
    So, in the storms of life$\ldots$
    
    May your heart soar like an Eagle.
    \item People say a lot.
    
    So, I watch what they do.
    \item Life is not about being rich, being popular, being highly educated or being perfect.
    
    It's about being real, humble, and kind.
    \item Isn't it scary, knowing that any time could be the last time you talk to someone?
    
    Always keep that in mind.
    \item I am too old or worry about who likes me and who dislikes me.
    
    I have more important things to do.
    
    If you love me, I love you.
    
    If you support me, I support you.
    
    If you hate me, I don't care.
    
    Life goes on with or without you.
    \item They say good things take time.
    
    That's why I'm always late.
    \item Silent people have the loudest minds.
    \item ``\textit{I have a theory that as long as you have 1 good friend, \textbf{1 real friend}, you can get through anything}.'' - Dana Reinhardt
    \item Your truest friends are the people who don't walk out the door when life gets hard.
    
    They actually pour some coffee and pull up a chair.
    \item Life is short.
    
    Spend it with people who make you laugh and feel loved.
    \item Nothing is more attractive than loyalty.
    \item He said, ``there are only 2 days in the year that nothing can be done.
    
    One is called \textit{yesterday} and the other is called \textit{tomorrow}, so today is the right day to love, believe, do and mostly live.'' - Dalai Lama
    \item Put me in a room with the same people who talk shit about me and watch how friendly they become.
    \item Give yourself enough respect to walk away from someone who doesn't see your worth.
    \item No one owns th water.
    
    No one owns the land.
    
    No one owns the oceans.
    
    No one owns the sand.
    
    These are given by our Mother:
    
    \textit{The Planet provides for free}.
    
    Only by the hands of the greedy, does the Earth require a fee.
    \item A wise man can always be found alone.
    
    A weak man can always be found in a crowd.
    \item We are all in the same game, just different levels, dealing with the same hell, just different devils.
    \item Tell me your deepest fantasy.
    
    8--12 hours of uninterrupted sleep.
    \item You know I'm comfortable with you when:
    \begin{itemize}
        \item[1.] I can be weird with you.
        \item[2.] I sing whatever song comes into my mind.
        \item[3.] I say what's on my mind.
        \item[4.] I talk nonsense.
    \end{itemize}
    \item Never expect.
    
    Never assume.
    
    Never ask.
    
    And Never demand.
    
    Just let it be, if it's meant to be, it will happen.
    \item Usually, the people with the best advice are the ones with most problems.
    \item You can't change how people feel about you, so don't try.
    
    Just live your life and be happy.
    \item Ladies, Please stop wasting your time looking for Mr. Right
    
    Just find Mr. Left and drag that idiot to the right.
    \item \textit{People who have been single for too long are the hardest to love}, because they have become so used to being single, independent, and self-sufficient that it takes something extraordinary to convince them that they need you in their life.
    \item Before you ask why someone hates you, ask yourself why you give a fuck.
    \item Friends who do not have a lot of pictures together have a lot of memories together!
    \item People who correct your grammar during a conversation always die alone.
    \item ``\textit{Why worry?}
    
    \textit{If you have done the very best you can}.
    
    \textit{Worrying won't make it any better}.
    
    \textit{If you want to be successful, respect 1 rule}.
    
    \textbf{\textit{Never let failure take control of you}.}
    
    \textit{Everybody has gone through something that has changed them in a way that they could never go back to person they once were}.
    
    \textit{Relationships are like electric currents, \textbf{wrong connection will give you shock throughout your life.}}
    
    \textbf{\textit{But the right one will light up your life}.}'' - Leonardo Dicaprio
    \item Everyone you meet is fighting a battle you know nothing about.
    
    Be kind.
    
    Always.
    \item Mood: wanna move to another city and start a new life.
    \item Don't be afraid to love again, not everyone is like your ex.
    \item \textbf{Stay single} until you meet the person that will take care of you the most.
    \item Once in your life you'll come across \textbf{a special person} that makes you happy, supports you, and makes you a better person.
    
    \textbf{Don't let them go.}
    \item If a friendship lasts longer than 7 years, psychologists say it will last a lifetime.
    \item ``\textit{If there's just 1 piece of advice I can give you, it's this when there's something you really want, fight for it, don't give up no matter how hopeless it seems}.
    
    \textit{And when you've lost hope, ask yourself if 10 years from now, you're gonna wish you gave it just 1 more shot, because they best things in life, they don't come free}.'' - Shonda Rhimes
    \item Love is temporary but friendship is forever$\ldots$
    \item \textbf{Always help someone.}
    
    You might be the only one that does.
    \item ``\textit{Stay kind}.
    
    \textit{It makes you beautiful}.'' - Najwa Zebian
    \item ``\textit{When she's important to you, you won't make excuses}.
    
    \textit{You'll make sacrifices}.'' - R. Solo
    \item Friendship isn't about who you've known the longest, it's about who walked into your life, said ``\textit{I'm here for you}'' \textit{and proved it}.
    \item  May all the negative energy trying to bring you down come to an end.
    
    May the dark thoughts, the overthinking, and the doubt exit your mind right now.
    
    May clarity replace confusion.
    
    May hope replace fear.
    
    May blessings and success fill your life.
    
    May the light of your spirit shine so bright that nothing can dim your glow.
    
    Shine one.
    \item Common sense is not a gift, it's a punishment.
    
    Because you have to deal with everyone who doesn't have it.
    \item A wise woman once said ``fuck this shit'' and she lived happily ever after.
    \item \textit{Challenge}: When a negative thought enters your mind, think 3 positive ones.
    
    Train yourself to flip the script.
    
    Accept the challenge!
    \item Strangers can become best friends just as easy as best friends can become strangers.
    \item Hold on to every genuine person you find.
    
    This generation has people driven by \textit{ego}, \textit{money} and \textit{status}.
    
    As a result good souls are ruined daily.
    
    Keep your head up and be conscious of the energy you give out and connect with.
    \item The quality of your thinking determines the quality of your life.
    \item A relationship means that you come together to \textbf{make each other better.}
    
    Believe in each other.
    
    Support each other.
    
    Build each other.
    
    \textbf{Be their peace, not their problem.}
    \item May everything I post confuse you until you learn to mind your business.
    \item People should seriously stop expecting normal from me$\ldots$
    
    We all know it's never gonna happen!
    \item Choose a partner who is good for you.
    
    Not good for your parents.
    
    Not good for your image.
    
    Not good for your bank account.
    
    Choose someone who's going to make your life emotionally fulfilling.
    \item Not all classrooms have 4 walls.
    \item Finding friends with the same mental disorder is \textit{priceless}.
    \item If you stay, stay forever.
    
    If you go, do it today.
    
    If you change, change for the better.
    
    And if you talk, make sure you mean what you say.
    \item Hugs are actually so underrated, especially those hugs that are so tight you can literally feel the other person's heartbeat and for a moment everything feels so calm and safe like nothing can hurt you.
    \item ``\textit{Hell is empty and all the devils are here}.'' - William Shakespeare
    \item Your relationship doesn't need to make sense to anyone, except you and your partner.
    
    It's a relationship.
    
    Not a community project.
    \item Never say ``That won't happen to me.''
    
    Life has a funny way of proving us wrong.
    \item When I put my earphones on, I enter the world of my own.
    \item I love people who gossip behind my back.
    
    That's exactly where they belong.
    
    Behind my back.
    \item Psychology says if you can't sleep at night it's all because you slept in the afternoon.
    
    Not everything is about love.
    \item You often feel tired, not because you've done too much, but because you've done too little of what sparks a light in you.
    \item Once in a lifetime we meet someone who changes everything.
    \item Being alone has a power that very few people can handle.
    \item Having a real friend is \textit{a blessing}.
    \item ``\textit{Do you know what happens whne you decide to stop worrying about what other people might think of you?}
    
    \textit{You get to dance}.
    
    \textit{You get to sing}.
    
    \textit{You get to laugh loudly, paint, write, and create}.
    
    \textit{You get to be yourself}.
    
    \textit{And you know what?}
    
    \textit{Some people won't like you}.
    
    \textit{Some will laugh or mock or point out flaws$\ldots$ but it just won't bother you all that much}.'' - Doe Zantamata
    \item Your brother will never say he loves you but he loves you more than anyone else in this world.
    \item If it involves mountain, breakfast food, coffee or campfires, I'm in.
    \item \textit{Don't be afraid of being different}, be afraid of being the same as everyone else.
    \item Have you ever given someone else a motivational speech while you were hurting on the inside?
    
    That's \textbf{strength}!
    \item There's always 3 best friends.
    
    The wild one, the one who got that IDGAF attitude, and the cute, innocent one.
    \item Without rain nothing grows.
    
    Learn to embrace the storms in your life.
    \item You don't always need a logical reason for doing everything in your life.
    
    Do it because you want to; because it's fun; because it makes you happy.
    \item Marry a girl who loves food more than expensive gifts.
    \item ``\textit{No amount of security is worth the suffering of a mediocre life chained to a routine that has killed your dreams}.'' - Maya Mendoza
    \item If you fix a time with your best friend to meet at 6 p.m and he is there at 6 p.m, then he is not your best friend.
    \item ``\textit{How to tell somebody is genuinely interested in you}:
    
    \textit{If you removed all of your effort from the equation would any communication remain between you?}
    
    \textit{If not, there is nothing there, and you deserve better}.'' - Beau Taplin
    \item ``\textit{Working hard for something we don't care about is called stress;}
    
    \textit{Working hard for something we love is called passion}.'' - Simon Sinek
    \item Life is not about the quantity of friends you have.
    
    It is about the quality of friends you have.
    \item I am a \textbf{strong} person.
    
    But once in a while, I need someone to hug me and tell me, ``Everything is going to be okay.''
    \item Tell the truth and run$\ldots$
    \item There is honestly no reason to lie to me.
    
    I'm too understanding.
    
    I get it.
    
    I get life
    
    I know that shit happens.
    
    Just be straight up with me.
    \item I love the smell of freshly brewed coffee in the morning.
    
    And I love the sound of no one talking to me while I drink it.
    \item Never stop working on yourself.
    \item It's amazing when your humor is exactly the same as someone else's and you both just spend the whole day laughing at everything you both say.
    \item Not every friend request is a friend request.
    
    Some are surveillance cameras.
    \item ``\textit{Magic happens when you don't give up, even though you want to}.
    
    \textit{The universe always falls in love with a stubborn heart}.'' - JmStorm
    \item Sometimes I delete my own posts because I'm not the same person I was 4 minutes ago.
    \item Someday someone will break you so badly that you'll become unbreakable.
    \item The right one will know all your weaknesses and never use them against you.
    \item ``\textit{No Girl will choose 6 pack over 6 cars}$\ldots$
    
    \textit{So stop going to the gym and Go to Work}. - Robert Mugabe (Former PM of Zimbabwe)
    \item ``\textit{When you find people who not only tolerate your quirks but celebrate them with glad cries of ``Me too!'' be sure to cherish them}.
    
    \textit{Because \textbf{those weirdos are your tribe}.}'' - A.J. Downey
    \item When you're happy, you enjoy the music but when you're sad, you understand the lyrics.
    \item When a Man makes money, he feels like he wants more women, but a Woman makes money she feels like she doesn't need a man.
    \item Dear Music, Thanks for always clearing my head, healing my heart, and lifting my spirits.
    \item I have these 2 neighbors, and they're married, and they gotta be in their late 30s, and I'm making dinner, and I look out the window, and they're running around outside in their pajamas and bare feet with water pistols soaking each other and laughing so loud it made me realize \textbf{I'm wasting so much time trying to make relationships perfect} when all that's really needed is someone who will laugh with me for the rest of my life.'' - on-cloud-mine
    \item ``\textit{Since I've been spending my time in nature with animals I found out that, we are the wild savages and they are the ones who are actually civilized}.'' - Jim Carrey
    \item The world is full of monsters with friendly faces and angels full of scars.
    \item Solve the problem or leave the problem.
    
    But do not sleep with the problem.
    \item When someone appears in your dreams, it's because that person misses you.
    \item I like poems
    
    museums
    
    real conversation
    
    genuine people
    
    Friday nights
    
    fantasies
    
    I like stories
    
    intimacy
    
    I like soul.
    \item \textsc{Decide}
    
    ``\textit{So, do it}.
    
    \textit{Decide}.
    
    \textit{Is this the life you want to live?}
    
    \textit{Is this the person you want to love?}
    
    \textit{Is this the best you can be?}
    
    \textit{Can you be stronger?}
    
    \textit{Kinder?}
    
    \textit{More Compassionate?}
    
    \textit{Decide}.
    
    \textit{Breathe in}.
    
    \textit{Breathe out and decide}.'' - Meredith Grey
    \item Be alone.
    
    Eat alone.
    
    Take yourself on dates.
    
    Sleep alone.
    
    Take the time to understand and love who you are.
    \item Girls with glasses are damn cute.
    \item A satisfied life is better than successful life.
    
    Because our success is measured by others, by our satisfaction is measured by our own \textit{soul}, \textit{mind} and \textit{heart}.
    \item Damaged people are \textsc{Strong} because they know how to survive.
    \item 3 most beautiful things in the world:
    \begin{itemize}
        \item[1.] Morning Sleep
        \item[2.] Afternoon Sleep
        \item[3.] Night Sleep
    \end{itemize}
    \item Everyone needs someone who will call and say, ``Get dressed, we're going on an adventure.''
    \item Life is short.
    
    Make sure you spend as much time as possible on the Internet arguing with strangers about politics.
    \item \textsc{Simple formula for living}
    \begin{itemize}
        \item Live beneath your means.
        \item Return everything you borrow.
        \item Stop blaming other people.
        \item Admit it when you make mistake.
        \item Give clothes not worn to charity.
        \item Do something nice and try not to get caught.
        \item Listen more; talk less.
        \item Everyday take a 30 min. walk.
        \item Strive for excellence, not perfection.
        \item Be on time.
        
        Don't make excuses.
        \item Don't argue.
        
        Get organized.
        \item Be kind to unkind people.
        \item Let someone cut ahead of you in line.
        \item Take time to be alone.
        \item Cultivate good manners.
        \item Be humble.
        \item Realize and accept that life isn't fair.
        \item Know when to keep your mouth shut.
        \item Go an entire day without criticizing anyone.
        \item Learn from the past.
        
        Plan for the future.
        \item Live in the present.
        \item Don't sweat the small stuff.
        \item It's all small stuff.
    \end{itemize}
    \item I would do it again!
    \item Listen to your best friends, sometimes they know you more than you know yourself.
    \item Notice the people who make an effort to stay in your life.
    \item Success is not always what you see.
    \item I don't care what other people think of me I enjoy my life with my own rules.
    \item Which pill would you take?
    \begin{itemize}
        \item \textsc{Red}: Love \& Happiness
        \item \textsc{Blue}: Money \& Power
    \end{itemize}
    \item Appreciate those who don't give up on you.
    \item Do not correct a fool or he will hate you.
    
    Correct a wise man and he will appreciate you.
    \item There comes a time in life, when you walk away from all the drama and people who create it.
    
    Surround yourself with people who make you laugh, forget the bad, \textbf{and focus on the good.}
    
    Love the people who treat you right.
    
    Pray for the ones who don't.
    
    Life is too short to by anything but happy.
    
    Falling down is part of life, getting back up is living.
    \item Don't be a parrot, in life, be an Eagle.
    
    A parrot talks way too much but can't fly high but an eagle is silent \& has the power to touch the sky.
    \item \textbf{The Full circle of life}: That moment when you realize the beginning and the end is same.
    \item Patience is when you're supposed to get mad, but you choose to understand.
    \item The goal is to laugh forever with someone you take seriously.
    \item You are a prisoner of your own mind.
    \item ``\textit{I think the saddest people always try their hardest to make people happy}.
    
    \textit{Because they know what it's like to feel \textbf{absolutely worthless} and they don't want anybody else to feel like that}.'' - Robin Williams (1951--2014)
    \item Someone who takes the time to listen when we are at our lowest is rare and should be valued.
    \item If you lose a friend because you're honest, it wasn't a good friend.
    \item You know what's sexy?
    
    A real conversation.
    \item Behind every strong, independent woman lies a broken little girl who had to learn how to get back up and to never depend on anyone.
    \item \textbf{Trust} [v] Something which I don't do anymore.
    \item Unexpected friendships are the best ones.
    \item The most common cause of stress nowadays is dealing with idiots.
    \item Your truest friends are the people who don't walk out the door when life gets hard.
    
    They actually pour some coffee and pull up a chair.
    \item Great Leaders don't tell you what to do.
    
    They show you how it's done.
    \item Caring about what people think of you is useless.
    
    Most people don't even know what they think of themselves.
    \item In the blink of an eye, everything can change.
    
    So \textbf{forgive often and love with all your heart.}
    
    You may never know when you may not have that chance again.
    \item We get so worried about \textit{being pretty}.
    
    Let's be pretty kind.
    
    Pretty funny.
    
    Pretty smart.
    
    Pretty strong.
    \item Never think that what you have to offer is insignificant.
    
    There will always be someone out there who needs what you have to give.
    \item Amazing Life-Hack: If you sleep till lunch time then you can save the Breakfast Money.
    \item Men, if you can't control your woman$\ldots$ then you've found the right one.
    \item Believe in yourself!
    \item Dear Heart, please stop getting involved in everything.
    
    Your job is to pump blood, that's it.
    \item Food will never break your heart.
    \item Your age doesn't define your maturity, your grades don't define your intelligence, and rumors don't define who you are.
    \item I don't go crazy.
    
    I am crazy, I just go normal from time to time.
    \item Before you marry someone, wait.
    
    Wait till you see them when they're angry.
    
    Are they abusive?
    
    Are they quiet?
    
    How do they deal with their anger?
    
    Because your children can't choose who gets to be their parents, but you can.
    
    Do it for them.
    \item ``\textit{You don't have to be positive all the time}.
    
    \textit{It's perfectly okay to feel sad, angry, annoyed, frustrated, scared, or anxious}.
    
    \textit{Having feeling doesn't make you a `negative person'}.
    
    \textit{It makes you human}.'' - Lori Deschene
    \item Brain: \textit{Be patient}
    
    Heart: \textit{Until when?}
    \item Trust yourself.
    
    You've survived a lot, and you'll survive whatever is coming.
    \item You know You're an adult when you get excited to just go home.
    \item Real eyes realize real lies.
    \item Sometimes you need bad things to happen to inspire you to change and grow.
    \item If your life were a book, what would the title be?
    \item If you remember anything of me, after I leave this world, remember that I loved even when it was foolish.
    
    That I cared even when it was unwanted.
    
    When my body is gone, remember my heart.
    \item When you fully trust someone without any doubt, you finally get 1 of 2 results:
    
    \textit{A person for life} or \textit{A lesson for life}$\ldots$
    \item I'm the kind of person who doesn't talk to others until they talk to me 1st.
    \item People wait all week for Friday, all year for summer, all life for happiness.
    \item No matter how many times a snake sheds its skin it will always be a snake.
    
    Remember that before allowing certain people back into your life.
    \item \textit{Good morning my wonderful friends}.
    
    May your day begin with a smile on your face, love in your heart and happiness within your soul.
    \item \textbf{The perfect man} [n] An unknown creature available only in books and movies.
    \item Sometimes a person needs to hear those 3 words$\ldots$
    
    No, not those words.
    
    These$\ldots$
    
    \textit{I've got you}.
    \item Kindness is not an act, it's a lifestyle.
    \item The most precious gift you can give someone is the gift of your time and attention.
    \item Just like seasons, people change.
    \item ``\textit{Straight roads do not make skillful drivers}.'' - Paulo Coelho
    \item If you have at least 1 person genuinely supporting you, you're blessed.
    \item I hate when people use their zodiac to justify their shitty behavior like, ``Sorry, I can't help it, I'm a Scorpio.''
    
    No Susan, you're just an asshole.
    \item Sometimes you just need \textbf{an adventure} to cleanse the bitter taste of life from your soul.
    \item The biggest lesson I learned this year is to not force anything; conversations, friendships, relationships, attention, love.
    
    Anything forced is just not worth fighting for, whatever flow flows, what crashes crashes.
    
    It is what it is.
    \item You make me laugh even when I am not in the mood to smile.
    \item If you cannot be the poet, be the poem.
    \item \st{Single}
    
    \st{Taken}
    
    Building an empire,
    
    Finding myself,
    
    Healing myself,
    
    Loving myself,
    
    Being passionate,
    
    Getting fit,
    
    Growing friendships,
    
    Meeting new people \& making memories.
    \item If friendship lasts longer than 10 years, you are no longer just friends.
    
    You are family.
    \item Give your heart to dogs.
    
    They will never break it.
    \item Ordinary is never beautiful.
    
    To be beautiful, You must be weird, different, strange.
    \item You will never know the value of a moment until it becomes a memory.
    \item ``\textit{People think I'm anti-social because I don't joint their conversations}.
    
    \textit{The truth is, I don't give a damn what they're talking about most of the time}.'' - Health Ledger
    \item \textbf{People} will always notice the change in your attitude towards them, but they will never notice it's their behavior that made your change.
    \item I asked God to send me a man who will always love me.
    
    So he gave me a son.
    \item Consider how hard it is to change yourself and you'll understand what little chance you have in trying to change others.
    \item Confuse them with your silence.
    
    Shock them with your actions.
    \item If I'm wrong, educate me.
    
    Don't belittle me.
    \item Chap. 1: ``If it is important enough to you, you will find a way.
    
    If it is not, you will find an excuse.''
    \item Kindness is a language the blind can see and the deaf can hear.
    \item I am definitely not the same person I was when this year started.
    \item People who stay in the car to listen to music a little bit longer are my kind of people.
    \item I don't know how people can fake an entire relationship.
    
    I can't even fake a hello to someone I don't like.
    \item Apparently when you treat people like they treat you, they get upset.
    \item Stay strong, make them wonder how you're still smiling.
    \item Fall in love with the person who enjoys your Madness.
    
    Not an idiot who forces you to be normal.
    \item Say what you feel.
    
    It's not being rude, it's called being \textit{real}.
    \item If you repeat a lie often enough, it becomes \st{truth} politics.
    \item Note to self:
    
    It is entirely possible to be a kind, loving person who refuses to tolerate bullshit.
    
    In fact, it's not only possible, it's absolutely necessary. - Nanea Hoffman
    \item \textit{That's when I realized what a true friend was}.
    
    \textit{Someone who would always love you - the imperfect you, the confused you, the wrong you - because that is what people are supposed to do}. - r. j. l.
    \item Always be in love with a soul, not a face.
    \item The reason I like staying up late so much is because between the hours of 1am to 5am, the world is quiet and no one experts anything from me.
    
    I could stare at my wall for 4 hours and there would be no consequences.
    
    It's so silent and calm.
    
    I love it.
    \item I don't believe in forever but I do believe in growing old with someone you love.
    \item Find someone who will love your soul more than your body.
    \item ``\textit{My mom said something}.
    
    \textit{``You can lie down for people to walk on you and they will still complain that you're not flat enough.''}
    
    \textit{Live you life}.'' - Mature Gambino
    \item ``\textit{Remember that sometimes not getting what you want is a wonderful stroke of luck}.'' - Dalai Lama
    \item Do you know what awesome feeling when you get into bed, fall right to sleep, stay asleep all night, and wake up feeling refreshed?
    
    Me neither.
    \item 1 day or day 1.
    
    You decide.
    \item Hugging is the silent way of saying ``You matter to me''.
    \item The most common cause of stress nowadays is dealing with idiots.
    \item So many years of education yet nobody ever taught us how to love ourselves and why it's so important.
    \item Sometimes my heart needs more time to accept what my mind already knows.
    \item We've been friends for so long I can't remember which 1 of us is the bad influence.
    \item The biggest communication problem is we do not listen to understand.
    
    We listen to reply.
    \item Dear bestie,
    
    No matter how many friends I have, no matter how much I talk to them and spend time with them, always remember that no one can replace you.
    
    You were, are and will always be irreplaceable.
    
    You will always have a special place in my heart.
    
    Someone yours till eternity.
    \item Sometimes the best therapy is a long drive and good music.
    \item \textit{Go for it}.
    
    Whether it ends good or bad, it was an experience.
    \item When you \textit{forgive}, you \textit{heal}.
    
    When you \textit{let go}, you \textit{grow}.
    \item The older I get, the more I realize I don't want to be around drama, conflict or stress.
    
    I want a cozy home, good food, and to be surrounded by happy people.
    \item A Saint was asked - ``What is Anger$\ldots$?''
    
    He gave a beautiful answer - ``It is a punishment we give to ourselves, for somebody else's mistake!''
    \item Find someone who wants to invest in you, learn from your, see you win, support your visions and fall in love with you daily.
    \item ``\textit{How amazing it is to find someone who wants to hear about all the things that go on in your head}.'' - Nina LaCour
    \item Travel.
    
    Money can always return.
    
    Time will not.
    \item Such a great feeling when someone just genuinely wants to talk to you and wants to know how your life is going.
    \item ``\textit{I think people spend \textbf{too much time staring at screens} and not enough time drinking wine, tongue kissing, and dancing under the moon}.'' - Rachel Wolchin
    \item Dear life, I've always been strong even since.
    
    But can you please be easy on me sometimes?
    \item \textsc{Understanding is the nature of Love}:
    
    ``\textit{Understanding someone's suffering is the best gift you can give another person}.
    
    \textit{Understanding is love's other name}.
    
    \textit{If you don't understand, you can't love}.'' - Thich Nhat Hanh
    \item ``\textit{Don't be the reason someone feels insecure}.
    
    \textit{Be the reason someone feels seen, heard, and supported by the whole universe}.'' - Cleo Wade
    \item Never let an old flame burn you twice.
    \item May your vibes shift the whole damn frequency of the room when you walk in.
    \item ``\textit{Stay away from negative people}.
    
    \textit{They have a problem for every solution}.'' - Albert Einstein
    \item ``\textit{I know people who graduated college at 21 and didn't get a salary job until they were 27, I know people who graduated at 25 and already had a salary job}.
    
    \textit{I know people who have children and are single, I know people who are married and had to wait 8--10 years to be parents}.
    
    \textit{I know people who are in a relationship and love someone else, I know people who love each other and aren't together, there are people waiting to love and be loved}.
    
    \textit{My point is, everything in life happens according to our time, our lock}.
    
    \textit{You may look at your friends and some may seem to be ahead or behind you, but they're not, they're living according to the pace of their clock, so be patient}.
    
    \textit{You're not falling behind, it's just not your time}$\ldots$'' - Julissa Loaiza
    \item Some days it's hard to find motivation$\ldots$
    
    Some days motivation finds you!
    \item ``\textit{So don't hurt her, don't change her, don't analyze and don't expect more than she can give}.
    
    \textit{Smile when she makes you happy, let her know when you makes you mad, and miss her when she's not there}.'' - Bob Marley
    \item ``\textit{Sometimes you have to love people from a distance}.'' - Robert Tew
    \item People who make me laugh until it hurts are my favorite kinds of people.
    \item When they call you crazy, remember that great ideas don't come from average minds.
    \item Life would be so much easier if we just said what we fucking feel.
    \item Forgive him, but let him go.
    \item I'm not going to complain about you not texting me back, if I'm not worth your attention then OK that's when I start to lose interest.
    \item Come lay with me for hours so we can talk thousands of nothings while it means Millions of somethings.
    \item Before you judge me, make sure you're perfect.
    
    If you're not, then shut up.
    \item ``\textit{Your heart will never lie to you, that's your mind's job}.'' - Medusa
    \item People keep telling me the right guy will come along, but I think mine got hit by a bus or something.
    \item Note to self: stop missing people who don't even waste a 2nd thinking about you.
    \item Learn to wait.
    
    There's always time for everything.
    \item You want to change your life?
    
    Change the way you think.
    \item Respect is for those who deserve it, not for those who demand it.
    \item My Super Power?
    
    I can look you dead in the face while you're talking and not hear a damn thing you said.
    \item Why is there no maximum wage?
    \item You're always 1 decision away from a totally different life.
    \item People don't care for you when you are alone.
    
    They just care for you when they are alone.
    \item Be your own driver for the race called Life.
    
    Never expect anything, Expectations ruin your Life.
    
    Just work hard and believe in yourself to achieve what your goals are.
    \item Don't work 8 hours for a company then go home \& not work on your own goals.
    
    You're not tired, \textit{you're uninspired}.
    \item Every girl has male best friend in her life with whom she can share everything, without any fear.
    \item Psychology says: Go with the choice that scares you the most, because that's the one that's going to help you grow.
    \item If you want to know where your \textit{heart} is, look where your \textit{mind} goes when it wanders.
    \item \textbf{The truth?}
    
    \textbf{I like you.}
    
    \textbf{A lot.}
    
    You make me happy.
    
    You're smart.
    
    You're different.
    
    You're a little crazy, and awkward, and your smile alone can make my day.
    \item Don't settle for being someone's sometimes.
    \item People who have done you wrong, will forever think your posts are about them.
    \item Choose your battles wisely because if you fight them all you'll be too tired to win the really important ones.
    \item ``\textit{If you tell the \textbf{truth}, you don't have to \textbf{remember} anything}.'' - Mark Twain
    \item 1 of the best feelings is finally losing your attachment to somebody that isn't good for you.
    \item The things that excite you are not random.
    
    They are connected to your purpose.
    
    Follow them.
    \item Never let your fear decide your future.
    \item The better you become, the better you attract.
    \item ``\textit{Great minds discuss ideas}.
    
    \textit{Average minds discuss events}.
    
    \textit{Small minds discuss people}.'' - Eleanor Roosevelt
    \item \textit{I wanna go on a roadtrip someday}.
    
    Alone or with someone I love.
    
    I wanna get away.
    
    Explore places.
    
    Sleep in the car.
    
    Stop a lot just to admire the view.
    
    Visit museums and try out coffee shops.
    
    Listen to my favorite albums while driving.
    
    Have a polaroid camera.
    
    Take pretty pictures of the sunrise.
    
    Take pictures of myself.
    
    Run through a forest.
    
    Chase fog.
    
    Chase the sun.
    
    Spend hours on a field making flower crowns.
    
    Feel the wind in my hairs.
    
    Buy souvenirs.
    
    Meet people.
    
    Take time to observe.
    
    \textit{I wanna make memories}.
    
    \textit{I wanna feel alive}.
    \item I had to forgive a person who wasn't even sorry$\ldots$ \textbf{that's strength.}
    \item If you don't like where you are, move.
    
    You are not a tree!
    \item I'm not heartless, I just learned to use my heart less.
    \item In the end, we all just want someone that chooses us$\ldots$
    
    Over everyone else, under any circumstances.
    \item Me: Alexa, can you check my bank balance and tell me which Apple product I can afford$\ldots$
    
    Alexa: Apple juice$\ldots$
    \item ``\textit{Live your life for you not for anyone else}.
    
    \textit{Don't let the fear of being judged, rejected or disliked stop you from being yourself}.'' - Sonya Parker
    \item ``\textit{No great mind has ever existed without a touch of madness}.'' - Aristotle
    \item As your sister, always remember$\ldots$
    
    I loved you yesterday.
    
    I love you today.
    
    I always have.
    
    \textbf{I always will.}
    \item \textit{My biggest problem?}
    
    I notice everything.
    \item When you truly don't care what the fuck anyone thinks of you, you have reached a dangerously awesome level of freedom$\ldots$!!
    \item ``\textit{Never wish them pain}.
    
    \textit{That's not who you are}.
    
    \textit{If they caused you pain, they must have pain inside}.
    
    \textit{Wish them healing}.'' - Najwa Zebian
    \item A good laugh and a long sleep are the 2 best cures for anything.
    \item Underestimate me.
    
    That'll be fun.
    \item \textbf{Love} [n] Giving someone the power to destroy you, and trusting them not to.
    \item \textit{Sadness is}$\ldots$ not meeting your best friend for a long time.
    \item ``\textit{Inner peace is the new success}.'' - Idil Ahmed
    \item ``\textit{You don't have a soul}.
    
    \textit{You are a soul}.
    
    \textit{You have a body}.'' - C. S. Lewis
\end{enumerate}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}