\documentclass{book}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/nguyen/1_NQBH/math/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\addtolength{\headheight}{0pt}% obsolete
\lhead{\itshape \scriptsize \chaptername~\thechapter}
\rhead{\itshape \scriptsize \nouppercase{\leftmark}} %\nouppercase !
\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\cfoot{\thepage}

\title{Journal}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
\tableofcontents

%------------------------------------------------------------------------------%

\part{Mathematical Journals}

\chapter{\href{https://www.degruyter.com/journal/key/ANS/html}{Advanced Nonlinear Studies}}

\textbf{Details.}
\begin{itemize}
    \item ISSN: 2169-0375
    \item Journal; English; De Gruyter; 4 Issues per Year
    \item 1st published: Aug 1, 2001
    \item Editor-in-chief: Shair Ahmad
    \item 5-year Impact Factor: 1.469
\end{itemize}

\section{Overview}

\subsection{About this journal}
Advanced Nonlinear Studies (ANS) is aimed at publishing scholarly articles on nonlinear problems, particularly those involving Differential and Integral Equations, Dynamical Systems, Calculus of Variations, and related areas.

It will also publish novel and interesting applications of these areas to problems in biology, engineering, materials sciences, physics and other sciences.

Papers submitted to this journal must contain original, timely, and significant results.

This journal will also publish exceptionally well written and timely expository articles written by authors with established expertise in the area.

Articles will generally, but not always, be published in the order when the final copies were received.

%
Advanced Nonlinear Studies is a quarterly journal and will be published in February, May, August, and November.

Papers published in this journal are peer-reviewed.

The publication formats for the Mathematical Reviews Database, published by the American Mathematical Society, include Mathematical reviews, Current Mathematical Publications, and the web based MathSciNet.

This journal is included in the following ISI products: Science Citation Index-Expanded (SCIE), including the Web of Science, Compumath Citation Index (CMCI), Research Alert.

\subsection{Abstracting \& Indexing}
\textit{Advanced Nonlinear Studies} is covered by the following services:
\begin{itemize}
    \item Baidu Scholar
    \item Cabells Journalytics
    \item CNKI Scholar (China National Knowledge Infrastructure)
    \item CNPIEC - cnpLINKer
    \item Dimensions
    \item EBSCO Discovery Service
    \item Genamics JournalSeek
    \item Google Scholar
    \item Japan Science and Technology Agency (JST)
    \item J-Gate
    \item Journal Citation Reports/Science Edition
    \item JournalTOCs
    \item KESLI-NDSL (Korean National Discovery for \item Science Leaders)
    \item Mathematical Reviews (MathSciNet)
    \item Microsoft Academic
    \item MyScienceWork
    \item Naver Academic
    \item Naviga (Softweco)
    \item Primo Central (ExLibris)
    \item Publons
    \item QOAM (Quality Open Access Market)
    \item ReadCube
    \item SCImago (SJR)
    \item SCOPUS
    \item Semantic Scholar
    \item Sherpa/RoMEO
    \item Summon (ProQuest)
    \item TDNet
    \item Ulrich's Periodicals Directory/ulrichsweb
    \item WanFang Data
    \item Web of Science - Current Contents/Physical, Chemical and Earth Sciences
    \item Web of Science - Science Citation Index Expanded
    \item WorldCat (OCLC)
    \item Yewno Discover
    \item zbMATH Open
\end{itemize}

\subsection{Topics}
\begin{itemize}
    \item Applied Mathematics
    \item Differential Equations \& Dynamical Systems
    \item Mathematics
\end{itemize}

\subsection{External links}
\begin{itemize}
    \item \href{https://mc.manuscriptcentral.com/dgans}{Online Submission of Manuscripts}.
\end{itemize}

\section{Latest issue}
\textbf{Volume 21 Issue 1 Feb 2021.}
\begin{itemize}
    \item \textit{On the Fractional NLS Equation and the Effects of the Potential Well's Topology}
    
    Silvia Cingolani, Marco Gallo Nov 12, 2020
    \item \textit{Multiple Periodic Solutions of a Class of Fractional Laplacian Equations}
    
    Ying-Xin Cui, Zhi-Qiang Wang Nov 20, 2020
    \item \textit{Liouville Results and Asymptotics of Solutions of a Quasilinear Elliptic Equation with Supercritical Source Gradient Term}
    
    Marie-Françoise Bidaut-Véron Oct 16, 2020
    \item \textit{Existence of Solutions for Choquard Type Elliptic Problems with Doubly Critical Nonlinearities}
    
    Yansheng Shen Sep 13, 2019
    \item \textit{Sharp Liouville Theorems}
    
     Salvador Villegas Nov 7, 2020
    \item \textit{Existence Results for the Conformal Dirac-Einstein System}
    
    Chiara Guidi, Ali Maalaoui, Vittorio Martino December 15, 2020
    \item \textit{On a Class of Reaction-Diffusion Equations with Aggregation}
    
    Li Chen, Laurent Desvillettes, Evangelos Latos May 12, 2020
    \item \textit{Existence and Concentration of Solutions for Choquard Equations with Steep Potential Well and Doubly Critical Exponents}
    
    Yong-Yong Li, Gui-Dong Li, Chun-Lei Tang Nov 4, 2020
    \item \textit{Deforming a Convex Hypersurface by Anisotropic Curvature Flows}
    
    HongJie Ju, BoYa Li, YanNan Liu Oct 2, 2020
    \item \textit{Reversed Stein–Weiss Inequalities with Poisson-Type Kernel and Qualitative Analysis of Extremal Functions}
    
    Chunxia Tao Nov 20, 2020
    \item \textit{Nonexistence of Solutions for Dirichlet Problems with Supercritical Growth in Tubular Domains}
    
    Riccardo Molle, Donato Passaseo Jan 12, 2021
    \item \textit{A Qualitative Study of $(p,q)$ Singular Parabolic Equations: Local Existence, Sobolev Regularity and Asymptotic Behavior}
    
    Jacques Giacomoni, Deepak Kumar, Konijeti Sreenadh Jan 21, 2021 
\end{itemize}

\section{Issues}
See \href{https://www.degruyter.com/journal/key/ANS/html}{De Gruyter/Advanced Nonlinear Studies/Issues}.

\section{Ranking}
\begin{itemize}
    \item 5-year Impact Factor: 1.469
    \item Cite Score: 2.9
    \item Impact Factor: 1.533
    \item Mathematical Citation Quotient: 1.19
    \item SCImago Journal Rank: 1.594
    \item Source Normalized Impact per Paper: 1.289
\end{itemize}

\section{Submit}

\subsection{Submitting an article}
$\ldots$ is done online to our ScholarOne online submission system at: \url{http://mc.manuscriptcentral.com/dgans}.

%
Contributions submitted to this journal must be written in clear and concise English.

Only unpublished material can be accepted, and authors may not republish their paper in the same or similar form.

The decision on the acceptance is taken after a peer-reviewing procedure.

The journal makes no page charges.

The authors can provide the names and e-mail addresses of up to 4 potential reviewers.

\subsection{Each article}
$\ldots$ must begin with an abstract of less than 150 words.

The abstract must be followed by an appropriate classification according to MSC 2010 (Mathematics Subject Classification) and a list of 3 to 6 keywords.

The short title for the running head should contain at most 80 characters, spaces inclusive.

%
Authors are asked to follow a few basic guidelines while formatting their paper for publication, in order for the journal to be as homogeneous in appearance as possible.

Download the LaTeX class file and the respective sample file including the instructions for formatting.

\subsection{Figures}
$\ldots$ must be of sufficiently high resolution (minimum 600 dpi).

Lettering of all figures within the article should be uniform in style (preferably a sans serif typeface like Helvetica) and of sufficient size (ca. 8 pt.).

Uppercase letters A, B, C, etc. should be used to identify parts of multi-part figures.

Cite all figures in the text in numerical order. Indicate the approximate placement of each figure.

Do not embed figures within the text body of the manuscript; submit figures in separate files.

In a TEX document EPS files may be incorporated into the TEX file by using 1 of the standard macro packages.

%
Authors are encouraged to submit illustrations in color if necessary for their scientific content.

Publication of color figures is provided free of charge both in online and print editions.

\subsection{References}
$\ldots$ should be collected at the end of the paper, numbered with Arabic numerals in alphabetical order of the authors' names and surrounded by square brackets.

Use abbreviations for titles of journals in accordance with those listed in Mathematical Reviews.
\begin{itemize}
    \item Articles in journals:
    
    [1] A. Carbonaro, G. Metafune and C. Spina, Parabolic Schrödinger operators, J. Math Anal. and Appl. \textbf{343} (2008), 965-974.
    \item Articles ahead of print:
    
    [2] D. Farley and L. Sabalka, Presentations of graph braid groups, Forum Math. (2010), doi: 10.1515/FORM.2011.086.
    \item Books and Monographs:
    
    [3] I. M. Isaacs, Character theory of finite groups, AMS Chelsea Publishing, 2006.
    \item Chapters:
    
    [4] T. De Medts, F. Haot, R. Knop and H. Van Maldeghem, On the uniqueness of the unipotent subgroups of some Moufang sets, in: Finite Geometries, Groups, and Computation, pp. 43-66, de Gruyter, Berlin, 2006.
\end{itemize}

\subsection{Galley proofs}
$\ldots$ will be sent to all authors in due time with a request to carefully check them.

\subsection{After publication}
$\ldots$ each author will receive a PDF file of the final version for private use only.

\subsection{Hybrid Open Access}
For complete details on hybrid open access publishing at De Gruyter please see: \url{https://www.degruyter.com/page/560}.

%
Effective Jan 1, 2016, authors from an institution affiliated with either the Association of Dutch Universities (VSNU), the Consortium of Swiss Academic Libraries or some UK libraries participating in the Jisc Collections SMP may publish primary research and review articles open access in any of De Gruyter's OnlineOpen journals at a discount of 90\% of the APC price.

For more information and to confirm whether your institution is eligible see the following:
\begin{itemize}
    \item The Netherlands: \url{https://www.degruyter.com/page/1461}
    \item The UK: \url{https://www.degruyter.com/page/1462}
    \item Switzerland: \url{https://www.degruyter.com/page/1463}
\end{itemize}

\section{Editorial}
See \href{https://www.degruyter.com/journal/key/ANS/html}{De Gruyter/Advanced Nonlinear Studies/Editorial}.\hfill$\square$

%------------------------------------------------------------------------------%

\part{Engineering Journals}



%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}