\documentclass{book}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/nguyen/1_NQBH/math/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\addtolength{\headheight}{0pt} % obsolete
\lhead{\scriptsize \chaptername~\thechapter}
\rhead{\scriptsize \nouppercase{\leftmark}} %\nouppercase !
\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\cfoot{\thepage}
\def\labelitemii{$\circ$}

\title{Wise}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
\tableofcontents

%------------------------------------------------------------------------------%

\part{Wikipedia's}

\chapter{\href{https://en.wikipedia.org/wiki/Wise_(company)}{Wikipedia/Wise (Company)}}

\textbf{Wise.}
\begin{itemize}
    \item \textbf{Type of site.} \href{https://en.wikipedia.org/wiki/Private_company}{Private company}.
    \item \textbf{Founded.} Jan 2011.
    \begin{itemize}
        \item Bryant, Martin. ``\href{https://www.bbc.co.uk/news/business-31639262}{\textit{Money may make the world go round, but at what cost?}}''. BBC. Retrieved Apr 1, 2015.
        \item ``\href{https://www.bloomberg.com/news/articles/2017-05-23/transferwise-steps-toward-banking-with-multicurrency-account}{\textit{TransferWise Steps Toward Banking With Multicurrency Account}}''. Bloomberg.com. 23 May 2017. Retrieved May 31, 2017.
        \item Bryant, Martin. ``\href{https://www.independent.co.uk/news/business/news/transferwise-valued-at-1bn-by-top-silicon-valley-venture-capital-fund-10002618.html}{\textit{TransferWise valued at bn by top Silicon Valley venture capital fund}}''. The Independent. Retrieved Apr 1, 2015.
    \end{itemize}
    \item \textbf{Headquarters.} \href{https://en.wikipedia.org/wiki/London}{London}, \href{https://en.wikipedia.org/wiki/United_Kingdom}{United Kingdom}.
    \item \textbf{Founder(s).}
    \begin{itemize}
        \item Taavet Hinrikus (\href{https://en.wikipedia.org/wiki/Executive_Chairman}{Executive Chairman}).
        \item Kristo Käärmann (\href{https://en.wikipedia.org/wiki/Chief_executive_officer}{CEO})
    \end{itemize} 
    \item \textbf{Services.} \href{https://en.wikipedia.org/wiki/Financial_services}{Financial services}, \href{https://en.wikipedia.org/wiki/Remittance}{Remittance}.
    \item \textbf{URL.} \url{https://wise.com/}.
\end{itemize}
\textit{Wise} (formerly \textit{TransferWise}) is a London-based \href{https://en.wikipedia.org/wiki/Financial_technology}{financial technology company} founded in Jan 2011 by \href{https://en.wikipedia.org/wiki/Estonians}{Estonians} Kristo Käärmann and Taavet Hinrikus.[1][2][3]

The company rebranded to Wise in February 2021, to reflect its expanded product offering beyond international money transfer.[``\href{https://sifted.eu/articles/transferwise-becomes-wise/}{\textit{TransferWise becomes Wise}}''. Sifted. Feb 22, 2021. Retrieved Feb 22, 2021.]

\section{History}
\textsf{Regular money transfer vs. peer-to-peer money transfer.}

Wise was founded in 2011 by Taavet Hinrikus, \href{https://en.wikipedia.org/wiki/Skype}{Skype}'s 1st employee,[Temperton, James (Nov 3, 2016). ``\href{https://www.wired.co.uk/article/taavet-hinrikus-transferwise-skype-estonia}{\textit{Has Skype lost its way? Taavet Hinrikus reveals what the company taught him}}''. Wired UK. ISSN 1357-0978. Retrieved Apr 2, 2019.] and financial consultant Kristo Käärmann.

%
Wise's system has been compared to the \href{https://en.wikipedia.org/wiki/Hawala}{hawala} money transfer system.
\begin{itemize}
    \item Price, Rob (Jan 27, 2015). ``\href{http://www.businessinsider.com.au/transferwises-similarities-to-hawala-2015-1}{\textit{London's \$1 Billion Finance Startup TransferWise Is Just Like An Ancient Islamic Money Transfer System}}''. Business Insider. Retrieved Oct 19, 2015.
    \item Marco della Cava (Aug 18, 2014). ``\href{https://www.usatoday.com/story/tech/2014/08/16/transferwise-takes-on-banking-money-transfer-fees/13622337/}{\textit{London's TransferWise aims to disrupt banking}}''. USA Today. Retrieved Oct 19, 2015.
    \item Bennett Voyles (Sep 9, 2015). ``\href{http://forbesindia.com/article/ckgsb/online-money-transfers-and-the-skype-of-money/40999/1}{\textit{Online money transfers and the ``Skype'' of money}}''. Forbes India. Retrieved Oct 19, 2015.
    \item Leander Bindewald (Jul 8, 2015). ``\href{https://www.theguardian.com/sustainable-business/2015/jul/07/banking-sector-credit-unions-local-currency-money}{\textit{You need never use a bank again. Here's why}}''. The Guardian. Retrieved Oct 19, 2015.
\end{itemize}
In its 1st year of operation, transactions through Wise amounted to €10 million.[Bryant, Martin. ``Peer-to-peer currency exchange service Transferwise handles \$13.4m in its 1st year''. Insider. Retrieved Oct 1, 2012.]

%
In 2012, Wise was named as one of ``East London's 20 hottest tech startups'' by \href{https://en.wikipedia.org/wiki/The_Guardian}{\textit{The Guardian}}, Start Up of the Week by \href{https://en.wikipedia.org/wiki/Wired_UK}{\textit{Wired UK}}, 1 of 5 ``start-ups to watch'' at Seedcamp's 2012 US Demo Day by TechCrunch, and made it to Startups.co.uk's list of the top 100 UK start-ups of 2012.
\begin{itemize}
    \item Silver, James (Jul 8, 2012). ``\href{https://www.theguardian.com/uk/2012/jul/08/east-london-20-hottest-tech-companies}{\textit{East London's 20 hottest tech startups}}''. The Guardian.
    \item ``\href{http://www.startups.co.uk/transferwise-taavet-hinrikus-and-kristo-k%C3%A4%C3%A4rmann-startups-100.html}{\textit{The 2012 Startups 100: revealed}}''. Startups. Retrieved Oct 1, 2012.
    \item Taylor, Colleen. ``\href{https://techcrunch.com/2012/02/29/5-startups-to-watch-from-seedcamps-2012-us-demo-day-tctv/}{\textit{5 Startups to Watch from Seedcamp's 2012 US Demo Day}}''. TechCrunch. Retrieved Oct 1, 2012.
\end{itemize}
In Apr 2013, Wise stopped letting users purchase \href{https://en.wikipedia.org/wiki/Bitcoin}{Bitcoin}, citing pressure from banking providers.[``\href{https://transferwise.com/help/questions/2759169/bitcoin}{Bitcoin $|$ TransferWise Support \& FAQ - International Money Transfer}''. transferwise.com. Retrieved May 31, 2017.]

%
Independent comparison site Monito revealed that Wise was on average 83\% cheaper than the big 4 UK banks on major currency ``routes'', but could be up to 90\% cheaper in certain specific cases.[``\href{https://www.monito.com/blog/transferwise-cheaper-banks/}{\textit{TransferWise – Are they that cheap or is it just (inaccurate) advertising?}}''. www.monito.com. Monito. 26 May 2016. Retrieved 21 November 2016.]

%
In May 2015, Wise was ranked No. 8 on \href{https://en.wikipedia.org/wiki/CNBC}{CNBC}'s 2015 Disruptor 50 list,[``\href{https://www.cnbc.com/id/102609977}{\textit{Meet the 2015 CNBC Disruptor 50 companies}}''. CNBC. Retrieved May12, 2015.] and in Aug 2015, the company was named a \href{https://en.wikipedia.org/wiki/World_Economic_Forum}{World Economic Forum} Tech Pioneer.[Barber, Lynsey (Aug 5, 2015). ``\href{http://www.cityam.com/221702/world-economic-forum-tech-pioneers-transferwise-and-mike-lynchs-darktrace-among-four-uk-firms}{\textit{4 UK firms named tech pioneers by WEF}}''. Retrieved Nov 13, 2016.]

%
On Apr 8, 2017, an internal memo from British bank \href{https://en.wikipedia.org/wiki/Santander_UK}{Santander} claimed the bank would lose 84\% of its revenue from its money transfer business if its charges were the same as Wise.[Collinson, Patrick (Apr 8, 2017). ``\href{https://www.theguardian.com/money/2017/apr/08/leaked-santander-international-money-transfers-transferwise}{\textit{Revealed: the huge profits earned by big banks on overseas money transfers}}''. The Guardian. ISSN 0261-3077. Retrieved May 31, 2017.]

%
Also in Apr 2017, the company announced the opening of its \href{https://en.wikipedia.org/wiki/Asia-Pacific}{APAC} hub in \href{https://en.wikipedia.org/wiki/Singapore}{Singapore}.[Russell, Jon. ``\href{https://techcrunch.com/2017/04/27/transferwise-asia-singapore-office/}{\textit{TransferWise moves into Asia Pacific with opening of regional HQ in Singapore $|$ TechCrunch}}''. Retrieved May 31, 2017.]

In 2019, the company announced opening an office in Brussels.[Megaw, Nicholas (Jan 10, 2019). ``\href{https://www.ft.com/content/bbe1487e-1438-11e9-a581-4ff78404524e}{\textit{Brexit drives TransferWise to open Belgian office}}''. www.ft.com. Retrieved Feb 10, 2021.]

%
In May 2017, the company announced its customers were sending over £1 billion every month using the service,[CNBC.com, Neil Ainger, Writer at (17 May 2017). ``\href{https://www.cnbc.com/2017/05/17/fintech-unicorn-transferwise-achieves-first-profit.html}{\textit{Fintech unicorn Transferwise achieves 1st profit}}''. CNBC. Retrieved May 31, 2017.] and that the company turned profitable 6 years after being founded.[``\href{https://www.bbc.co.uk/news/business-39943651}{\textit{TransferWise becomes profitable six years after being founded}}''. BBC News. May 17, 2017. Retrieved May 31, 2017.]

%
On Jan 21, 2021, \href{https://en.wikipedia.org/wiki/Sky_News}{Sky News} reported that Wise had appointed \href{https://en.wikipedia.org/wiki/Goldman_Sachs}{Goldman Sachs} and \href{https://en.wikipedia.org/wiki/Morgan_Stanley}{Morgan Stanley} as joint global co-ordinators for its planned \href{https://en.wikipedia.org/wiki/Initial_public_offering}{initial public offering}.[Kleinman, Mark (21 January 2021). ``\href{https://news.sky.com/story/transferwise-picks-banks-to-spearhead-blockbuster-flotation-12194386}{\textit{TransferWise picks banks to spearhead blockbuster flotation}}''. Sky News. Retrieved Feb 11, 2021.]

%
On Feb 22, 2021, the company rebrands from TransferWise to Wise.
\begin{itemize}
    \item O'Hear, Steve (Feb 22, 2021). ``\href{https://techcrunch.com/2021/02/21/wise/}{\textit{TransferWise rebrands as Wise ahead of an expected IPO}}''. TechCrunch. Retrieved Feb 22, 2021.
    \item Pratty, Freya; Lewin, Amy (Feb 22, 2021). ``\href{https://sifted.eu/articles/transferwise-becomes-wise/}{\textit{TransferWise becomes Wise}}''. Sifted. Retrieved Feb 22, 2021.
    \item Musgrove, Annie (Feb 22, 2021). ``\href{https://tech.eu/brief/transferwise-rebrands-to-wise/}{\textit{Transferwise rebrands to Wise}}''. Tech.eu. Retrieved Feb 22, 2021.
\end{itemize}
As part of this rebranding the company also launched a new website \href{https://en.wikipedia.org/wiki/Domain_name}{domain}.[25]

\section{Funding}
Wise received seed funding amounting to \$1.3 million from a consortium including venture firms IA Ventures and \href{https://en.wikipedia.org/wiki/Index_Ventures}{Index Ventures}, IJNR Ventures, NYPPE as well as individual investors such as \href{https://en.wikipedia.org/wiki/PayPal}{PayPal} co-founder \href{https://en.wikipedia.org/wiki/Max_Levchin}{Max Levchin}, former \href{https://en.wikipedia.org/wiki/Betfair}{Betfair} CEO \href{https://en.wikipedia.org/wiki/David_Yu}{David Yu}, and \href{https://en.wikipedia.org/wiki/Wonga.com}{Wonga.com} co-founder Errol Damelin.[Johnson, Bobbie. ``\href{http://gigaom.com/europe/currency-startup-transferwise-unveils-superstar-backers/}{\textit{Transferwise unveils Levchin, other superstar backers}}''. Gigaoam. Retrieved Oct 1, 2012.]

%
Wise also received investment after being named one of \href{https://en.wikipedia.org/wiki/Seedcamp}{Seedcamp} 2011's winners.[``\href{http://www.seedcamp.com/companies/transferwise}{TransferWise}''. Seedcamp. Retrieved Oct 1, 2012.]

%
In May 2013, it was announced that Wise had secured a \$6 million investment round led by \href{https://en.wikipedia.org/wiki/Peter_Thiel}{Peter Thiel}'s \href{https://en.wikipedia.org/wiki/Valar_Ventures}{Valar Ventures}.[O'Hear, Steve. ``\href{https://techcrunch.com/2013/05/13/tw/}{P2P Currency Exchange TransferWise Raises \$6M Led By Peter Thiel's Valar Ventures, With Participation From SV Angel, Others}''. TechCrunch. Retrieved Jul 12, 2013.]

%
Wise raised a further \$25 million in Jun 2014, adding \href{https://en.wikipedia.org/wiki/Richard_Branson}{Richard Branson} as an investor.[O'Hear, Steve. ``Now Backed By Sir Richard Branson, TransferWise Raises \$25M For Cheaper Money Transfers''. TechCrunch. Retrieved Jun 9, 2014.]

%
In January 2015, it was announced that Wise had raised a \$58 million Series C round, led by investors \href{https://en.wikipedia.org/wiki/Andreessen_Horowitz}{Andreessen Horowitz}.[Price, Rob. ``\textit{London Cash Startup TransferWise Is Now Worth \$1 Billion}''. BusinessInsider. Retrieved Mar 31, 2015.]

%
In May 2016, Wise secured a funding of \$26 million.

This raised the company's valuation to \$1.1 billion.

As of May 2016, Wise has raised a total of \$117 million in funding.[O'Hear, Steve. ``\href{https://techcrunch.com/2016/05/25/money-transfer-company-transferwise-raises-further-26m-at-1-1b-valuation/}{\textit{Money transfer company TransferWise raises further \$26M at \$1.1B valuation}}''. TechCrunch. Retrieved May 25, 2016.]

%
In Nov 2017, the company raised a \$280 million Series E led by Old Mutual Global Investors and IVP, as well as \href{https://en.wikipedia.org/wiki/Sapphire_Ventures}{Sapphire Ventures}, Japanese Mitsui \& Co, and World Innovation Lab.[``\href{https://www.ivp.com/news/press-release/transferwise-announces-280m-investment-round-as-company-focuses-on-new-products-and-apac-expansion/}{\textit{TransferWise Announces \$280M Investment Round as Company Focuses on New Products and APAC Expansion}}''. IVP. Nov 2, 2017.]

%
In May 2019, the company had the secondary investment round of \$292 million and reached the total valuation of \$3.5 billion, more than double the valuation Wise achieved in late 2017 at the time of its \$280 million \href{https://en.wikipedia.org/wiki/Venture_round}{Series E round}.[``\href{http://social.techcrunch.com/2019/05/21/three-point-five-unicorns/}{\textit{TransferWise now valued at \$3.5B following a new \$292M secondary round}}''. TechCrunch. Retrieved May 22, 2019.]

\section{Criticism}
In May 2016, Wise's claim ``you save up to 90\% against banks'' was considered as misleading by the \href{https://en.wikipedia.org/wiki/Advertising_Standards_Authority_(United_Kingdom)}{Advertising Standards Authority}.[``\href{https://www.asa.org.uk/rulings/transferwise-ltd-a15-312197.html}{\textit{ASA Ruling on TransferWise Ltd}}''. May 4, 2016. Retrieved Nov 21, 2016.]

%
In Jun 2020, after experts raised ethical and privacy concerns around the digital \href{https://en.wikipedia.org/wiki/COVID-19}{COVID-19} \href{https://en.wikipedia.org/wiki/Immunity_passport}{immunity passports} Wise was helping develop, the company conceded immunity passports aren't a ``perfect solution'' and co-founder Hinrikus said they would not be launched publicly until there was scientific consensus on COVID-19 immunity.[Browne, Ryan (Jun 3, 2020). ``\href{https://www.cnbc.com/2020/06/03/coronavirus-experts-warn-digital-immunity-passports-are-unethical.html}{\textit{Start-ups are racing to develop Covid-19 `immunity passports' — but experts warn they're unethical}}''. CNBC. Retrieved Feb 11, 2021.]

\section{See also}
\begin{itemize}
    \item \href{https://en.wikipedia.org/wiki/CurrencyFair}{CurrencyFair}
    \item \href{https://en.wikipedia.org/wiki/Financial_technology}{Financial technology}
    \item \href{https://en.wikipedia.org/wiki/Xoom_Corporation}{Xoom Corporation}
    \item \href{https://en.wikipedia.org/wiki/WorldRemit}{WorldRemit}
    \item \href{https://en.wikipedia.org/wiki/Midpoint_(company)}{Midpoint (company)}
    \item \href{https://en.wikipedia.org/wiki/TransferMate}{TransferMate}\hfill$\square$
\end{itemize}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}