\documentclass[onsided]{book}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{~/1_NQBH/math/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\addtolength{\headheight}{0pt}% obsolete
\lhead{\itshape \small \chaptername~\thechapter}
\rhead{\itshape \small \nouppercase{\leftmark}} %\nouppercase !
\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\cfoot{\thepage}

\title{Movie Quotes}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
\tableofcontents

%------------------------------------------------------------------------------%

\part{TV Series}

\chapter{Ano hi mita hana no namae o bokutachi wa mada shiranai $\star$ Anohana: The Flower We Saw That Day (2011)}

%------------------------------------------------------------------------------%

\chapter{Berserk (1997--1998)}

%------------------------------------------------------------------------------%

\chapter{BoJack Horseman (2014--2020)}

\section{BoJack Horseman}
\begin{enumerate}
    \item ``He's so stupid he doesn't realize how miserable he should be. I envy that.''
    \item ``Ow, crap. I hate this. Running is terrible. Everything is the worst.''
    \item ``You know sometimes I feel like I was born with a leak, and any goodness I started with just slowly spilled out of me, and now it's all gone. And I'll never get it back in me. It's too late. ... Life is a series of closing doors, isn't it?''
    \item``You are a horrible person, and you not understanding that you're a horrible person, doesn't make you less of a horrible person.''
    \item ``I'm responsible for my own happiness? I can't even be responsible for my own breakfast!''
    \item ``You sleep on my couch and you don't pay rent. I've had tapeworms that are less parasitic. I don't even remember why I let you stay with me in the first place.''
    \item ``Man, I really regret buying those muffins and then eating them all on the drive home.''
    \item ``I feel like my life is just a series of unrelated wacky adventures''
    \item ``I don't understand how people... live. It's amazing to me that people wake up every morning and say: `Yeah, another day, let's do it.' How do people do it? I don't know how.''
    \item ``Look, for a lot of people, life is just one long hard kick in the urethra.''
    \item ``Same thing that always happens. You didn't know me and then you fell in love with me. And now you know me.''
    \item ``Now if you excuse me, I need to go take a shower so I can't tell if I'm crying or not.''
    \item ``I spend a lot of time with the real me and believe me, nobody's gonna love that guy.''
    \item ``Yeah, I ate them all in one sitting because I have no self-control and I hate myself.''
    \item ``nothing on the outside, nothing on the inside'' as he runs his hand over the oven
    \item ``Settle. Because otherwise you're just gonna get older and harder, and more alone. And you're gonna do everything you can to fill that hole, with friends, and your career, and meaningless sex, but the hole doesn't get filled. One day, you're gonna look around and you're going to realize that everybody loves you, but nobody likes you. And that is the loneliest feeling in the world.''
    \item ``I want to feel good about myself. The way you do. And I don't know how. I don't know if I can.''
    \item ``Fool me once, shame on you. But teach a man to fool me and I'll be fooled for the rest of my life'' 
    \item ``Dead on the inside, dead on the outside.''
    \item ``I have half a mind...''
    \item ``I can't say no to people because I want everyone to like me.''
    \item ``It's made up by Steven Spielberg to sell movie tickets. It's like the happiness or the Munich Olympics, it doesn't exist in the real world. The only thing to do now is just to keep living forward.''
    \item ``We're just two lonely people trying to hate ourselves a little less.''
    \item ``your girlfriend has a stripper name''
    \item ``Kelsey, in this terrifying world, all we have are the connections that we make. I'm sorry I got you fired, I'm sorry I never called you after.''
    \item ``Being a movie star is the hardest job, and we get no recognition!''
    \item ``Todd, your good-hearted naivety has once again conspired with outrageous happenstance to completely dick me over!''
    \item ``Why 12 steps? Nobody wants to do 12 of anything. Did you see `12 Years A Slave' and think `that's a short number of years to be a slave?'''
    \item ``When your father dies, you ask yourself a lot of questions, like `Wait, did you say he died in a duel?' and `Who dies in a duel?'''
    \item ``Is it possible that this whole time, I've been an amazing feminist hero and nobody knew it?''
    \item ``Diane, I am a famous -- everyone gives me everything I want all the time. It is an existential curse, but a huge day-to-day convenience.''
    \item ``You've got a nice set of pipes. You're like a Josh Groban who doesn't also think he's funny.''
    \item ``I want to do things with you. Fully clothed, sober, in daylight hours.''
    \item ``How can I put this... imagine if the Holocaust happened every four years like the Olympics. I would rather THAT happen than your rock opera.''
    \item ``Goddammit, Honeydew? Jesus, why does Cantaloupe think every time it gets invited to a party it can bring along its dumb friend Honeydew? You don't get a plus one Cantaloupe.''
    \item ``There's that old saying: Liquor before beer, never fear, don't do heroin.''
    \item ``Why put the skip ad button so late? I'm not skipping now. I'm invested.''
    \item ``I drove, but I moved my arm a bunch so the Fitbit counted the miles.''
    \item ``That was, and I don't say this lightly, worse than 100 September 11ths.''
    \item ```Here's the thing about most long distance plans.' I hate that. `Here's the thing.' It's so stupid! Just say the thing! You don't need to introduce the concept that there's going to be a thing.''
    \item ``Todd! Get in the car. It is time to get serious about autoerotic asphyxiation.''
    \item ``Also, are you wearing comfortable shoes? Because I feel like that was a long road to walk to get to that punchline.''
    \item ``I'm sorry I accused you of murder, American TV legend Henry Winkler.''
    \item ``Slap my salami, the guy's a commie.''
    \item ``I told you, that's not what was happening that time! I was masturbating to what the picture represented! And you came in at the worst possible time!''
    \item ```Neigh way, Jose!' I improvised that line. I mean, it was scripted, but I gave it the ol' BoJack spin.''
    \item ``Well, that was another in a long series of regrettable life choices.''
    \item ``It's so sad that when you see someone as they really are, it ruins them.''
    \item ``I've had a lot of what I thought were rock bottoms, only to discover another, rockier bottom underneath.''
    \item ``Just pretend you are happy, and eventually you'll forget you're pretending.''
    \item ``I have poison inside me, and I destroy everything I touch.''
    \item ``The only thing to do now is just to keep living forward.''
    \item ``My darling mother gave the eulogy. My entire life, I never heard her say a kind word to OR about my father, but at his funeral, she said, ``My husband is dead, and everything is worse now.''
    \item ``I'm gonna say ``Hello. I am BoJack Horseman. Obviously, you know who I am, because I'm very famous, and also because we called ahead. And I am here because ... I need help.''
    \item ``I don't hate the troops, I just hate one specific troop. I don't even hate him, I just think he's wrong about the muffins.''
    \item ``All I learned about being good, I learned from TV. And in TV, flawed characters are constantly showing people they care with these surprising grand gestures. And I think that part of me still believes that's what love is.''
    \item ``I told you, that's not what was happening that time. I was masturbating to what the picture represented! You walked in at the worst possible moment.''
    \item ``You turn yourself around. That's what it's all about.''
    \item ``I don't cry in front of other people.''
    \item ``But in real life, the big gesture isn't enough. You need to be consistent, you need to be dependably good. You need to do it every day, which is so ... hard.''
    \item ``What more do you want?! What else could the Universe possibly owe you?''
\end{enumerate}

\section{Diane Nguyen}
\begin{enumerate}
    \item ``It's not about being happy, that is the thing. I'm just trying to get through each day. I can't keep asking myself `Am I happy?' It just makes me more miserable. I don't know If I believe in it, real lasting happiness, All those perky, well-adjusted people you see in movies and TV shows ? I don't think they exist.''
    \item ``Idea for a new app: an undo button that can undo long amounts of time. Three months. A year. A life.''
    \item ``Even if no one appreciates you, it's important that you don't stop being good.''
    \item ``When I was a kid, I used to watch you on TV, and you know I didn't have the best family. Things weren't that great for me. But, for half an hour, every week, I got to watch this show about four people who had nobody, who came together and became a family, and for half an hour, every week, I had a home, and it helped me survive.''
    \item ``Well, that's the problem with life, right? Either you know that you want and then you don't get what you want, or you get what you want, and then you don't know what you want.''
    \item ``Most people don't even get to do the brady bunch version of the thing they want to do with their lives.''
    \item ``You're responsible for your own happiness, you know?''
    \item ``You learn that you can survive being alone.''
    \item ``Sometimes life's a bitch and you keep living.''
    \item ``There are people in your life that help you become the person you end up being, and you can be grateful for them, even if they were never meant to be in your life forever.''
    \item ``Every happy ending has the day after the happy ending.''
    \item ``There's no deep down, I believe that all we are is what we do.''
    \item ``There's no such thing as ``bad guys'' or ``good guys''. We're all just... guys, who do good stuff sometimes and bad stuff sometimes. And all we can do is try to do less bad stuff and more good stuff, but you're never going to be good because you're not bad.''
\end{enumerate}

\section{Todd Chavez}
\begin{enumerate}
    \item ``Things don't become traditions because they're good, BoJack, they become good because they're traditions.''
    \item ``You can't keep doing shitty things, and then feel bad about yourself like that makes it okay! You need to be better!''
    \item ``You are all the things that are wrong with you.''
\end{enumerate}

\section{Princess Carolyn}
\begin{enumerate}
    \item ``Oh fish. Of course! Why would you ever make things easy for me when instead you could make things incredibly difficult. Laura! Clear our my schedule! I have to push a boulder up a hill and then have it roll over me time and time again with no regard for my well-being!''
    \item ``That woman can knock a drink back like a Kennedy at a wake for another Kennedy, but I'll be damned if she doesn't get shit done!''
    \item ``BoJack, I'm gonna level with you, honey. This whole you-hating-the-troops thing is not great.''
    \item ``Because my life is a mess right now and I compulsively take care of other people.''
    \item ``I got into this business because I love stories. They comfort us, they inspire us, they create a context for how we experience the world. But also, you have to be careful, because if you spend a lot of time with stories, you start to believe that life is just stories, and it's not. Life is life, and that's so sad because there's so little time and ... what are we doing with it?''
    \item ``So, are you available for Tuesday, or are you gonna be too busy masturbating to old pictures of yourself.``
    \item ``Stop pissing off the orphans. A lot of them grow up to be serial killers.''
\end{enumerate}

\section{Mr. Peanutbutter}
\begin{enumerate}
    \item ``Sweetie, you know I support you, whatever you want to do, but you're not gonna find what you're looking for in these awful made-up places. The universe is a cruel, uncaring void. The key to being happy isn't the search for meaning; it's just to keep yourself busy with unimportant nonsense, and eventually, you'll be dead.''
    \item ``Everybody deserves to be loved''
    \item ``Crack an egg on your head. Let the yolk drip down.''
    \item ``All I ever wanted was to be your friend ... and you treat me like a big joke. You think I don't notice? Why don't you like me?''
    \item ``I told you I don't know where it is. Don't put things in my butt if you want them back.''
\end{enumerate}

\section{Herb Kazzaz}
\begin{enumerate}
    \item ``You know what your problem is? You want to think of yourself as the good guy. Well, I know you better than anyone, and I can tell you that you're not. In fact, you'd probably sleep a lot better at night if you just admitted to yourself that you're a selfish goddamn coward who just takes whatever he wants and doesn't give a shit about who he hurts. That's you. That's BoJack Horseman.''
    \item ``There is no other side.''
\end{enumerate}

\section{Beatrice Horseman}
\begin{enumerate}
    \item ``I'm punishing you for being alive.''
    \item ``I just wanted to tell you that I know. I know you want to be happy, but you won't be... and I'm sorry.''
    
    BoJack: ``... What?''
    
    ``It's not just you, you know. Your father and I, we, well... you come by it honestly, the ugliness inside you. You were born broken, that's your birthright. And now you can fill your life with projects your books and your movies and your little girlfriends but... that won't make you whole. You're BoJack Horseman. There's no cure for that.''
    \item ``Don't throw your dreams away for this child. Don't let that man poison your life the way he did mine. You are going to finish your schooling and become a nurse. You'll meet a man, a good man, and you'll have a family, but please, believe me, you don't want this. Please, Henrietta, you have to believe me. Please ... don't do what I did.''
    \item ``Henrietta, don't use a foreign language in front of the child, she'll get ideas!''
\end{enumerate}

\section{Henry Winkler}
\begin{enumerate}
    \item ``There is no shame in dying for nothing. That's why most people die.''
\end{enumerate}

\section{Hollyhock}
\begin{enumerate}
    \item ``That voice, the one that tells you you're stupid, worthless, and ugly... it goes away right? It's just a dumb teenage girl thing, but... then it goes away?'', ``Yeah''
\end{enumerate}

\section{Secretariat}
\begin{enumerate}
    \item ``BoJack, when I was your age, I got sad. A lot. I didn't come from such a great home, but one day, I started running, and that seemed to make sense, so then I just kept running. BoJack, when you get sad you run, straight ahead and you keep running forward no matter what there are people in your life that who are going to try to hold you back , slow you down , but you don't let them. Don't you ever stop running, don't ever look behind you there is nothing for you behind you. All that exists is whats ahead.'' 
\end{enumerate}

\section{Wanda Pierce}
\begin{enumerate}
    \item ``When you look at someone through rose-colored glasses, all the red flags just look like flags.''
\end{enumerate}

\section{Amanda Hannity}
\begin{enumerate}
    \item ''When we know what we know about a monster like that and we still put him on TV every week, we're teaching a generation of young boys and girls that a man's reputation is more important than the lives of the women he's ruined.''
\end{enumerate}

\section{Fuzzy Whiskers}
\begin{enumerate}
    \item ``I don't know what to tell you. I'm happy for the first time in my life and I'm not gonna feel bad about it. It takes a long time to realise how truly miserable you are and even longer to see it doesn't have to be that way. Only after you give up everything can you begin to find a way to be happy.''
\end{enumerate}

\section{Sextina Aquafina}
\begin{enumerate}
    \item ``I think about my child's heartbeat and oh, it makes me weep. I hope and pray to God my little fetus has a soul. Because I want it to feel pain when I eject it from my hole.''
\end{enumerate}

\section{Miscellaneous}
\begin{enumerate}
    \item ``There has been another mass shooting. I am totally unqualified to cover a news story this important, but as a straight white male, I will plow forward with confidence and assume I'm doing fine!''
    \item ``It gets easier. Every day it gets a little easier. But you gotta do it every day -- that's the hard part. But it does get easier''
    \item ``When you do bad things, you have something you can point to when people eventually leave you. It's not you, you tell yourself, it's that bad thing you did. Do you often keep people at arm's length? Are you afraid of being known and knowing others?''
    \item ``He's probably just razzin' ya. But he's a good dog. All bark, no bite. Oh, sorry! That's a labrador expression. I guess in human terms it would be: he's all talk, no shooting you with an assault rifle.''
    \item ``Do you ever get the feeling that to know you more is to love you less?''
    \item ``It doesn't matter where you are, it's who you are, that's not gonna change whether you're in California or Maine or New Mexico. You know, you can't escape you.''
\end{enumerate}

%------------------------------------------------------------------------------%

\chapter{DARLING in the FRANXX (2018)}

\textbf{References.}
\begin{itemize}
    \item \href{https://darling-in-the-franxx.fandom.com/wiki/List_of_Quotes}{Fandom. A list of quotes from DARLING in the FRANXX}
\end{itemize}

\section{Hiro $\star$ 016}
\begin{enumerate}
    \item ``Our sole reason to exist is to become FRANXX and fight. We were born only for that purpose, raised only for that purpose. But to me, the failure of the group, that's something I can't forgive. With nowhere I can feel like I belong... that's when I met you.'' - Hiro's first narrative in Episode 01 preview
    \item ``It was like I was put under a spell. Her two alluring horns, and my first ever look at a naked girl's body, left me transfixed, unable to take my eyes off of her.'' - Hiro to himself, upon his first meeting with Zero Two
    \item ``I feel myself going deeper inside you!'' - Hiro to Zero Two, piloting Strelizia for the second time 
    \item ``Getting mocked for my lack of a partner would be one thing. But the fact that there was still a part of me that still hoped for one made me feel pathetic.'' - Hiro to himself after first meeting Zero Two
    \item ``The act of bringing two pairs of lips together, which we'd never heard of, was called a kiss, she said.'' - Hiro remembering what a kiss means from Zero Two
    \item  ''My wings exist for you. I'm your partner. I'm not going to leave you alone.'' - Hiro promising to be Zero Two's partner
    \item ``It's only I met you, that I can stand here right now.'' - Hiro to Zero Two
    \item ``When we first met, I couldn't take my eyes off you. You were confident, held your head high, and I found that beautiful.'' - Hiro as he confesses to Zero Two
    \item ``You and I; alone and lonesome. With the wings that I tore apart with my own hands, it will never be possible again to fly away. The days that I could spend as your wing, the sky that I dreamed we could take to some day. It has all started fading away, far, far away.'' - Hiro about wanting to fly with Zero Two
    \item ``And I called you a monster! Now we're even!'' - Hiro to Zero Two when they reconcile
    \item ``I feel the same! Zero Two, I love you too!'' - Hiro declaring his love to Zero Two
    \item ``If Zero Two can't smile for me, than I might as well be dead.'' - Hiro to Goro after announcing he is going to space 
    \item ``Liar. If that were true, than why is your last page left blank? Don't fly off on your own. Please, Zero Two, let me stay with you. Together... let's rewrite that story.'' - Hiro to Zero Two before going into the warp gate
    \item ``The distant skies. Beyond time and distance. An overwhelmingly long journey just for the two of us. You're a part of me. I'm a part of you. I'll remember your warmth, along with the memories we've made together. I'll never let you go again!'' - Hiro and Zero Two about their bond
    \item ``I love you!'' - Hiro confessing to Zero Two
    \item ``Zero Two... we're becoming one. Now you are me... We're probably going to disappear soon but the path we walked on, others will take it from here... I love you, Zero two'' - Hiro's last words, to Zero Two before his death
\end{enumerate}

\section{Zero Two $\star$ 002}
\begin{enumerate}
    \item ``The Jian, also known as “the bird that shares wings,” only possesses one wing. Unless a male and female pair lean on each other and act as one, they're incapable of flight. They're imperfect, incomplete creatures. But, for some reason, their way of life, struck me as profoundly beautiful. It was beautiful, I felt.'' - Zero Two's narrative debut
    \item ``I'm always alone, too. Thanks to these horns.''
    \item ``I'm always alone, too. Thanks to these horns.'' - Zero Two talking to Hiro
    \item ``If you don't belong here, just build a place where you do. If you don't have a partner, find one you. And if you can't, take one by force.'' - Zero Two after Hiro explains his lack of partner
    \item ``Your taste makes my heart race. It bites and lingers... the taste of danger.'' - Zero Two after licking Hiro
    \item ``I think I've taken a liking to you. How would you feel about being my darling?'' - Zero Two to Hiro
    \item ``Let me get a taste of you. After all... you are now my darling!'' - Zero Two declaring Hiro as her darling
    \item ``Its been a long time since I last saw a human cry.'' - Zero Two after seeing Hiro's tears
    \item ``Found you, my darling.'' - Zero Two about Hiro
    \item ``Once we die, we'll just become a statistic. It won't matter what we were called. Just look at this lifeless city. There are no skies or oceans here. It's a one-way street to nowhere. A dead-end life.'' - Zero Two to Hiro while life looking at the city
    \item ``Darling, wanna run away with me? I can get you out of here.'' - Zero Two to Hiro as they look at the inner city
    \item ``You're mine. I love the way you taste. It's true: all the weaklings before died. That's to be expected. But you're special. Believe in me, okay, Darling?'' - Zero Two about Hiro
    \item ``What is human to you people?'' - Zero Two to Ichigo
    \item ``A kiss is something you share with your special someone. Is the one you kissed special to you?'' - Zero Two talking about a kiss with Ichigo
    \item ``Don't worry. We'll always be together until we die.'' - Zero Two to Hiro after the partner shuffle
    \item ``The weak ones die. Big deal.'' - Zero Two to 090 after his confrontation
    \item ``If you have anything you wanna say, you better spit it out while you can. Because you're all going to die sooner or later.'' - Zero Two to Ichigo, Miku, Kokoro and Ikuno during the Boys x Girls fight
    \item ``They're tiny fragments of memories. It's white, cold and filled with things unknown, but it's beautiful. In the outside world that I felt for the first time, I heard a voice calling to me from somewhere. And before my eyes, a warm hand was reaching out for me.'' - Zero Two about her childhood memories
    \item ``The leaving something behind part. My body can't do that. It's wonderful. You're all wonderful. You have the ability to decide your futures with your own hearts.'' - Zero Two to Squad 13
    \item ``The distant skies. Beyond time and distance. An overwhelmingly long journey just for the two of us. You're a part of me. I'm a part of you. I'll remember your warmth, along with the memories we've made together. I'll never let you go again!'' - Zero Two and Hiro about their bond
    \item ``And you are me... It doesn't matter how long it takes, as long as we have souls, I'm sure I will meet you again on Earth and we'll pick up where we left off... I feel the same. I love you, darling.'' - Zero Two's last words before her death 
\end{enumerate}

\section{Ichigo $\star$ 015}
\begin{enumerate}
    \item ``Maybe we can't win alone, but the two of us together can!'' - Ichigo to Goro after swimming in the klaxosaur to rescue him
    \item ``She puts a curse to drain the life out of stamen, the one who's not human. I can't consider her as one of us anymore. Even if she's considered as the key to save the world, even if as a result that I'm hated by the person that I consider the most important.'' - Ichigo about Zero Two devouring her partners
    \item ``I... I love you, Hiro!'' - Ichigo to Hiro when she confessed
    \item ``I'm happy for them.'' - Ichigo about Hiro and Zero Two
    \item ``But, you know, no one can swim in the same river water twice. We must choose our own path. Until then, our lives have just begun.'' - Ichigo to Goro
\end{enumerate}

\section{Goro $\star$ 056}
\begin{enumerate}
    \item ``You are such a complete, utter fool!'' - Goro to Hiro about Hiro choosing to keep riding with Zero Two
    \item ``Presents bring out special feelings. The feeling inspired by receiving something from someone. The feeling you get when you want to give something to someone. And then, there's that feeling when in the end, you couldn't give. I have been wondering for a long time what that emotion was, but it has finally hit me.'' - Goro about his feelings for Ichigo
    \item ``Things keep changing. Until now, food, shelter, a reason to live, even a place to die, everything was only given to us. But, that's over too. Now, we have woken up from that long dream and we will never go back to being CHILDREN. That choice, we made it on our own.'' - Goro after the parasites are abandoned
    \item ``What do you mean, you feel bad? What do you mean, you want us to understand? Have you ever spared a thought for how the people close to you feel? You've always been like this: you run off, make your own decisions, and never stop to think about us. Papa and the adults are gone, but forget freedom, we've got our hands tied on every single thing. Have you stopped thinking about how worried we are about this? When someone collapses, all we can do is watch. And yet we're trying our damnest hard to live on! But here you are, running off to die. There are people doing all they can to support you because they want you to live. And you're trampling their feelings as you leave! What the hell do you want us to understand about you?!'' - Goro snapping at Hiro for wanting to go on a death mission
    \item ``That was never about how everyone felt. It was all about me. I'm the one scared of the world that I chose myself.'' - Goro to Ichigo, about his rant earlier on
    \item ``This is the path we've chosen in order to live. You don't get to fight it, either.'' - Goro to Hiro, after forgiving him
\end{enumerate}

\section{Zorome $\star$ 666}
\begin{enumerate}
    \item ``I've never directly spoken with an adult. But they are always looking after us. If we do our best, they praise us and even give us rewards. If I continue fighting for them, I'm sure that someday I can also become an adult. I've been dreaming forever for this day to come.'' - Zorome about his dream of becoming an adult
    \item ``Hey! Nobody asked for a calm and composed analysis!'' - Zorome to Mitsuru
    \item ``Whoa, Whoa! Did you guys just try to press your bodies together? Are you imitating Hiro and Zero Two?'' - To Kokoro and Mitsuru
    \item ``We won't become ADULTS. We just keep on fighting until we die. That's what we were born and raised for. But, we're about to decide our own future. We're finally ready to take flight. Now is the time for us, nestlings, to leave the nest.'' - Zorome about the future
\end{enumerate}

\section{Miku $\star$ 039}
\begin{enumerate}
    \item ``No ordinary person has horns.'' - Miku to Kokoro about Zero Two
    \item ``Better than the boys we're stuck with'' - To Kokoro about the stamen of Squad 26
    \item ``The beginning of adolescence, it's just a small realization. You can't understand the others, and they misunderstand you. It seems to me there's such a huge divide between us. A change of heart like this makes a lot of things that were obvious, not obvious anymore.'' - Miku prior to the Boys vs Girls conflict
    \item ``Nana... Please help us?!'' - To Nana when the crops die
    \item ``Do you think we will find that? Something that means the world to us, that we'd choose it over everything else?'' - To Zorome, Futoshi, Mitsuru, and Ikuno
\end{enumerate}

\section{Mitsuru $\star$ 326}
\begin{enumerate}
    \item ``I say Hiro's showing real integrity. It must be hard for him to face us after what happened, too. Spare a thought for how he feels.'' - Mitsuru speaking sarcastically about Hiro
    \item ``If you place your hopes in anything, they will be betrayed. Promises will go unfulfilled and faith will let you down.'' - Mitsuru to himself after dreaming about his broken childhood promise
    \item ``In a name, there's power. While we were still young, when all we had was code numbers, you could say we lacked individuality. The one who gave us names was the young Hiro. At that time, you were a guidepost to us. But that's in the past. Right now, I don't expect anything from you.'' - Mitsuru about Hiro
    \item ``Sure, Hiro was special. Even among the double-digits, he was a cut above. We were all certain that he'd be our leader and show us the way. But reality had other ideas. Hiro couldn't become a parasite. He's not who he used to be. We must give up and cut our losses. I don't want to see this pathetic side of Hiro anymore.'' - Mitsuru talking about Hiro to Ichigo
    \item ``Doesn't it scare you to trust someone so much? You're unbelievable.'' - Mitsuru to Kokoro during their first sortie together in Genista
    \item ``I promise to protect Kokoro from now on.'' - Mitsuru's promise to Futoshi to protect Kokoro
    \item ``That's what this is... I love Kokoro?'' - Mitsuru realizing his feelings for Kokoro
    \item ``Don't decide things on your own. I'm here. Lean on me more, Kokoro. I want to make you happy.'' - Mitsuru to Kokoro
    \item ``If I look up, I see thousands of lives shining deep in the sky. So far away and although I try, I just can't reach out. Yet, even with broken wings, we'll fly away once again. For the promise we must keep. For the future we must pass on.'' - Mitsuru and Kokoro about their future
    \item ``Please don't say you have nothing. If you don't have anything, neither do I! But that's not true. I've found my reason to live... in you. I want to protect you and that baby's future. I'm a weakling! And I don't really understand what love is, I don't. But if I'm with you, I can keep walking. That's what I've come to believe. Even without our memories, we can start over.'' - Mitsuru reconciling with Kokoro
    \item ``I read an old book that said how you two feel means you love each other. We want our child, our Ai, to inherit that bond.'' - Mitsuru to Hiro and Zero Two about Ai
\end{enumerate}

\section{Kokoro $\star$ 556}
\begin{enumerate}
    \item ``We were raised in an institution and now we live in the Bird Cage. There are so many things we don't know. Take the sea, for example. How it actually smells, the sound of the washing waves, the fact that sea water tastes salty. All those things that we'd only seen in books and all those scenes, it was all kept intact there.'' - Kokoro about going to the beach for the first time
    \item ``Do you know the language of flowers?'' - Kokoro to Mitsuru during the Boys x Girls fight
    \item ``In the past, having a baby was the most natural thing.'' - Kokoro to Mitsuru about why humans stopped having children
    \item ``Mitsuru, you can rely on others more, you know. I'll believe in you, so believe in me too.'' - Kokoro trying to comfort Mitsuru during their first sortie in Genista together
    \item ``We lived our lives believing our only purpose was to ride the FRANXX into battle. But you know what? That might not be true! We could carry new lives and leave them for the future. When I learned about that, I was very happy.'' - Kokoro to Squad 13
    \item ``Is it wrong to create a new life? Are we not allowed to think of the future?'' - Kokoro to Mitsuru about having a child
    \item ``Together, we'll find happiness.'' - Kokoro to Mitsuru on their wedding day
    \item ``f I look up, I see thousands of lives shining deep in the sky. So far away and although I try, I just can't reach out. Yet, even with broken wings, we'll fly away once again. For the promise we must keep. For the future we must pass on.'' - Kokoro and Mitsuru about their future
    \item ``Why do you call my name even when it hurts you?'' - Kokoro to Mitsuru during the thunderstorm
    \item ``Look, there's your papa.'' - Kokoro to Ai, when Mitsuru arrives
    \item ``Thanks to Hiro and Zero Two, this planet go go back and start over from square one.'' - Kokoro after Hiro and Zero Two's sacrifice
\end{enumerate}

\section{Futoshi $\star$ 214}
\begin{enumerate}
    \item ``If the Plantation has no magma-fuel, the city can't maintain function. That's why there's mining facilities everywhere, to keep digging out the fuel. It's the same with us and food. When we eat lots, it gives us strength. So, let's eat!'' - Futoshi about magma energy providing necessities 
    \item ``Will you promise to be my partner forever?!'' - Futoshi to Kokoro
    \item ``The birdcage is still here. And so are we. The people we were supposed to protect aren't there anymore and our wings are still being repaired, but we believe that the ten of us will all be able to fly again soon. Until then, we'll make it through on our own.'' - Futoshi after Squad 13 is left to fend for themselves
\end{enumerate}

\section{Ikuno $\star$ 196}
\begin{enumerate}
    \item ``We're always fighting our fears. We might end up dragging down our partner. We might end up failing to move our FRANXX correctly. We might become unnecessary. Will we be able to overcome our fears someday?'' - Ikuno about failing as a parasite
    \item ``Everything comes to an end. The only difference is whether it comes sooner or later. And even though we sense the end is near, we spend another day, idling in the cradle of our lives.'' - Ikuno about life
    \item ``A pain. Even if it is, so what?!'' - Ikuno to Alpha after he berated Kokoro
    \item ``I love you, Ichigo.'' - Ikuno confessing her love to Ichigo
    \item ``We will leave the birdcage when the sakura blooms. Together, we had our laughs and we had our clashes. All these memories are as delicate as glass, but they're also the beautiful and invaluable testimony that we were here.'' - Ikuno before leaving Mistilteinn
    \item ``It's different for each person'' - To Miku about finding the most important thing in the world to them
    \item ``I've decided to make most of the time I have left'' - To Hiro and Zero Two about her accelerated aging process
\end{enumerate}

\section{Nana}
\begin{enumerate}
    \item ``These children... they're not like the others.'' - Nana describing the squad after the bathroom fight
    \item ``Don't think you and your squad are special'' - To Kokoro about her breaking the rules
    \item ``CHILDREN are usually obedient, docile and rational. They only live to fulfil their mission: fight the Klaxosaur and protect the ADULTS. Not these kids. Too many times, they don't know what to do with their emotions and get carried away. That's definitely not CHILD-like.'' - Nana talking about Squad 13 
\end{enumerate}

\section{Hachi}
\begin{enumerate}
    \item ``The numbers conclusively prove Code 016's compatibility with Strelizia'' - Upon Hiro suffering no damage from his second sortie
    \item ``Code 016 is hereby appointed Code 002's partner'' - Declaring Hiro as Zero Two's partner
    \item ``We call the training facility for Parasites, the Garden. There, CHILDREN are brought up to develop a sound mind and learn how to behave properly as CHILDREN, they are then pruned and only the Parasites who leave the Garden's nest with their heads high have the honor of defending ADULTS against the Klaxosaur threat.'' - Hachi prior to taking Squad 13 to the Garden
    \item ``That feeling... is having it that wrong? The children's sole purpose in life was to fight and if this is truly their final battle, and in their lives after, Don't you think they'll need someone like you.'' - Hachi to Nana about the children's future
\end{enumerate}

\section{Papa}
\begin{enumerate}
    \item ``My children, you have had the good fortune to be chosen as parasites. Your great predecessors defended out cities and our people, and they splendidly took flight as our representatives. Turn your life into a blaze of glory and shed every last drop of blood you have. I pray that you will become a shining ray of hope for humanity.'' - Papa to Squad 13 During the graduation ceremony
    \item ``VIRM will not die. I'm certain we will face each other again in the apex of evolution. As long as this universe contains a flicker of life.'' - Papa to Hiro and Zero Two after the VIRM planet is destroyed
\end{enumerate}

\section{9'a}
\begin{enumerate}
    \item ``Humans have evolved and cast their reproductive functions aside in the process. Reject that, and we'll all have to go back to conforming to one gender'' - Alpha to Kokoro about her `leaving a mark'
    \item ``The gender is a pain, an annoyance that's only tolerated to operate the FRANXX? And that's all it really is.'' - Alpha criticizing Kokoro
    \item ``Humans have also cast away their tendency to be ruled by emotions like that because it serves zero purpose in life.'' - Alpha to Ikuno after she slaps him
    \item ``Meeting you guys helped me a little about being human, but my home is the battlefield and nowhere else. Tell Iota hi for me. It's been kinda fun, I guess.'' - Alpha's last words before his death
\end{enumerate}

\section{Dr. FRANXX}
\begin{enumerate}
    \item ``Fate is cruel. In order to obtain something, one must lose something else. Humanity has searched and sacrificed a lot for what lies beyond life which is our limit. And then we met her, such a beautiful and perfect being.'' - Dr.FRANXX about humanity and 001
    \item ``You two become the children's new adults'' - Dr. FRANXX's final wish upon his death
\end{enumerate}

\section{New Nana}
\begin{enumerate}
    \item ``It is considered impossible for a pregnant woman to pilot a FRANXX. Past records indicate abortion, that is the removal of the fetus as one of the possible treatments. Please let me know what you decide to do.'' - To Kokoro upon informing her of her pregnancy
\end{enumerate}

\section{Naomi}
\begin{enumerate}
    \item ``I hope you find a good partner, Hiro'' - To Hiro when she leaves
\end{enumerate}

\section{Klaxosaur Princess}
\begin{enumerate}
    \item ``I smell my brethren on you. It seems your sins are far too grave to be punished with mere death.''
    \item ``Ah, the fake that the humans created. A duplicate that does not realize it is being used by the invaders.''
    \item ``I shall devour your body and soul.''
    \item ``Damned human wannabes.'' - 001 about VIRM after killing Tarsier
    \item ``We are the defenders of this planet. Once upon a time, during a long battle, we turned ourselves into an immortal weapon. Except fighting, all was lost, so we went to sleep at the bottom of death, our bond was to prepare for the returning invader.'' - 001 about the war with VIRM
    \item ``Perhaps some lives only shine when in unison with others.'' - 001 about Hiro and Zero Two's bond
    \item ``Is this what living is for you? We believed that by abandoning our ties and embracing solitude, we could perfect ourselves, make ourselves stronger. Very well. I shall give you every ounce of strength that remains within me. Whether or not you can take over the controls will be up to you. I stake this planet's future on you two.''
    \item ``Decide whether you want to fight or accept your ruin.'' - 001's last words before she sacrifices herself and entrusts the world's fate to Hiro and Zero Two
\end{enumerate}

\section{Ai}
\begin{enumerate}
    \item ``Papa, Darling, darling!'' - Ai to Mitsuru
    \item ``Mama?'' - Ai to Kokoro
\end{enumerate}


%------------------------------------------------------------------------------%

\chapter{Hagane no renkinjutsushi $\star$ Fullmetal Alchemist: Brotherhood (2009--2012)}

%------------------------------------------------------------------------------%

\chapter{Kimetsu No Yaiba $\star$ Demon Slayer (2019--)}

\section{Tanjiro Kamado}

\section{Inosuke Hashibira}

\section{Zenitsu Agatsuma}

\section{Nezuko Kamado}

\section{Akaza}
\begin{enumerate}
    \item ``\textit{I hate weak people. Weak people$\ldots$ never fight face to face. And poison well. Despicable. Weak people$\ldots$ They lack patience. They will quickly reap what they sow. I killed people with these `protecting fists'. I blooded my master's previous Soryuu style. I couldn't respect my father's last words. That's right. That's who I wanted to kill.}'' - Akaza, before committing suicide, Chap. 155.
\end{enumerate}

%------------------------------------------------------------------------------%

\chapter{Shigatsu wa kimi no uso $\star$ Your Lie in April (2014--2015)}

\section{Arima Kousei}
\begin{enumerate}
    \item ``Spring will be here soon. Spring, the season I met you, is coming. A Spring without you... is coming.''
    \item ``Everything you say and do... it all sparkles so brightly. It's too blinding for me, and I end up closing my eyes. But I can't help aspiring to be like you.''
    \item ``Maybe... just maybe, the light can reach even the bottom of a dark ocean.''
    \item ``Music speaks louder than words.''
    \item ``For you, I am casting about for an excuse again.''
    \item ``This silence belongs to us... and every single person out there, is waiting for us to fill it with something.''
    \item ``This silence belongs to us. Every single person here... is waiting for us to start producing sounds.''
    \item ``Isn't it funny how the most unforgettable scenes can be so trivial?''
    \item ``A lump of steel, like a shooting star. Just seeing the same sky as you makes familiar scenery look different. I swing between hope and despair at your slightest gesture, and my heart starts to play a melody. What kind of feeling is this again? What do they call this kind of feeling? I think it's probably... called love. I'm sure this is what they call love.''
    \item ``Sure I'm okay. Because that's how I was built, after all.''
    \item ``The piano is meant to be played like you're embracing it, right?''
    \item ``You know, I discovered something. Everyone has something... Something deep inside their hearts. For some, it might have been enmity. For others, admiration. Wishes, a craving for the spotlight, feelings that one wants to deliver, feelings for one's mother. Everyone was supported by their own feelings. I realize now that, perhaps, no one can stand alone on stage.''
    \item ``You're like a cat. If I get close, you'll ignore me and go far away. If I get hurt, you'll play around to share the pain.''
    \item ``The more I concentrate, the more I get consumed by my performance. The sounds I play fade away from my reach, tangling up like flowers seized by the spring wind and vanish.''
    \item ``But I wasn't slacking at all. I practiced till I passed out. I gave my all. And if I still turned in a sloppy performance after all that, then... That's who I am now. It's the current me, playing with everything I've got.''
    \item ``The moment I met her, my life changed. Everything I saw, everything I heard, everything I felt, all the scenery around me... started to take on color.''
    \item ``I want to hear it again, yet I don't want to hear it. I want to see her, yet I don't want to see her. What do you call this kind of feeling again?''
    \item ``As if you can see right through me, into my heart... Always, out of nowhere, you... just show up.''
    \item ``She's merciless. That unbending gaze even from the back, she won't let me give up. The one who was being supported... was me. Thank you. Thank you.''
    \item ``A single petal that drifted into my life. The worst first impression ever. The girl who likes my best friend. Will it reach her? I hope it reaches her.''
    \item ``You exist inside a spring that can't be replaced.''
    \item ``Naïve... Bizarre... it's like I'm on a rollercoaster. I'm being jerked around and around. It's like this girl herself is the journey with no clear destination. You're freedom itself.''
    \item ``That moment, when my music reached them... there's no way I could ever forget that. Because I'm a musician, just like you.''
    \item ``She moves me. With such power, like the pounding of my heart. I can hear your sound. You're here.''
    \item ``Because of music, I was given the chance to meet others. I was moved by those encounters. There are people I got to meet. I got to discover emotion. These are all... memories that my mother, who taught me how to play the piano, left me.''
    \item ``Just one person matters to me. Only you matter.''
    \item ``You're in love with food, you're in love with the violin, you're in love with music. I guess that's why you sparkle.''
    \item ``After struggling, losing my way, and suffering... the answer I arrived at was so laughably simple...''
    \item ``This is all your fault. Because you put me back on the stage. Always... you move me. I'm going to prove it. That I'm incredible. That Kaori Miyazono, who's named me to be her accompanist... is even more incredible.''
    \item ``I'm... going on a journey. The applause raining down. Pursuing that moment when my music reached them. Pursuing that sight of her with her back to me. until one day, for sure, I've pulled even with her... until that day comes.''
    \item ``There's no-one who'd ever fall in love with me.''
    \item ``To me it all looks like it's in monotone. Just like music scores... just like a keyboard.''
    \item ``Setbacks come with the territory of becoming a superstar.''
    \item ``Even it the depths of the darkest oceans, some light always pierces through.''
    \item ``I knew all along. The ghost of my mother was a shadow of my own creation. An excuse for me to run away. My own weakness. Mom isn't there anymore. Mom... is inside me.''
    \item ``The way I touch the keys, the way I move my fingers, my habit of squeezing the pedals, my tastes, the order that I eat... Mom's in every little gesture of mine. We're... Mom and I... are connected.''
    \item ``I'm a guy who hurled my precious sheet music away. I don't deserve to be a performer.''
    \item ``I look like I'm suffering, huh? That's not good... but of course I'd be suffering. I mean, I'm gonna sail in charted waters, right? Both, taking on a challenge and creating something. It is painful, but it's fulfilling. So thank you. For sweeping away the dust that had collected on my body. .. For encountering me... ever since that day... my world, even the keyboard... became colorful.''
    \item ``Spring will be here soon. Spring, the season I met you, is coming. A Spring without you... is coming.''
    \item ``That smile of yours, who flew threw the window and died. I'll never forget it.''
    \item ``I couldn't ask her the reason for her tears.''
    \item ``How can I forgot about you, when everything about you, already became a part of me?''
    \item ``Hold on. Don't go! Let's argue again. I'll bribe you with a canalé. I'll call you to kill time. I don't mind being Friend A. Please don't go. Please don't go. Please don't go...please don't leave me behind...''
    \item ``Nothing better than memories and there's nothing worse than them.''
    \item ``You gave me forever within the numbered days. And I'm grateful.''
    \item ``I just really wish that you were her. To compose another memory. I promise when I'm missing you. I'll play your symphony.''
    \item ``What did you have in your heart? What did you lean on?''
    \item ``One day in April, I met a really weird violist. Totally outrageous. Self-righteous. But the smile she shows to people, she is like angelic.''
    \item ``I have no idea what lies ahead, but I've taken the first steps.''
    \item ``Ever since that day I met you, the world became colorful.''
\end{enumerate}

\section{Kaori Miyazono}
\begin{enumerate}
    \item ``Was I able to live inside someone's heart? Was I able to live inside your heart? Do you think you'll remember me at least a little? You'd better not hit ''reset!'' Don't forget me, okay? That's a promise, okay? I'm glad it's you, after all. Will I reach you? I hope I can reach you.''
    \item ``If you can't move with your hands then play with your feet! If you don't have enough fingers, then use your nose as well. Whether you're sad, you're a mess, or you've hit rock bottom, you still have to play. That's how people like us survive.''
    \item ``Maybe there's only a dark road ahead. But you still have to believe and keep going. Believe that the stars will light your path, even a little bit. Come on... Let's go on a journey!''
    \item ``Mozart's telling us from up the sky... ``Go on a journey,'' he's saying.''
    \item So ephemeral and weak. But it's shining with all its might. Thump, Thump, like a heartbeat. This is the light of life.''
    \item ``Such a cruel boy. Telling me to dream one more time. I thought I was satisfied because my dream had come true... And I'd told myself it was enough... Yet here you are, watering this withered heart again.''
    \item ``We're all afraid, you know.. to get up on stage. Maybe you'll mess up. Maybe they'll totally reject you. Even so, you grit your teeth and get up on stage anyway.''
    \item ``Music is freedom.''
    \item ``By exchanging notes, you get to know one another, to understand one another. As if your souls were connected and your hearts were overlapping. It's a conversation through instruments. A miracle that creates harmony. In that moment, music transcends words.''
    \item ``Do you think you will be able to forget?''
    \item ``Memories are always special. Sometimes we smile by remembering the days we cried. And cry by remembering the days we smile.''
    \item ``Even if you know what's coming you're never prepared for the feels.''
    \item ``You were my greatest hello. My saddest goodbye and the biggest what if. I will question it for the rest of my life.''
    \item ``Don't worry if you're rejected. They maybe the people you want not the people you need.''
    \item ``I want you to smile everytime you think about me and in your smile I will live forever.''
    \item ``Sometimes laughter is not the best medicine it can also be a disguise.''
    \item ``You and I, we have music in our bones.''
\end{enumerate}

\section{Tsubaki Sawabe}
\begin{enumerate}
    \item ``Even though I'm bitter over losing, even though I'm depressed, even though my ankle hurts, and my eyes are smeared with tears... even though I've never felt worse... I wonder why the stars are sparkling like this.'' 
    \item ``Sure, I know that I have no right to be feeling this way. But I still don't like it! I just don't like it. We were always together. I was always by his side. During time of joy, and grief as well. But... I realize he's far away from me now... I'm not by his side... there's somebody else there.''
    \item ``The boy I took for granted would always be my side, the boy I want to be by my side forever. I'm such an idiot.''
    \item ``I wish time would just stand still.''
\end{enumerate}

\section{Hiroko Seto}
\begin{enumerate}
    \item ``We're all connected. Just like the notes are intermittently connected. It's shared by us all. Through music, with the people you know, with the people you don't know, with all the people in this world.''
    \item ``Defiance toward parents is an establishment of one's self; it's a sign of independence.''
    \item It's not just allies who support each other. From your enemies, you learn so much and gain so much. Until the day you meet again... Just knowing they exist helps you to withstand the loneliness. Those who compete, even if they're enemies, help each other out.''
    \item ``As a musician, in the process of learning from a teacher, your differences breed a sense of discomfort which you should cherish. It's because of those differences that we have individuality.''
    \item ``When you say you ``can't hear the sound'', doesn't it really mean you ``aren't restrained by the sound''? Rather than the sound you hear with your ears, an image inside you is boiling up from the depths of yourself and taking over without you even knowing. The sound inside, the landscape in your heart, your wishes, a sound loaded with your thoughts; didn't you feel it, even for a moment? ``Not being able to hear the sound.'' That is a gift.''
\end{enumerate}

\section{Watari Ryouta}
\begin{enumerate}
    \item ``Setbacks are inevitable to superstars. Adversity is what separates the good from the great. After all, stars can only shine during the night.''
    \item ``It's only natural for the girl you're crushing on to be in love with someone else. Since you're in love with her, she sparkles in your eyes. That's why people fall so irrationally in love.''
\end{enumerate}

\section{Emi Igawa}
\begin{enumerate}
    \item``The music felt like April.''
\end{enumerate}

%------------------------------------------------------------------------------%

\chapter{Shingeki no kyojin $\star$ Attack on Titan (2013--2021)}

%------------------------------------------------------------------------------%

\chapter{Monster (2004--2010)}

%------------------------------------------------------------------------------%

\chapter{Mushoku-tensei $\sim$Isekai ittara honki dasu $\star$ Mushoku Tensei: Jobless Reincarnation (2021--)}

%------------------------------------------------------------------------------%

\chapter{Stein; Gate (2011--2015)}

%------------------------------------------------------------------------------%

\chapter{Violet Evergarden (2018)}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}