\documentclass{book}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/nguyen/1_NQBH/math/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\addtolength{\headheight}{0pt} % obsolete
\lhead{\scriptsize \chaptername~\thechapter}
\rhead{\itshape \scriptsize \nouppercase{\leftmark}} %\nouppercase !
\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\cfoot{\thepage}
\def\labelitemii{$\circ$}

\title{Legends}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
\tableofcontents

%------------------------------------------------------------------------------%

\part{\href{https://en.wikipedia.org/wiki/Mathematician}{Mathematicians}}

\chapter{\href{https://en.wikipedia.org/wiki/Laurent_Schwartz}{Wikipedia/Laurent Schwartz}}

\textbf{Laurent Schwartz.}
\begin{itemize}
    \item \textbf{Born.} Mar 5, 1915. \href{https://en.wikipedia.org/wiki/Paris}{Paris}, France.
    \item \textbf{Died.} Jul 4, 2002 (aged 87). Paris, France.
    \item \textbf{Nationality.} French
    \item \textbf{Alma mater.} \href{https://en.wikipedia.org/wiki/%C3%89cole_Normale_Sup%C3%A9rieure}{Ecole Normale Sup\'erieure}
    \item \textbf{Known for.}
    \begin{itemize}
        \item \href{https://en.wikipedia.org/wiki/Distribution_(mathematics)}{Theory of Distributions}
        \item \href{https://en.wikipedia.org/wiki/Schwartz_kernel_theorem}{Schwartz kernel theorem}
        \item \href{https://en.wikipedia.org/wiki/Schwartz_space}{Schwartz space}
        \item \href{https://en.wikipedia.org/wiki/Schwartz-Bruhat_function}{Schwartz-Bruhat function}
        \item \href{https://en.wikipedia.org/wiki/Radonifying_function}{Radonifying operator}
        \item \href{https://en.wikipedia.org/wiki/Cylinder_set_measure}{Cylinder set measure}
    \end{itemize}
    \item \textbf{Awards.} \href{https://en.wikipedia.org/wiki/Fields_Medal}{Fields Medal} (1950)
\end{itemize}
\textbf{Scientific career.}
\begin{itemize}
    \item \textbf{Fields.} Mathematics
    \item \textbf{Institutions.}
    \begin{itemize}
        \item \href{https://en.wikipedia.org/wiki/University_of_Strasbourg}{University of Strashbourg}
        \item \href{https://en.wikipedia.org/wiki/University_of_Nancy}{University of Nancy}
        \item \href{https://en.wikipedia.org/wiki/University_of_Grenoble}{University of Grenoble}
        \item \href{https://en.wikipedia.org/wiki/%C3%89cole_Polytechnique}{\'Ecole Polytechnique}
        \item \href{https://en.wikipedia.org/wiki/Universit%C3%A9_de_Paris_VII}{Universit\'e de Paris VII}
    \end{itemize}
    \item \textbf{Doctoral advisor.} \href{https://en.wikipedia.org/wiki/Georges_Valiron}{Georges Valiron}
    \item \textbf{Doctoral students.}
    \begin{itemize}
        \item \href{https://en.wikipedia.org/wiki/Maurice_Audin}{Maurice Audin}
        \item \href{https://en.wikipedia.org/wiki/Georges_Glaeser}{Georges Glaeser}
        \item \href{https://en.wikipedia.org/wiki/Alexander_Grothendieck}{Alexander Grothendieck}
        \item \href{https://en.wikipedia.org/wiki/Jacques-Louis_Lions}{Jacques-Louis Lions}
        \item \href{https://en.wikipedia.org/wiki/Bernard_Malgrange}{Bernard Malgrange}
        \item \href{https://en.wikipedia.org/wiki/Andr%C3%A9_Martineau}{Andr\'e Martineau}
        \item \href{https://en.wikipedia.org/wiki/Bernard_Maurey}{Bernard Maurey}
        \item \href{https://en.wikipedia.org/wiki/Leopoldo_Nachbin}{Leopoldo Nachbin}
        \item \href{https://en.wikipedia.org/wiki/Henri_Hogbe_Nlend}{Henri Hogbe Nlend}
        \item \href{https://en.wikipedia.org/wiki/Gilles_Pisier}{Gilles Pisier}
        \item \href{https://en.wikipedia.org/wiki/Fran%C3%A7ois_Treves}{Fran\c{c}ois Treves}
    \end{itemize}
    \item \textbf{Influenced.} \href{https://en.wikipedia.org/wiki/Per_Enflo}{Per Enflo}
\end{itemize}
\textit{Laurent-Moïse Schwartz} (Mar 5, 1915 - Jul 4, 2002) was a French mathematician.

He pioneered the \href{https://en.wikipedia.org/wiki/Theory}{theory} of \href{https://en.wikipedia.org/wiki/Distribution_(mathematics)}{distributions}, which gives a well-defined meaning to objects such as the \href{https://en.wikipedia.org/wiki/Dirac_delta_function}{Dirac delta function}.

He was awarded the \href{https://en.wikipedia.org/wiki/Fields_Medal}{Fields Medal} in 1950 for his work on the \href{https://en.wikipedia.org/wiki/Distribution_(mathematics)}{theory of distributions}.

For several years he taught at the \href{https://en.wikipedia.org/wiki/%C3%89cole_polytechnique}{École polytechnique}.
    
\section{Biography}

\subsection{Family}
Laurent Schwartz came from a Jewish family of \href{https://en.wikipedia.org/wiki/Alsace}{Alsatian} origin, with a strong scientific background: his father was a well-known \href{https://en.wikipedia.org/wiki/Surgeon}{surgeon}, his uncle \href{https://en.wikipedia.org/wiki/Robert_Debr%C3%A9}{Robert Debré} (who contributed to the creation of \href{https://en.wikipedia.org/wiki/UNICEF}{UNICEF}) was a famous \href{https://en.wikipedia.org/wiki/Pediatrics}{pediatrician}, and his great-uncle-in-law, \href{https://en.wikipedia.org/wiki/Jacques_Hadamard}{Jacques Hadamard}, was a famous mathematician.

%
During his training at \href{https://en.wikipedia.org/wiki/Lyc%C3%A9e_Louis-le-Grand}{Lycée Louis-le-Grand} to enter the \href{https://en.wikipedia.org/wiki/%C3%89cole_Normale_Sup%C3%A9rieure}{École Normale Supérieure}, he fell in love with \href{https://en.wikipedia.org/wiki/Marie-H%C3%A9l%C3%A8ne_Schwartz}{Marie-Hélène Lévy}, daughter of the probabilist \href{https://en.wikipedia.org/wiki/Paul_L%C3%A9vy_(mathematician)}{Paul Lévy} who was then teaching at the \href{https://en.wikipedia.org/wiki/%C3%89cole_polytechnique}{École polytechnique}.
    
Later they would have 2 children, Marc-André and Claudine.

Marie-Hélène was gifted in mathematics as well, as she contributed to the geometry of singular analytic spaces and taught at the \href{https://en.wikipedia.org/wiki/Universit%C3%A9_Lille_Nord_de_France}{University of Lille}.

%
Angelo Guerraggio describes ``Mathematics, politics and butterflies'' as his ``3 great loves''.[1]

\subsection{Education}
According to his teachers, Schwartz was an exceptional student.

He was particularly gifted in Latin, Greek and mathematics.

1 of his teachers told his parents: ``\textit{Beware, some will say your son has a gift for languages, but he is only interested in the scientific and mathematical aspect of languages: he should become a mathematician}.''

%
In 1934, he was admitted at the École Normale Supérieure, and in 1937 he obtained the \href{https://en.wikipedia.org/wiki/Agr%C3%A9gation}{agrégation} (with rank 2).

\subsection{World War II}
As a man of \href{https://en.wikipedia.org/wiki/Trotskyism}{Trotskyist} affinities and \href{https://en.wikipedia.org/wiki/Jew}{Jewish} descent, life was difficult for Schwartz during \href{https://en.wikipedia.org/wiki/World_War_II}{World War II}.

He had to hide and change his identity to avoid being \href{https://en.wikipedia.org/wiki/Deportation}{deported} after Nazi Germany overran France.

He worked for the \href{https://en.wikipedia.org/wiki/University_of_Strasbourg}{University of Strasbourg} (which had been relocated in \href{https://en.wikipedia.org/wiki/Clermont-Ferrand}{Clermont-Ferrand} because of the war) under the name of Laurent-Marie Sélimartin, while Marie-Hélène used the name Lengé instead of Lévy.

Unlike other mathematicians at Clermont-Ferrand such as \href{https://en.wikipedia.org/wiki/Jacques_Feldbau}{Feldbau}, the couple managed to escape the Nazis.

\subsection{Later career}
Schwartz taught mainly at \href{https://en.wikipedia.org/wiki/%C3%89cole_Polytechnique}{École Polytechnique}, from 1958 to 1980.
    
At the end of the war, he spent one year in \href{https://en.wikipedia.org/wiki/Grenoble}{Grenoble} (1944), then in 1945 joined the University of \href{https://en.wikipedia.org/wiki/Nancy,_France}{Nancy} on the advice of \href{https://en.wikipedia.org/wiki/Jean_Delsarte}{Jean Delsarte} and \href{https://en.wikipedia.org/wiki/Jean_Dieudonn%C3%A9}{Jean Dieudonné}, where he spent 7 years.
    
He was both an influential researcher and teacher, with students such as \href{https://en.wikipedia.org/wiki/Bernard_Malgrange}{Bernard Malgrange}, \href{https://en.wikipedia.org/wiki/Jacques-Louis_Lions}{Jacques-Louis Lions}, \href{https://en.wikipedia.org/wiki/Fran%C3%A7ois_Bruhat}{François Bruhat} and \href{https://en.wikipedia.org/wiki/Alexander_Grothendieck}{Alexander Grothendieck}.
    
He joined the science faculty of the \href{https://en.wikipedia.org/wiki/University_of_Paris}{University of Paris} in 1952.

In 1958 he became a teacher at the \href{https://en.wikipedia.org/wiki/%C3%89cole_polytechnique}{École polytechnique} after having at 1st refused this position.
    
From 1961 to 1963 the École polytechnique suspended his right to teach, because of his having signed the \href{https://en.wikipedia.org/wiki/Manifesto_of_the_121}{Manifesto of the 121} about the \href{https://en.wikipedia.org/wiki/Algerian_war}{Algerian war}, a gesture not appreciated by Polytechnique's military administration.

However, Schwartz had a lasting influence on mathematics at the École polytechnique, having reorganized both teaching and research there.

In 1965 he established the \href{https://en.wikipedia.org/wiki/Centre_de_math%C3%A9matiques_Laurent-Schwartz}{Centre de mathématiques Laurent-Schwartz} (CMLS) as its 1st director.

%
In 1973 he was elected corresponding member of the \href{https://en.wikipedia.org/wiki/French_Academy_of_Sciences}{French Academy of Sciences}, and was promoted to full membership in 1975.

\section{Mathematical legacy}
In 1950 at the \href{https://en.wikipedia.org/wiki/International_Congress_of_Mathematicians}{International Congress of Mathematicians}, Schwartz was a plenary speaker[Schwartz, Laurent (1950). ``\textit{Théorie des noyaux}'' (PDF). In: Proceedings of the International Congress of Mathematicians, Cambridge, Massachusetts, U.S.A., Aug 30--Sep 6, 1950. vol. 1. pp. 220--230.] and was awarded the \href{https://en.wikipedia.org/wiki/Fields_Medal}{Fields Medal} for his work on \href{https://en.wikipedia.org/wiki/Distribution_(mathematics)}{distributions}.

He was the 1st French mathematician to receive the Fields medal.

Because of his sympathy for \href{https://en.wikipedia.org/wiki/Trotskyism}{Trotskyism}, Schwartz encountered serious problems trying to enter the United States to receive the medal; however, he was ultimately successful.

%
The theory of distributions clarified the (then) mysteries of the \href{https://en.wikipedia.org/wiki/Dirac_delta_function}{Dirac delta function} and \href{https://en.wikipedia.org/wiki/Heaviside_step_function}{Heaviside step function}.

It helps to extend the theory of \href{https://en.wikipedia.org/wiki/Fourier_transform}{Fourier transforms} and is now of critical importance to the theory of \href{https://en.wikipedia.org/wiki/Partial_differential_equation}{partial differential equations}.

\section{Popular science}
Throughout his life, Schwartz actively worked to promote science and bring it closer to the general audience.

Schwartz said:
\begin{quotation}\it
    ``What are mathematics helpful for? Mathematics are helpful for physics.
    
    Physics helps us make fridges.
    
    Fridges are made to contain spiny lobsters, and spiny lobsters help mathematicians who eat them and have hence better abilities to do mathematics, which are helpful for physics, which helps us make fridges which$\ldots$''[3]
\end{quotation}

\section{Entomology}
\textsf{\href{https://en.wikipedia.org/wiki/Clanis_schwartzi}{Clanis schwartzi} Paratype \href{https://en.wikipedia.org/wiki/MHNT}{MHNT}.}

His mother, who was passionate about natural science, passed on her taste for \href{https://en.wikipedia.org/wiki/Entomology}{entomology} to Laurent.

His personal collection of 20,000 \href{https://en.wikipedia.org/wiki/Lepidoptera}{Lepidoptera} specimens, collected during his various travels was bequeathed to the \href{https://en.wikipedia.org/wiki/National_Museum_of_Natural_History_(France)}{Muséum national d'histoire naturelle}), the \href{https://en.wikipedia.org/wiki/Mus%C3%A9e_des_Confluences}{Science Museum of Lyon}, the \href{https://en.wikipedia.org/wiki/Mus%C3%A9um_de_Toulouse}{Museum of Toulouse} and the Museo de Historia Natural Alcide d'Orbigny in \href{https://en.wikipedia.org/wiki/Cochabamba}{Cochabamba} (Bolivia).
    
Several species discovered by Schwartz bear his name.

\section{Personal ideology}
Apart from his scientific work, Schwartz was a well-known outspoken \href{https://en.wikipedia.org/wiki/Intellectual}{intellectual}.

As a young socialist influenced by \href{https://en.wikipedia.org/wiki/Leon_Trotsky}{Leon Trotsky}, Schwartz opposed the \href{https://en.wikipedia.org/wiki/Totalitarianism}{}{totalitarianism} of the \href{https://en.wikipedia.org/wiki/Soviet_Union}{Soviet Union}, particularly under \href{https://en.wikipedia.org/wiki/Joseph_Stalin}{Joseph Stalin}.

Schwartz ultimately rejected \href{https://en.wikipedia.org/wiki/Trotskyism}{Trotskyism} for \href{https://en.wikipedia.org/wiki/Democratic_socialism}{democratic socialism}.

%
On his religious views, Schwartz called himself an atheist.[4]

\section{Books}

\subsection{Research articles}
\begin{itemize}
    \item \textit{Œuvres scientifiques. I}.
    
    With a general introduction to the works of Schwartz by Claude Viterbo and an appreciation of Schwartz by Bernard Malgrange.
    
    With 1 DVD.
    
    Documents Mathématiques (Paris), 9. Société Mathématique de France, Paris, 2011. x+523 pp. ISBN 978-2-85629-317-1
    \begin{quotation}
        the 1st half of his works in analysis and partial differential equations.
        
        After a preface by Claude Viterbo, which includes a few photos, one will find a note by Schwartz himself about his works, followed by a few original documents (letters, course notes), a presentation by Bernard Malgrange of the theory of distributions for which Schwartz received the Fields Medal in 1950, and a selection of articles covering the period 1944--1954.
    \end{quotation}
    \item \textit{Œuvres scientifiques. II}.
    
    With an appreciation of Schwartz by Alain Guichardet.
    
    With 1 DVD.
    
    Documents Mathématiques (Paris), 10.
    
    Société Mathématique de France, Paris, 2011. x+507 pp. ISBN 978-2-85629-318-8
    \begin{quotation}
        the 2nd half of his works in analysis and partial differential equations.
        
        After a note by Alain Guichardet on Schwartz and his seminars, one will find a selection of articles covering the period 1954--1966.
    \end{quotation}
    \item \textit{Œuvres scientifiques. III}.
    
    With appreciations of Schwartz by Gilles Godefroy and Michel Émery.
    
    With 1 DVD.
    
    Documents Mathématiques (Paris), 11. Société Mathématique de France, Paris, 2011. x+619 pp. ISBN 978-2-85629-319-5
    \begin{quotation}
        his works on Banach space theory (1968--1987), introduced by Gilles Godefroy, and on probability theory (1970--1996), presented by Michel Émery, as well as some articles of a historical nature (1955--1994).
    \end{quotation}
\end{itemize}

\subsection{Technical books}
\begin{itemize}
    \item \textit{Analyse hilbertienne}. Collection Méthodes. Hermann, Paris, 1979. ii+297 pp. ISBN 2-7056-5897-1
    \item \textit{Application of distributions to the theory of elementary particles in quantum mechanics}. Gordon and Breach, New York, NY, 1968. 144pp. ISBN 9780677300900
    \item \textit{Cours d'analyse. 1}. 2nd edition. Hermann, Paris, 1981. xxix+830 pp. ISBN 2-7056-5764-9
    \item \textit{Cours d'analyse. 2}. 2nd edition. Hermann, Paris, 1981. xxiii+475+21+75 pp. ISBN 2-7056-5765-7
    \item [5] \textit{Étude des sommes d'exponentielles. 2ième éd}. Publications de l'Institut de Mathématique de l'Université de Strasbourg, V. Actualités Sci. Ind., Hermann, Paris 1959 151 pp.
    \item \textit{Geometry and probability in Banach spaces}. Based on notes taken by Paul R. Chernoff. Lecture Notes in Mathematics, 852. Springer-Verlag, Berlin-New York, 1981. x+101 pp. ISBN 3-540-10691-X
    \item \textit{Lectures on complex analytic manifolds}. With notes by M. S. Narasimhan. Reprint of the 1955 edition. Tata Institute of Fundamental Research Lectures on Mathematics and Physics, 4. Published for the Tata Institute of Fundamental Research, Bombay; by Springer-Verlag, Berlin, 1986. iv+182 pp. ISBN 3-540-12877-8
    \item \textit{Mathematics for the physical sciences}. Hermann, Paris; Addison-Wesley Publishing Co., Reading, Mass.-London-Don Mills, Ont. 1966 358 pp.
    \item \textit{Radon measures on arbitrary topological spaces and cylindrical measures}. Tata Institute of Fundamental Research Studies in Mathematics, No. 6. Published for the Tata Institute of Fundamental Research, Bombay by Oxford University Press, London, 1973. xii+393 pp.
    \item \textit{Semimartingales and their stochastic calculus on manifolds}. Edited and with a preface by Ian Iscoe. Collection de la Chaire Aisenstadt. Presses de l'Université de Montréal, Montreal, QC, 1984. 187 pp. ISBN 2-7606-0660-0
    \item \textit{Semi-martingales sur des variétés, et martingales conformes sur des variétés analytiques complexes}. Lecture Notes in Mathematics, 780. Springer, Berlin, 1980. xv+132 pp. ISBN 3-540-09749-X
    \item Les tenseurs. \textit{Suivi de ``Torseurs sur un espace affine'' by Y. Bamberger and J.-P. Bourguignon}. 2nd edition. Hermann, Paris, 1981. i+203 pp. ISBN 2-7056-1376-5
    \item [6] \textit{Théorie des distributions}. Publications de l'Institut de Mathématique de l'Université de Strasbourg, No. IX-X. Nouvelle édition, entiérement corrigée, refondue et augmentée. Hermann, Paris 1966 xiii+420 pp.
\end{itemize}

\subsection{Seminar notes}
\begin{itemize}
    \item \textit{Séminaire Schwartz in Paris 1953 bis 1961}. Online edition: [1]
\end{itemize}

\subsection{Popular books}
\begin{itemize}
    \item \textit{Pour sauver l'université.} Editions du Seuil, 1983. 122 pp. ISBN 2020065878
    \item \textit{A mathematician grappling with his century}. Translated from the 1997 French original by Leila Schneps. Birkhäuser Verlag, Basel, 2001. viii+490 pp. ISBN 3-7643-6052-6
\end{itemize}

\section{See also}
\begin{itemize}
    \item \href{https://en.wikipedia.org/wiki/Schwartz_distribution}{Schwartz distribution}
    \item \href{https://en.wikipedia.org/wiki/Schwartz_kernel_theorem}{Schwartz kernel theorem}
    \item \href{https://en.wikipedia.org/wiki/Schwartz_space}{Schwartz space}
    \item \href{https://en.wikipedia.org/wiki/Schwartz-Bruhat_function}{Schwartz-Bruhat function}
    \item \href{https://en.wikipedia.org/wiki/Nicolas_Bourbaki}{Nicolas Bourbaki}\hfill$\square$
\end{itemize}

%------------------------------------------------------------------------------%

\chapter{\href{https://en.wikipedia.org/wiki/Jacques-Louis_Lions}{Wikipedia/Jacques-Louis Lions}}

\textbf{Jacques-Louis Lions.}
\begin{itemize}
    \item \textbf{Born.} May 3, 1928. \href{https://en.wikipedia.org/wiki/Grasse}{Grasse}, \href{https://en.wikipedia.org/wiki/Alpes-Maritimes}{Alpes-Maritimes}, \href{https://en.wikipedia.org/wiki/France}{France}.
    \item \textbf{Died.} May 17, 2001 (aged 73)
    \item \textbf{Nationality.} French.
    \item \textbf{Alma mater.} \href{https://en.wikipedia.org/wiki/University_of_Nancy}{University of Nancy}.
    \item \textbf{Known for.} PDEs.
    \item \textbf{Awards.} \href{https://en.wikipedia.org/wiki/Japan_Prize}{Japan Prize} (1991).
\end{itemize}
\textbf{Scientific career.}
\begin{itemize}
    \item \textbf{Fields.} Mathematics.
    \item \textbf{Institutions.}
    \begin{itemize}
        \item \href{https://en.wikipedia.org/wiki/%C3%89cole_Polytechnique}{\'Ecole Polytechnique}
        \item \href{https://en.wikipedia.org/wiki/Coll%C3%A8ge_de_France}{Coll\`ege de France}
    \end{itemize}
    \item \textbf{\href{https://en.wikipedia.org/wiki/Doctoral_advisor}{Doctoral advisor}.} \href{https://en.wikipedia.org/wiki/Laurent_Schwartz}{Laurent Schwartz}
    \item \textbf{Doctoral students.}
    \begin{itemize}
        \item \href{https://en.wikipedia.org/wiki/Alain_Bensoussan}{Alain Bensoussan}
        \item \href{https://en.wikipedia.org/wiki/Jean-Michel_Bismut}{Jean-Michel Bismut}
        \item \href{https://en.wikipedia.org/wiki/Ha%C3%AFm_Brezis}{Ha\"im Brezis}
        \item \href{https://en.wikipedia.org/wiki/Erol_Gelenbe}{Erol Gelenbe}
        \item \href{https://en.wikipedia.org/wiki/Roland_Glowinski}{Roland Glowinski}
        \item \href{https://en.wikipedia.org/wiki/Roger_Temam}{Roger Temam}
    \end{itemize}
\end{itemize}
\textit{Jacques-Louis Lions} ([1] 3 May 1928 - May 17, 2001) was a French mathematician who made contributions to the theory of \href{https://en.wikipedia.org/wiki/Partial_differential_equation}{partial differential equations} and to \href{https://en.wikipedia.org/wiki/Stochastic_processes}{stochastic control}, among other areas.

He received the \href{https://en.wikipedia.org/wiki/Society_for_Industrial_and_Applied_Mathematics}{SIAM}'s \href{https://en.wikipedia.org/wiki/John_von_Neumann_Lecture}{John von Neumann Lecture} prize in 1986 and numerous other distinctions.[2][3]

Lions is listed as an \href{https://en.wikipedia.org/wiki/ISI_highly_cited_researcher}{ISI highly cited researcher}.[4] 

\section{Biography}
After being part of the French Résistance in 1943 and 1944, J.-L. Lions entered the \href{https://en.wikipedia.org/wiki/%C3%89cole_Normale_Sup%C3%A9rieure}{École Normale Supérieure} in 1947.
    
He was a professor of mathematics at the Université of Nancy, the Faculty of Sciences of Paris, and the \href{https://en.wikipedia.org/wiki/%C3%89cole_polytechnique}{École polytechnique}.
    
%
In 1966 he sent an invitation to \href{https://en.wikipedia.org/wiki/Gury_Marchuk}{Gury Marchuk}, the soviet mathematician to visit Paris.

This was hand delivered by \href{https://en.wikipedia.org/wiki/General_De_Gaulle}{General De Gaulle} during his visit to \href{https://en.wikipedia.org/wiki/Akademgorodok}{Akademgorodok} in June of that year.[5]

%
He joined the prestigious \href{https://en.wikipedia.org/wiki/Coll%C3%A8ge_de_France}{Collège de France} as well as the French Academy of Sciences in 1973.
    
In 1979, he was appointed director of the Institut National de la Recherche en Informatique et Automatique (\href{https://en.wikipedia.org/wiki/INRIA}{INRIA}), where he taught and promoted the use of numerical simulations using finite elements integration.

Throughout his career, Lions insisted on the \textit{use of mathematics in industry}, with a particular involvement in the French space program, as well as in domains such as energy and the environment.

This eventually led him to be appointed director of the Centre National d'Etudes Spatiales (\href{https://en.wikipedia.org/wiki/CNES}{CNES}) from 1984 to 1992.

%
Lions was elected President of the \href{https://en.wikipedia.org/wiki/International_Mathematical_Union}{International Mathematical Union} in 1991 and also received the \href{https://en.wikipedia.org/wiki/Japan_Prize}{Japan Prize} and the \href{https://en.wikipedia.org/wiki/Harvey_Prize}{Harvey Prize} that same year.[3]

In 1992, the \href{https://en.wikipedia.org/wiki/University_of_Houston}{University of Houston} awarded him an honorary doctoral degree.

He was elected president of the \href{https://en.wikipedia.org/wiki/French_Academy_of_Sciences}{French Academy of Sciences} in 1996 and was also a Foreign Member of the \href{https://en.wikipedia.org/wiki/Royal_Society}{Royal Society} (ForMemRS)[6] and numerous other foreign academies.[2][3]

%
He has left a considerable body of work, among this more than 400 scientific articles, 20 volumes of mathematics that were translated into English and Russian, and major contributions to several collective works, including the 4000 pages of the monumental \textit{Mathematical analysis and numerical methods for science and technology} (in collaboration with Robert Dautray), as well as the \textit{Handbook of numerical analysis} in 7 volumes (with \href{https://en.wikipedia.org/wiki/Philippe_G._Ciarlet}{Philippe G. Ciarlet}).

%
His son \href{https://en.wikipedia.org/wiki/Pierre-Louis_Lions}{Pierre-Louis Lions} is also a well-known mathematician who was awarded a \href{https://en.wikipedia.org/wiki/Fields_Medal}{Fields Medal} in 1994.[7]

Both father and son have received honorary doctorates from \href{https://en.wikipedia.org/wiki/Heriot-Watt_University}{Heriot-Watt University} in 1986 and 1995 respectively.[8]

\section{Books}
\begin{itemize}
    \item with Enrico Magenes: \textit{Problèmes aux limites non homogènes et applications}. 3 vols., 1968, 1970
    \item \textit{Contrôle optimal de systèmes gouvernés par des équations aux dérivées partielles}. 1968
    \item with L. Cesari: \textit{Quelques méthodes de résolution des problèmes aux limites non linéaires}. 1969
    \item with Roger Dautray: \textit{Mathematical analysis and numerical methods for science and technology}. 9 vols., 1984/5
    \item with Philippe Ciarlet: \textit{Handbook of numerical analysis}. 7 vols.
    \item with Alain Bensoussan, Papanicolaou: \textit{Asymptotic analysis of periodic structures}. North Holland 1978
    \item \textit{Controlabilité exacte, perturbations et stabilisation de systèmes distribués}[9]
    \item with John E. Lagnese: \textit{Modelling Analysis and Control of Thin Plates}.\hfill$\square$
\end{itemize}

%------------------------------------------------------------------------------%

\chapter{\href{https://en.wikipedia.org/wiki/Roger_Temam}{Wikipedia/Roger Temam}}

\textbf{Roger Meyer Temam.}
\begin{itemize}
    \item \textbf{Born.} May 19, 1940 (age 80)
    \item \textbf{Nationality.} French
    \item \textbf{Alma mater.} \href{https://en.wikipedia.org/wiki/University_of_Paris}{University of Paris}
    \item \textbf{Known for.} \href{https://en.wikipedia.org/wiki/Navier-Stokes_equations}{Navier-Stokes equations}
\end{itemize}
\textbf{Scientific career.}
\begin{itemize}
    \item \textbf{Fields.} \href{https://en.wikipedia.org/wiki/Applied_mathematics}{Applied mathematics.}
    \item \textbf{Institutions.}
    \begin{itemize}
        \item \href{https://en.wikipedia.org/wiki/Paris-Sud_University}{Paris-Sud University} (\href{https://en.wikipedia.org/wiki/Orsay}{Orsay}).
        \item \href{https://en.wikipedia.org/wiki/Indiana_University_Bloomington}{Indiana University}
    \end{itemize}
    \item \textbf{Doctoral advisor.} \href{https://en.wikipedia.org/wiki/Jacques-Louis_Lions}{Jacques-Louis Lions}
    \item \textbf{Doctoral students.}
    \begin{itemize}
        \item \href{https://en.wikipedia.org/wiki/Etienne_Pardoux}{Etienne Pardoux}
        \item \href{https://en.wikipedia.org/wiki/Denis_Serre}{Denis Serre}
    \end{itemize}
\end{itemize}
\textit{Roger Meyer Temam} (born May 19, 1940) is a French applied mathematician working in \href{https://en.wikipedia.org/wiki/Numerical_analysis}{numerical analysis}, \href{https://en.wikipedia.org/wiki/Nonlinear_partial_differential_equations}{nonlinear partial differential equations} and \href{https://en.wikipedia.org/wiki/Fluid_mechanics}{fluid mechanics}.

He graduated from the \href{https://en.wikipedia.org/wiki/University_of_Paris}{University of Paris} - the \href{https://en.wikipedia.org/wiki/Sorbonne}{Sorbonne} in 1967, completing a doctorate (\textit{thèse d'Etat}) under the direction of \href{https://en.wikipedia.org/wiki/Jacques-Louis_Lions}{Jacques-Louis Lions}.

He has published over 400 articles, as well as 12 (authored or co-authored) books.

\section{Scientific work}
The 1st work of Temam in his thesis dealt with the \textit{fractional steps method}.

Thereafter, ``he has continually explored and developed new directions and techniques'':[2]
\begin{itemize}
    \item \href{https://en.wikipedia.org/wiki/Calculus_of_variations}{calculus of variations}, and the notion of \textit{duality} (book \#7), developing the mathematical framework for \textit{discontinuous} (in displacement) \textit{solutions}; a concept later used for his works on the \textit{mathematical theory of plasticity} (book \#5);
    \item mathematical formulation of the equilibrium of a plasma in a cavity, expressed as a nonlinear \href{https://en.wikipedia.org/wiki/Free_boundary_problem}{free boundary problem};[R. Temam, A nonlinear eigenvalue problem: the shape at equilibrium of a confined plasma, \textit{Arch. Rational Mech. Anal.}, 60, 1975, 51--73.]
    \item \href{https://en.wikipedia.org/wiki/Korteweg-de_Vries_equation}{Korteweg-de Vries equation};[R. Temam, Sur un problème non linéaire, \textit{J. Math. Pures Appl.}, 48, 1969, 159--172.]
    \item \href{https://en.wikipedia.org/wiki/Kuramoto%E2%80%93Sivashinsky_equation}{Kuramoto-Sivashinsky equation};[5]
    \item \href{https://en.wikipedia.org/wiki/Euler_equations}{Euler equations} in a bounded domain;[R. Temam, On the Euler equations of incompressible perfect fluids, \textit{J. Funct. Anal.}, 20, 1975, 32--43.]
    \item infinite-dimensional \href{https://en.wikipedia.org/wiki/Dynamical_systems}{dynamical systems} theory.
    
    In particular, he studied the existence of the finite-dimensional global \href{https://en.wikipedia.org/wiki/Attractor}{attractor} for many dissipative equations of mathematical physics, including the incompressible \href{https://en.wikipedia.org/wiki/Navier-Stokes_equations}{Navier-Stokes equations}.[P. Constantin, C. Foias, O. Manley and R. Temam, Determining modes and fractal dimension of turbulent flows, \textit{J. Fluid Mech.}, 150, 1985, 427--440.][C. Foias, O.P. Manley and R. Temam, Physical estimates of the number of degrees of freedom in free convection, \textit{Phys. Fluids}, 29, 1986, 3101--3103.]
    
    He was also the co-founder of the notion of \textit{inertial manifolds}[C. Foias, G.R. Sell and R. Temam, Inertial manifolds for nonlinear evolutionary equations, \textit{J. Diff. Equ.}, 73, 1988, 309--353.] together with Ciprian Foias and \href{https://en.wikipedia.org/wiki/George_R._Sell}{George R. Sell} and of exponential attractors[A. Eden, C. Foias, B. Nicolaenko and R. Temam, \textit{Exponential attractors for dissipative evolution equations}, Collection Recherches en Mathématiques Appliquées, Masson, Paris, and John Wiley, England, 1994.] together with Alp Eden, Ciprian Foias and Basil Nicolaenko;[2]
    \item \href{https://en.wikipedia.org/wiki/Optimal_control}{optimal control} of the incompressible Navier-Stokes equations as a tool for the \textit{control of turbulence};[F. Abergel and R. Temam, On some control problems in fluid mechanics, \textit{Theoret. Comput. Fluid Dynamics}, 1, 1990, 303--325.]
    \item \href{https://en.wikipedia.org/wiki/Boundary_layer}{boundary layer} phenomena for incompressible flows.[12]
\end{itemize}
Temam's main activities concern the study of \href{https://en.wikipedia.org/wiki/Geophysical_fluid_dynamics}{geophysical flows}, the atmosphere and oceans.[2]

This started in the 1990s by collaboration with Jacques-Louis Lions and Shouhong Wang.[J.L. Lions, R. Temam and S. Wang, New formulations of the primitive equations of the atmosphere and applications, \textit{Nonlinearity}, 5, 1992, 237--288.][J.L. Lions, R. Temam and S. Wang, On the equations of the large-scale ocean, \textit{Nonlinearity}, 5, 1992, 1007--1053.][M. Coti Zelati, M. Frémond, R. Temam and J. Tribbia, Uniqueness, regularity and maximum principles for the equations of the atmosphere with humidity and saturation, \textit{Physica D}, 264, 2013, 49-65, https://doi.org/10.1016/j.physd.2013.08.007][Y. Cao, M. Hamouda, R. Temam, J. Tribbia and X. Wang, The equations of the multi-phase humid atmosphere expressed as a quasi variational inequality, \textit{Nonlinearity}, 31, 2018, 4692-4723, https://doi.org/10.1088/1361-6544/aad525.]

%
According to the \href{https://en.wikipedia.org/wiki/Mathematical_Genealogy_Project}{Mathematical Genealogy Project} database,[17][18] he holds the first position in the top 50 advisors.

More than 30 of his students are now full professors all over the world, and have themselves many descendants.[19]

\section{Administrative activities}
Temam became a professor at the \href{https://en.wikipedia.org/wiki/Paris-Sud_University}{Paris-Sud University} at Orsay in 1968.

There, he co-founded the Laboratory of Numerical and Functional Analysis which he directed from 1972 to 1988.

He was also a \textit{Maître de Conférences} at the \href{https://en.wikipedia.org/wiki/Ecole_Polytechnique}{Ecole Polytechnique} in Paris from 1968 to 1986.[20]

%
In 1983, Temam co-founded the French \href{https://en.wikipedia.org/wiki/Soci%C3%A9t%C3%A9_de_Math%C3%A9matiques_Appliqu%C3%A9es_et_Industrielles}{Société de Mathématiques Appliquées et Industrielles} (SMAI), analogous to the \href{https://en.wikipedia.org/wiki/Society_for_Industrial_and_Applied_Mathematics}{Society for Industrial and Applied Mathematics} (SIAM), and served as its 1st president.[21]
    
He was also 1 of the founders of the \href{https://en.wikipedia.org/wiki/International_Congress_on_Industrial_and_Applied_Mathematics}{International Congress on Industrial and Applied Mathematics} (ICIAM) series and was the chair of the steering committee of the 1st ICIAM meeting held in Paris in 1987; and the chair of the standing committee of the 2nd ICIAM meeting held in Washington, D.C., in 1991.[22]

He was the Editor-in-Chief of the mathematical journal M2AN[23] from 1986 to 1997.

%
Temam has been the Director of the Institute for Scientific Computing \& Applied Mathematics (ISCAM)[24] at \href{https://en.wikipedia.org/wiki/Indiana_University_Bloomington}{Indiana University} since 1986 (co-director with Ciprian Foias from 1986 to 1992).

He is also a College Professor (part-time till 2003) and he has been a Distinguished Professor since 2014.[25]

\section{Books}
\begin{enumerate}
    \item (with G.-M. Gie, M. Hamouda and C.-Y. Jung): \textit{Singular perturbations and boundary layers}, Springer-Verlag, New-York, 2018.
    \item (with A. Miranville): \textit{Mathematical Modelling in Continuum Mechanics}, Cambridge University Press, 2001. French Translation, Springer-Verlag France, 2002. Chinese Translation, Tsinghua University Press, 2004. 2nd English Edition 2005. Russian translation, Moskva Linom, 2013.
    \item (with T. Dubois and F. Jauberteau): \textit{Dynamic, multilevel methods and the numerical simulation of turbulence}; Cambridge University Press, 1999.
    \item \textit{Infinite Dimensional Dynamical Systems in Mechanics and Physics}, Springer-Verlag, New-York, Applied Mathematical Sciences Series, vol. 68, 1988. 2nd augmented edition, 1997. Reprinted in China by Beijing World Publishing Corp., 2000.
    \item \textit{Mathematical Problems in Plasticity}, Gauthier-Villars, Paris, 1983 (in French). English Transl., Gauthier-Villars, New-York, 1985. Russian Transl., Nauk, Moscow, 1991. ``Republished by Dover books in Physics, 2018.''
    \item \textit{Navier-Stokes Equations}, North-Holland Pub. Company, in English, 1977, 500 pages. Revised editions 1979, 1984 and 1985. Russian Translation, Mir, Moscow, 1981. ``Republished in the AMS-Chelsea Series, AMS, Providence, 2001.''
    \item (with I. Ekeland): \textit{Convex Analysis and Variational Problems}. Dunod, Paris, 1974, 350 pages (in French). English Translation, North-Holland, Amsterdam, 1976. Russian Translation, Mir, Moscow, 1979. ``English version republished in the Series 'Classics in Applied Mathematics', SIAM, Philadelphia, 1999.''
\end{enumerate}

\section{Awards \& honors}
\begin{itemize}
    \item Fellow of the \href{https://en.wikipedia.org/wiki/American_Academy_of_Arts_and_Sciences}{American Academy of Arts and Sciences} (2015),[26] of the \href{https://en.wikipedia.org/wiki/American_Mathematical_Society}{American Mathematical Society} (2013),[27] of the American Association for the Advancement of Science (2011),[28] of the Society for Industrial and Applied Mathematics (2009).[29]
    \item Knight of the \href{https://en.wikipedia.org/wiki/Legion_of_Honour}{Legion of Honor}, France, 2012.[30]
    \item Member of the \href{https://en.wikipedia.org/wiki/French_Academy_of_Sciences}{French Academy of Sciences} since 2007.[31]\hfill$\square$
\end{itemize}

%------------------------------------------------------------------------------%

\chapter{\href{https://en.wikipedia.org/wiki/Haim_Brezis}{Wikipedia/Ha\"im Brezis}}

\textbf{Haïm Brezis.}
\begin{itemize}
    \item \textbf{Born.} Jun 1m 1944 (age 76). Riom-ès-Montagnes, Cantal, France.
    \item \textbf{Nationality.} French
    \item \textbf{Alma mater.} \href{https://en.wikipedia.org/wiki/University_of_Paris}{University of Paris}
    \item \textbf{Known for.}
    \begin{itemize}
        \item \href{https://en.wikipedia.org/wiki/Brezis-Gallouet_inequality}{Brezis-Gallouet inequality}
        \item \href{https://en.wikipedia.org/wiki/Bony-Brezis_theorem}{Bony-Brezis theorem}
        \item \href{https://en.wikipedia.org/wiki/Brezis-Lieb_lemma}{Brezis-Lieb lemma}
    \end{itemize}
\end{itemize}
\textbf{Scientific career.}
\begin{itemize}
    \item \textbf{Fields.} Mathematics
    \item \textbf{Institutions.} \href{https://en.wikipedia.org/wiki/Pierre_and_Marie_Curie_University}{Pierre and Marie Curie University}
    \item \textbf{Doctoral advisor.}
    \begin{itemize}
        \item \href{https://en.wikipedia.org/wiki/Gustave_Choquet}{Gustave Choquet}
        \item \href{https://en.wikipedia.org/wiki/Jacques-Louis_Lions}{Jacques-Louis Lions}
    \end{itemize}
    \item \textbf{Doctoral students.}
    \begin{itemize}
        \item \href{https://en.wikipedia.org/wiki/Abbas_Bahri}{Abbas Bahri}
        \item \href{https://en.wikipedia.org/wiki/Jean-Michel_Coron}{Henri Berestycki}
        \item \href{https://en.wikipedia.org/wiki/Jean-Michel_Coron}{Jean-Michel Coron}
        \item \href{https://en.wikipedia.org/wiki/Jes%C3%BAs_Ildefonso_D%C3%ADaz}{Jes\'us Ildefonso D\'iaz}
        \item \href{https://en.wikipedia.org/wiki/Pierre-Louis_Lions}{Pierre-Louis Lions}
        \item \href{https://en.wikipedia.org/wiki/Juan_Luis_V%C3%A1zquez_Su%C3%A1rez}{Juan Luis V\'azquez Su\'arez}
    \end{itemize}
\end{itemize}
\textit{Haïm Brezis} (born Jun 1, 1944) is a French mathematician who works in \href{https://en.wikipedia.org/wiki/Functional_analysis}{functional analysis} and \href{https://en.wikipedia.org/wiki/Partial_differential_equation}{partial differential equations}.

\section{Biography}
Born in \href{https://en.wikipedia.org/wiki/Riom-%C3%A8s-Montagnes}{Riom-ès-Montagnes}, \href{https://en.wikipedia.org/wiki/Cantal}{Cantal}, France.
    
Brezis is the son of a Romanian immigrant father, who came to France in the 1930s, and a Jewish mother who fled from the Netherlands.

His wife, \href{https://en.wikipedia.org/wiki/Michal_Govrin}{Michal Govrin}, a native Israeli, works as a novelist, poet, and theater director.[1]

Brezis received his Ph.D. from the \href{https://en.wikipedia.org/wiki/University_of_Paris}{University of Paris} in 1972 under the supervision of \href{https://en.wikipedia.org/wiki/Gustave_Choquet}{Gustave Choquet}.

He is currently a Professor at the \href{https://en.wikipedia.org/wiki/Pierre_and_Marie_Curie_University}{Pierre and Marie Curie University} and a Visiting Distinguished Professor at \href{https://en.wikipedia.org/wiki/Rutgers_University}{Rutgers University}.

He is a member of the \href{https://en.wikipedia.org/wiki/Academia_Europaea}{Academia Europaea} (1988) and a foreign associate of the \href{https://en.wikipedia.org/wiki/United_States_National_Academy_of_Sciences}{United States National Academy of Sciences} (2003).

In 2012 he became a fellow of the \href{https://en.wikipedia.org/wiki/American_Mathematical_Society}{American Mathematical Society}.[2]

He holds honorary doctorates from several universities including \href{https://en.wikipedia.org/wiki/National_Technical_University_of_Athens}{National Technical University of Athens}.[3]

Brezis is listed as an \href{https://en.wikipedia.org/wiki/ISI_highly_cited_researcher}{ISI highly cited researcher}.[4]

\section{Works}
\begin{itemize}
    \item \textit{Opérateurs maximaux monotones et semi-groupes de contractions dans les espaces de Hilbert} (1973)
    \item \textit{Analyse Fonctionnelle}. Théorie et Applications (1983)
    \item \textit{Haïm Brezis. Un mathématicien juif}. Entretien Avec Jacques Vauthier. Collection Scientifiques \& Croyants. Editions Beauchesne, 1999. ISBN 978-2-7010-1335-0, ISBN 2-7010-1335-6
    \item \textit{Functional Analysis, Sobolev Spaces and Partial Differential Equations}, Springer; 1st Edition. edition (November 10, 2010), ISBN 978-0-387-70913-0, ISBN 0-387-70913-4
\end{itemize}

\section{See also}
\begin{itemize}
    \item \href{https://en.wikipedia.org/wiki/Bony%E2%80%93Brezis_theorem}{Bony-Brezis theorem}
    \item \href{https://en.wikipedia.org/wiki/Brezis%E2%80%93Gallouet_inequality}{Brezis-Gallouet inequality}
    \item \href{https://en.wikipedia.org/wiki/Brezis%E2%80%93Lieb_lemma}{Brezis-Lieb lemma}\hfill$\square$
\end{itemize}

%------------------------------------------------------------------------------%



%------------------------------------------------------------------------------%

\part{Physicists}

%------------------------------------------------------------------------------%

\part{Computer Scientists}

%------------------------------------------------------------------------------%

\part{Psychologists}

%------------------------------------------------------------------------------%

\part{Philosophers}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}