\documentclass{article}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/nguyen/1_NQBH/math/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}} % for sections, if you really want! (It is default in report and book class (So you may not need it).
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{notation}{Notation}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem{ques}{Question}
\newtheorem{problem}{Problem}
\newtheorem{conjecture}{Conjecture}
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\def\labelitemii{$\circ$}

\title{Don Miguel Ruiz. \textit{The Four Agreements: A Practical Guide to Personal Freedom}. A Toltec Wisdom Book}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
\tableofcontents
\vspace{5mm}
\noindent
\textbf{Note.} The term ``black magic'' is not meant to convey racial connotation; it is merely used to describe the use of magic for adverse or harmful purposes.

\begin{quotation}
    To the \textit{Circle of Fire}; those who have gone before, those who are present, and those who have yet to come.
\end{quotation}

%------------------------------------------------------------------------------%

\section*{Acknowledgments}
\textsc{I would like to humbly acknowledge my} mother Sarita, who taught me unconditional love; my father Jose Luis, who taught me discipline; my grandfather Leonardo Macias, who gave me the key to unlock the Toltec mysteries; and my sons Miguel, Jose Luis, and Leonardo.

%
I wish to express my deep affection and appreciation to the dedication of Gaya Jenkins and Trey Jenkins.

%
I would like to extend my profound gratitude to Janet Mills - publisher, editor, believe.

I am also abidingly grateful to Ray Chambers for lighting the way.

%
I would like to honor my dear friend Gini Gentry, an amazing ``brain'' whose faith touched my heart.

%
I would like to pay tribute to the many people who have given freely of their time, hearts, and resources to support the teachings.

A partial list includes: $\ldots$ (see \textbf{[Ruiz2018]}).

\section*{The Toltec}
\textsc{Thousands of years ago, the Toltec were} known throughout southern Mexico as ``women and men of knowledge.''

Anthropologists have spoken of the Toltec as a nation or a race, but, in fact, the Toltec were scientists and artists who formed a society to explore and conserve the spiritual knowledge and practices of the ancient ones.

They came together as masters (\textit{naguals}) and students oat Teotihuacan, the ancient city of pyramids outside Mexico City known as the place where ``\textit{Man Becomes God}.''

%
Over the millennia, the \textit{naguals} were forced to conceal the ancestral wisdom and maintain its existence in obscurity.

European conquest, coupled with rampant misuse of personal power by a few of the apprentices, made it necessary to shield the knowledge from those who were not prepare to use it wisely or who might intentionally misuse it for personal gain.

%
Fortunately, the esoteric Toltec knowledge was embodied and passed on through generations by different lineages of \textit{naguals}.

Though it remained veiled in secrecy for hundreds of years, ancient prophecies foretold the coming of an age when it would be necessary to return the wisdom to the people.

Now, don Miguel Ruiz, a \textit{nagual} from the Eagle Knight lineage, has been guided to share with us the powerful teachings of the Toltec.

%
Toltec knowledge arises from the same essential unity of truth as all the sacred esoteric traditions found around the world.

Though it is not a religion, it honors all the spiritual masters who have taught on the earth.

While it does embrace spirit, it is most accurately described as a way of life, distinguished by the ready accessibility of happiness and love.

\section*{Introduction: The Smokey Mirror}
\textsc{3000 years ago, there was a human} just like you and me who lived near a city surrounded by mountains.

The human was studying to become a medicine man, to learn the knowledge of his ancestors, but he didn't completely agree with everything he was learning.

In his heart, he felt there must be something more.

%
1 day, as he slept in a cave, he dreamed that he saw his own body sleeping.

He came out of the cave on the night of a new moon.

The sky was clear, and he could see millions of stars.

Then something happened inside of him that transformed his life forever.

He looked at his hands, he felt his body, and he heard his own voice say, ``I am made of light; I am made of stars.''

%
He looked at the stars again, and he realized that it's not the stars that create light, but rather light that creates the stars.

``Everything is made of light,'' he said, ``and the space in-between isn't empty.''

And he knew that everything that exists in 1 \textit{living being}, and that light is the messenger of life, because it is alive and contains all information.

%
Then he realized that although he was made of stars, he was not those stars.

``I am in-between the stars,'' he thought.

So he called the stars the \textit{tonal} and the light between the stars the \textit{nagual}, and he knew that what created the harmony and space between the two is Life or Intent.

Without Life, the \textit{tonal} and the \textit{nagual} could not exist.

Life is the force of the absolute, the supreme, the Creator who creates everything.

%
This is what he discovered: Everything in existence is a manifestation of the one living being we call God.

\textit{Everything is God}.

And he came to the conclusion that human perception is merely light perceiving light.

He also saw that matter is a mirror - everything is a mirror that reflects light and creates images of that light - and the world of illusion, the \textit{Dream}, is just like smoke which doesn't allow us to see what we really are.

``The real us is pure love, pure light,'' he said.

%
This realization changed his life.

Once he knew what he really was, he looked around at other humans and the rest of nature, and he was amazed at what he saw.

He saw himself in everything - in every human, in every animal, in every tree, in the water, in the rain, in the clouds, in the earth.

And he saw that Life mixed the \textit{tonal} and the \textit{nagual} in different ways to create billions of manifestations of Life.

%
In those few moments he comprehended everything.

He was very excited, and his heart was filled with peace.

He could hardly wait to tell his people what he had discovered.

But there were no words to explain it.

He tried to tell the others, but they could not understand.

They would see that he had changed, that something beautiful was radiating from his eyes and his voice.

They noticed that he no longer had judgment about anything or anyone.

He was no longer like anyone else.

%
He could understand everyone very well, but no one could understand him.

They believed that he was an incarnation of God, and he smiled when he heard this and he said,
\begin{quotation}
    ``It is true.
    
    I am God.
    
    But you are also God.
    
    We are the same, you and I.
    
    We are images of light.
    
    We are God.''
\end{quotation}
But still the people didn't understand him.

%
He had discovered that he was a mirror for the rest of the people, a mirror in which he could see himself.

``Everyone is a mirror,'' he said.

He saw himself in everyone, but nobody saw him as themselves.

And he realized that everyone was dreaming, but without awareness, without knowing what they really are.

They couldn't see him as themselves because there was a wall of fog or smoke between the mirrors.

And that wall of fog was made by the interpretation of images of light - the \fbox{\textit{Dream} of humans}.

%
Then he knew that he would soon forget all that he had learned.

He wanted to remember all the visions he had had, so he decided to call himself the Smokey Mirror so that he would always know that matter is a mirror and the smoke in-between is what keeps us from knowing what we are.

He said, ``I am the Smokey Mirror, because I am looking at myself in all of you, but we don't recognize each other because of the smoke in-between us.

That smoke is the \textit{Dream}, and the mirror is you, the dreamer.''
\begin{quotation}
    ``\textit{Living is easy with eyes closed, misunderstanding all you see}$\ldots$'' - John Lennon
\end{quotation}

\section{Domestication \& the Dream of the Planet}
\textsc{What you are seeing and hearing right now is} nothing but a dream.

You are dreaming right now in this moment.

You are dreaming with the brain awake.

%
Dreaming is the main function of the mind, and the mind dreams 24 hours a day.

It dreams when the brain is awake, and it also dreams when the brain is asleep.

The difference is that when the brain is awake, there is a material frame that makes us perceive things in a linear way.

When we go to sleep we do not have the frame, and the dream has the tendency to change constantly.

%
Humans are dreaming all the time.

Before we were born the humans before us created a big outside dream that we will call society's dream or \textit{the dream of the planet}.

The dream of the planet is the collective dream of billions of smaller, personal dreams, which together create a dream of a family, a dream of a community, a dream of a city, a dream of a country, and finally a dream of the whole humanity.

The dream of the planet includes all of society's rules, its beliefs, its laws, its religions, its different cultures and ways to be, its governments, schools, social events, and holidays.

%
We are born with the capacity to learn how to dream, and the humans who live before us teach us how to dream the way society dreams.

The outside dream has so many rules that when a new human is born, we hook the child's attention and introduce these rules into his or her mind.

The outside dream uses Mom and Dad, the schools, the religion to teach us how to dream.

%
\fbox{\textit{Attention} is the ability we have to discriminate} and to focus only on that which we want to perceive.

We can perceive millions of things simultaneously, but using our attention, we can hold whatever we want to perceive in the foreground of our mind.

The adults around us hooked our attention and put information into our minds through repetition.

That is the way we learned everything we know.

%
By using our attention we learned a whole reality, a whole dream.

We learned how to behave in society: what to believe and what not to believe; what is acceptable and what is not acceptable; what is good and what is bad; what is beautiful and what is ugly; what is right and what is wrong.

It was all there already - all that knowledge, all those rules and concepts about how to behave in the world.

%
When you were in school, you sat in a little chair and put your attention on what the teacher was teaching you.

When you went to church, you put your attention on what the priest or minister was telling you.

It is the same dynamic with Mom and Dad, brothers and sisters: They were all trying to hook your attention.

We also learn to hook the attention of other humans, and we develop a need for attention which can become very competitive.

Children compete for the attention of their parents, their teachers, their friends.

``Look at me!

Look at what I'm doing!

Hey, I'm here.''

The need for attention becomes very strong and continues into adulthood.

%
The outside dream hooks our attention and teaches us what to believe, beginning with the language that we speak.

Language is the code for understanding and communication between humans.

Every letter, every word in each language is an agreement.

We call this a page in a book; the word \textit{page} is an agreement that we understand.

\fbox{\textit{Once we understand the code, our attention is hooked and the energy is transferred from 1 person to another}.}

%
It was not your choice to speak English.

You didn't choose your religion or your moral values - they were already there before you were born.

We never had the opportunity to choose what to believe or what not to believe.

We never chose even the smallest of these agreements.

We didn't even chose our own name.

%
As children, we didn't have the opportunity to choose our beliefs, but we \textit{agreed} with the information that was passed to us from the dream of the planet via other humans.

\fbox{\textit{The only way to store information is by agreement}.}

The outside dream may hook our attention, but if we don't agree, we don't store that information.

As soon as we agree, we \textit{believe} it, and this is called \textit{faith}.

\fbox{\textit{To have faith is to believe unconditionally}.}

%
That's how we learn as children.

Children believe everything adults say.

We agree with them, and our faith is so strong that \textit{the belief system controls our whole dream of life}.

We didn't choose these beliefs, and we may have rebelled against them, but we were not strong enough to win the rebellion.

The result is surrender to the beliefs with out \textit{agreement}.

%
I call this process \textit{the domestication of humans}.

And though this domestication we learn how to live and how to dream.

In human domestication, the information from the outside dream is conveyed to the inside dream, creating our whole belief system.

1st the child is taught the names of things: Mom, Dad, milk, bottle.

Day by day, at home, at school, at church, and from television, we are told how to live, what kind of behavior is acceptable.

\textit{The outside dream teaches us how to be a human}.

We have a whole concept of what a ``woman'' is and what a ``man'' is.

And we also learn to judge: We judge ourselves, judge other people, judge the neighbors.

%
Children are domesticated the same way that we domesticate a dog, a cat, or any other animal.

In order to teach a dog we punish the dog and we give it rewards.

We train our children whom we love so much the same way that we train any domesticated animal: with a system of punishment and reward.

We are told, ``You're a good boy,'' or ``You're a good girl,'' when we do what Mom and Dad want us to do.

When we don't, we are ``a bad girl'' or ``a bad boy.''

%
When we went against the rules we were punished; when we went along with the rules we got a reward.

We were punished many times a day, and we were also rewarded many times a day.

Soon we became afraid of being punished and also afraid of not receiving the reward.

The reward is the attention that we got from our parents or from other people like siblings, teachers, and friends.

\textit{We soon develop a need to hook other people's attention in order to get the reward}.

%
The rewards feels good, and we keep doing what others want us to do in order to get the reward.

With that fear of being punished and that fear of not getting the reward, we start pretending to be what we are not, just to please others, just to be good enough for someone else.

We try to please Mom and Dad, we try to please the teachers at school, we try to please the church, and so we start acting.

\textit{We pretend to be what we are not because we are afraid of being rejected}.

The fear of being rejected becomes the fear of not being good enough.

Eventually we become someone that we are not.

We become a copy of Mamma's beliefs, Daddy's beliefs, society's beliefs, and religion's beliefs.

%
All our normal tendencies are lost in the process of domestication.

And when we are old enough for our mind to understand, we learn the word \textit{no}.

The adults say, ``Don't do this and don't do that.''

We rebel and say, ``No!''

\fbox{\textit{We rebel because we are defending our freedom}.}

We want to be ourselves, but we are very little, and the adults are big and strong.

After a certain time we are afraid because we know that every time we do something wrong we are going to be punished.

%
The domestication is so strong that at a certain point in our lives we no longer need anyone to domesticate us.

We don't need Mom or Dad, the school or the church to domesticate us.

\textit{We are so well trained that we are our own domesticator}.

We are an autodomesticated animal.

We can now domesticate ourselves according to the same belief system we were given, and using the same system of punishment and reward.

We punish ourselves when we don't follow the rules according to our belief system; we reward ourselves when we are the ``good boy'' or ``good girl.''

%
The belief system is like a Book of Law that rules our mind.

Without question, whatever is in that Book of Law, is our truth.

We base all of our judgments according to the Book of Law, even if these judgments go against our own inner nature.

Even moral laws like the Ten Commandments are programmed into our mind in the process of domestication.

1 by 1, all these agreements go into the Book of Law, and these agreements rule our dream.

%
There is something in our minds that judges everybody and everything, including the weather, the dog, the cat - everything.

The \fbox{\textit{inner judges}} uses what is in our Book of Law to judge everything we do and don't do, everything we think and don't think, and everything we feel and don't feel.

Everything lives under the tyranny of this Judge.

Every time we do something that goes against the Book of Law, the Judge says we are guilty, we need to be punished, we should be ashamed.

This happens many times a day, day after day, for all the years of our lives.

%
There is another part of us that receives the judgments, and this part is called the Victim.

The Victim carries the blame, the guilt, and the shame.

It is the part of us that says, ``Poor me, I'm not good enough, I'm not intelligent enough, I'm not attractive enough, I'm not worthy of love, poor me.''

The bid Judge agrees and says, ``yes, you are not good enough.''

And this is all based on a belief system that we never chose to believe.

These beliefs are so strong, that even years later when we are exposed to new concepts and try to make our own decisions, we find that these beliefs still control our lives.

%
Whatever goes against the Book of Law will make you feel a funny sensation in your solar plexus, and it's called \textit{fear}.

Breaking the rules in the Book of Law opens your \textit{emotional wounds}, and \textit{your reaction is to create emotional poison}.

Because everything that is in the Book of Law has to be true, anything that challenges what you believe is going to make you \textit{feel unsafe}.

Even if the Book of Law is wrong, it makes you \textit{feel safe}.

%
That is why we need a great deal of courage to challenge our own beliefs.

Because even if we know we didn't choose all these beliefs, it is also true that we agreed to all of them.

The agreement is so strong that even if we understand the concept of it not being true, we feel the blame, the guilt, and the shame that occur if we go against these rules.

%
Just as the government has a book of laws that rule the society's dream, our belief system is the Book of Laws that rules our personal dream.

All these laws exist in our mind, we believe them, and the Judge inside us bases everything on these rules.

The Judge decrees, and the Victim suffers the guilt and punishment.

But who says there is justice in this dream?

True justice is paying only once for each mistake.

True \textit{injustice} is paying more than once for each mistake.

%
How many times do we pay for 1 mistake?

The answer is thousands of times.

The human is the only animal on earth that pays a thousand times for the same mistake.

The rest of the animals pay once for every mistake they make.

But not us.

We have a powerful memory.

We make a mistake, we judge ourselves, we find ourselves guilty, and we punish ourselves.

If justice exists, then that was enough; we don't need to do it again.

But every time we remember, we judge ourselves again, we are guilty again, and we punish ourselves again, and again, and again.

If we have a wife or husband he or she also reminds us of the mistake, so we can judge ourselves again, punish ourselves again, and find ourselves guilty again.

\textit{Is this fair?}

%
How many times do we make our spouse, our children, or our parents pay for the same mistake?

Every time we remember the mistake, we blame them again and send them all the emotional poison we feel at the justice, and then we make them pay again for the same mistake.

\textit{Is that justice?}

The Judge in the mind is wrong because the belief system, the Book of Law, is wrong.

The whole dream is based on false law.

95\% of the beliefs we have stored in our minds are nothing but lies, and we suffer because we believe al these lies.

%
In the dream of the planet it is normal for humans to suffer, to live in fear, and to create \textit{emotional dramas}.

The outside dream is not a pleasant dream; it is a dream of violence, a dream of fear, a dream of war, a dream of injustice.

The personal dream of humans will vary, but globally it is mostly a nightmare.

If we look at human society we see a place so difficult to live in because it is ruled by fear.

Throughout the world we see human suffering, anger, revenge, addictions, violence in the street, and tremendous injustice.

It may exist at different levels in different counties around the world, but fear is controlling the outside dream.

%
If we compare the dream of human society with the description of hell that religions all around the world have promulgated, we find they are exactly the same.

Religions say that hell is a place of punishment, a place of fear, pain, and suffering, a place where the fire burns you.

Fire is generated by emotions that come from fear.

Whenever we feel the emotions of anger, jealousy, envy, or hate, we experience a fire burning within us.

We are living in a dream of hell.

%
\textit{If you consider hell as a state of mind, the hell is all around us}.

Others may warn us that if we don't do what they say we should do, we will go to hell.

Bad news!

\fbox{\textit{We are already in hell, including the people who tell us that}.}

No human can condemn another to hell because we are already there.

Others can put us into a deeper hell, true.

But only if we allow this to happen.

%
Every human has his or her own personal dream, and just like the society dream, it is \textit{often ruled by fear}.

We learn to dream hell in our own life, in our personal dream.

The same fears manifest in different ways for each person, of course, but we experience anger, jealousy, hate, envy, and other negative emotions.

Our personal dream can also become an ongoing nightmare where we suffer and live in a state of fear.

But we don't need to dream a nightmare.

It is possible to enjoy a pleasant dream.

%
All of humanity is searching for truth, justice, and beauty.

We are on an eternal search for the truth because we only believe in the lies we have stored in our mind.

We are searching for justice because in the belief system we have, there is no justice.

We search for beauty because it doesn't matter how beautiful a person is, we don't believe that person has beauty.

We keep searching and searching, when everything is already within us.

There is no truth to find.

Wherever we turn our heads, all we see is the truth, but with the agreements and beliefs we have stored in our mind, we have no eyes for this truth.

%
\textit{We don't see the truth because we are blind}.

\textit{What blinds us are all those false beliefs we have in our mind}.

We have the need to be right and to make others wrong.

We trust what we believe, and our beliefs set us up for suffering.

It is as if we live in the middle of a fog that doesn't let us see any further than our own nose.

We live in a fog that is not even real.

This fog is a dream, your personal dream of life - what you believe, all the concepts you have about what you are, all the agreements you have made with others, with yourself, and even with God.

%
Your whole mind is a fog which the Toltecs called a \textit{mitote} (pronounced MIH-TOE' -TAY).

\fbox{\textit{Your mind is a dream where a thousand people at the same time, and nobody understands each other}.}

This is the condition of the human kind - a big \textit{minote}, and with that big \textit{mitote} you cannot see what you really are.

In India they call the \textit{mitote maya}, which means \textit{illusion}.

It is the personality's notion of ``I am.''

Everything you believe about yourself and the world, all the concepts and programming you have in your mind, are all the \textit{minote}.

\textit{We cannot see who we truly are; we cannot see that we are not free}.

%
\textit{That is why humans resist life}.

\textit{To be alive is the biggest fear humans have}.

Death is not the biggest fear we have; \textit{our biggest fear is taking the risk to be alive - the risk to be alive and express what we really are}.

Just being ourselves is the biggest fear of humans.

We have learned to live our lives trying to satisfy other people's demands.

We have learned to live by other people's points of view because of the fear of not being accepted and of not being good enough for someone else.

%
During the process of domestication, we form an image of what perfection is in order to try to be good enough.

We create an image of how we should be in order to be accepted by everybody.

We especially try to please the ones who love us, like Mom and Dad, big brothers and sisters, the priests and the teacher.

Trying to be good enough for them, we create an image of perfection, but we don't fit this image.

We create this image, but this image is not real.

We are never going to be perfect from this point of view.

\textit{Never!}

%
\fbox{\textit{Not being perfect, we reject ourselves}.}

And the level fo self-rejection depends upon how effective the adults were in breaking our integrity.

After domestication it is no longer about being good enough for anybody else.

We are not good enough for ourselves because we don't fit with our own image of perfection.

We cannot forgive ourselves for not being what we wish to be, or rather what we \textit{believe} we should be.

We cannot forgive ourselves for not being perfect.

%
We know we are not what we believe we are supposed to be and so we feel false, frustrated, and dishonest.

We try to hide ourselves, and we pretend to be what we are not.

The result is that we feel unauthentic and wear social masks to keep others from noticing this.

We are so afraid that somebody else will notice that we are not what we pretend to be.

We judge others according to our image of perfection as well, and naturally they fall short of our expectations.

%
\textit{We dishonor ourselves just to please other people}.

We even do harm to our physical bodies just to be accepted by others.

You see teenagers taking drugs just to avoid being rejected by other teenagers.

They are not aware that the problem is that they don't accept themselves.

They reject themselves because they are not what they pretend to be.

They wish to be a certain way, but they are not, and for this they carry shame and guilt.

Humans punish themselves endlessly for not being what they believe they should be.

They become very self-abusive, and they use other people to abuse themselves as well.

%
\textit{But nobody abuses us more than we abuse ourselves}, and it is the Judge, the Victim, and the belief system that make us do this.

True, we find people who say their husband or wife, or mother or father, abused them, but you know that we abuse ourselves much more than that.

\textit{The way we judge ourselves is the worst judge that ever existed}.

If we make a mistake in front of people, we try to deny the mistake and cover it up.

But as soon as we are alone, the Judge becomes so strong, the guilt is so strong, and we feel so stupid, or so bad, or so unworthy.

%
\textit{In your whole life nobody has ever abused you more than you have abused yourself}.

\fbox{\textit{And the limit of your self-abuse is exactly the limit that you will tolerate from someone else}.}

\textit{If someone abuses you a little more than you abuse yourself, you will probably walk away from that person}.

\textit{But if someone abuses you a little less than you abuse yourself, you will probably stay in the relationship and tolerate it endlessly}.

%
If you abuse yourself very badly, you can even tolerate someone who beats you up, humiliates you, and treats you like dirt.

\textit{Why?}

Because in your belief system you say, ``I deserve it.

This person is doing me a favor by being with me.

I'm not worthy of love and respect.

I'm not good enough.''

%
\textit{We have the need to be accepted and to be loved by others, but we cannot accept and love ourselves}.

The more self-love we have, the less we will experience self-abuse.

\textit{Self-abuse comes from self-rejection, and self-rejection comes from having an image of what it means to be perfect and never measuring up to that ideal}.

Our image of perfection is the reason we reject ourselves; it is why we don't accept ourselves the way we are, and why we don't accept others the way they are.
\begin{center}
    \textsc{Prelude to a New Dream}
\end{center}
There are thousands of agreements you have made with yourself, with other people, with your dream of life, with God, with society, with your parents, with your spouse, with your children.

But the most important agreements are the ones you made with yourself.

In these agreements you tell yourself \textit{who you are, what you feel, what you believe}, and \textit{how to behave}.

The result is what you call \textit{your personality}.

In these agreements you say, ``This is what I am.

This is what I believe.

I can do certain things, and some things I cannot do.

This is reality, that is fantasy; this is possible, that is impossible.''

%
1 single agreement is not such a problem, but we have many agreements that make us suffer, that make us fail in life.

If you want to live a life of joy and fulfillment, you have to find the courage to break those agreements that are fear-based and claim your \textit{personal power}.

The agreements that come from fear require us to expend a lot of energy, but the agreements that come from love help us to conserve energy and even gain extra energy.

%
Each of us is born with a certain amount of personal power that we rebuild every day after we rest.

Unfortunately, we spend all our personal power 1st to create all these agreements and then to keep these agreements.

Our personal power is dissipated by all the agreements we have created, and the result is that we feel powerless.

We have just enough power to survive each day, because most of it is used to keep the agreements that trap us in the dream of the planet.

\textit{How can we change the entire dream of our life when we have no power to change even the smallest agreement?}

%
If we can see it is our agreements that rule our own life, and we don't like the dream of our life, we need to change the agreements.

When we are finally ready to change our agreements, there are 4 very powerful agreements that will help us to break those agreements that come from fear and deplete our energy.

%
Each time you break an agreement, all the power you used to create it returns to you.

If you adopt these 4 new agreements, they will create enough personal power for you to change the entire system of your old agreements.

%
You need a very strong will in order to adopt The Four Agreements - but if you can begin to live your life with these agreements, the transformation in your life will be amazing.

You will see the drama of hell disappear right before your very eyes.

Instead of living in a dream of hell, you will be creating a new dream - your personal dream of heaven.

\section{\textsc{The 1st Agreement}: Be Impeccable with your Word}
\textsc{The 1st agreement is the most important one} and also the \textit{most difficult} one to honor.

It is so important that with just this 1st agreement you will be able to transcend to the level of existence I call \fbox{\textit{heaven on earth}}.

%
The 1st agreement is to \textit{be impeccable with your word}.

It sounds very simple, but it is very, very powerful.

%
\textit{Why your word?}

\textit{Your word is the power that you have to create}.

\textit{Your word is the gift that comes directly from God}.

The Gospel of John in the Bible, speaking of the creation of the universe, says, ``\textit{In the beginning was the word, and the word was with God, and the word is God}.''

Through the word you express your creative power.

It is through the word that you manifest everything.

Regardless of what language you speak, your intent manifests through the word.

What you dream, what you feel, and what you really are, will all be manifested through the word.

%
The word is not just a sound or written symbol.

\textit{The word is a force}; it is the \textit{power you have to express and communicate, to think, and thereby to create the events in your life}.

You can speak.

What other animal on the planet can speak?

\textit{The word is the most powerful tool you have as a human; it is the tool of magic}.

But like \fbox{\textit{a sword with 2 edges}}, your word can create the most beautiful dream, or your word can destroy everything around you.

\textit{1 edge is the misuse of the word, which creates a living hell}.

\textit{The other edge is the impeccability of the word, which will only create beauty, love, and heaven on earth}.

Depending upon how it is used, the word can set you free, or it can enslave you even more than you know.

\textit{All the magic you possess is based on your word}.

\textit{Your word is pure magic, and misuse of your word is black magic}.

%
The word is so powerful that 1 word can change a life or destroy the lives of millions of people.

Some years ago 1 man in Germany, by the use of the word, manipulated a whole country of the most intelligent people.

He led them into a world war with just the power of his word.

He convinced others to commit the mos atrocious acts of violence.

He activated people's fear with the word, and like a big explosion, there was killing and war all around the world.

All over the world humans destroyed other humans because they were afraid of each other.

\textit{Hitler's word, based on fear-generated beliefs and agreements, will be remembered for centuries}.

%
\textit{The human mind is like a fertile ground where seeds are continually being planted}.

\textit{The seeds are opinions, ideas, and concepts}.

\textit{You plant a seed, a thought, and it grows}.

The word is like a seed, and the human mind is so fertile!

\textit{The only problem is that too often it is fertile for the seeds of fear}.

Every human mind is fertile, but only for those kinds of seeds it is prepared for.

What is important is to see which kind of seeds our mind is fertile for, and to prepare it to receive the seeds of love.

%
Take the example of Hitler: He sent out all those seeds of fear, and they grew very strong and beautifully achieved massive destruction.

Seeing the awesome power of the word, we must understand what power comes out of our mouths.

1 fear or doubt planted in our mind can create an endless drama of events.

1 word is like a spell, and humans use the word like black magicians, thoughtlessly putting spells on each other.

%
\textit{Every human is a magician}, and we can either put a spell on someone with our word or we can release someone from a spell.

\textit{We cast spells all the time with our opinions}.

An example: I see a friend and give him an opinion that just popped into my mind.

I say, ``Hmmm! I see that kind of color in your face in people who are going to get cancer.''

If he listens to the word, and if he agrees, he will have cancer in $< 1$ year.

That is the power of the word.

%
During our domestication, our parents and siblings gave their opinions about us without even thinking.

We believed these opinions and we lived in fear over these opinions, like not being good at swimming, or sports, or writing.

Someone gives an opinion and says, ``Look, this girl is ugly!''

The girl listens, believes she is ugly, and grows up with the idea that she is ugly.

It doesn't matter how beautiful she is; as long as she has that agreement, she will believe that she is ugly.

That is the spell she is under.

%
By hooking our attention, the word can enter our mind and change a whole belief for better or for worse.

Another example: You may believe you are stupid, and you may have believed this for as long as you can remember.

This agreement can be very tricky, causing you to do a lot of things just to ensure that you are stupid.

You may do something and think to yourself, ``I wish I were smart, but I must be stupid or I wouldn't have done that.''

The mind goes in hundreds of different directions, and we could spend days getting hooked by just that 1 belief in our own stupidity.

%
Then 1 day someone hooks your attention and using the word, lets you know that you are not stupid.

You believe what the person says and make a new agreement.

As a result, you no longer feel or act stupid.

A whole spell is broken, just by the power of the word.

Conversely, if you believe you are stupid, and someone hooks your attention and says, ``Yes, you are really the most stupid person I have ever met,'' the agreement will be reinforced and become even stronger.
\begin{center}
    $\star$
\end{center}

Now let us see what the word \textit{impeccability} means.

\textit{Impeccability} means ``without sin.''

\textit{Impeccable} comes from the Latin \textit{pecatus}, which means ``sin.''

The \textit{im} in impeccable means ``without,'' so \textit{impeccable} means ``without sin.''

Religions talk about sin and sinners, but let's understand what it really means to sin.

\textit{A sin is anything that you do which goes against yourself}.

\textit{Everything you feel or believe or say that goes against yourself is a sin}.

You go against yourself when you judge or blame yourself for anything.

Being without sin is exactly the opposite.

Being impeccable is not going against yourself.

When you are impeccable, you take responsibility for your actions, but you do not judge or blame yourself.

%
From this point of view, the whole concept of sin changes from something moral or religious to something commonsense.

\textit{Sin begins with rejection of yourself}.

\fbox{\textit{Self-rejection is the biggest sin that you commit}.}

In religious terms self-rejection is a ``mortal sin,'' which leads to death.

Impeccability, on the other hand, leads to life.

%
\textit{Being impeccable with your word is not using the word against yourself}.

If I see you in the street and I call you stupid, it appears that I'm using the word against you.

But really I'm using my word against myself, because you're going to hate me for this, and your hating me is not good for me.

Therefore, if I get angry and with my word send all that emotional poison to you, I'm using the word against myself.

%
If I love myself I will express that love in my interactions with you, and then I am being impeccable with the word, because that action will produce a like reaction.

If I love you, then you will love me.

If I insult you, you will insult me.

If I have gratitude for you, you will have gratitude for me.

If I'm selfish with you, you will be selfish with me.

\textit{If I use the word to put a spell on you, you are going to put a spell on me}.

%
Being impeccable with your word is the correct use of your energy; it means to use your energy in the direction of truth and love for yourself.

If you make an agreement with yourself to be impeccable with your word, just with that intention, the truth will manifest through you and clean all the emotional poison that exists within you.

But making this agreement is difficult because we have learned to do precisely the opposite.

We have learned to lie as a habit of our communication with others and more importantly with ourselves.

We are not impeccable with the word.

%
\textit{The power of the word is completely misused in hell}.

We use the word to curse, to blame, to find guilt, to destroy.

Of course, we also use it in the right way, but not too often.

Mostly we use the word to spread our personal poison - to express anger, jealousy, envy, and hate.

\fbox{\textit{The word is pure magic - the most powerful gift we have as humans - and we use it against ourselves}.}

We plan revenge.

We create chaos with the word.

We use the word to create hate between different races, between different people, between families, between nations.

We misuse the word so often, and this misuse is how we create and perpetuate the dream of hell.

Misuse of the word is how we pull each other down and keep each other in a state of fear and doubt.

\textit{Because the word is the magic that humans possess and misuse of the word is black magic, we are using black magic all the time without knowing that our word is magic at all}.

%
There was a woman, e.g., who was intelligent and had a very good heart.

She has a daughter whom she adored and loved very much.

1 night she came home from a very bad day at work, tired, full of emotional tension, and with a terrible headache.

She wanted peace and quiet, but her daughter was singing and jumping happily.

The daughter was unaware of how her mother was feeling; she was in her own world, in her own dream.

She felt so wonderful, and she was jumping and singing louder and louder, expressing her joy and her love.

She was singing so loud that it made her mother's headache even worse, and at a certain moment, the mother lost control.

Angrily she looked at her beautiful little girl and said, ``Shut up!

You have an ugly voice.

Can you just shut up!''

%
The truth is that the mother's tolerance for any noise was nonexistent; it was not that the little girl's voice was ugly.

But the daughter believed what her mother said, and in that moment she made an agreement with herself.

After that she no longer sang, because she believed her voice was ugly and would bother anyone who heard it.

She became shy at school, and if she was asked to sing, she refused.

Even speaking to others became difficult for her.

Everything changed in the little girl because of this new agreement: She believed she must repress her emotions in order to be accepted and loved.

%
Whenever we hear an opinion and believe it, we make an agreement, and it becomes part of our belief system.

This little girl grew up, and even though she had a beautiful voice, she never sang again.

She developed a whole complex from 1 spell.

This spell was cast upon her by the one who loved her the most: her own mother.

Her mother didn't notice what she did with her word.

She didn't notice that she used black magic and put a spell on her daughter.

She didn't know the power of her word, and therefore she isn't to blame.

She did what her own mother, father, and others had done to her in many ways.

They misused the word.

%
How many times do we do this with our own children?

We give them these types of opinions and our children carry that black magic for years and years.

People who love us do black magic on us, but they don't know what they do.

\textit{That is why we must forgive them; they don't know what they do}.

%
Another example: You awake in the morning feeling very happy.

You feel so wonderful, you stay 1 or 2 hours in front of the mirror, making yourself beautiful.

Well, 1 of your best friends says, ``What has happened to you?

You look so ugly.

Look at the dress you are wearing; you look ridiculous.''

That's it; that is enough to put you all the way down in hell.

Maybe this girlfriend just told you this to hurt you.

And, she did.

She gave you an opinion with all the power of her word behind it.

If you accept the opinion, it becomes an agreement now, and you put all your power into that opinion.

That opinion becomes black magic.

%
These types of spells are difficult to break.

The only thing that can break a spell is to make a new agreement based on truth.

The truth is the most important part of being impeccable with your word.

On 1 side of the sword are the lies which create black magic, and on the other side of the sword is the truth which has the power to break the spell of black magic.

\fbox{\textit{Only the truth will set us free}.}
\begin{center}
    $\star$
\end{center}

Looking at everyday human interactions, imagine how many times we cast spells on each other with our word.

Over time this interaction has become the worst form of black magic, and we call it \textit{gossip}.

%
\fbox{\textit{Gossip is black magic at its very worst because it is pure poison}.}

We learned how to gossip by agreement.

When we were children, we heard the adults around us gossiping all the time, openly giving their opinions about other people.

They even had opinions about people they didn't know.

Emotional poison was transferred along with the opinions, and we learned this as the normal way to communicate.

%
\textit{Gossiping has become the main form of communication in human society}.

It has become the way we feel close to each other, because it makes us feel better to see someone else feel as badly as we do.

There is an old expression that says, ``\textit{Misery likes company},'' and people who are suffering in hell don't want to be all alone.

Fear and suffering are an important part of the dream of the planet; they are how the dream of the planet keeps us down.

%
Using the analogy of the human mind as a computer, gossip can be compared to a computer virus.

A computer virus is a piece of computer language written in the same language all the other codes are written in, but with a harmful intent.

This code is inserted into the program of your computer when you least expect it and most of the time without your awareness.

After this code has been introduced, your computer doesn't work quite right, or it doesn't function at all because the codes get so mixed up with so many conflicting messages that it stops producing good results.

%
Human gossip works exactly the same way.

E.g., you are beginning a new class with a new teacher and you have looked forward to it for a long time.

On the 1st day of class, you run into someone who took the class before, who tells you, ``Oh that instructor was such a pompous jerk!

He didn't know what he was talking about, and he was a pervert too, so watch out!''

%
You are immediately imprinted with the word and the emotional code the person had when saying this, but what you are not aware of is his or her motivation in telling you.

This person could be angry for failing the class or simply making an assumption based on fears and prejudices, but because you have learned to ingest information like a child, some part of you believes the gossip, and you go on to the class.

As the teacher speaks, you feel the poison come up inside you and you don't realize you see the teacher through the eyes of the person who gave you that gossip.

Then you start talking to other people in the class about this, and they start to see the teacher in the same way: as a jerk and a pervert.

You really hate the class, and soon you decide to drop out.

You blame the teacher, but \b{\textit{it is gossip that is to blame}.}

%
All of this mess can be caused by 1 little computer virus.

1 little piece of misinformation can break down communication between people, causing every person it touches to become infected and contagious to others.

Imagine that every single time others gossip to you, they insert a computer virus into your mind, causing you to think a little less clearly every time.

Then imagine that in an effort to clean up your own confusion and get some relief from the poison, you gossip and spread these viruses to someone else.

%
Now imagine this pattern going on in a never-ending chain between all the humans on earth.

The result is a world full of humans who can only read information through circuits that are clogged with a poisonous, contagious virus.

Once again, this poisonous virus is what the Toltecs called the \textit{mitote}, the chaos of a thousand different voices all trying to talk at once in the mind.

%
Even worse are the black magicians or ``computer hackers'' who intentionally spread the virus.

Think back to a time when you or someone you know was angry with someone else and desired revenge.

In order to seek revenge you said something to or about that person with the intention of spreading poison and making that person feel bad about him- or herself.

As children we do this quite thoughtlessly, but as the grow older we become much more calculated in our efforts to bring other people down.

Then we live to ourselves and say that person received a just punishment for their wrongdoing.

%
\fbox{\textit{When we see the world through a computer virus, it is easy to justify the cruelest behavior}.}

What we don't see is that misuse of our word is putting us deeper into hell.
\begin{center}
    $\star$
\end{center}
For years we have received the gossip and spells from the words of others, but also from the way we use our word with ourselves.

We talk to ourselves constantly and most of the time we say things like, ``Oh, I look fat, I look ugly.

I'm getting old, I'm losing my hair.

I'm stupid, I never understand anything.

I will never be good enough, and I'm never going to be perfect.''

\textit{Do you see how we use the word against ourselves?}

\fbox{We must begin to understand what the word \textit{is} and what the word \textit{does}.}

If you understand the 1st agreement, \textit{be impeccable with your word}, you begin to see all the changes that can happen in your life.

Changes 1st in the way you deal with yourself, and later in the way you deal with other people, especially those you love the most.

%
Consider how many times you have gossiped about the person you love the most to gain the support of others for your point of view.

How many times you have hooked other people's attention, and spread poison about your loved on in order to make your opinion right?

\textit{Your opinion is nothing but your point of view}.

It is not necessarily true.

\textit{Your opinion comes from your beliefs, your own ego, and your own dream}.

We create all this poison and spread it to others just so we can feel right about our own point of view.

%
If we adopt the 1st agreement, and become impeccable with our word, any emotional poison will eventually be cleaned from our mind and from our communication in our personal relationships, including with our pet dog or cat.

%
\textit{Impeccability of the word will also give you immunity from anyone putting a negative spell on you}.

You will only receive a negative idea if your mind is fertile ground for that idea.

When you become impeccable with your word, your mind is no longer fertile ground for words that come from black magic.

Instead, it is fertile for the words that come from love.

\textit{You can measure the impeccability of your word by your level of self-love}.

How much you love yourself and how you feel about yourself are directly proportionate to the quality and integrity of your word.

\textit{When you are impeccable with your word, you feel good; you feel happy and at peace}.

%
You can transcend the dream of hell just by making the agreement to be impeccable with your word.

Right now I am planting that seed in your mind.

Whether or not the seed grows depends upon how fertile your mind is for the seeds of love.

It is up to you to make this agreement with yourself: \textit{I am impeccable with my word}.

Nurture this seed, and as it grows in your mind, it will generate more seeds of love to replace the seeds of fear.

This 1st agreement will change the kind of seeds your mind is fertile for.

%
\textit{Be impeccable with your word}.

This is the 1st agreement that you should make if you want to be free, if you want to be happy, if you want to transcend the level of existence that is hell.

It is very powerful.

Use the word in the correct way.

Use the word to share your love.

Use white magic, beginning with yourself.

Tell yourself how wonderful you are, how great you are.

Tell yourself how much you love yourself.

Use the word to break all those teeny, tiny agreements that make you suffer.

%
It is possible.

It is possible because I did it, and I am no better than you.

No, we are exactly the same.

We have the same kind of brain, the same kind of bodies; we are humans.

If I was able to break those agreements and create new agreements, then you can do the same.

If I can be impeccable with my word, why not you?

Just this 1 agreement can change your whole life.

Impeccability of the word can lead you to personal freedom, to huge success and abundance; it can take away all fear and transform it into joy and love.

%
Just imagine what you can create with impeccability of the word.

With the impeccability of the word you can transcend the dream of fear and live a different life.

\fbox{\textit{You can live in heaven in the middle of thousands of people living in hell because you are immune to that hell}.}

You can attain the kingdom of heaven from this 1 agreement: \textit{Be impeccable with your word}.

\section{\textsc{The 2nd Agreement}: Don't Take Anything Personally}
\textsc{The next 3 agreements are really born} from the 1st agreement.

The 2nd agreement is \textit{don't take anything personally}.

%
Whatever happens around you, don't take it personally.

Using an earlier example, if I see you on the street and I say, ``Hey, you are so stupid,'' without knowing you, it's not about you; it's about me.

If you take it personally, then perhaps you believe you are stupid.

Maybe you think to yourself, ``How does he know?

Is he clairvoyant, or can everybody see how stupid I am?''

%
You take it personally because you agree with whatever was said.

As soon as you agree, the poison goes through you, and you are trapped in the dream of hell.

What causes you to be trapped is what we call \fbox{\textit{personal importance}}.

Personal importance, or taking things personally, is the maximum expression of selfishness because we make the assumption that everything is about ``me.''

During the period of our education, or our domestication, we learn to take everything personally.

\textit{We think we are responsible for everything}.

Me, me, me, always me!

%
Nothing other people do is because of you.

It is because of themselves.

All people live in their own dream, in their own mind; they are in a completely different world from the one we live in.

When we take something personally, we make the assumption that they know what is in our world, and we try to impose our world on their world.

%
Even when a situation seems so personal, even if others insult you directly, it has nothing to do with you.

\textit{What they say, what they do, and the opinions they give are according to the agreements they have in their own minds}.

Their point of view comes from all the programming they received during domestication.

%
If someone gives you an opinion and says, ``Hey, you look so fat,'' don't take it personally, because the truth is that this person is dealing with his or her own feelings, beliefs, and opinions.

That person tried to send poison to you and if you take it personally, then you take that poison and it becomes yours.

Taking things personally makes your easy prey for these predators, the black magicians.

They can hook you easily with 1 little opinion and feed you whatever poison they want, and because you take it personally, you eat it up.

%
You eat all their emotional garbage, and now it becomes your garbage.

But if you do not take it personally, you are immune in the middle of hell.

\fbox{\textit{Immunity to poison in the middle of hell}} is the gift of this agreement.

%
When you take things personally, then you feel offended, and your reaction is to defend your beliefs and create conflicts.

You make something big out of something so little, because you have the need to be right and make everybody else wrong.

You also try hard to be right by giving them your own opinions.

In the same way, whatever you feel and do is just a projection of your own personal dream, a reflection of your own agreements.

What you say, what you do, and the opinions you have are according to the agreements you have made - and these opinions have nothing to do with me.

%
It is not important to me what you think about me, and I don't take what you think personally.

I don't take it personally when people say, ``Miguel, you are the best,'' and I also don't take it personally when they say, ``Miguel, you are the worst.''

I know that when you are happy you will tell me, ``Miguel, you are such an angel!''

But, when you are mad at me you will say, ``Oh, Miguel, you are such a devil!

You are so disgusting.

How can you say those things?''

Either way, it does not affect me because I know what I am.

I don't have the need to be accepted.

I don't have the need to have someone tell me, ``Miguel, you are doing so good!'' or ``How dare you do that!''

%
No, I don't take it personally.

Whatever you think, whatever you feel, I know is your problem and not my problem.

\textit{It is the way you see the world}.

It is nothing personal, because you are dealing with yourself, not with me.

Others are going to have their own opinion according to their belief system, so nothing they think about me is really about me, but it is about them.

%
You may even tell me, ``Miguel, what you are saying is hurting me.''

But it is not what I am saying that is hurting you; it is that you have wounds that I touch by what I have said.

You are hurting yourself.

There is no way that I can take this personally.

Not because I don't believe in you or don't trust you, but because I know that you see the world with different eyes, with your eyes.

\textit{You create an entire picture or movie in your mind, and in that picture you are the director, you are the producer, you are the main actor or actress}.

\textit{Everyone else is a secondary actor or actress}.

It is your movie.

%
The way you see that movie is according to the agreements you have made with life.

Your point of view is something personal to you.

It is no one's truth but yours.

Then, if you get mad at me, I know you are dealing with yourself.

I am the excuse for you to get mad.

And \fbox{\textit{you get mad because you are afraid, because you are dealing with fear}.}

If you are not afraid, there is no way you will get mad at me.

If you are not afraid, there is no way you will hate me.

If you are not afraid, there is no way you will be jealous or sad.

%
If you live without fear, if you love, there is no place for any of those emotions.

If you don't feel any of those emotions, it is logical that you will feel good.

\fbox{\textit{When you feel good, everything around you is good}.}

\textit{When everything around you is great, everything makes you happy}.

\textit{You are loving everything that is around you, because you are loving yourself}.

\textit{Because you like the way you are}.

\textit{Because you are content with you}.

Because you are happy with your life.

You are happy with the movie that you are producing, happy with your agreements with life.

You are at peace, and you are happy.

You live in that state of bliss where everything is so wonderful, and everything is so beautiful.

In that state of bliss you are making love all the time with everything that you perceive.
\begin{center}
    $\star$
\end{center}

Whatever people do, feel, think, or say, \textit{don't take it personally}.

If they tell you how wonderful you are, they are not saying that because of you.

You know you are wonderful.

It is not necessary to believe other people who tell you that you are wonderful.

Don't take \textit{anything} personally.

Even if someone got a gun and shot you in the head, it was nothing personal.

Even at that extreme.

%
Even the opinions you have about yourself are not necessarily true; therefore, you don't need to take whatever you hear in your own mind personally.

\textit{The mind has the ability to talk to itself, but it also has the ability to hear information that is available from other realms}.

Sometimes you hear a voice in your mind, and you may wonder where it came from.

This voice may have come from another reality in which there are living beings very similar to the human mind.

The Toltecs called these beings \textit{Allies}.

In Europe, Africa, and India they called them the Gods.

%
Our mind also exists in the level of the Gods.

\textit{Our mind also lives in that reality and can perceive that reality}.

\textit{The mind sees with the eyes and perceives this waking reality}.

But the mind also sees and perceives without the eyes, although the reason is hardly aware of this perception.

The mind lives in more than 1 dimension.

There may be times when you have ideas that don't originate in your mind, but you are perceiving them with your mind.

You have the right to believe or not believe these voices and the right not to take what they say personally.

We have a choice whether or not to believe the voices we hear within our own minds, just as we have a choice of what to believe and agree with in the dream of the planet.

%
\textit{The mind can also talk and listen to itself}.

\textit{The mind is divided as your body is divided}.

Just as you can say, ``I have 1 hand, and I can shake my other hand and feel the other hand,'' the mind can talk to itself.

\textit{Part of the mind is speaking, and the other part is listening}.

\textit{It is a big problem when a thousand parts of your mind are all speaking at the same time}.

This is called a \textit{minote}, remember?

%
The \textit{mitote} can be compared to a huge marketplace where thousands of people are talking and bartering at the same time.

Each one has different thoughts and feelings; each one has a different point of view.

The programming in the mind - all of those agreements we have made - are not necessarily compatible with each other.

\textit{Every agreement is like a separate living being; it has its own personality and its own voice}.

There are conflicting agreements that go against other agreements and on and on until it becomes a big war in the mind.

The \textit{minote} is the reason humans hardly know what they want, how they want it, or when they want it.

They don't agree with themselves because there are parts of the mind that want 1 thing, and other parts that want exactly the opposite.

%
Some part of the mind has objections to certain thoughts and actions, and another part supports the actions of the opposing thoughts.

All these little living beings create inner conflict because they are alive and they each have a voice.

Only by making an inventory of our agreements will we uncover all of the conflicts in the mind and eventually make order out of the chaos of the \textit{minote}.
\begin{center}
    $\star$
\end{center}

\textit{Don't take anything personally} because by taking things personally you set yourself up to suffer for nothing.

\textit{Humans are addicted to suffering at different levels and to different degrees, and we support each other in maintaining these addictions}.

Humans agree to help each other suffer.

\textit{If you have the need to be abused, you will find it easy to be abused by others}.

Likewise, if you are with people who need to suffer, something in you makes you abuse them.

It is as if they have a note on their back that says, ``Please kick me.''

They are asking for justification for their suffering.

\textit{Their addiction to suffering is nothing but an agreement that is reinforced everyday}.

%
Wherever you go you will find people lying to you, and as your awareness grows, you will notice that you also lie to yourself.

Do not expect people to tell you the truth because they also lie to themselves.

You have to trust yourself and choose to believe or not to believe what someone says to you.

%
\textit{When we really see other people as they are without taking it personally, we can never be hurt by what they say or do}.

Even if others lie to you, it is okay.

They are lying to you because they are afraid.

They are afraid you will discover that they are not perfect.

It is painful to take that social mask off.

If others say 1 thing, but do another, you are lying to yourself if you don't listen to their actions.

\textit{But if you are truthful with yourself, you will save yourself a lot of emotional pain}.

\textit{Telling yourself the truth about it may hurt, but you don't need to be attached to the pain}.

Healing is on the way, and it's just a matter of time before things will be better for you.

%
If someone is not treating you with love and respect, it is a gift if they walk away from you.

If that person doesn't walk away, you will surely endure many years of suffering with him or her.

Walking away may hurt for a while, but your heart will eventually heal.

Then you can choose what you really want.

You will find that you don't need to trust others as much as you need to trust yourself to make the right choices.

%
When you make it a strong habit not to take anything personally, you avoid many upsets in your life.

Your anger, jealousy, and envy will disappear, and even your sadness will simply disappear if you don't take things personally.

%
If you can make this 2nd agreement a habit, you will find that nothing can put you back into hell.

There is a huge amount of freedom that comes to you when you take nothing personally.

You become immune to black magicians, and no spell can affect you regardless of how strong it may be.

The whole world can gossip about you, and if you don't take it personally you are immune.

Someone can intentionally send emotional poison, and if you don't take it personally, you will not eat it.

When you don't take the emotional poison, it becomes even worse in the sender, but not in you.

%
You can see how important this agreement is.

Taking nothing personally helps you to break many habits and routines that trap you in the dream of hell and cause needless suffering.

Just by practicing this 2nd agreement you begin to break dozens of teeny, tiny agreements that cause you to suffer.

And if you practice the 1st 2 agreements, you will break 75\% of the teeny, tiny agreements that keep you trapped in hell.

%
Write this agreement on paper, and put it on your refrigerator to remind you all the time: \textit{Don't take anything personally}.

%
As you make a habit of not taking anything personally, you won't need to place your trust in what others do or say.

You will only need to trust yourself to make responsible choices.

You are never responsible for the actions of others; you are only responsible for you.

When you truly understand this, and refuse to take things personally, you can hardly be hurt by the careless comments or actions of others.

%
If you keep this agreement, you can travel around the world with your heart completely open and no one can hurt you.

You can say, ``I love you,'' without fear of being ridiculed or rejected.

You can ask for what you need.

You can say yes, or you can say no - whatever you choose - without guilt or self-judgment.

You can choose to follow your heart always.

Then you can be in the middle of hell and still experience inner peace and happiness.

You can stay in your state of bliss, and hell will not affect you at all.

\section{\textsc{The 3rd Agreement}: Don't Make Assumptions}
\textsc{The 3rd agreement is: Don't make assumptions}.

%
We have the tendency to make assumptions about everything.

The problem with making assumptions is that we \textit{believe} they are the truth.

We could swear they are real.

We make assumptions about what other are doing or thinking - we take it personally - then we blame them and react by sending emotional poison with our word.

That is \textit{why whenever we make assumptions, we're asking for problems}.

We make an assumption, we misunderstand, we take it personally, and we end up creating a whole big drama for nothing.

%
All the sadness and drama you have lived in your life was rooted in making assumptions and taking things personally.

Take a moment to consider the truth of this statement.

The whole war of control between humans is about making assumptions and taking things personally.

Our whole dream of hell is based on that.

%
We create a lot of emotional poison just by making assumptions and taking it personally, because usually we start gossiping about our assumptions.

Remember, gossiping is the way we communicate to each other in the dream of hell and transform poison to one another.

Because we are afraid to ask for clarification, we make assumptions, and believe we are right about the assumptions; then we defend our assumptions and try to make someone else wrong.

\textit{It is always better to ask questions than to make an assumption, because assumptions set us up for suffering}.

%
The big \textit{mitote} in the human mind creates a lot of chaos which causes us to misinterpret everything and misunderstand everything.

We only see what we want to see, and hear what we want to hear.

We don't perceive things the way they are.

We have the habit of dreaming with no basis in reality.

We literally dream things up in our imaginations.

Because we don't understand something, we make an assumption about the meaning, and when the truth comes out, the bubble of our dream pops and we find out it was not what we thought it was at all.

%
An example: You are walking in the mall, and you see a person you like.

That person turns to you and smiles, and then walks away.

You can make a lot of assumptions just because of this 1 experience.

With these assumptions you can create a whole fantasy.

And you really want to believe this fantasy and make it real.

A whole dream begins to form just from your assumptions, and you can believe, ``Oh, this person really likes me.''

In your mind a whole relationship begins from that.

Maybe you even get married in this fantasyland.

But the fantasy is in \textit{your} mind, in your personal dream.

%
Making assumptions in our relationships is really asking for problems.

Often we make the assumption that our partners know what we think and that we don't have to say what we want.

We assume they are going to do what we want, because they know us so well.

If they don't do what we assume they should do, we feel so hurt and say, ``You should have known.''

%
Another example: You decide to get married, and you make the assumption that your partner sees marriage the same way that you do.

Then you live together and you find out this is not true.

This creates a lot of conflict, but you still don't try to clarify your feelings about marriage.

The husband comes home from work and the wife is mad, and the husband doesn't know why.

Maybe it's because the wife made an assumption.

Without telling him what she wants, she makes an assumption that he knows her so well, that he knows what she wants, as if he can read her mind.

She gets so upset because he fails to meet her expectations.

Making assumptions in relationships leads to a lot of fights, a lot of difficulties, a lot of misunderstandings with people we supposedly love.

%
In any kind of relationship we can make the assumption that others know what we think, and we don't have to say what we want.

They are going to do what we want because they know us so well.

If they don't do what we want, what we assume they should do, we feel hurt and think, ``How could you do that? You should know.''

Again, we make the assumption that the other person knows what we want.

A whole drama is created because we make this assumption and then put more assumptions on top of it.

%
It is very interesting how the human mind works.

\fbox{\textit{We have the need to justify everything, to explain and understand everything, in order to feel safe}.}

We have millions of questions that need answers because there are so many things that the reasoning mind cannot explain.

\fbox{\textit{It is not important if the answer is correct; just the answer itself makes us feel safe}.}

This is why we make assumptions.

%
If others tell us something, we make assumptions, and if they don't tell us something we make assumptions to fulfill our need to know and to replace the need to communicate.

Even if we hear something and we don't understand, we make assumptions about what it means and then believe the assumptions.

\textit{We make all sorts of assumptions because we don't have the courage to ask questions}.

%
These assumptions are made so fast and unconsciously most of the time because we have agreements to communicate this way.

We have agreed that it is not safe to ask questions; we have agreed that if people love us, they should know what we want or how we feel.

When we believe something, we assume we are right about it to the point that we will destroy relationships in order to defend our position.

%
We make the assumption that everyone sees life the way \textit{we} do.

We assume that others think the way we think, feel the way we feel, judge the way we judge, and abuse the way we abuse.

This is the biggest assumption that humans make.

And this is \textit{why we have a fear of being ourselves around others}.

Because we think everyone else will judge us, victimize us, abuse us, and blame us as we do ourselves.

So even before others have a chance to reject us, we have already rejected ourselves.

That is the way the human mind works.

%
We also make assumptions about ourselves, and this creates a lot of inner conflict.

``I think I am able to do this.''

You make this assumption, e.g., then you discover you aren't able to do it.

You overestimate or underestimate yourself because you haven't taken the time to ask yourself questions and to answer them.

Perhaps you need to gather more facts about a particular situation.

Or maybe you need to stop lying to yourself about what you truly want.

%
Often when you go into a relationship with someone you like, you have to justify why you like that person.

You only see what you want to see and you deny there are things you don't like about that person.

You lie to yourself just to make yourself right.

Then you make assumptions, and 1 of the assumptions is ``My love will change this person.''

But this is no true.

Your love will not change anybody.

If others change, it's because they want to change, not because you can change them.

Then something happens between the 2 of you, and you get hurt.

Suddenly you see what you didn't want to see before, only now it is amplified by your emotional poison.

Now you have to justify your emotional pain and blame them for your choices.

%
We don't need to justify love; it is there or not there.

Real love is accepting other people the way they ware without trying to change them.

If we try to change them, this means we don't really like them.

Of course, if you decide to live with someone, if you make that agreement, it is always better to make that agreement with someone who is exactly the way you want him or her to be.

Find someone whom you don't have to change at all.

It is much easier to find someone who is already the way you want him or her to be, instead of trying to change that person.

Also, that person must love you just the way you are, so he or she doesn't have to change you at all.

If others feel they have to change you, that means they really don't love you just the way you are.

So why be with someone if you're not the way he or she wants you to be?

%
We have to be what we are, so we don't have to present a false image.

If you love me the way I am, ``Okay, take me.''

If you don't love me the way I am, ``Okay, bye-bye.

Find someone else.''

It may sound harsh, but this kind of communication means the personal agreements we make with others are clear and impeccable.

%
Just imagine the day that you stop making assumptions with your partner and eventually with everyone else in your life.

You way of communicating will change completely, and your relationships will no longer suffer from conflicts created by mistaken assumptions.

%
The way to keep yourself from making assumptions is to ask questions.

Make sure the communication is clear.

If you don't understand, ask.

Have the courage to ask questions until you are clear as you can be, and even then do not assume you know all there is to know about a given situation.

Once you hear the answer, you will not have to make assumptions because you will know the truth.

%
Also, find your voice to ask for what you want.

Everybody has the right to tell you no or yes, but you always have the right to ask.

Likewise, everybody has the right to ask you, and you have the right to say yes or no.

%
If you don't understand something, it is better for you to ask and be clear, instead of making an assumption.

The day you stop making assumptions you will communicate cleanly and clearly, free of emotional poison.

Without making assumptions you word becomes impeccable.

%
With clear communication, all of your relationships will change, not only with your partner, but with everyone else.

You won't need to make assumptions because everything becomes so clear.

This is what I want; this is what you want.

If we communicate in this way, our word becomes impeccable.

If all humans could communicate in this way, with impeccability of the word, there would be no wars, no violence, no misunderstandings.

\textit{All human problems would be resolved if we could just have good, clear communication}.

%
This, then, is the 3rd Agreement: \textit{Don't make assumptions}.

Just saying this sounds easy, but I understand that it is difficult to do.

It is difficult because we so often do exactly the opposite.

We have all these habits and routines that we are not even aware of.

Becoming aware of these habits and understanding the importance of this agreement is the 1st step.

But understanding its importance is not enough.

Information or an idea is merely the seed in your mind.

What will really make the difference is action.

Taking the action over and over again strengthens your will, nurtures the seed, and establishes a solid foundation for the new habit to grow.

After many repetitions these new agreements will become 2nd nature, and you will see how the magic of your word transforms you from a black magician into a white magician.

%
A white magician uses the word for creation, giving, sharing, and loving.

By making this 1 agreement a habit, your whole life will be completely transformed.

%
\textit{When you transform your whole dream, magic just happens in your life}.

What you need comes to you easily because spirit moves freely through you.

This is \textit{the mastery of intent, the mastery of the spirit, the mastery of love, the mastery of gratitude}, and \textit{the mastery of life}.

This is the goal of the Toltec.

\textit{This is the path to personal freedom}.

\section{\textsc{The 4th Agreement}: Always Do Your Best}

\section{\textsc{The Toltec Path to Freedom}: Breaking Old Agreements}

\section{\textsc{The New Dream}: Heaven on Earth}

\section{Prayers}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}