\documentclass{book}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/nguyen/1_NQBH/math/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\addtolength{\headheight}{0pt}% obsolete
\lhead{\itshape \scriptsize \chaptername~\thechapter}
\rhead{\itshape \scriptsize \nouppercase{\leftmark}} %\nouppercase !
\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\cfoot{\thepage}

\title{Susan Cain. Quiet: The Power of Introverts in a World That Can't Stop Talking. 2014}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\setcounter{secnumdepth}{6}
\setcounter{tocdepth}{6}
\tableofcontents

%------------------------------------------------------------------------------%

\section*{More advance noise for \textit{quiet}}

\begin{quotation}
    ``\textbf{\textit{An intriguing and potentially life-altering examination of the human psyche that is sure to benefit both introverts and extroverts alike}.}'' - \textit{Kirkus Reviews} (starred review)
\end{quotation}

\begin{quotation}
    ``\textit{Gentle is powerful $\ldots$ Solitude is socially productive $\ldots$ These important counterintuitive ideas are among the many reasons to \textbf{take \emph{Quiet} to a quiet corner and absorb its brilliant, thought-provoking message.}}'' - Rosabeth Moss Kanter, professor at Harvard Business School, author of \textit{Confidence} and \textit{SuperCorp}
\end{quotation}

\begin{quotation}
    ``\textit{\textbf{An informative, well-researched book} on the power of quietness and the virtues of having a rich inner life. It dispels the myth that you have to be extroverted to be happy and successful.}'' - Judith Orloff, M.D., author of \textit{Emotional Freedom}
\end{quotation}

\begin{quotation}
    ``\textit{If this \textbf{engaging and beautifully written} book, Susan Cain makes a powerful case for the wisdom of introspection. She also warns us ably about the downside to our culture's noisiness, including all that it risks drowning out. Above the din, \textbf{Susan's own voice remains a compelling presence-thoughtful, generous, calm, and eloquent. \emph{Quiet} deserves a very large readership.}}'' - Christopher Lane, author of \textit{Shyness: How Normal Behavior Became a Sickness}
\end{quotation}

\begin{quotation}
    ``\textit{Susan Cain's quest to understand introversion, \textbf{a beautifully wrought journey from the lab bench to the motivational speaker's hall}, offers convincing evidence for valuing substance over style, steak over sizzle, and qualities that are, in America, often derided. \textbf{This book is brilliant, profound, full of feeling and brimming with insights.}}'' - Sheri Fink, M.D., author of \textit{War Hospital}
\end{quotation}

\begin{quotation}
    ``\textit{\textbf{Brilliant, illuminating, empowering!} \emph{Quiet} gives not only a voice, but a path to homecoming for so many who've walked through the better part of their lives thinking the way they engage with the world is something in need of fixing.}'' - Jonathan Fields, author of \textit{Uncertainty: Turning Fear and Doubt into Fuel for Brilliance}
\end{quotation}

\begin{quotation}
    ``\textit{Shatters misconceptions$\ldots$ Cain consistently holds the reader's interest by presenting individual profiles $\ldots$ and reporting on the latest studies. \textbf{Her diligence, research, and passion for this important topic has richly paid off.}}'' - Adam M. Grant, Ph.D., associate professor of management, the Wharton School of Business
\end{quotation}

\section*{Still more advance noise for \textit{quiet}}

\begin{quotation}
    ``\textit{\textbf{Once in a blue moon, a book comes along that gives us startling new insights}. \emph{Quiet} is that book: it's part page-turner, part cutting-edge science. The implications for business are especially valuable: \emph{Quiet} offers tips on how introverts can lead effectively, give winning speeches, avoid burnout, and choose the right roles. \textbf{This charming, gracefully written, thoroughly researched book is simply masterful.}}'' - Publishers Weekly
\end{quotation}

\begin{quotation}
    ``\textit{\textbf{Quiet} elevates the conversation about introverts in our outwardly oriented society to new heights. \textbf{I think that many introverts will discover that, even though they didn't know it, they have been waiting for this book all their lives.}}'' - Adam S. McHugh, author of Introverts in the Church
\end{quotation}

\begin{quotation}
    ``\textit{Susan Cain's \emph{Quiet} is wonderfully informative about the culture of the extrovert ideal and the psychology of a sensitive temperament, and she is helpfully perceptive about how introverts can make the most of their personality preferences in all aspects of life. \textbf{Society needs introverts, so everyone can benefit from the insights in this important book.}}'' - Jonathan M. Cheek, professor of psychology at Wellesley College, co-editor of \textit{Shyness: Perspectives on Research and Treatment}
\end{quotation}

\begin{quotation}
    ``\textit{\textbf{A brilliant, important, and personally affecting book}. Cain shows that, for all its virtue, America's Extrovert Ideal takes up way too much oxygen. \textbf{Cain herself is the perfect person to make this case - with winning grace and clarity she shows us what it looks like to think outside the group.}}'' - Christine Kenneally, author of \textit{The First Word}
\end{quotation}

\begin{quotation}
    ``\textit{What Susan Cain understands - and readers of this fascinating volume will soon appreciate - is something that psychology and our fast-moving and fast-talking society have been all too slow to realize: \textbf{Not only is there really nothing wrong with being quiet, reflective, shy, and introverted, but there are distinct advantages to being this way.}} - Jay Belsky, Robert M. and Natalie Reid Dorn Professor, Human and Community Development, University of California, Davis
\end{quotation}

\begin{quotation}
    ``\textit{\textbf{Author Susan Cain exemplifies her own quiet power in this exquisitely written and highly reader page-turner.} She brings important research and poignant, personal examples into the light, greatly deepening our understanding of the introvert experience.''} - Jennifer B. Kahnweiler, Ph.D., author of \textit{The Introverted Leader}
\end{quotation}

%------------------------------------------------------------------------------%

\begin{quotation}
    ``\textit{}''
\end{quotation}

\part{}

\chapter{}

\section{}


%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}